### 2021-11-25 code/R/Analysis.Rmd
Added readme.

---
title: "Analysis"
author: "Hjalte Mann"
date: "25/11/2021"
output: html_document
---
---
title: "IndividualPhenology"
author: "Hjalte Mann"
date: "2/2/2021"
output: html_document
---

```{r setup, include=FALSE}
library(ggplot2)
library(dplyr)
library(stringr)
knitr::opts_chunk$set(echo = TRUE)
```

Import ground truth tracked data

```{r}
thul01 = read.csv("../../data/annotations/2020_05_15_NorwayAnnotations_THUL-01_IndividualAnnotations_FRCNN.csv", header = F)

colnames(thul01) = c("FileName", "x_min", "y_min", "x_max", "y_max", "Class")

thul01 = thul01 %>%
  mutate(FileName = str_extract(FileName, regex("[A-Z]{4,5}-\\d{0,3}_\\d{6}.JPG")))

nars04 = read.csv("../../data/annotations/2020_05_17_NorwayAnnotations_NARS-04_IndividualAnnotations_FRCNN.csv", header = F)

colnames(nars04) = c("FileName", "x_min", "y_min", "x_max", "y_max", "Class")

nars04 = nars04 %>%
  mutate(FileName = str_extract(FileName, regex("[A-Z]{4,5}-\\d{0,3}_\\d{6}.JPG")))

nars13 = read.csv("../../data/annotations/2020_04_30_NorwayAnnotations_NARS-13_IndividualAnnotations_FRCNN.csv", header = F)

colnames(nars13) = c("FileName", "x_min", "y_min", "x_max", "y_max", "Class")

nars13 = nars13 %>%
  mutate(FileName = str_extract(FileName, regex("[A-Z]{4,5}-\\d{0,3}_\\d{6}.JPG")))

nyaa04 = read.csv("../../data/annotations/2020_05_23_NorwayAnnotations_Nyaa-04_IndividualAnnotations_FRCNN.csv", header = F)

colnames(nyaa04) = c("FileName", "x_min", "y_min", "x_max", "y_max", "Class")

nyaa04 = nyaa04 %>%
  mutate(FileName = str_extract(FileName, regex("[A-Z]{4,5}-\\d{0,3}_\\d{6}.JPG")))

```

Import metadata for the 1-h subsets

```{r}

meta_2018 = read.csv("../../data/annotations/DetectionPhenology_2018_Meta_Clean_Prepped.csv")

thul01_meta = merge(thul01, meta_2018, by.x = "FileName")

thul01_meta$time = thul01_meta$DOY + (thul01_meta$Hour/24)

thul01_meta %>% 
ggplot()+
  geom_point(aes(x = time, y = Class))

thul01_meta = thul01_meta %>%
  group_by(Class) %>% 
  mutate(flower_start = min(time),
    flower_length = max(time) - min(time))

thul01_meta %>% 
ggplot()+
  geom_point(aes(x = flower_start, y = flower_length))

nars04_meta = merge(nars04, meta_2018, by.x = "FileName")

nars04_meta$time = nars04_meta$DOY + (nars04_meta$Hour/24)

nars04_meta %>% 
ggplot()+
  geom_point(aes(x = time, y = Class))

nars04_meta = nars04_meta %>%
  group_by(Class) %>% 
  mutate(flower_start = min(time),
    flower_length = max(time) - min(time))

nars04_meta %>% 
ggplot()+
  geom_point(aes(x = flower_start, y = flower_length))

meta_2019 = read.csv("../../data/annotations/DetectionPhenology_2019_Meta_Clean_Prepped.csv")

nars13_meta = merge(nars13, meta_2019, by.x = "FileName")

nars13_meta$time = nars13_meta$DOY + (nars13_meta$Hour/24)

nars13_meta %>% 
ggplot()+
  geom_point(aes(x = time, y = Class))

nars13_meta = nars13_meta %>%
  group_by(Class) %>% 
  mutate(flower_start = min(time),
    flower_length = max(time) - min(time))

nars13_meta %>% 
ggplot()+
  geom_point(aes(x = flower_start, y = flower_length))

```

```{r}

nyaa04_meta = merge(nyaa04, meta_2019, by.x = "FileName")

nyaa04_meta$time = nyaa04_meta$DOY + (nyaa04_meta$Hour/24)

nyaa04_meta %>% 
ggplot()+
  geom_point(aes(x = time, y = Class))

nyaa04_meta = nyaa04_meta %>%
  group_by(Class) %>% 
  mutate(flower_start = min(time),
    flower_length = max(time) - min(time))

nyaa04_meta %>% 
ggplot()+
  geom_point(aes(x = flower_start, y = flower_length))

```

```{r}

all = bind_rows(thul01_meta[,c("flower_start","flower_length", "CameraID")], nars04_meta[,c("flower_start","flower_length", "CameraID")], nars13_meta[,c("flower_start","flower_length", "CameraID")], nyaa04_meta[,c("flower_start","flower_length", "CameraID")])

all %>% 
ggplot()+
  geom_point(aes(x = flower_start, y = flower_length, colour = CameraID)) +
geom_smooth(aes(x = flower_start, y = flower_length), method=lm)

t.test(nyaa04_meta$flower_length, nars13_meta$flower_length)

```

```{r}
all %>% 
ggplot()+
  geom_point(aes(x = flower_start, y = flower_length, colour = CameraID))
```

### 2021-11-25 text/ManuscriptDraft.Rmd
Added readme.

---
new_session: FALSE
title: |
  **Monitoring the phenology of individual flowers using deep learning and automatic tracking**
author:
- Hjalte M. R. Mann
- Alexandros Iosifidis
- Toke T. Høye
date: "`r format(Sys.time(), '%B %d, %Y')`"
output:
  word_document: default
  pdf_document:
    fig_caption: yes
    keep_tex: yes
    latex_engine: pdflatex
  html_document:
    df_print: paged
fontfamily: mathpazo
fontsize: 12pt
geometry: margin = 1in
header-includes:
- \usepackage{setspace}\doublespacing
- \usepackage[left]{lineno}
- \linenumbers
- \usepackage{dcolumn}
- \usepackage{caption}
- \usepackage{float}
- \usepackage{afterpage}
- \usepackage{siunitx}
- \usepackage{amsmath}
keywords: Citizen science, Arctic, insects
bibliography: ./library.bib
csl: ./journal-of-ecology.csl
abstract: ABSTRACT | Often simple variables will be used to describe the flowering phenology of a population of plants, e.g. onset or peak of flowering and to infer respones to climate change. Here we show that image-based monitoring of field plots at very high temporal resolution can return information on flowering phenology at the level of indiviuals. Further, we present an automatic flower tracking algorithm.
---

```{r eval = FALSE, echo=FALSE}
library(knitr)
knit('ManuscriptDraft.Rmd')
rmarkdown::render("ManuscriptDraft.Rmd", output_file = "ManuscriptDraft.docx")
system2("open","ManuscriptDraft.docx")
```

\newpage
**NOTES**

What questions do we want to answer?

*Flower information*

\pagebreak

# Introduction

The flowering phenology of a population may mask responses at the individual level. For example,

Does flower visitation rates and/or reproductive success depend on the timing of flowering for the indiviual flower?

# Material and methods

## Study site

## The image series

## Flower annotations

# Results

**Figure 1:** Figure text....

# Discussion

## Flower phenology

# Acknowledgements

# Data availability

The code that supports the results in this paper will be made openly available at https://github.com/TECOLOGYxyz. Raw data as well as the trained flower detection model will be archived on https://zenodo.org/.

### 2021-11-25 FlowerCentroidTracker_Object_v1.py

@@ -66,9 +66,11 @@ class tracker():
         self.disappeared[self.nextObjectID] = 0 # Set number of times the new object has disappeared to zero. 
         self.nextObjectID += 1 # Add 1 to the objectID counter so it's ready for the next point
 
+        self.length_dict = {key: len(value) for key, value in self.objects.items()}
+
     def update(self, frame):
         frame_detections = self.get_frame_detections(frame)#  Get the detections for the current frame
-        print("FRAME: ", frame)
+        print(f'FRAME {frame}. Contains {len(frame_detections)} points.')
         """
         If a frame has no detections, we +1 to disappeared of all objects being tracked. 
         If this takes any objects above the max_disappeared threshold, we remove them from the objects.
@@ -93,16 +95,19 @@ class tracker():
         
         inputCentroids = frame_detections.to_numpy()
 
-
         if not self.objects: # if Objects is empty, we are currently not tracking any objects and take the input centroids and register each of them
             print("Not tracking objects. Initiating tracking on current detections.")
             for i in range(0, len(inputCentroids)):
                 print("Added detection ", i)
                 self.register(inputCentroids[i])
+
             print("Current objects: ", self.objects)
-        
+            print("Object 0: ",self.objects[0])
+            print("Current lengths: ", self.length_dict)
+            
+            
         else: # We are already tracking objects, so let's see if we can associate any frame detections with objects that are being tracked.
-            print("Hello")    
+            print(br,"We are tracking existing objects.")    
             objectIDs = list(self.objects.keys()) # grab the set of object IDs and corresponding centroids
             objectCentroids = list(self.objects.values())
             
@@ -114,7 +119,11 @@ class tracker():
 
             print("Distance matrix: \n", D)
             
-            # D_passed = D[D.distance <= max_distance]
+            rows = D.min(axis=1).argsort()
+            print(rows)
+            # For 
+            
+            #D_passed = D[D.distance <= max_distance]
             # print(D_passed)
 #             D_passed = D_passed.loc[D_passed.groupby('id').distance.idxmin()] # If the there are more than one frame point that is close to an object point, use the one closest. TO-DO: What happens if two frame points have the same distance to the object point?
 #             print("Frame points and closest object point, below threshold: \n:", D_passed)
@@ -172,8 +181,8 @@ class tracker():
[…]

### 2021-11-25 MOTA.py

@@ -1,12 +1,0 @@
-# -*- coding: utf-8 -*-
-"""
-Created on Wed Jan  6 12:22:49 2021
-
-@author: au309263
-"""
-
-import pandas as pd
-
-
-dat = pd.read_csv("MOTA_DummyData.csv")
-

### 2021-11-30 FlowerCentroidTracker_Object_v1.py
Added hand drawn figures

@@ -1,0 +1,226 @@
+# -*- coding: utf-8 -*-
+"""
+Created on Thu Feb  4 16:36:42 2021
+
+
+
+Custom flower tracker
+
+Multiobject tracker tracker
+
+Max distance, max dissapeared, running mean
+
+
+Detections are in the format: 
+
+@author:
+Hjalte Mann
+TECOLOGY.xyz
+
+"""
+import pandas as pd # Replace with cudf if performance is too slow?
+from collections import OrderedDict
+import numpy as np
+from scipy.spatial import distance as dist
+#from scipy.ndimage.filters import uniform_filter1d
+import matplotlib.pyplot as plt
+import time
+
+br = "\n"
+
+"""
+TO-DO
+
+- Record tracks and output a result file
+- Create running mean solution
+- Plot results
+- Create option to disregard max_distance (to use on non-fixed objects)
+
+"""
+
+
+#### SETTINGS ####
+
+resultFilename = "trackResults.csv"
+
+max_disappeared = 4 # Maximum number of frames the algorithm should continue to look for an object
+max_distance = 200 # Maximum distance to a point before it is forced to be registered as a new ID instead of associated with the closest point
+running_mean_threshold = 5 # Number of frames for calculating the running mean of the position of an object
+
+######
+
+
+#### PATH TO DETECTIONS ####
+detections = pd.read_csv('Dummy_fortracking2.csv')
+detections['x_c'] = (detections['x_min'] + detections['x_max']) / 2
+detections['y_c'] = (detections['y_min'] + detections['y_max']) / 2
+
+frames = list(set(detections['frame'].tolist()))
+
+print("Here are the detections",br,detections,br)
+
+
+with open(resultFilename, 'a') as resultFile:
+    header = 'frame,filename,x_c,y_c,objectID\n'
+    resultFile.write(header)  
+
+
+class tracker():
+    def __init__(self, max_disappeared, frames):
+        self.nextObjectID = 0 # Counter for object ids
+        self.objects = OrderedDict() # Dictionary. objectID is the key, centroid is the content
+        self.disappeared = OrderedDict() # Keeps track of how long an objectID has been lost
+
+        self.max_disappeared = max_disappeared # store max_disappeared number of frames
+        self.frames = frames
+        
+        self.tracks = []
+        
+    def store_tracking_results(self, frame, centroid, objectID):
+        #print(br, f'Object ID {objectID} with centroid {centroid} in frame {frame} stored.', br)
[…]

### 2021-11-30 text/ManuscriptDraft.Rmd
Updated manuscript

@@
 
+We built a framework for tracking, filtering, and evaluating tracking of objects in time-lapse image series.
+
+# Automatic tracking
+
+Our algorithm tracks objects based on distances between centroids of bounding boxes.
+
+The tracking algorithm has a set of user adjusted parameters that can optimise tracking accuracy.
+
+The parameters are particularly relevant for optimal tracking of objects that are constrained to a specific area such as flowers.
+
+It is important to note, however, that the tracking algorithm can be used to track any objects.
+
+The tracking algorithm can be applied both offline (on a set of detections/annotations that have already been produced) or online (realtime tracking frame per frame).
+
+### User parameters
+
+**Max distance threshold**, **running mean**, **max disappeared**,
+
+As the wind shifts, the flower heads changes direction.
+This can happen instantaneously (i.e.
+between two censecutive frames).
+As they are constrained by their stalk, there is a limit to the distance they can move.
+
+Establishing associations between points based on just the distance between points in the current and the previous frame can cause errors when flowers are in close vicinity of each other.
+
+The flowers move around a center point because of their stalk.
+We base the tracking on the distance between a point in the current frame and the running mean of the positions of the previous X points in a track.
+
+# Evaluating tracking perfomance
+
+Mota counts shifts in tracking.
+
+To derive flowering length, in theory we just need to track the most extreme points correctly and don't care about other points (although we filter by length when overlap).
+
+To associate other information to the flower, for example flower visits, we want as much as possible of the track to be correct.
+
+## Filtering tracks
+
+Tracks that overlap have significant risk of errors.
+Overlapping tracks can be caused by a single flower that was erroneously assigned to several tracks, two flowers that were located sufficiently close to each other that there areas overlapped (e.g.
+when wind moves the flowers around), false positive detections close to a flower.Best case is two flowers that flowered in the same area but were separated by time.
+Here we will remove overlapping tracks to reduce the risk of error.
+
+We disregard single points that were not associated to a track.
+For tracks consisting of two points, we establish the straight line between the points.
+For tracks consisting of three points, we establish the triangle from the points.
+For tracks consisting of more than three points, we calculate the convex hull of all the points included in the track and derive the polygon from the vertices of the convex hull.
+
+We then check if any two lines intersect (for tracks with two points), if any lines intersect with any polygons, and if any polygons overlap with other polygons, and remove tracks that overlap.
+
+We evaluate the accuracy of the remaining tracks.
+
+![](../figures/figure_1.png){ width=100% }
+
+**Figure 1:** Simple centroid tracking may produce erronous associations when objects move between frames.
+Blue shows detections in the current frame (bounding box and centroid point).
+Red shows centroid points for the detections in the previous frame.
+
 # Results
-
-**Figure 1:** Figure text....
 

### 2021-11-30 FlowerCentroidTracker_Object_v1.py
Added solution for standardizing coordinates based on image dimensions

@@ -44,14 +44,18 @@ TO-DO
 resultFilename = "trackResults.csv"
 
 max_disappeared = 4 # Maximum number of frames the algorithm should continue to look for an object
-max_distance = 200 # Maximum distance to a point before it is forced to be registered as a new ID instead of associated with the closest point
+max_distance = 1500 # Maximum distance to a point before it is forced to be registered as a new ID instead of associated with the closest point
 running_mean_threshold = 5 # Number of frames for calculating the running mean of the position of an object
 
 ######
 
 
 #### PATH TO DETECTIONS ####
-detections = pd.read_csv('Dummy_fortracking2.csv')
+#detections = pd.read_csv('Dummy_fortracking2.csv')
+detections = pd.read_csv(r"U:\BITCue\Projekter\TrackingFlowers\data\annotations\2020_04_30_NorwayAnnotations_NARS-13_IndividualAnnotations_FRCNN_Metrics.csv")
+
+detections['frame'] = detections['filename'].str.extract('(\d{6})')
+
 detections['x_c'] = (detections['x_min'] + detections['x_max']) / 2
 detections['y_c'] = (detections['y_min'] + detections['y_max']) / 2
 
@@ -90,7 +94,7 @@ class tracker():
                 x_c = t[1]
                 y_c = t[2]
                 objectID = t[3]
-                filename = detections.loc[detections['frame'] == 1, 'filename'].iloc[0]
+                filename = detections.loc[detections['frame'] == frame, 'filename'].iloc[0]
                 resultFile.write(f'{frame},{filename},{x_c},{y_c},{objectID}{br}')
         endtime = time.time()
         print(f'Writing done. That took {round(endtime-starttime, 4)} seconds.')
@@ -209,6 +213,12 @@ class tracker():
         endtime = time.time()
         print(f'Tracking done. That took {round(endtime-starttime, 4)} seconds.')
 
+
+
+
+
+
+
 ### RUN ###
 t = tracker(max_disappeared, frames) # Instantiate the class instance and pass in the threshold for max_disappeared and the list of frames.
 
@@ -218,9 +228,14 @@ t.write_tracks_file()
 
 
 ### PLOT STUFF ###
-plt.scatter(detections['x_c'],detections['y_c'], c = detections['frame'])
-plt.plot(detections['x_c'],detections['y_c'])
[…]

### 2021-11-30 filter.py
First version of the track filtering solution

@@ -1,0 +1,256 @@
+# -*- coding: utf-8 -*-
+"""
+Created on Sun Nov 28 20:14:12 2021
+
+@author: au309263
+"""
+from shapely.geometry import Polygon
+import pandas as pd # Replace with cudf if performance is too slow?
+from collections import OrderedDict
+import numpy as np
+from scipy.spatial import distance as dist
+from scipy.spatial import ConvexHull
+#from scipy.ndimage.filters import uniform_filter1d
+import matplotlib.pyplot as plt
+import time
+
+br = "\n"
+
+
+
+"""
+TO-DO
+
+- Remmove single points
+- Establish straight line between two-point tracks
+- Establish triangle between three-point tracks
+- Establish polygon from vertices of the convex hull of n-point tracks (n > 3)
+- Check if any lines overlap and remove both if they do
+- Check if any remaining lines overlap with any polygons, remove both line and polygon if so
+- Check if any polygons overlap and remove both if they do
+- Output filtered track file
+
+"""
+
+#### PATH TO TRACKS ####
+tracks = pd.read_csv('trackResults.csv')
+print(tracks)
+
+#### NORMALIZE X AND Y ####
+#tracks['x_c'] = tracks['x_c']/6080
+#tracks['y_c'] = tracks['y_c']/3420
+
+
+#### FUNCTIONS ####
+def convex_hull(points): # Calculate the convex hull of points
+    print("Calculating convex hulls")
+    #points = np.array([[1, 2],[1, 4],[4, 1], [2,2.5], [3.5,3.5]])
+    hulls = []
+    for p in points:
+        #print(p)
+        p = np.array(p)
+        print("Points: ", p)
+        hull = ConvexHull(p)
+        #print("Hull vertices: ", p[hull.vertices])
+        hulls.append(p[hull.vertices])
+        
+    return hulls
+
+def cart_product(listOfPolygons):
+    print("Calculating cartesian products of polygons")
+    cartProduct = []
+    #print("List of polygons: ",br, listOfPolygons)
[…]

### 2021-12-05 code/analyze.py
Added code for analyzing tracks

@@ -1,0 +1,75 @@
+# -*- coding: utf-8 -*-
+"""
+Created on Wed Dec  1 15:08:22 2021
+
+@author:
+    Hjalte M. R. Mann
+    TECOLOGY.xyz
+
+In this script we'll do some analyses of the ground truth individual annotations.
+We'll use the results for setting the parameters of our tracking algorithm.
+
+"""
+
+import pandas as pd
+import numpy as np
+
+# Input ground truth annotations
+df = pd.read_csv(r"data\annotations\2020_05_23_NorwayAnnotations_NYAA-04_IndividualAnnotations_FRCNN_Metrics.csv")
+
+#print(thul01)
+
+df['x_c'] = (df['x_min'] + df['x_max']) / 2
+df['y_c'] = (df['y_min'] + df['y_max']) / 2
+
+
+#print(thul01)
+
+
+a = np.array([3.4,4])
+b = np.array([6,9])
+
+def dist(a,b):
+    d = np.linalg.norm(a-b)
+    return d
+
+#print(dist)
+
+df_grouped = df.groupby(['id_gt'])
+
+largestDist = 0
+
+for i, g in df_grouped:
+    #print(g)
+    g_points = []
+    for i,l in g.iterrows():
+        g_points.append(np.array([l['x_c'], l['y_c']]))
+        
+    for p1 in g_points:
+        for p2 in g_points:
+            d = dist(p1, p2)
+            if d > largestDist:
+                print(f'{d} is larger than {largestDist}. Storing {d}')
+                print(f'Distance calculated between {p1} and {p2}')
+                largestDist = d
+            
+            
+print(f'Largest dist found: {largestDist}')
+
+# THUL-01: 449.00334074481003
+# NYAA-04: 2307.41034495384
+# NARS-13: 782.0206199327483
+# NARS-04: 557.953851138246
+
+    
+
+# Find the largest distance between two points in a track. This will help us set max_distance in the tracking algorithm.
+
+
+# Find the largest number of frames a track was lost and then regained. This will help us set max_disappeared in the tracking algorithm.
+
+
+# 
+
+
+# 

### 2021-12-05 code/filter.py
Started code for filtering tracks based on polygon overlap

@@ -1,0 +1,261 @@
+# -*- coding: utf-8 -*-
+"""
+Created on Sun Nov 28 20:14:12 2021
+
+@author: au309263
+"""
+from shapely.geometry import Polygon
+import pandas as pd # Replace with cudf if performance is too slow?
+from collections import OrderedDict
+import numpy as np
+from scipy.spatial import distance as dist
+from scipy.spatial import ConvexHull
+#from scipy.ndimage.filters import uniform_filter1d
+import matplotlib.pyplot as plt
+import time
+
+br = "\n"
+
+
+
+"""
+TO-DO
+
+- Remmove single points
+- Establish straight line between two-point tracks
+- Establish triangle between three-point tracks
+- Establish polygon from vertices of the convex hull of n-point tracks (n > 3)
+- Check if any lines overlap and remove both if they do
+- Check if any remaining lines overlap with any polygons, remove both line and polygon if so
+- Check if any polygons overlap and remove both if they do
+- Output filtered track file
+
+"""
+
+#### PATH TO TRACKS ####
+tracks = pd.read_csv(r'testResults/trackResults7.csv')
+print(tracks)
+
+#### NORMALIZE X AND Y ####
+#tracks['x_c'] = tracks['x_c']/6080
+#tracks['y_c'] = tracks['y_c']/3420
+
+
+#### FUNCTIONS ####
+def convex_hull(points): # Calculate the convex hull of points
+    print("Calculating convex hulls")
+    #points = np.array([[1, 2],[1, 4],[4, 1], [2,2.5], [3.5,3.5]])
+    hulls = []
+    for p in points:
+        #print(p)
+        p = np.array(p)
+        print("Points: ", p)
+        hull = ConvexHull(p)
+        #print("Hull vertices: ", p[hull.vertices])
+        hulls.append(p[hull.vertices])
+        
+    return hulls
+
+def cart_product(listOfPolygons):
+    print("Calculating cartesian products of polygons")
+    cartProduct = []
+    #print("List of polygons: ",br, listOfPolygons)
[…]

### 2021-12-05 code/track.py
Renamed script for tracking

@@ -1,0 +1,250 @@
+# -*- coding: utf-8 -*-
+"""
+Created on Thu Feb  4 16:36:42 2021
+
+
+
+Custom flower tracker
+
+Multiobject tracker tracker
+
+Max distance, max dissapeared, running mean
+
+
+Detections are in the format: 
+
+@author:
+Hjalte Mann
+TECOLOGY.xyz
+
+"""
+import pandas as pd # Replace with cudf if performance is too slow?
+from collections import OrderedDict
+import numpy as np
+from scipy.spatial import distance as dist
+#from scipy.ndimage.filters import uniform_filter1d
+import matplotlib.pyplot as plt
+import time
+
+br = "\n"
+
+"""
+TO-DO
+
+- Record tracks and output a result file
+- Create running mean solution
+- Plot results
+- Create option to disregard max_distance (to use on non-fixed objects)
+
+"""
+
+
+#### SETTINGS ####
+
+resultFilename = "testResults/trackResults7.csv"
+
+max_disappeared = 4 # Maximum number of frames the algorithm should continue to look for an object
+max_distance = 500 # Maximum distance to a point before it is forced to be registered as a new ID instead of associated with the closest point
+running_mean_threshold = 5 # Number of frames for calculating the running mean of the position of an object
+
+######
+
+
+#### PATH TO DETECTIONS ####
+#detections = pd.read_csv('Dummy_fortracking2.csv')
+detections = pd.read_csv(r"U:\BITCue\Projekter\TrackingFlowers\data\annotations\2020_05_23_NorwayAnnotations_NYAA-04_IndividualAnnotations_FRCNN_Metrics.csv")
+
+detections['frame'] = detections['filename'].str.extract('(\d{6})')
+
+detections['x_c'] = (detections['x_min'] + detections['x_max']) / 2
+detections['y_c'] = (detections['y_min'] + detections['y_max']) / 2
+
+detections['frame'] = detections['frame'].astype('int')
+
+frames = list(set(detections['frame'].tolist()))
+frames = sorted([int(i) for i in frames])
+print(frames)
+
+
+print("Here are the detections",br,detections,br)
+
+
+with open(resultFilename, 'a') as resultFile:
+    header = 'frame,filename,x_c,y_c,objectID\n'
+    resultFile.write(header)  
+
+
+class tracker():
+    def __init__(self, max_disappeared, frames):
+        self.nextObjectID = 0 # Counter for object ids
[…]

### 2021-12-05 code/visualize.py
Started code for visualizing tracking

@@ -1,0 +1,547 @@
+# # -*- coding: utf-8 -*-
+# """
+# Created on Wed Dec  1 19:40:45 2021
+
+# @author: au309263
+# """
+# # -*- coding: utf-8 -*-
+# """
+# Created on Thu Nov  7 11:07:43 2019
+
+# @author: Hjalte Mann
+
+# This script stitches images in a folder together to a time-lapse video
+# """
+
+# import cv2
+# import os
+# from PIL import Image
+# from PIL.ExifTags import TAGS
+# from pathlib import Path
+
+# rootdir = r'U:\BITCue\Projekter\TrackingFlowers\data\images'
+
+# #Set parameters for writing text on image
+# font                   = cv2.FONT_HERSHEY_DUPLEX
+# bottomLeftCornerOfText = (20,20)
+# fontScale              = 0.4
+# fontColor              = (255,255,255)
+# lineType               = 0
+# rectangle_bgr = (0, 0, 0)
+
+# #A function that retrieves exif information from an image
+# def get_exif(fn):
+#     ret = {}
+#     i = Image.open(fn)
+#     info = i._getexif()
+#     for tag, value in info.items():
+#         decoded = TAGS.get(tag, tag)
+#         ret[decoded] = value
+#     return ret
+
+# #A function that stitches images in a folder together to a time-lapse video and writes capture time from metadata and filepath on each image
+# def stitch(image_folder):
+# 	images = [img for img in os.listdir(image_folder) if img.endswith(".JPG")]
+# 	images.sort(key = lambda x: int(x[-10:-4])) # Sort list of images according to value in filename
+# 	frame = cv2.imread(os.path.join(image_folder, images[0])) # Get the first image
+# 	dim = frame.shape # Get the dimensions of the first image
+# 	height, width, layers = frame.shape
+# 	video_name = str(image_folder) + '_TimeLapse6.avi'
+# 	video = cv2.VideoWriter(video_name, 0, 5, (int(width/7),int(height/7)))
+# 	x = int(dim[1]/7)
+# 	y = int(dim[0]/7)
+
+# 	for image in images:
+# 		img_path = os.path.join(image_folder, image)
+# 		image = cv2.imread(img_path)
+# 		imager = cv2.resize(image, (x,y))
+# 		meta = get_exif(img_path)
+
+# 		DTO = meta['DateTimeOriginal']
+# 		text = str(DTO) + str(img_path)
[…]

### 2021-12-05 code/R/Analysis.Rmd
Added images to gitignore and renamed scripts

@@
 ```{r}
+thul01$x_c = (thul01$x_min + thul01$x_max) / 2
+thul01$y_c = (thul01$y_min + thul01$y_max) / 2
+
+thul01 %>% 
+  ggplot() +
+  geom_point(aes(x = x_c, y = y_c, color = Class)) + 
+  xlim(0,6080) +
+  ylim(0,3420)
+
+nyaa04$x_c = (nyaa04$x_min + nyaa04$x_max) / 2
+nyaa04$y_c = (nyaa04$y_min + nyaa04$y_max) / 2
+
+nyaa04 %>% 
+  ggplot() +
+  geom_point(aes(x = x_c, y = y_c, color = Class), size = 1) + 
+  xlim(0,6080) +
+  ylim(0,3420) + 
+  theme(legend.position = "none")
+
+```
+
+```{r}
+tracksThul01 = read.csv('../../trackResults4.csv',stringsAsFactors = FALSE)
+
+tracksThul01$x_c = as.numeric(tracksThul01$x_C)
+
+tracksThul01 %>% 
+  ggplot() +
+  geom_point(aes(x = x_c, y = y_c, color = as.factor(objectID))) + 
+  xlim(0,6080) +
+  ylim(0,3420)
+```
+
+```{r}
 

### 2021-12-05 EvaluatingMLPerformance_SingleClass_NewPredictionFormat_Tracking_v0.2.py

@@ -1,294 +1,0 @@
-# -*- coding: utf-8 -*-
-"""
-Created on Mon Feb 10 15:48:25 2020
-
-Evaluating ML performance - SingleClass
-
-@author: Hjalte Mann, AU
-"""
-
-
-import pandas as pd
-
-iou_threshold = 0.5
-
-# Get the ground truth annotations on the format [filename, x_min, y_min, x_max, y_max, id_gt]
-gt = pd.read_csv(r"U:\BITCue\Projekter\NorwayAnnotations\Experiments\2020_06_30_DetectionPhenology\Rmd\ModelQ_Results\DT_GT\2020_02_01_NorwayAnnotations_AllTestSeries_IndividualAnnotations_FRCNN_Metrics.csv", sep=",")
-#print(gt)
-# Get the tracked detections on the format [filename, x_min, y_min, x_max, y_max, id_dt]
-dt = pd.read_csv(r"U:\BITCue\Projekter\NorwayAnnotations\Experiments\2020_06_30_DetectionPhenology\Rmd\ModelQ_Results\DT_GT\submit_boxes_NewFormat_AllTestSeries_Sampled.csv", sep=",")
-
-print(dt)
-
-#dt['x_min'] = dt['x_min'] * 8
-#dt['y_min'] = dt['y_min'] * 8
-#dt['x_max'] = dt['x_max'] * 8
-#dt['y_max'] = dt['y_max'] * 8
-
-#print(dt)
-
-
-
-P = 0 # Positives (meaning total number of objects (e.g. flowers) across all images)
-CP = 0 # Correct positives (meaning total number of objects that were correclty detected)
-FP = 0 # False positives
-FN = 0 # False negatives
-MM = 0 # Tracking mismatches
-
-
-
-#
-##
-###
-####
-#Make sure that the rows/images are in the correct ascending order by ordering by the image number in the filename
-
-gt['frame'] = gt['filename'].str.extract('(\d{6})')
-dt['frame'] = dt['filename'].str.extract('(\d{6})')
-
-gt = gt.sort_values('frame')
-dt = dt.sort_values('frame')
-
-#print(gt)
-#print(dt)
-
-
-
-####
-###
-##
-#
-
-
-#
-##
-###
-####
-def calculate_iou(boxA, boxB):
-	"""
-	Calculates the intersection over union between two bounding boxes. The format of a box is: [x_min, y_min, x_max, y_max]
-	"""
[…]

### 2021-12-05 FlowerCentroidTracker_Object_v1.py

@@ -1,250 +1,0 @@
-# -*- coding: utf-8 -*-
-"""
-Created on Thu Feb  4 16:36:42 2021
-
-
-
-Custom flower tracker
-
-Multiobject tracker tracker
-
-Max distance, max dissapeared, running mean
-
-
-Detections are in the format: 
-
-@author:
-Hjalte Mann
-TECOLOGY.xyz
-
-"""
-import pandas as pd # Replace with cudf if performance is too slow?
-from collections import OrderedDict
-import numpy as np
-from scipy.spatial import distance as dist
-#from scipy.ndimage.filters import uniform_filter1d
-import matplotlib.pyplot as plt
-import time
-
-br = "\n"
-
-"""
-TO-DO
-
-- Record tracks and output a result file
-- Create running mean solution
-- Plot results
-- Create option to disregard max_distance (to use on non-fixed objects)
-
-"""
-
-
-#### SETTINGS ####
-
-resultFilename = "trackResults4.csv"
-
-max_disappeared = 3 # Maximum number of frames the algorithm should continue to look for an object
-max_distance = 500 # Maximum distance to a point before it is forced to be registered as a new ID instead of associated with the closest point
-running_mean_threshold = 5 # Number of frames for calculating the running mean of the position of an object
-
-######
-
-
-#### PATH TO DETECTIONS ####
-#detections = pd.read_csv('Dummy_fortracking2.csv')
-detections = pd.read_csv(r"U:\BITCue\Projekter\TrackingFlowers\data\annotations\2020_05_15_NorwayAnnotations_THUL-01_IndividualAnnotations_FRCNN_Metrics.csv")
-
-detections['frame'] = detections['filename'].str.extract('(\d{6})')
-
-detections['x_c'] = (detections['x_min'] + detections['x_max']) / 2
-detections['y_c'] = (detections['y_min'] + detections['y_max']) / 2
-
-detections['frame'] = detections['frame'].astype('int')
-
-frames = list(set(detections['frame'].tolist()))
-frames = sorted([int(i) for i in frames])
-print(frames)
-
-
-print("Here are the detections",br,detections,br)
-
-
-with open(resultFilename, 'a') as resultFile:
-    header = 'frame,filename,x_c,y_c,objectID\n'
-    resultFile.write(header)  
-
-
-class tracker():
-    def __init__(self, max_disappeared, frames):
-        self.nextObjectID = 0 # Counter for object ids
-        self.objects = OrderedDict() # Dictionary. objectID is the key, centroid is the content
[…]

### 2021-12-05 filter.py

@@ -1,256 +1,0 @@
-# -*- coding: utf-8 -*-
-"""
-Created on Sun Nov 28 20:14:12 2021
-
-@author: au309263
-"""
-from shapely.geometry import Polygon
-import pandas as pd # Replace with cudf if performance is too slow?
-from collections import OrderedDict
-import numpy as np
-from scipy.spatial import distance as dist
-from scipy.spatial import ConvexHull
-#from scipy.ndimage.filters import uniform_filter1d
-import matplotlib.pyplot as plt
-import time
-
-br = "\n"
-
-
-
-"""
-TO-DO
-
-- Remmove single points
-- Establish straight line between two-point tracks
-- Establish triangle between three-point tracks
-- Establish polygon from vertices of the convex hull of n-point tracks (n > 3)
-- Check if any lines overlap and remove both if they do
-- Check if any remaining lines overlap with any polygons, remove both line and polygon if so
-- Check if any polygons overlap and remove both if they do
-- Output filtered track file
-
-"""
-
-#### PATH TO TRACKS ####
-tracks = pd.read_csv('trackResults.csv')
-print(tracks)
-
-#### NORMALIZE X AND Y ####
-#tracks['x_c'] = tracks['x_c']/6080
-#tracks['y_c'] = tracks['y_c']/3420
-
-
-#### FUNCTIONS ####
-def convex_hull(points): # Calculate the convex hull of points
-    print("Calculating convex hulls")
-    #points = np.array([[1, 2],[1, 4],[4, 1], [2,2.5], [3.5,3.5]])
-    hulls = []
-    for p in points:
-        #print(p)
-        p = np.array(p)
-        print("Points: ", p)
-        hull = ConvexHull(p)
-        #print("Hull vertices: ", p[hull.vertices])
-        hulls.append(p[hull.vertices])
-        
-    return hulls
-
-def cart_product(listOfPolygons):
-    print("Calculating cartesian products of polygons")
-    cartProduct = []
-    #print("List of polygons: ",br, listOfPolygons)
[…]

### 2021-12-05 code/track.py
Small edit

@@ -31,17 +31,17 @@ br = "\n"
 """
 TO-DO
 
-- Record tracks and output a result file
+- Done: Record tracks and output a result file
 - Create running mean solution
 - Plot results
 - Create option to disregard max_distance (to use on non-fixed objects)
-
+S
 """
 
 
 #### SETTINGS ####
 
-resultFilename = "testResults/trackResults7.csv"
+resultFilename = "testResults/trackResults8.csv"
 
 max_disappeared = 4 # Maximum number of frames the algorithm should continue to look for an object
 max_distance = 500 # Maximum distance to a point before it is forced to be registered as a new ID instead of associated with the closest point

### 2021-12-06 code/track.py
Replaced conversion pandas -> numpy with pandas -> list

@@ -53,33 +53,32 @@ running_mean_threshold = 5 # Number of frames for calculating the running mean o
 ######
 
 
-objects = OrderedDict()
-objects[0] = [[2,3]]
-objects[1] = [[4,5]]
-objects[1].append([6,7])
+# objects = OrderedDict()
+# objects[0] = [[2,3]]
+# objects[1] = [[4,5]]
+# objects[1].append([6,7])
 
-means = OrderedDict()
+# means = OrderedDict()
 
-def runningMean(dic):
-    for key, value in dic.items():
-        print(key, value)
-        print(len(value))
-        if len(value) == 1:
-            means[key] = value
-        if len(value) > 1:
-            c_m = [mean([i[0] for i in value]), mean([i[1] for i in value])]
-            means[key] = c_m
+# def runningMean(dic):
+#     for key, value in dic.items():
+#         print(key, value)
+#         print(len(value))
+#         if len(value) == 1:
+#             means[key] = value
+#         if len(value) > 1:
+#             c_m = [mean([i[0] for i in value]), mean([i[1] for i in value])]
+#             means[key] = c_m
             
        
-runningMean(objects)
+# runningMean(objects)
 
-print("Objects: ", br, objects)
-print("Means: ", br, means)
+# print("Objects: ", br, objects)
+# print("Means: ", br, means)
 
 #### PATH TO DETECTIONS ####
 detections = pd.read_csv(r'../Dummy_fortracking2.csv')
 #detections = pd.read_csv(r"U:\BITCue\Projekter\TrackingFlowers\data\annotations\2020_05_15_NorwayAnnotations_THUL-01_IndividualAnnotations_FRCNN_Metrics.csv")
-
 
 
 #detections['frame'] = detections['filename'].str.extract('(\d{6})')
@@ -169,20 +168,19 @@ class tracker():
             print("Length is equal to running mean threshold.")
             print("Deleting first item in ",self.objects[objectID], "(",self.objects[objectID][0],")")
             del self.objects[objectID][0]
-            print("Appending ", [centroid])
+            print("Appending ", centroid)
             
-            self.objects[objectID].append([centroid][0])
+            self.objects[objectID].append(centroid)
             print("Updated: ",self.objects[objectID] )
         
     def update_means(self):
         print("Updating means dictionary")
         for key, value in self.objects.items():
             print("Key ", key, "value", value)
-            if len(value) == 1:
-                print("Length is one")
-                print("Setting ", self.means[key], " to ", value)
[…]

### 2021-12-06 text/ManuscriptDraft.Rmd
Manuscript updates (methods)

@@
 
+As winds shift, flowers close to the edge of the image ..
+may move in and out of view.
+If a flower reappears in the same area as a flower is already being tracked after disappearing in a few frames, it is a reasonable assumption that it is the same individual and not that the old flower wilted/disappeared and a new one developed.
+The parameter **max disappeared** sets the number of frames a track can be lost before a new track is initiated for points appearing in the same area.
+
+Similarly, this deals with potential false negatives. If a given flower has not been annotated in a single frame, it should not be assigned a new track.
+
+Setting **max disappeared** to 0 tracks objects based on the coordinates of the points in the previous frame.
+
 # Evaluating tracking perfomance
@@
 
+# Results
+
 ![](../figures/figure_1.png){ width=100% }
@@
 # Results
+
+**Figure 1:** Figure text....
 

### 2021-12-06 code/track.py
Fixed error with counting disappeared objects

@@ -77,11 +77,11 @@ running_mean_threshold = 5 # Number of frames for calculating the running mean o
 # print("Means: ", br, means)
 
 #### PATH TO DETECTIONS ####
-detections = pd.read_csv(r'../Dummy_fortracking2.csv')
-#detections = pd.read_csv(r"U:\BITCue\Projekter\TrackingFlowers\data\annotations\2020_05_15_NorwayAnnotations_THUL-01_IndividualAnnotations_FRCNN_Metrics.csv")
+#detections = pd.read_csv(r'../Dummy_fortracking2.csv')
+detections = pd.read_csv(r"U:\BITCue\Projekter\TrackingFlowers\data\annotations\2020_05_15_NorwayAnnotations_THUL-01_IndividualAnnotations_FRCNN_Metrics.csv")
 
 
-#detections['frame'] = detections['filename'].str.extract('(\d{6})')
+detections['frame'] = detections['filename'].str.extract('(\d{6})')
 
 detections['x_c'] = (detections['x_min'] + detections['x_max']) / 2
 detections['y_c'] = (detections['y_min'] + detections['y_max']) / 2
@@ -155,8 +155,9 @@ class tracker():
     def deregister(self, objectID): # deregister object by deleting it from the objects dict and removing the associated counter from the disappeared dict.
         del self.objects[objectID]
         del self.disappeared[objectID]
+        del self.means[objectID]
     
-    def update(self, objectID, centroid):
+    def update_object(self, objectID, centroid):
         # First we'll update the dictionary storing the object centroids
         print("Received in update", br, "Object id: ", objectID, br, "Centroid: ", centroid, br)
         if len(self.objects[objectID]) < running_mean_threshold:
@@ -177,14 +178,8 @@ class tracker():
         print("Updating means dictionary")
         for key, value in self.objects.items():
             print("Key ", key, "value", value)
-            # if len(value) == 1:
-            #     print("Length is one")
-            #     print("Setting ", self.means[key], " to ", value)
-            #     self.means[key] = value[0]
-                
             if len(value) > 1:
                 print("Length is more than one")
-
                 c_m = [mean([i[0] for i in value]), mean([i[1] for i in value])]
                 print("Means calculated to: ", c_m)
                 self.means[key] = c_m
@@ -201,6 +196,7 @@ class tracker():
 
         if frame_detections.empty: # If the frame has no detections, we will add 1 to disappeared for all objects that are being tracked.
[…]

### 2021-12-06 code/track.py
Run check

@@ -44,7 +44,7 @@ TO-DO
 
 #### SETTINGS ####
 
-resultFilename = "../testResults/trackResults7.csv"
+resultFilename = "../testResults/trackResults8.csv"
 
 max_disappeared = 4 # Maximum number of frames the algorithm should continue to look for an object
 max_distance = 500 # Maximum distance to a point before it is forced to be registered as a new ID instead of associated with the closest point

### 2021-12-06 code/visualize.py
Updates to code

@@ -434,17 +434,17 @@ from matplotlib.animation import FuncAnimation, PillowWriter
 import pandas as pd
 
 
-df = pd.read_csv(r"U:\BITCue\Projekter\TrackingFlowers\data\annotations\2020_04_30_NorwayAnnotations_NARS-13_IndividualAnnotations_FRCNN_Metrics.csv")
-
+#df = pd.read_csv(r"U:\BITCue\Projekter\TrackingFlowers\data\annotations\2020_04_30_NorwayAnnotations_NARS-13_IndividualAnnotations_FRCNN_Metrics.csv")
+df = pd.read_csv(r'../testResults/trackResults8.csv')
 print(df)
 
 
-df['frame'] = df['filename'].str.extract('(\d{6})')
+#df['frame'] = df['filename'].str.extract('(\d{6})')
 
-df['x_c'] = (df['x_min'] + df['x_max']) / 2
-df['y_c'] = (df['y_min'] + df['y_max']) / 2
+#df['x_c'] = (df['x_min'] + df['x_max']) / 2
+#df['y_c'] = (df['y_min'] + df['y_max']) / 2
 
-df['frame'] = df['frame'].astype('int')
+#df['frame'] = df['frame'].astype('int')
 
 
 dfGrouped = df.groupby(['frame'])
@@ -513,8 +513,9 @@ time_text = ax.text(0.10, 1.1, '', transform=ax.transAxes)
 # ----------------------------------------------------------------------------
 
 ax.set(xlim=(0, 6080), ylim=(0, 3420))
-scat = ax.scatter(x,y, s = 202)
 scat2 = ax.scatter(x,y, s = 100)
+scat = ax.scatter(x,y, s = 202)
+
 
 for i,c in enumerate(x):
     line, = ax.plot(x[i],y[i])
@@ -530,9 +531,12 @@ def animate(i):
     x_i_pre = np.array(points[i-1][0])
     y_i_pre = np.array(points[i-1][1])
     
-    scat.set_offsets(np.c_[x_i, y_i])
+    #x_i_prepre = np.array(points[i-2][0])
+    #y_i_prepre = np.array(points[i-2][1])
     scat2.set_offsets(np.c_[x_i_pre, y_i_pre])
+    scat.set_offsets(np.c_[x_i, y_i])
     
+    #scat2.set_offsets(np.c_[x_i_pre, y_i_pre])
     #line.set_data(x_i[:i][0], y_i[:i][1])
[…]

### 2021-12-06 code/visualize.py
Manuscript update (methods)

@@ -539,9 +539,10 @@ def animate(i):
     #scat2.set_offsets(np.c_[x_i_pre, y_i_pre])
     #line.set_data(x_i[:i][0], y_i[:i][1])
     #line.set_data(300, 500)
+    time_text.set_text(f'Frame {t[i]}')
     for i,c in enumerate(x_i):
         line.set_data = ax.plot(x_i[i],y_i[i])
-    time_text.set_text(f'Frame {t[i]}')
+    
 
 
 # ----------------------------------------------------------------------------

### 2021-12-09 code/evaluate.py
Updated code so it can run directly on tracking output

@@ -13,11 +13,14 @@ import pandas as pd
 iou_threshold = 0.5
 
 # Get the ground truth annotations on the format [filename, x_min, y_min, x_max, y_max, id_gt]
-gt = pd.read_csv(r"U:\BITCue\Projekter\NorwayAnnotations\Experiments\2020_06_30_DetectionPhenology\Rmd\ModelQ_Results\DT_GT\2020_02_01_NorwayAnnotations_AllTestSeries_IndividualAnnotations_FRCNN_Metrics.csv", sep=",")
+gt = pd.read_csv(r"U:\BITCue\Projekter\TrackingFlowers\data\annotations\2020_05_15_NorwayAnnotations_THUL-01_IndividualAnnotations_FRCNN_Metrics.csv", sep=",")
 #print(gt)
 # Get the tracked detections on the format [filename, x_min, y_min, x_max, y_max, id_dt]
-dt = pd.read_csv(r"U:\BITCue\Projekter\NorwayAnnotations\Experiments\2020_06_30_DetectionPhenology\Rmd\ModelQ_Results\DT_GT\submit_boxes_NewFormat_AllTestSeries_Sampled.csv", sep=",")
+#dt = pd.read_csv(r"U:\BITCue\Projekter\NorwayAnnotations\Experiments\2020_06_30_DetectionPhenology\Rmd\ModelQ_Results\DT_GT\submit_boxes_NewFormat_AllTestSeries_Sampled.csv", sep=",")
+#dt = pd.read_csv(r"U:\BITCue\Projekter\NorwayAnnotations\Experiments\2020_06_30_DetectionPhenology\Rmd\ModelQ_Results\DT_GT\2020_02_01_NorwayAnnotations_AllTestSeries_IndividualAnnotations_FRCNN_Metrics_copy.csv", sep=",")
 
+dt = pd.read_csv(r"../testResults/5testing_THUL-01_maxDis_4_runMean_5.csv")
+dt.rename(columns={'objectID': 'id_tr'}, inplace=True)
 print(dt)
 
 #dt['x_min'] = dt['x_min'] * 8
@@ -244,6 +247,7 @@ P = gt.shape[0]
 PR = dt.shape[0]
 CP =  PR - FP
 
+print("\n#####", " RESULTS ", "#####\n")
 print("Number of annotated objects: ", P)
 print("Number of predictions made: ", PR )
 print("Correct positives: ",CP)
[…]

### 2021-12-09 code/track.py
Updated code to produce the right output for evaluate.py

@@ -35,10 +35,10 @@ br = "\n"
 
 verbose = True # Set to True if you want tracking process printed to screen and False if not
 
-prefix_results_filename = "testing"
+prefix_results_filename = "5testing_THUL-01"
 
 list_max_disappeared = [4] # Maximum number of frames a track can be lost before new points will be forces into a new track.
-list_running_mean_threshold = [0] # Maximum number of frames for calculating the running mean of the position of an object If there are less than this number of frames currently in the track, a mean over what is in the track will be used..
+list_running_mean_threshold = [5] # Maximum number of frames for calculating the running mean of the position of an object If there are less than this number of frames currently in the track, a mean over what is in the track will be used..
 list_max_distance = [0] # If set to 0, this parameter will be ignored. If not zero, points that have a distance to tracked objects higher than this parameter will be forced into new tracks.
 
 list_of_parameters = [(x,y,z) for x in list_max_disappeared for y in list_running_mean_threshold for z in list_max_distance] # A list of all combinations of the above parameters.
@@ -46,7 +46,7 @@ list_of_parameters = [(x,y,z) for x in list_max_disappeared for y in list_runnin
 
 #### PATH TO DETECTIONS ####
 #detections = pd.read_csv(r'../Dummy_fortracking2.csv')
-detections = pd.read_csv(r"U:\BITCue\Projekter\TrackingFlowers\data\annotations\2020_04_30_NorwayAnnotations_NARS-13_IndividualAnnotations_FRCNN_Metrics.csv")
+detections = pd.read_csv(r"U:\BITCue\Projekter\TrackingFlowers\data\annotations\2020_05_15_NorwayAnnotations_THUL-01_IndividualAnnotations_FRCNN_Metrics.csv")
 detections['frame'] = detections['filename'].str.extract('(\d{6})')
 detections['x_c'] = (detections['x_min'] + detections['x_max']) / 2
 detections['y_c'] = (detections['y_min'] + detections['y_max']) / 2
@@ -75,7 +75,7 @@ class tracker():
         self.tracks = [] # Create a list for storing tracking results as we go
 
         with open(self.results_filename, 'a') as resultFile: # Write the header of the output file
-            header = 'frame,filename,x_c,y_c,objectID\n'
+            header = 'frame,filename,x_min,x_max,y_min,y_max,x_c,y_c,objectID\n'
             resultFile.write(header)
 
[…]

### 2021-12-09 text/ManuscriptDraft.Rmd
Manuscript update (methods)
Updated manuscript

@@
 
-The tracking algorithm can be applied both offline (on a set of detections/annotations that have already been produced) or online (realtime tracking frame per frame).
+The tracking algorithm can be applied both offline (on a set of detections/annotations that have already been produced) or online (real-time tracking frame per frame).
+
+The speed of the tracking algorithm depends on the computational power available as well as the number of objects that are being tracked.
 
@@
 
+The counter for number of disappeared frames is reset when a new point is associated with the track within the threshold.
+
 # Evaluating tracking perfomance
 
-Mota counts shifts in tracking.
+Mota counts tracking mismathces (shifts in a track).
 
@@
 To associate other information to the flower, for example flower visits, we want as much as possible of the track to be correct.
+
+Lastly, we may be interested in the number of flowers that existed in a plot.
+Therefore, a final way of evaluating automatic flower tracking performance is to compare the number of tracks identified by the automatic tracking with the true number of flowers in a series.
+These should ideally be equal.
 

### 2021-12-09 text/ManuscriptDraft.Rmd
Updated manuscript

@@
 
+![](../figures/figure_1.png){ width=100% }
+**Figure 1:** Simple centroid tracking may produce erronous associations when objects move between frames.
+Blue shows detections in the current frame (bounding box and centroid point).
+Red shows centroid points for the detections in the previous frame.
+
 ### User parameters
@@
 
+![](../figures/figure_22.jpg){ width=100% }
+
 The flowers move around a center point because of their stalk.
@@
 The parameter **max disappeared** sets the number of frames a track can be lost before a new track is initiated for points appearing in the same area.
+
+![](../figures/figure_3.png){ width=100% }
 
@@
 
-![](../figures/figure_1.png){ width=100% }
-
-**Figure 1:** Simple centroid tracking may produce erronous associations when objects move between frames.
-Blue shows detections in the current frame (bounding box and centroid point).
-Red shows centroid points for the detections in the previous frame.
-
-# Results
-
 **Figure 1:** Figure text....
@@
 
-The code that supports the results in this paper will be made openly available at https://github.com/TECOLOGYxyz.
+The code that supports the results in this paper will be made openly available at https://github.com/TECOLOGYxyz/FlowerTracking.
 Raw data as well as the trained flower detection model will be archived on https://zenodo.org/.

### 2021-12-11 code/evaluate.py
Experiment run

@@ -13,13 +13,13 @@ import pandas as pd
 iou_threshold = 0.5
 
 # Get the ground truth annotations on the format [filename, x_min, y_min, x_max, y_max, id_gt]
-gt = pd.read_csv(r"U:\BITCue\Projekter\TrackingFlowers\data\annotations\2020_05_15_NorwayAnnotations_THUL-01_IndividualAnnotations_FRCNN_Metrics.csv", sep=",")
+gt = pd.read_csv(r"U:\BITCue\Projekter\TrackingFlowers\data\annotations\2020_04_30_NorwayAnnotations_NARS-13_IndividualAnnotations_FRCNN_Metrics.csv", sep=",")
 #print(gt)
 # Get the tracked detections on the format [filename, x_min, y_min, x_max, y_max, id_dt]
 #dt = pd.read_csv(r"U:\BITCue\Projekter\NorwayAnnotations\Experiments\2020_06_30_DetectionPhenology\Rmd\ModelQ_Results\DT_GT\submit_boxes_NewFormat_AllTestSeries_Sampled.csv", sep=",")
 #dt = pd.read_csv(r"U:\BITCue\Projekter\NorwayAnnotations\Experiments\2020_06_30_DetectionPhenology\Rmd\ModelQ_Results\DT_GT\2020_02_01_NorwayAnnotations_AllTestSeries_IndividualAnnotations_FRCNN_Metrics_copy.csv", sep=",")
 
-dt = pd.read_csv(r"../testResults/5testing_THUL-01_maxDis_4_runMean_5.csv")
+dt = pd.read_csv(r"../testResults/2testing_NARS-13_maxDis_4_runMean_0.csv")
 dt.rename(columns={'objectID': 'id_tr'}, inplace=True)
 print(dt)
 

### 2021-12-11 code/track.py

@@ -35,10 +35,10 @@ br = "\n"
 
 verbose = True # Set to True if you want tracking process printed to screen and False if not
 
-prefix_results_filename = "5testing_THUL-01"
+prefix_results_filename = "2testing_NARS-13"
 
 list_max_disappeared = [4] # Maximum number of frames a track can be lost before new points will be forces into a new track.
-list_running_mean_threshold = [5] # Maximum number of frames for calculating the running mean of the position of an object If there are less than this number of frames currently in the track, a mean over what is in the track will be used..
+list_running_mean_threshold = [0] # Maximum number of frames for calculating the running mean of the position of an object If there are less than this number of frames currently in the track, a mean over what is in the track will be used..
 list_max_distance = [0] # If set to 0, this parameter will be ignored. If not zero, points that have a distance to tracked objects higher than this parameter will be forced into new tracks.
 
 list_of_parameters = [(x,y,z) for x in list_max_disappeared for y in list_running_mean_threshold for z in list_max_distance] # A list of all combinations of the above parameters.
@@ -46,7 +46,7 @@ list_of_parameters = [(x,y,z) for x in list_max_disappeared for y in list_runnin
 
 #### PATH TO DETECTIONS ####
 #detections = pd.read_csv(r'../Dummy_fortracking2.csv')
-detections = pd.read_csv(r"U:\BITCue\Projekter\TrackingFlowers\data\annotations\2020_05_15_NorwayAnnotations_THUL-01_IndividualAnnotations_FRCNN_Metrics.csv")
+detections = pd.read_csv(r"U:\BITCue\Projekter\TrackingFlowers\data\annotations\2020_04_30_NorwayAnnotations_NARS-13_IndividualAnnotations_FRCNN_Metrics.csv")
 detections['frame'] = detections['filename'].str.extract('(\d{6})')
 detections['x_c'] = (detections['x_min'] + detections['x_max']) / 2
 detections['y_c'] = (detections['y_min'] + detections['y_max']) / 2
@@ -252,11 +252,12 @@ for i in list_of_parameters:
     
     for f in frames:
         t.track(f)
+    endtime = time.time()
+    print(f'Tracking done. That took {round(endtime-starttime, 3)} seconds. That is {round((endtime-starttime)/len(frames), 3)} seconds per frame.')
     t.write_tracks_file()
     tracks = pd.read_csv(resultsFilename)
[…]

### 2021-12-11 text/ManuscriptDraft.Rmd
Updated manuscript (introduction)

@@
 - \usepackage{amsmath}
-keywords: Citizen science, Arctic, insects
+keywords: Tracking, flowering phenology, Arctic, 
 bibliography: ./library.bib
@@
 
-What questions do we want to answer?
-
-*Flower information*
+Online tracking:
+SORT
+DeepSORT
 
@@
 
-The flowering phenology of a population may mask responses at the individual level. For example,
+* **Flowering phenology**
+  + Importance of studying flowering phenology
+  + Responses to climate change
+  + Phenology of communities, populations, individuals
+  + Traditional methods for studying
+  + Onset of flowering says little about true distribution
+    - Even true distribition of community says little about flowering lengths of individuals and for example how it varies accross the season.
+  + Difficult to study at the individual level - requires high temporal resolution and keeping track of individuals
+* **Image based monitoring**
+  + Automatic, high temporal resolution, remote sites
+  + High temporal resolution means that we can annotate individuals and get phenology of the individual
 
-Does flower visitation rates and/or reproductive success depend on the timing of flowering for the indiviual flower?
+* **Tracking**
+  + Offline and online
+  + Online often coupled with CNNs that attempt to distinguish individuals from each other and recognize them through frames
+
+  + Flowers appear very similar and 
+  + Many methods for offline tracking
+  + Hungarian/Kahlmann filter
+  ++ May not be applicable for objects that move weirdly, e.g. change directions between frames.
+  + Tracking based on distance.
+    - Good but has some problems
+    - Two points always associated disregarding absolute distance
+    - Tracks lost when objects disappear
+    - Objects close to each other may swap tracks
+
+* **Our solution**
+  + Here we demonstrate a framework for automatic flower tracking and evualutation of tracking performance
+  + Ground truth tracks
+
+\newpage
+
+For the individual plant, timing of flowering is of utmost importance.
+Precocious flowering means that the plant has failed to exploit the whole temporal window for accumulating resources before allocating energy to flowering.
+On the other hand, flowering too late limit the time for reproduction before the end of the growing season [@ELZINGA2007].
+Further, flowering may need to be synchronous with pollinator activity for successful reproduction.
+Flowering phenology may plastically change as a response to abiotic cues in the environment, such as timing of spring, temperature, and photoperiod, but variation in flowering phenology is partly heritable and shaped by selective forces from the abiotic and biotic environment.
+
+Monitoring of flower phenology at high temporal resolution is laboursome and time-consuming, particularly in logistically challenging environment such as the Arctic.
+Consequently, simple variables are often used to represent the flowering phenology of a population, such as the date for onset of flowering, often derived from weekly observations of sample plots.
+
+Automatic image-based monitoring of flowering phenology can return phenology data for specific species at very high temporal resolution (Mann et al., in prep).
+
+For example, phenological responses at the individual level may be indiscernible regardless of the temporal resolution of the data at population level.
+For example, a shortening of flower longevity may not be directly obvious at the population level.
+Further, investigating the association between reproductive success and timing of flowering and flower longevity requires phenology data at the level of individuals.
+
+Does flower visitation rates and/or reproductive success depend on the timing of flowering for the individual flower?
 
@@
 ![](../figures/figure_1.png){ width=100% }
+
 **Figure 1:** Simple centroid tracking may produce erronous associations when objects move between frames.
@@
 
-### User parameters
-
-**Max distance threshold**, **running mean**, **max disappeared**,
+## User parameters
 
@@
 This can happen instantaneously (i.e.
-between two censecutive frames).
+between two consecutive frames).
 As they are constrained by their stalk, there is a limit to the distance they can move.
@@
 
-As winds shift, flowers close to the edge of the image ..
-may move in and out of view.
+As winds shift, flowers close to the edge of the frame may move in and out of view.
 If a flower reappears in the same area as a flower is already being tracked after disappearing in a few frames, it is a reasonable assumption that it is the same individual and not that the old flower wilted/disappeared and a new one developed.
@@
 
-Setting **max disappeared** to 0 tracks objects based on the coordinates of the points in the previous frame.
-
+Setting **max disappeared** to 0 tracks objects based on the coordinates of the points in only the previous frame.
 The counter for number of disappeared frames is reset when a new point is associated with the track within the threshold.
@@
 
-Mota counts tracking mismathces (shifts in a track).
+Mota counts tracking mismatches (shifts in a track).
 
@@
 
-**Figure 1:** Figure text....
-
 # Discussion
@@
 Raw data as well as the trained flower detection model will be archived on https://zenodo.org/.
+
+\newpage
+# References\

### 2021-12-11 code/evaluate.py
Showing mismatches

@@ -13,13 +13,13 @@ import pandas as pd
 iou_threshold = 0.5
 
 # Get the ground truth annotations on the format [filename, x_min, y_min, x_max, y_max, id_gt]
-gt = pd.read_csv(r"U:\BITCue\Projekter\TrackingFlowers\data\annotations\2020_04_30_NorwayAnnotations_NARS-13_IndividualAnnotations_FRCNN_Metrics.csv", sep=",")
+gt = pd.read_csv(r"U:\BITCue\Projekter\TrackingFlowers\data\annotations\2020_05_23_NorwayAnnotations_NYAA-04_IndividualAnnotations_FRCNN_Metrics.csv", sep=",")
 #print(gt)
 # Get the tracked detections on the format [filename, x_min, y_min, x_max, y_max, id_dt]
 #dt = pd.read_csv(r"U:\BITCue\Projekter\NorwayAnnotations\Experiments\2020_06_30_DetectionPhenology\Rmd\ModelQ_Results\DT_GT\submit_boxes_NewFormat_AllTestSeries_Sampled.csv", sep=",")
 #dt = pd.read_csv(r"U:\BITCue\Projekter\NorwayAnnotations\Experiments\2020_06_30_DetectionPhenology\Rmd\ModelQ_Results\DT_GT\2020_02_01_NorwayAnnotations_AllTestSeries_IndividualAnnotations_FRCNN_Metrics_copy.csv", sep=",")
 
-dt = pd.read_csv(r"../testResults/2testing_NARS-13_maxDis_4_runMean_0.csv")
+dt = pd.read_csv(r"../testResults/6testing_NYAA-04_maxDis_8_runMean_10.csv")
 dt.rename(columns={'objectID': 'id_tr'}, inplace=True)
 print(dt)
 

### 2021-12-15 code/R/Analysis.Rmd
Added plotting of evaluation results

@@
 library(stringr)
+library(GGally)
+library(plotly)
 knitr::opts_chunk$set(echo = TRUE)
@@
 ```
+
+```{r}
+d = read.csv("U:/BITCue/Projekter/TrackingFlowers/code/7temptest_eval.csv")
+
+d %>% 
+  ggplot() +
+  geom_point(aes(x=runMean, y = maxDisap, size = mm, colour = mm)) + 
+  facet_wrap(vars(maxDist))+scale_color_gradient(low="blue", high="red")
+```
+
+```{r}
+d1 = read.csv("U:/BITCue/Projekter/TrackingFlowers/code/8temptest_eval_THUL01.csv")
+
+d1 %>% 
+  ggplot() +
+  geom_point(aes(x=runMean, y = maxDisap, size = mm, colour = mm)) + 
+  facet_wrap(vars(maxDist))+scale_color_gradient(low="blue", high="red")
+
+```
+
+```{r}
+remotes::install_github("tylermorganwall/rayshader")
+library(rayshader)
+library(tidyverse)
+measles = read_csv("https://tylermw.com/data/measles_country_2011_2019.csv")
+melt_measles = reshape2::melt(measles, id.vars = c("Year", "Country", "Region", "ISO3"))
+melt_measles$Month = melt_measles$variable
+melt_measles$cases = melt_measles$value
+melt_measles %>% 
+  group_by(Year, Month) %>%
+  summarize(totalcases = sum(cases,na.rm = TRUE)) %>% 
+  mutate(totalcases = ifelse(Year == 2019 & !(Month %in% c("January","February","March")), NA, totalcases)) %>%
+  ggplot() + 
+  geom_tile(aes(x=Year, y=Month, fill=totalcases,color=totalcases),size=1,color="black") + 
+  scale_x_continuous("Year", expand=c(0,0), breaks = seq(2011,2019,1)) +
+  scale_y_discrete("Month", expand=c(0,0)) +
+  scale_fill_viridis("Total\nCases") +
+  ggtitle("Reported Worldwide Measles Cases") +
+  labs(caption = "Data Source: WHO") +
+  theme(axis.text = element_text(size = 12),
+        title = element_text(size = 12,face="bold"),
+        panel.border= element_rect(size=2,color="black",fill=NA)) -> 
+measles_gg
+
+plot_gg(measles_gg, multicore = TRUE, width = 6, height = 5.5, scale = 300, 
+background = "#afceff",shadowcolor = "#3a4f70")
+```

### 2021-12-15 code/evaluate.py
Added solution to run over multiple files

@@ -16,7 +16,7 @@ iou_threshold = 0.5
 
 def write_header(results_filename):
     with open(results_filename, 'a') as resultFile: # Write the header of the output file
-        header = f'maxDisap,maxDist,runMean,mm,mmratio{br}'
+        header = f'maxDisap,maxDist,runMean,mm,mota,numberOfObjects{br}'
         resultFile.write(header)
 
 
@@ -37,16 +37,19 @@ class evaluator():
         self.dt_filename = dt
         basename = os.path.basename(self.dt_filename)
         print(basename)
-        runMeanSearch = re.search('runMean_(.+?)_', basename)
-        maxDistSearch = re.search('maxDist_(.+?).', basename)
-        maxDisapSearch = re.search('runMean_(.+?)_', basename)
+        runMeanSearch = re.search('runMean_(.+?)_max', basename)
+        maxDistSearch = re.search('maxDist_(.+?).csv', basename)
+        maxDisapSearch = re.search('maxDisap_(.+?)_run', basename)
         
         if runMeanSearch:
             self.runMean = runMeanSearch.group(1)
+            print(f'runMean: {self.runMean}')
         if maxDistSearch:
             self.maxDist = maxDistSearch.group(1)
+            print(f'maxDist: {self.maxDist}')
         if maxDisapSearch:
             self.maxDisap = maxDisapSearch.group(1)
+            print(f'maxDisap: {self.maxDisap}')
             
             
     def set_up_data(self):
@@ -72,7 +75,9 @@ class evaluator():
         #dt['x_max'] = dt['x_max'] * 8
         #dt['y_max'] = dt['y_max'] * 8
 
-        return dt, gt
+        number_of_objects = len(dt['id_tr'].unique())
+
+        return dt, gt, number_of_objects
         
     
     def calculate_iou(self, boxA, boxB):
@@ -143,14 +148,14 @@ class evaluator():
     	return MOTA
     
         
-    def write_results_file(self, mm, mmratio):
+    def write_results_file(self, mm, mota, number_of_objects):
         with open(self.results_filename, 'a') as resultFile:
-            resultFile.write(f'{self.maxDisap},{self.maxDist},{self.runMean},{mm},{mmratio}{br}')
[…]

### 2021-12-15 code/track.py
Minor update

@@ -50,6 +50,7 @@ class tracker():
         
         self.tracks = [] # Create a list for storing tracking results as we go
 
+
         with open(self.results_filename, 'a') as resultFile: # Write the header of the output file
             header = 'frame,filename,x_min,x_max,y_min,y_max,x_c,y_c,objectID\n'
             resultFile.write(header)

### 2021-12-15 code/run.py
Parameter combination testing

@@ -13,14 +13,13 @@ import time
 
 ### SETTINGS ###
 
-prefix_results_filename = "parameterTest_NYAA-04_"
+prefix_results_filename = "parameterTest_NYAA-04"
 
-list_max_disappeared = [0,5,10,15,20,25,30] # Maximum number of frames a track can be lost before new points will be forces into a new track.
-list_running_mean_threshold = [0,5,10,15,20,25,30] # Maximum number of frames for calculating the running mean of the position of an object If there are less than this number of frames currently in the track, a mean over what is in the track will be used..
+list_max_disappeared = [0,5,10,15,20,25,30,35,40,45,50] # Maximum number of frames a track can be lost before new points will be forces into a new track.
+list_running_mean_threshold = [0,5,10,15,20,25,30,35,40,45,50] # Maximum number of frames for calculating the running mean of the position of an object If there are less than this number of frames currently in the track, a mean over what is in the track will be used..
 list_max_distance = [0, 100, 200, 300, 400, 500] # If set to 0, this parameter will be ignored. If not zero, points that have a distance to tracked objects higher than this parameter will be forced into new tracks.
 
 list_of_parameters = [(x,y,z) for x in list_max_disappeared for y in list_running_mean_threshold for z in list_max_distance] # A list of all combinations of the above parameters.
-
 
 #### PATH TO DETECTIONS ####
 #detections = pd.read_csv(r'../Dummy_fortracking2.csv')
@@ -41,7 +40,7 @@ verbose = False # Set to True if you want tracking process printed to screen and
 
 for i in list_of_parameters:
     starttime = time.time()
-    resultsFilename = f'../testResults/_parameterTest_NYAA-04_1/{prefix_results_filename}_maxDisap_{i[0]}_runMean_{i[1]}_maxDist_{i[2]}.csv'    
+    resultsFilename = f'../testResults/_parameterTest_NYAA-04_2/{prefix_results_filename}_maxDisap_{i[0]}_runMean_{i[1]}_maxDist_{i[2]}.csv'    
     
     t = tracker(i[0], i[2], i[1], resultsFilename, frames, detections, verbose)
     

### 2021-12-16 code/R/Analysis.Rmd
Plot evaluation results

--- 2021-12-05 code/R/Analysis.Rmd
@@
 ```
+
+```{r}
+d = read.csv("U:/BITCue/Projekter/TrackingFlowers/code/7temptest_eval.csv")
+
+d %>% 
+  ggplot() +
+  geom_point(aes(x=runMean, y = maxDisap, size = mm, colour = mm)) + 
+  facet_wrap(vars(maxDist))+scale_color_gradient(low="blue", high="red")
+```
+
+```{r}
+d1 = read.csv("U:/BITCue/Projekter/TrackingFlowers/code/8temptest_eval_THUL01.csv")
+
+d1 %>% 
+  ggplot() +
+  geom_point(aes(x=runMean, y = maxDisap, size = mm, colour = mm)) + 
+  facet_wrap(vars(maxDist))+scale_color_gradient(low="blue", high="red")
+
+```
+
+```{r}
+d2 = read.csv("U:/BITCue/Projekter/TrackingFlowers/testResults/parameterTest_2_NYAA-04_Evaluate.csv")
+
+d2 %>% 
+  ggplot() +
+  geom_point(aes(x=runMean, y = maxDisap, size = mm, colour = mm)) + 
+  facet_wrap(vars(maxDist))+scale_color_gradient(low="blue", high="red")
+
+```

### 2021-12-16 code/run.py
Parameter testing

@@ -13,18 +13,23 @@ import time
 
 ### SETTINGS ###
 
-prefix_results_filename = "parameterTest_NYAA-04"
+prefix_results_filename = "parameterTest_NARS-13"
 
-list_max_disappeared = [0,5,10,15,20,25,30,35,40,45,50] # Maximum number of frames a track can be lost before new points will be forces into a new track.
-list_running_mean_threshold = [0,5,10,15,20,25,30,35,40,45,50] # Maximum number of frames for calculating the running mean of the position of an object If there are less than this number of frames currently in the track, a mean over what is in the track will be used..
-list_max_distance = [0, 100, 200, 300, 400, 500] # If set to 0, this parameter will be ignored. If not zero, points that have a distance to tracked objects higher than this parameter will be forced into new tracks.
+list_max_disappeared = [0,10,20,30,40,50,60,70,80,90,100,110,120,130,140,150,160] # Maximum number of frames a track can be lost before new points will be forces into a new track.
+list_running_mean_threshold = [0,10,20,30,40,50,60,70,80,90,100,110,120,130,140,150,160] # Maximum number of frames for calculating the running mean of the position of an object If there are less than this number of frames currently in the track, a mean over what is in the track will be used..
+list_max_distance = [0, 100, 200, 300, 400, 500, 600, 700, 800, 900, 1000] # If set to 0, this parameter will be ignored. If not zero, points that have a distance to tracked objects higher than this parameter will be forced into new tracks.
+
+# list_max_disappeared = [100] # Maximum number of frames a track can be lost before new points will be forces into a new track.
+# list_running_mean_threshold = [5] # Maximum number of frames for calculating the running mean of the position of an object If there are less than this number of frames currently in the track, a mean over what is in the track will be used..
+# list_max_distance = [400] # If set to 0, this parameter will be ignored. If not zero, points that have a distance to tracked objects higher than this parameter will be forced into new tracks.
+
 
 list_of_parameters = [(x,y,z) for x in list_max_disappeared for y in list_running_mean_threshold for z in list_max_distance] # A list of all combinations of the above parameters.
 
 #### PATH TO DETECTIONS ####
[…]

### 2021-12-16 code/evaluate.py
Parameter testing

@@ -277,17 +277,17 @@ class evaluator():
 #dt = pd.read_csv(r"U:\BITCue\Projekter\NorwayAnnotations\Experiments\2020_06_30_DetectionPhenology\Rmd\ModelQ_Results\DT_GT\submit_boxes_NewFormat_AllTestSeries_Sampled.csv", sep=",")
 #dt = pd.read_csv(r"U:\BITCue\Projekter\NorwayAnnotations\Experiments\2020_06_30_DetectionPhenology\Rmd\ModelQ_Results\DT_GT\2020_02_01_NorwayAnnotations_AllTestSeries_IndividualAnnotations_FRCNN_Metrics_copy.csv", sep=",")
 
-gt = r"U:\BITCue\Projekter\TrackingFlowers\data\annotations\2020_05_15_NorwayAnnotations_THUL-01_IndividualAnnotations_FRCNN_Metrics.csv"
+gt = r"U:\BITCue\Projekter\TrackingFlowers\data\annotations\2021_12_13_NorwayAnnotations_NYAA-04_IndividualAnnotations_FRCNN_Metrics.csv"
 #dt = r"../testResults/6testing_NYAA-04_maxDis_8_runMean_10.csv"
 #dt = r"../testResults\_parameterTest_NYAA-04_1\parameterTest_NYAA-04_maxDisap_30_runMean_30_maxDist_500.csv"
 #dt = r'../testResults\_parameterTest_NYAA-04_1\parameterTest_NYAA-04__maxDisap_30_runMean_30_maxDist_500.csv'
 
 
-path = r'../testResults\_parameterTest_THUL-01_1'
+path = r'../testResults\_parameterTest_NYAA-04_2/1'
 files = [os.path.join(path,f) for f in os.listdir(path) if f.endswith('.csv')]
 print(files)
 
-results_filename = "8temptest_eval_THUL01.csv"
+results_filename = "../testResults/maxDisap_100_parameterTest_2_NYAA-04_Evaluate.csv"
 
 
 write_header(results_filename)

### 2021-12-16 text/ManuscriptDraft.Rmd
Note about visualization
Minor update
Manuscript update

@@
 ```
-
 \newpage
-**NOTES**
-
-Online tracking:
-SORT
-DeepSORT
-
-\pagebreak
 
@@
 * **Our solution**
-  + Here we demonstrate a framework for automatic flower tracking and evualutation of tracking performance
+  + Here we demonstrate a framework for automatic flower tracking and evaluation of tracking performance
   + Ground truth tracks
@@
 Monitoring of flower phenology at high temporal resolution is laboursome and time-consuming, particularly in logistically challenging environment such as the Arctic.
-Consequently, simple variables are often used to represent the flowering phenology of a population, such as the date for onset of flowering, often derived from weekly observations of sample plots.
+Consequently, simple variables are often used as proxies for the flowering phenology of a population, such as the date for onset of flowering, often derived from weekly observations of sample plots.
+Such proxies may fail to reveal dynamics in flowering phenology for example caused by changes in climate.
 
-Automatic image-based monitoring of flowering phenology can return phenology data for specific species at very high temporal resolution (Mann et al., in prep).
+Automatic image-based monitoring of flowering phenology can return phenology data for specific species at very high temporal resolution (Mann et al., in prep), but phenological responses at the individual level may be indiscernible regardless of the temporal resolution of the data at population level.
+For example, a shortening of individual flower longevity may not be directly obvious at the population level.
+Many research questions can only be explored on the basis of individual phenology data.
+For example, investigating the association between reproductive success and timing of flowering and flower longevity requires phenology data at the level of individuals.
+Similarly, such data is necessary for investigating whether flower visitation rates and/or reproductive success depends on the timing of flowering for the individual flower.
 
-For example, phenological responses at the individual level may be indiscernible regardless of the temporal resolution of the data at population level.
-For example, a shortening of flower longevity may not be directly obvious at the population level.
-Further, investigating the association between reproductive success and timing of flowering and flower longevity requires phenology data at the level of individuals.
-
-Does flower visitation rates and/or reproductive success depend on the timing of flowering for the individual flower?
+Here, we show that information on phenology at the level of individuals can be derived from image-based monitoring of flower phenology.
+Further, we present and evaluate an automatic flower tracking algorithm.
 
@@
 
-We built a framework for tracking, filtering, and evaluating tracking of objects in time-lapse image series.
+The annotation process is described in Mann et al., in prep, but in short: We manually annotated all flowers in the sampled image series using the rectangular bounding box tool in the VIA VGG annotation software.
+Further, we assigned each individual a unique ID.
+These annotations constitue our ground truth tracks.
 
@@
 
+We built a framework for tracking, filtering, and evaluating tracking of objects in time-lapse image series.
+
 Our algorithm tracks objects based on distances between centroids of bounding boxes.
 
-The tracking algorithm has a set of user adjusted parameters that can optimise tracking accuracy.
-
-The parameters are particularly relevant for optimal tracking of objects that are constrained to a specific area such as flowers.
-
-It is important to note, however, that the tracking algorithm can be used to track any objects.
-
-The tracking algorithm can be applied both offline (on a set of detections/annotations that have already been produced) or online (real-time tracking frame per frame).
-
-The speed of the tracking algorithm depends on the computational power available as well as the number of objects that are being tracked.
-
+# 
 ![](../figures/figure_1.png){ width=100% }
 
-**Figure 1:** Simple centroid tracking may produce erronous associations when objects move between frames.
+**Figure 1:** Simple centroid tracking may produce erroneous associations when objects move between frames.
 Blue shows detections in the current frame (bounding box and centroid point).
@@
 
+# 
+![](../figures/figure_22.jpg){width=100%}
+
+**Figure 2:** Simple centroid tracking may produce erroneous associations when objects move between frames.
+Basing the association on the running mean of the positions of the previous n number of tracks may alleviate this issue.
+In this case basing the association on only the previous point would produce a wrong results while basing it on the running mean would produce a correct result.
+Red points: Centroids for bounding boxes in current frame, blue points: Centroids for bounding boxes in previous frame; grey points: Centroids for bounding boxes in a number of frames before t-1; green points: Running mean of the previous n points.
+Circles delimit the two individuals.
+
+# 
+![](../figures/figure_3.png){ width=100% }
+
+**Figure 3:** Simple centroid tracking may produce erroneous associations when objects disappear periodically from the frame.
+Here the top flower moves out of frame and the bottom flower would be assigned to the track of the top flower in frame t-1.
+
 ## User parameters
+
+The tracking algorithm has a set of user adjusted parameters that can optimize tracking accuracy.
+The parameters are particularly relevant for optimal tracking of objects that are constrained to a specific area such as flowers.
+It is important to note, however, that the tracking algorithm can be used to track any objects.
+The tracking algorithm can be applied both offline (on a set of detections/annotations that have already been produced) or online (real-time tracking frame per frame).
+The speed of the tracking algorithm depends on the computational power available as well as the number of objects that are being tracked.
 
@@
 
-![](../figures/figure_22.jpg){ width=100% }
-
 The flowers move around a center point because of their stalk.
@@
 
-![](../figures/figure_3.png){ width=100% }
-
 Similarly, this deals with potential false negatives. If a given flower has not been annotated in a single frame, it should not be assigned a new track.
@@
 
+### Identifying optimal user parameters
+
+To explore the effect of the user parameters and to identify the optimal combination of parameters for our case of tracking flowers, we followed a step wise approach.
+First, we ran the tracking algorithm on each of the four image series with every combination of the following settings (3.179 combinations):
+
+max_disappeared = [0,10,20,30,40,50,60,70,80,90,100,110,120,130,140,150,160]
+
+running_mean_threshold = [0,10,20,30,40,50,60,70,80,90,100,110,120,130,140,150,160]
+
+max_distance = [0, 100, 200, 300, 400, 500, 600, 700, 800, 900, 1000]
+
+Note that max_distance set to zero.
+We identified the setting(s) that returned the lowest number of track mismatches and performed a second run with finer scaled settings for each series.
+Finally, we compare the tracking results between optimal settings and all parameters set to zero.
+
 # Evaluating tracking perfomance
 
-Mota counts tracking mismatches (shifts in a track).
+The optimal way of quantifying tracking performance depends on the goal of the tracking.
+To associate other information obtained in the images to the individual flower, for example flower visits, we want as much as possible of the track to be correct.
+To derive flowering length in theory we just need to track the most extreme points correctly and don't care about the intermediate points.
+Lastly, we may be interested in the number of flowers that existed in a plot, in which case we want the number of tracks obtained by automatic tracking to be as close as possible to the actual number of individuals in the series.
 
-To derive flowering length, in theory we just need to track the most extreme points correctly and don't care about other points (although we filter by length when overlap).
+The multiple object tracking accuracy (MOTA) score quantifies tracking performance based on counts of tracking mismatches [@bernardin2008].
+Mismatches occur when objects swap track identity because they are in close vicinity to each other or when an object periodically disappears and is assigned a new track identity when it reappears.
+Only the shifts in tracking identity are counted as mismatches while the number of points assigned to each track is not considered.
 
-To associate other information to the flower, for example flower visits, we want as much as possible of the track to be correct.
+For each series, we calculate the ratio of flowers for which the automatic tracking algorithm returns the correct flowering length compared to the ground truth tracks.
 
-Lastly, we may be interested in the number of flowers that existed in a plot.
-Therefore, a final way of evaluating automatic flower tracking performance is to compare the number of tracks identified by the automatic tracking with the true number of flowers in a series.
-These should ideally be equal.
+Finally, we compare the number of tracks identified by the automatic tracking with the true number of flowers in a series. These should ideally be equal.
 
@@
 
+One way to plot?: Time on x-axis, gt id on y axis, tr id as colour. This will show mismatches.
+
 # Discussion
 
-## Flower phenology
+All three parameters make a difference.
+
+In cases where we are tracking perfectly with maxDisap = 0, setting it any value will not make a difference. Not quite right. Explore more...
 
@@
 The code that supports the results in this paper will be made openly available at https://github.com/TECOLOGYxyz/FlowerTracking.
+A publicly available web application can be accessed from the Github repository through which users can run the tracking algorithm on their own data.
 Raw data as well as the trained flower detection model will be archived on https://zenodo.org/.
@@
 \newpage
+**NOTES**
+
+Online tracking:
+SORT
+DeepSORT
+
+\newpage
 # References\

### 2021-12-17 code/evaluate.py
Parameter testing

@@ -277,17 +277,17 @@ class evaluator():
 #dt = pd.read_csv(r"U:\BITCue\Projekter\NorwayAnnotations\Experiments\2020_06_30_DetectionPhenology\Rmd\ModelQ_Results\DT_GT\submit_boxes_NewFormat_AllTestSeries_Sampled.csv", sep=",")
 #dt = pd.read_csv(r"U:\BITCue\Projekter\NorwayAnnotations\Experiments\2020_06_30_DetectionPhenology\Rmd\ModelQ_Results\DT_GT\2020_02_01_NorwayAnnotations_AllTestSeries_IndividualAnnotations_FRCNN_Metrics_copy.csv", sep=",")
 
-gt = r"U:\BITCue\Projekter\TrackingFlowers\data\annotations\2021_12_13_NorwayAnnotations_NYAA-04_IndividualAnnotations_FRCNN_Metrics.csv"
+gt = r"U:\BITCue\Projekter\TrackingFlowers\data\annotations\2020_05_15_NorwayAnnotations_THUL-01_IndividualAnnotations_FRCNN_Metrics.csv"
 #dt = r"../testResults/6testing_NYAA-04_maxDis_8_runMean_10.csv"
 #dt = r"../testResults\_parameterTest_NYAA-04_1\parameterTest_NYAA-04_maxDisap_30_runMean_30_maxDist_500.csv"
 #dt = r'../testResults\_parameterTest_NYAA-04_1\parameterTest_NYAA-04__maxDisap_30_runMean_30_maxDist_500.csv'
 
 
-path = r'../testResults\_parameterTest_NYAA-04_2/1'
+path = r'../testResults\_parameterTest_THUL-01_3/'
 files = [os.path.join(path,f) for f in os.listdir(path) if f.endswith('.csv')]
 print(files)
 
-results_filename = "../testResults/maxDisap_100_parameterTest_2_NYAA-04_Evaluate.csv"
+results_filename = "../testResults/parameterTest_3_THUL-01_Evaluate.csv"
 
 
 write_header(results_filename)

### 2021-12-17 code/R/Analysis.Rmd
Parameter testing

@@
 ```
+
+parameterTest_3
+```{r}
+d1 = read.csv("U:/BITCue/Projekter/TrackingFlowers/testResults/parameterTest_3_NYAA-04_Evaluate.csv")
+
+d1 %>% 
+  ggplot() +
+  geom_point(aes(x=runMean, y = maxDisap, size = mm, colour = mm)) + 
+  facet_wrap(vars(maxDist))+scale_color_gradient(low="blue", high="red")
+
+d1 %>% 
+  arrange(mm, ascending = T) %>% 
+  slice_head(n = 10) %>% 
+  ggplot() +
+  geom_point(aes(x=runMean, y = maxDisap, size = mm, colour = mm)) + 
+  facet_wrap(vars(maxDist))+scale_color_gradient(low="blue", high="red")
+
+```
+```{r}
+# NYAA-04 gt:
+nyaa04_gt = read.csv("U:/BITCue/Projekter/TrackingFlowers/data/annotations/2021_12_13_NorwayAnnotations_NYAA-04_IndividualAnnotations_FRCNN_Metrics.csv")
+
+length(unique(nyaa04_gt$id_gt))
+
+```
+```{r}
+# Minimum mm: 10,300,60-70,29,0.9835414,38
+# All zero: 0,0,0,68,0.9614075,31
+# Correct noo with minimum mm:  70-160,800-1000,10,36,0.9795687,21
+
+minmm = read.csv("U:/BITCue/Projekter/TrackingFlowers/testResults/_parameterTest_NYAA-04_3/parameterTest_NYAA-04_maxDisap_10_runMean_60_maxDist_300.csv")
+
+allzero = read.csv("U:/BITCue/Projekter/TrackingFlowers/testResults/_parameterTest_NYAA-04_3/parameterTest_NYAA-04_maxDisap_0_runMean_0_maxDist_0.csv")
+
+cornoo = read.csv("U:/BITCue/Projekter/TrackingFlowers/testResults/_parameterTest_NYAA-04_3/parameterTest_NYAA-04_maxDisap_70_runMean_10_maxDist_800.csv")
+
+```

### 2021-12-17 code/filter.py
Put code in class. Added stepwise filtering pipeline.

@@ -4,7 +4,7 @@ Created on Sun Nov 28 20:14:12 2021
 
 @author: au309263
 """
-from shapely.geometry import Polygon
+from shapely.geometry import Polygon, LineString
 import pandas as pd # Replace with cudf if performance is too slow?
 from collections import OrderedDict
 import numpy as np
@@ -33,8 +33,9 @@ TO-DO
 """
 
 #### PATH TO TRACKS ####
-tracks = pd.read_csv(r'../testResults/trackResults8.csv')
+tracks = pd.read_csv(r'../testResults/_parameterTest_NARS-04_3\parameterTest_NARS-04_maxDisap_0_runMean_0_maxDist_0.csv')
 print(tracks)
+
 
 #### NORMALIZE X AND Y ####
 #tracks['x_c'] = tracks['x_c']/6080
@@ -42,51 +43,278 @@ print(tracks)
 
 
 #### FUNCTIONS ####
-def convex_hull(points): # Calculate the convex hull of points
-    print("Calculating convex hulls")
-    #points = np.array([[1, 2],[1, 4],[4, 1], [2,2.5], [3.5,3.5]])
-    hulls = []
-    for p in points:
-        #print(p)
-        p = np.array(p)
-        print("Points: ", p)
-        hull = ConvexHull(p)
-        #print("Hull vertices: ", p[hull.vertices])
-        hulls.append(p[hull.vertices])
+
+class sieve():
+    def __init__(self, tracks):
+        self.tracks = tracks
         
-    return hulls
-
-def cart_product(listOfPolygons):
-    print("Calculating cartesian products of polygons")
-    cartProduct = []
-    #print("List of polygons: ",br, listOfPolygons)
-    listOfPolygons = [i.tolist() for i in listOfPolygons]
-    #print("List of polygons after converstion: ",br, listOfPolygons)
-    for i in listOfPolygons:
-        #print("i", i)
-        for j in listOfPolygons:
-            if not i == j and [j,i] not in cartProduct:
-            #if (i == j).all():# and [j,i] not in cartProduct:
-                #print("jello")
-                cartProduct.append([i,j])
-    return cartProduct
-
-
-tracksGrouped = tracks.groupby(['objectID'])
-
-polygons = []
-
-for i,group in tracksGrouped:
-    if not len(group) <= 3:
-        polygon = []
-        for i,line in group.iterrows():
-            point = [line['x_c'], line['y_c']]
-            polygon.append(point)
-        polygons.append(polygon)
     
[…]

### 2021-12-17 code/visualize.py
Parameter test

@@ -15,26 +15,27 @@ import pandas as pd
 
 
 #df = pd.read_csv(r"U:\BITCue\Projekter\TrackingFlowers\data\annotations\2020_04_30_NorwayAnnotations_NARS-13_IndividualAnnotations_FRCNN_Metrics.csv")
-df = pd.read_csv(r'../testResults/testing_NYAA-04_maxDis_4_runMean_0.csv')
-df2 = pd.read_csv(r'../testResults/9testing_NYAA-04_maxDis_12_runMean_5.csv')
+#df = pd.read_csv(r'../testResults/_parameterTest_NYAA-04_3/parameterTest_NYAA-04_maxDisap_10_runMean_60_maxDist_300.csv')
+df2 = pd.read_csv(r'../testResults/_parameterTest_NYAA-04_3/parameterTest_NYAA-04_maxDisap_10_runMean_60_maxDist_300.csv')
+df3 = pd.read_csv(r'../testResults/_parameterTest_NYAA-04_3/parameterTest_NYAA-04_maxDisap_0_runMean_0_maxDist_0.csv')
 gt = pd.read_csv(r'../data\annotations\2021_12_13_NorwayAnnotations_NYAA-04_IndividualAnnotations_FRCNN_Metrics.csv')
 
+df4 = pd.read_csv(r'../testResults/_parameterTest_NYAA-04_3/parameterTest_NYAA-04_maxDisap_70_runMean_10_maxDist_800.csv')
 
-m = pd.merge(df, gt, on = ['filename', 'x_min', 'x_max', 'y_min', 'y_max'], how = 'inner')
+
+#m = pd.merge(df, gt, on = ['filename', 'x_min', 'x_max', 'y_min', 'y_max'], how = 'inner')
 m2 = pd.merge(df2, gt, on = ['filename', 'x_min', 'x_max', 'y_min', 'y_max'], how = 'inner')
+m3 = pd.merge(df3, gt, on = ['filename', 'x_min', 'x_max', 'y_min', 'y_max'], how = 'inner')
+m4 = pd.merge(df4, gt, on = ['filename', 'x_min', 'x_max', 'y_min', 'y_max'], how = 'inner')
 
 
-plt.scatter(m['frame'], m['id_gt'], c = m['objectID'], s = 15)
-plt.show()
+#plt.scatter(m['frame'], m['id_gt'], c = m['objectID'], s = 15)
+#plt.show()
 
[…]

### 2021-12-17 text/ManuscriptDraft.Rmd
Manusscript update

@@
 The speed of the tracking algorithm depends on the computational power available as well as the number of objects that are being tracked.
+The method is fast, however.
+Tracking of a series containing 85 objects ran at 0.02 seconds per frame.
 
@@
 
-Tracks that overlap have significant risk of errors.
+Manually annotating a subset of the objects in the image series could help derive a good set of parameters.
+
+When deploying the tracking algorithm on naive data that has not been manually tracked, it is not possible to
+
+From our results we see that erroneous tracks often occur as small sections consisting of only few points within the main tracks.
+
+Therefore, tracks that overlap have significant risk of errors.
 Overlapping tracks can be caused by a single flower that was erroneously assigned to several tracks, two flowers that were located sufficiently close to each other that there areas overlapped (e.g.
-when wind moves the flowers around), false positive detections close to a flower.Best case is two flowers that flowered in the same area but were separated by time.
+when wind moves the flowers around), false positive detections close to a flower.
+Best case is two flowers that flowered in the same area but were separated by time.
 Here we will remove overlapping tracks to reduce the risk of error.
+
+We could compare number of points in the overlapping tracks. If one is less than e.g. 10% of the other, we'll remove the small track. If more than 10% we'll remove both?
 

### 2021-12-20 code/R/Analysis.Rmd
Added results from two series

@@
 
-```{r}
-d2 = read.csv("U:/BITCue/Projekter/TrackingFlowers/testResults/parameterTest_2_NYAA-04_Evaluate.csv")
+# parameterTest_3
 
-d2 %>% 
-  ggplot() +
-  geom_point(aes(x=runMean, y = maxDisap, size = mm, colour = mm)) + 
-  facet_wrap(vars(maxDist))+scale_color_gradient(low="blue", high="red")
+## NYAA 04
 
-```
-
-parameterTest_3
 ```{r}
@@
   geom_point(aes(x=runMean, y = maxDisap, size = mm, colour = mm)) + 
+  ggtitle("NYAA-04 Parameter combination") +
   facet_wrap(vars(maxDist))+scale_color_gradient(low="blue", high="red")
@@
   geom_point(aes(x=runMean, y = maxDisap, size = mm, colour = mm)) + 
+  ggtitle("NYAA-04 Ten Lowest Mismatches") +
   facet_wrap(vars(maxDist))+scale_color_gradient(low="blue", high="red")
@@
 ```
+
+## THUL 01
+
+```{r}
+d2 = read.csv("U:/BITCue/Projekter/TrackingFlowers/testResults/parameterTest_3_THUL-01_Evaluate.csv")
+
+d2 %>% 
+  ggplot() +
+  geom_point(aes(x=runMean, y = maxDisap, size = mm, colour = mm)) + 
+  ggtitle("THUL-01 Parameter combination") +
+  facet_wrap(vars(as.factor(maxDist)))
+
+d2 %>% 
+  arrange(mm, ascending = T) %>% 
+  filter(mm == 0) %>% 
+  ggplot() +
+  geom_point(aes(x=runMean, y = maxDisap, size = mm)) + 
+  ggtitle("THUL-01 All zero mismatches") +
+  facet_wrap(vars(maxDist))+scale_color_gradient(low="blue", high="red")
+
+```
+```{r}
+# NYAA-04 gt:
+thul01_gt = read.csv("U:/BITCue/Projekter/TrackingFlowers/data/annotations/2020_05_15_NorwayAnnotations_THUL-01_IndividualAnnotations_FRCNN_Metrics.csv")
+
+length(unique(thul01_gt$id_gt))
+
+```
+```{r}
+# Minimum mm: 0-160, 0 + 200-1000, 10, 0, 1, 6
+# All zero: 0, 0, 0, 8, 0.9885551, 6
+# Correct noo with minimum mm: 0-160, 0 + 200-1000, 10, 0, 1, 6
+
+minmm = read.csv("U:/BITCue/Projekter/TrackingFlowers/testResults/_parameterTest_THUL-01_3/parameterTest_THUL-01_maxDisap_0_runMean_10_maxDist_0.csv")
+
+allzero = read.csv("U:/BITCue/Projekter/TrackingFlowers/testResults/_parameterTest_THUL-01_3/parameterTest_THUL-01_maxDisap_0_runMean_0_maxDist_0.csv")
+
+cornoo = read.csv("U:/BITCue/Projekter/TrackingFlowers/testResults/_parameterTest_THUL-01_3/parameterTest_THUL-01_maxDisap_0_runMean_10_maxDist_0.csv")
+
+```
+
+## NARS-04
+
+```{r}
+d3 = read.csv("U:/BITCue/Projekter/TrackingFlowers/testResults/parameterTest_3_NARS-04_Evaluate.csv")
+
+d3 %>% 
+  ggplot() +
+  geom_point(aes(x=runMean, y = maxDisap, size = mm, colour = mm)) + 
+  ggtitle("NARS-04 Parameter combination") +
+  facet_wrap(vars(as.factor(maxDist))) +
+  scale_color_gradient(low="blue", high="red")
+
+d3 %>% 
+  arrange(mm, ascending = T) %>% 
+  slice_head(n = 10) %>% 
+  ggplot() +
+  geom_point(aes(x=runMean, y = maxDisap, size = mm)) + 
+  ggtitle("NARS-04 Ten lowest mismatches") +
+  facet_wrap(vars(maxDist))+scale_color_gradient(low="blue", high="red")
+
+```
+```{r}
+# NYAA-04 gt:
+NARS04_gt = read.csv("U:/BITCue/Projekter/TrackingFlowers/data/annotations/2020_05_17_NorwayAnnotations_NARS-04_IndividualAnnotations_FRCNN_Metrics.csv")
+
+length(unique(NARS04_gt$id_gt))
+
+```
+
+```{r}
+# Minimum mm: 10, 500, 10, 20, 0.9883991, 32
+# All zero: 0, 0, 0, 85, 0.9506961, 31
+# Correct noo with minimum mm: 10, 500, 10, 20, 0.9883991, 32
+
+minmm = read.csv("U:/BITCue/Projekter/TrackingFlowers/testResults/_parameterTest_NARS-04_3/parameterTest_NARS-04_maxDisap_10_runMean_10_maxDist_500.csv")
+
+allzero = read.csv("U:/BITCue/Projekter/TrackingFlowers/testResults/_parameterTest_NARS-04_3/parameterTest_NARS-04_maxDisap_0_runMean_0_maxDist_0.csv")
+
+cornoo = read.csv("U:/BITCue/Projekter/TrackingFlowers/testResults/_parameterTest_NARS-04_3/parameterTest_NARS-04_maxDisap_10_runMean_10_maxDist_500.csv")
+
+```
+
+## NARS-13
+
+```{r}
+d3 = read.csv("U:/BITCue/Projekter/TrackingFlowers/testResults/parameterTest_3_NARS-04_Evaluate.csv")
+
+d3 %>% 
+  ggplot() +
+  geom_point(aes(x=runMean, y = maxDisap, size = mm, colour = mm)) + 
+  ggtitle("NARS-04 Parameter combination") +
+  facet_wrap(vars(as.factor(maxDist))) +
+  scale_color_gradient(low="blue", high="red")
+
+d3 %>% 
+  arrange(mm, ascending = T) %>% 
+  slice_head(n = 10) %>% 
+  ggplot() +
+  geom_point(aes(x=runMean, y = maxDisap, size = mm)) + 
+  ggtitle("NARS-04 Ten lowest mismatches") +
+  facet_wrap(vars(maxDist))+scale_color_gradient(low="blue", high="red")
+
+```
+```{r}
+# NYAA-04 gt:
+NARS04_gt = read.csv("U:/BITCue/Projekter/TrackingFlowers/data/annotations/2020_05_17_NorwayAnnotations_NARS-04_IndividualAnnotations_FRCNN_Metrics.csv")
+
+length(unique(NARS04_gt$id_gt))
+
+```
+
+```{r}
+# Minimum mm: 10, 500, 10, 20, 0.9883991, 32
+# All zero: 0, 0, 0, 85, 0.9506961, 31
+# Correct noo with minimum mm: 10, 500, 10, 20, 0.9883991, 32
+
+minmm = read.csv("U:/BITCue/Projekter/TrackingFlowers/testResults/_parameterTest_NARS-04_3/parameterTest_NARS-04_maxDisap_10_runMean_10_maxDist_500.csv")
+
+allzero = read.csv("U:/BITCue/Projekter/TrackingFlowers/testResults/_parameterTest_NARS-04_3/parameterTest_NARS-04_maxDisap_0_runMean_0_maxDist_0.csv")
+
+cornoo = read.csv("U:/BITCue/Projekter/TrackingFlowers/testResults/_parameterTest_NARS-04_3/parameterTest_NARS-04_maxDisap_10_runMean_10_maxDist_500.csv")
+
+```

### 2021-12-20 code/evaluate.py
Ran parameter test 3 evaluation

@@ -277,17 +277,17 @@ class evaluator():
 #dt = pd.read_csv(r"U:\BITCue\Projekter\NorwayAnnotations\Experiments\2020_06_30_DetectionPhenology\Rmd\ModelQ_Results\DT_GT\submit_boxes_NewFormat_AllTestSeries_Sampled.csv", sep=",")
 #dt = pd.read_csv(r"U:\BITCue\Projekter\NorwayAnnotations\Experiments\2020_06_30_DetectionPhenology\Rmd\ModelQ_Results\DT_GT\2020_02_01_NorwayAnnotations_AllTestSeries_IndividualAnnotations_FRCNN_Metrics_copy.csv", sep=",")
 
-gt = r"U:\BITCue\Projekter\TrackingFlowers\data\annotations\2020_05_15_NorwayAnnotations_THUL-01_IndividualAnnotations_FRCNN_Metrics.csv"
+gt = r"U:\BITCue\Projekter\TrackingFlowers\data\annotations\2020_04_30_NorwayAnnotations_NARS-13_IndividualAnnotations_FRCNN_Metrics.csv"
 #dt = r"../testResults/6testing_NYAA-04_maxDis_8_runMean_10.csv"
 #dt = r"../testResults\_parameterTest_NYAA-04_1\parameterTest_NYAA-04_maxDisap_30_runMean_30_maxDist_500.csv"
 #dt = r'../testResults\_parameterTest_NYAA-04_1\parameterTest_NYAA-04__maxDisap_30_runMean_30_maxDist_500.csv'
 
 
-path = r'../testResults\_parameterTest_THUL-01_3/'
+path = r'../testResults\_parameterTest_NARS-13_3/'
 files = [os.path.join(path,f) for f in os.listdir(path) if f.endswith('.csv')]
 print(files)
 
-results_filename = "../testResults/parameterTest_3_THUL-01_Evaluate.csv"
+results_filename = "../testResults/parameterTest_3_NARS-13_Evaluate.csv"
 
 
 write_header(results_filename)

### 2021-12-20 code/filter.py
Implemented sieve class. Ready to merge.

@@ -270,17 +270,20 @@ class sieve():
                     polygonB = Polygon(pointsB)
                     
                     if polygonA.intersects(polygonB):
-                        if polygon_lengths[t] == polygon_lengths[t]: # If polygons contain the same number of points, remove both
+                        if polygon_lengths[t] == polygon_lengths[p]: # If polygons contain the same number of points, remove both
                             polygons_to_remove.append(t)
                             polygons_to_remove.append(p)
                         else:
-                            if self.get_change(polygon_lengths[t], polygon_lengths[p]) < 20: # If difference is less than 10%, remove both
+                            if self.get_change(polygon_lengths[t], polygon_lengths[p]) < 30: # If difference is less than 10%, remove both
+                                print(f'Difference is less than 20% ({self.get_change(polygon_lengths[t], polygon_lengths[p])}). Removing {t} and {p}')
                                 polygons_to_remove.append(t)
                                 polygons_to_remove.append(p)
                             else:                                
                                 if polygon_lengths[t] > polygon_lengths[p]: # If difference is more than 20%, remove the smallest
+                                    print(f'{polygon_lengths[t]} for {t} more than {polygon_lengths[p]} for {t}. Removing smallest ({p})')
                                     polygons_to_remove.append(p)
                                 else:
+                                    print(f'{polygon_lengths[p]} for {p} more than {polygon_lengths[t]} for {t}. Removing smallest ({t})')
                                     polygons_to_remove.append(t)
                                     
         print(f'Removing {len(set(polygons_to_remove))} polygons ovelapping with polygons')
@@ -315,9 +318,9 @@ class sieve():
         ### Return the objectIDs that have made their way through the filter
         tracks_to_keep = []
         
-        if flines.keys():
-            for o in flines.keys():
-                tracks_to_keep.append(o)
+        # if flines.keys():
+        #     for o in flines.keys():
+        #         tracks_to_keep.append(o)
         
         if ftriangles.keys():
             for o in ftriangles.keys():
@@ -339,7 +342,7 @@ s = sieve(tracks)
 d,p = s.run()
 
 tracks_filtered = tracks[tracks['objectID'].isin(d)]
[…]

### 2021-12-20 code/visualize.py
Visualized more results

@@ -16,24 +16,30 @@ import pandas as pd
 
 #df = pd.read_csv(r"U:\BITCue\Projekter\TrackingFlowers\data\annotations\2020_04_30_NorwayAnnotations_NARS-13_IndividualAnnotations_FRCNN_Metrics.csv")
 #df = pd.read_csv(r'../testResults/_parameterTest_NYAA-04_3/parameterTest_NYAA-04_maxDisap_10_runMean_60_maxDist_300.csv')
-df2 = pd.read_csv(r'../testResults/_parameterTest_NYAA-04_3/parameterTest_NYAA-04_maxDisap_10_runMean_60_maxDist_300.csv')
-df3 = pd.read_csv(r'../testResults/_parameterTest_NYAA-04_3/parameterTest_NYAA-04_maxDisap_0_runMean_0_maxDist_0.csv')
 gt = pd.read_csv(r'../data\annotations\2021_12_13_NorwayAnnotations_NYAA-04_IndividualAnnotations_FRCNN_Metrics.csv')
 
-df4 = pd.read_csv(r'../testResults/_parameterTest_NYAA-04_3/parameterTest_NYAA-04_maxDisap_70_runMean_10_maxDist_800.csv')
+df2 = pd.read_csv(r'../testResults/_parameterTest_NYAA-04_3/parameterTest_NYAA-04_maxDisap_0_runMean_0_maxDist_0.csv')
+df3 = pd.read_csv(r'../testResults/filtered4_NYAA-04_maxDisap_0_runMean_0_maxDist_0.csv')
+
 
 
 #m = pd.merge(df, gt, on = ['filename', 'x_min', 'x_max', 'y_min', 'y_max'], how = 'inner')
 m2 = pd.merge(df2, gt, on = ['filename', 'x_min', 'x_max', 'y_min', 'y_max'], how = 'inner')
 m3 = pd.merge(df3, gt, on = ['filename', 'x_min', 'x_max', 'y_min', 'y_max'], how = 'inner')
-m4 = pd.merge(df4, gt, on = ['filename', 'x_min', 'x_max', 'y_min', 'y_max'], how = 'inner')
+#m4 = pd.merge(df4, gt, on = ['filename', 'x_min', 'x_max', 'y_min', 'y_max'], how = 'inner')
 
 
 #plt.scatter(m['frame'], m['id_gt'], c = m['objectID'], s = 15)
 #plt.show()
 
[…]

### 2021-12-20 webapp.py
Added sieve option

@@ -8,6 +8,7 @@ Created on Mon Dec 13 15:30:13 2021
 import streamlit as st
 import pandas as pd
 from code.track import tracker
+from code.filter import sieve
 import time
 import matplotlib.pyplot as plt
 from PIL import Image
@@ -164,6 +165,24 @@ if trackButton:
 
         st.pyplot(fig)
         
+        
+        filterkButton = st.sidebar.button('FILTER')
+
+if filterkButton:   
+    
+    s = sieve(p)
+    d,polyhulls = s.run()
+    tracks_filtered = p[p['objectID'].isin(d)]
+
+
+    with plot2:
+        
+        st.header("Results")
+    
+        fig, ax2 = plt.subplots()
+        ax2.scatter(x = tracks_filtered['x_c'], y = tracks_filtered['y_c'], c = tracks_filtered['objectID'], s = 15)
+
+        st.pyplot(fig)
 
 
     

### 2021-12-20 code/filtering.py
Renamed script

@@ -1,0 +1,566 @@
+# -*- coding: utf-8 -*-
+"""
+Created on Sun Nov 28 20:14:12 2021
+
+@author: au309263
+"""
+from shapely.geometry import Polygon, LineString
+import pandas as pd # Replace with cudf if performance is too slow?
+from collections import OrderedDict
+import numpy as np
+from scipy.spatial import distance as dist
+from scipy.spatial import ConvexHull
+#from scipy.ndimage.filters import uniform_filter1d
+import matplotlib.pyplot as plt
+import time
+
+br = "\n"
+
+
+
+"""
+TO-DO
+
+- Remmove single points
+- Establish straight line between two-point tracks
+- Establish triangle between three-point tracks
+- Establish polygon from vertices of the convex hull of n-point tracks (n > 3)
+- Check if any lines overlap and remove both if they do
+- Check if any remaining lines overlap with any polygons, remove both line and polygon if so
+- Check if any polygons overlap and remove both if they do
+- Output filtered track file
+
+"""
+
+#### PATH TO TRACKS ####
+tracks = pd.read_csv(r'../testResults/_parameterTest_NYAA-04_3\parameterTest_NYAA-04_maxDisap_0_runMean_0_maxDist_0.csv')
+#print(tracks)
+
+
+#### NORMALIZE X AND Y ####
+#tracks['x_c'] = tracks['x_c']/6080
+#tracks['y_c'] = tracks['y_c']/3420
+
+
+fig, ax0 = plt.subplots(figsize=(15,10))
+ax0.set_xlim(0, 6080)
+ax0.set_ylim(0, 3420)
+scat0 = ax0.scatter(tracks['x_c'], tracks['y_c'], c = tracks['objectID'], s = 40)
+
+
+
+
+#### FUNCTIONS ####
+
+class sieve():
+    def __init__(self, tracks):
+        self.tracks = tracks
+        
+        
+    def gen(self):
+        t = tracks.groupby('objectID')
+        for i,g in t: # Generator that yields next group in tracks dataframe
+            oid = g['objectID'].iloc[0]
+            x_cs = g['x_c'].tolist()
+            y_cs = g['y_c'].tolist()
+            p = list(map(list, zip(x_cs, y_cs)))
+            yield oid, p
+    
+    
+    def separate(self, tracks):
+        
+        points = {}
+        lines = {}
+        triangles = {}
+        polygons = {}
+        polyLengths = {}
+        
[…]

### 2021-12-20 code/run.py
Added filtering to run

@@ -7,6 +7,7 @@ Created on Tue Dec 14 12:50:10 2021
 
 import pandas as pd
 from track import tracker
+from filtering import sieve
 import time
 
 # ===================== Run tracking on several combinations of parameters ====
@@ -58,3 +59,34 @@ for i in list_of_parameters:
     print(f'maxDis: {i[0]}, runMean: {i[1]} - Number of tracks found: {len(tracks.objectID.unique())}')
 
 # =============================================================================
+
+
+
+
+
+### FILTER ###
+
+s = sieve(tracks)
+
+d,p = s.run()
+
+tracks_filtered = tracks[tracks['objectID'].isin(d)]
+tracks_filtered.to_csv(r'U:\BITCue\Projekter\TrackingFlowers\testResults/filtered4_NYAA-04_maxDisap_0_runMean_0_maxDist_0.csv')
+
+
+fig, ax1 = plt.subplots(figsize=(15,10))
+ax1.set_xlim(0, 6080)
+ax1.set_ylim(0, 3420)
+
+scat1 = ax1.scatter(tracks_filtered['x_c'], tracks_filtered['y_c'], c = tracks_filtered['objectID'], s = 40)
+
+print(p)
+
+# for h in p:
+#     print(p[h])
+    
+#     #print(h)
+#     ax1.scatter(p[h][::1,0], p[h][::1,1], zorder = -1, s = 2)
+#     # ax1.fill(h[::1,0], h[::1,1], zorder = -1)
+
+hulls = [[[5,2],[1,3],[8,4]],[[9,4],[4,6],[10,20]] ]

### 2021-12-20 webapp.py
Fixed sieve import

@@ -8,7 +8,7 @@ Created on Mon Dec 13 15:30:13 2021
 import streamlit as st
 import pandas as pd
 from code.track import tracker
-from code.filter import sieve
+from code.filtering import sieve
 import time
 import matplotlib.pyplot as plt
 from PIL import Image

### 2021-12-20 code/filtering.py
Removed data import

@@ -32,24 +32,6 @@ TO-DO
 
 """
 
-#### PATH TO TRACKS ####
-tracks = pd.read_csv(r'../testResults/_parameterTest_NYAA-04_3\parameterTest_NYAA-04_maxDisap_0_runMean_0_maxDist_0.csv')
-#print(tracks)
-
-
-#### NORMALIZE X AND Y ####
-#tracks['x_c'] = tracks['x_c']/6080
-#tracks['y_c'] = tracks['y_c']/3420
-
-
-fig, ax0 = plt.subplots(figsize=(15,10))
-ax0.set_xlim(0, 6080)
-ax0.set_ylim(0, 3420)
-scat0 = ax0.scatter(tracks['x_c'], tracks['y_c'], c = tracks['objectID'], s = 40)
-
-
-
-
 #### FUNCTIONS ####
 
 class sieve():
@@ -58,7 +40,7 @@ class sieve():
         
         
     def gen(self):
-        t = tracks.groupby('objectID')
+        t = self.tracks.groupby('objectID')
         for i,g in t: # Generator that yields next group in tracks dataframe
             oid = g['objectID'].iloc[0]
             x_cs = g['x_c'].tolist()

### 2021-12-21 code/R/Analysis.Rmd
Code update

@@
 
-nyaa04 = read.csv("../../data/annotations/2020_05_23_NorwayAnnotations_Nyaa-04_IndividualAnnotations_FRCNN.csv", header = F)
+nyaa04 = read.csv("../../data/annotations/2021_12_13_NorwayAnnotations_NYAA-04_IndividualAnnotations_FRCNN_Metrics.csv", header = T)
 
@@
 ```{r}
-tracksThul01 = read.csv('../../trackResults4.csv',stringsAsFactors = FALSE)
-
-tracksThul01$x_c = as.numeric(tracksThul01$x_C)
-
-tracksThul01 %>% 
-  ggplot() +
-  geom_point(aes(x = x_c, y = y_c, color = as.factor(objectID))) + 
-  xlim(0,6080) +
-  ylim(0,3420)
-```
-
-```{r}
 
@@
 ggplot()+
-  geom_point(aes(x = flower_start, y = flower_length, colour = CameraID)) +
-geom_smooth(aes(x = flower_start, y = flower_length), method=lm)
+  geom_point(aes(x = flower_start, y = flower_length, colour = CameraID))# +
+#geom_smooth(aes(x = flower_start, y = flower_length), method=lm)
 
@@
   geom_point(aes(x = flower_start, y = flower_length, colour = CameraID))
-```
-
-```{r}
-d = read.csv("U:/BITCue/Projekter/TrackingFlowers/code/7temptest_eval.csv")
-
-d %>% 
-  ggplot() +
-  geom_point(aes(x=runMean, y = maxDisap, size = mm, colour = mm)) + 
-  facet_wrap(vars(maxDist))+scale_color_gradient(low="blue", high="red")
 ```
@@
 
-d2 %>% 
+THUL01_p1 = d2 %>% 
   ggplot() +
@@
 
-d2 %>% 
+THUL01_p2 = d2 %>% 
   arrange(mm, ascending = T) %>% 
@@
 
+THUL01_p1
+THUL01_p2
+
+ggsave('U:/BITCue/Projekter/TrackingFlowers/testResults/figures/THUL-01_AllCombinations.png', THUL01_p1)
+ggsave('U:/BITCue/Projekter/TrackingFlowers/testResults/figures/THUL-01_AllZero.png', THUL01_p2)
 ```
@@
 ```{r}
-d3 = read.csv("U:/BITCue/Projekter/TrackingFlowers/testResults/parameterTest_3_NARS-04_Evaluate.csv")
+d4 = read.csv("U:/BITCue/Projekter/TrackingFlowers/testResults/parameterTest_3_NARS-13_Evaluate.csv")
 
-d3 %>% 
+NARS13_p1 = d4 %>% 
   ggplot() +
   geom_point(aes(x=runMean, y = maxDisap, size = mm, colour = mm)) + 
-  ggtitle("NARS-04 Parameter combination") +
+  ggtitle("NARS-13 Parameter combination") +
   facet_wrap(vars(as.factor(maxDist))) +
@@
 
-d3 %>% 
+NARS13_p2 = d4 %>% 
   arrange(mm, ascending = T) %>% 
@@
   geom_point(aes(x=runMean, y = maxDisap, size = mm)) + 
-  ggtitle("NARS-04 Ten lowest mismatches") +
+  ggtitle("NARS-13 Ten lowest mismatches") +
   facet_wrap(vars(maxDist))+scale_color_gradient(low="blue", high="red")
+
+NARS13_p1
+NARS13_p2
+
+ggsave('U:/BITCue/Projekter/TrackingFlowers/testResults/figures/NARS-13_AllCombinations.png', NARS13_p1)
+ggsave('U:/BITCue/Projekter/TrackingFlowers/testResults/figures/NARS-13_TenLowest.png', NARS13_p2)
 
@@
 # NYAA-04 gt:
-NARS04_gt = read.csv("U:/BITCue/Projekter/TrackingFlowers/data/annotations/2020_05_17_NorwayAnnotations_NARS-04_IndividualAnnotations_FRCNN_Metrics.csv")
+NARS13_gt = read.csv("U:/BITCue/Projekter/TrackingFlowers/data/annotations/2020_04_30_NorwayAnnotations_NARS-13_IndividualAnnotations_FRCNN_Metrics.csv")
 
-length(unique(NARS04_gt$id_gt))
+length(unique(NARS13_gt$id_gt))
 
@@
 
-minmm = read.csv("U:/BITCue/Projekter/TrackingFlowers/testResults/_parameterTest_NARS-04_3/parameterTest_NARS-04_maxDisap_10_runMean_10_maxDist_500.csv")
+minmm = read.csv("U:/BITCue/Projekter/TrackingFlowers/testResults/_parameterTest_NARS-13_3/parameterTest_NARS-13_maxDisap_0_runMean_0_maxDist_0.csv")
 
-allzero = read.csv("U:/BITCue/Projekter/TrackingFlowers/testResults/_parameterTest_NARS-04_3/parameterTest_NARS-04_maxDisap_0_runMean_0_maxDist_0.csv")
+allzero = read.csv("U:/BITCue/Projekter/TrackingFlowers/testResults/_parameterTest_NARS-13_3/parameterTest_NARS-13_maxDisap_0_runMean_0_maxDist_0.csv")
 
-cornoo = read.csv("U:/BITCue/Projekter/TrackingFlowers/testResults/_parameterTest_NARS-04_3/parameterTest_NARS-04_maxDisap_10_runMean_10_maxDist_500.csv")
+cornoo = read.csv("U:/BITCue/Projekter/TrackingFlowers/testResults/_parameterTest_NARS-13_3/parameterTest_NARS-13_maxDisap_10_runMean_10_maxDist_0.csv")
 

### 2021-12-21 code/filter.py
Changed function for filtering overlapping polygons

@@ -1,589 +1,0 @@
-# -*- coding: utf-8 -*-
-"""
-Created on Sun Nov 28 20:14:12 2021
-
-@author: au309263
-"""
-from shapely.geometry import Polygon, LineString
-import pandas as pd # Replace with cudf if performance is too slow?
-from collections import OrderedDict
-import numpy as np
-from scipy.spatial import distance as dist
-from scipy.spatial import ConvexHull
-#from scipy.ndimage.filters import uniform_filter1d
-import matplotlib.pyplot as plt
-import time
-
-br = "\n"
-
-
-
-"""
-TO-DO
-
-- Remmove single points
-- Establish straight line between two-point tracks
-- Establish triangle between three-point tracks
-- Establish polygon from vertices of the convex hull of n-point tracks (n > 3)
-- Check if any lines overlap and remove both if they do
-- Check if any remaining lines overlap with any polygons, remove both line and polygon if so
-- Check if any polygons overlap and remove both if they do
-- Output filtered track file
-
-"""
-
-#### PATH TO TRACKS ####
-tracks = pd.read_csv(r'../testResults/_parameterTest_NYAA-04_3\parameterTest_NYAA-04_maxDisap_0_runMean_0_maxDist_0.csv')
-#print(tracks)
-
-
-#### NORMALIZE X AND Y ####
-#tracks['x_c'] = tracks['x_c']/6080
-#tracks['y_c'] = tracks['y_c']/3420
-
-
-fig, ax0 = plt.subplots(figsize=(15,10))
-ax0.set_xlim(0, 6080)
-ax0.set_ylim(0, 3420)
-scat0 = ax0.scatter(tracks['x_c'], tracks['y_c'], c = tracks['objectID'], s = 40)
-
-
-
-
-#### FUNCTIONS ####
-
-class sieve():
-    def __init__(self, tracks):
-        self.tracks = tracks
-        
-        
-    def gen(self):
-        t = tracks.groupby('objectID')
-        for i,g in t: # Generator that yields next group in tracks dataframe
-            oid = g['objectID'].iloc[0]
-            x_cs = g['x_c'].tolist()
-            y_cs = g['y_c'].tolist()
-            p = list(map(list, zip(x_cs, y_cs)))
-            yield oid, p
-    
-    
-    def separate(self, tracks):
-        
-        points = {}
-        lines = {}
-        triangles = {}
-        polygons = {}
-        polyLengths = {}
-        
[…]

### 2021-12-21 code/filtering.py

@@ -13,6 +13,8 @@ from scipy.spatial import ConvexHull
 #from scipy.ndimage.filters import uniform_filter1d
 import matplotlib.pyplot as plt
 import time
+import itertools
+
 
 br = "\n"
 
@@ -246,39 +248,120 @@ class sieve():
 
     def filter_polygons_on_polygons(self, list_of_polygons, polygon_lengths):
         polygons_to_remove = []
+        print(f'Polygon lengths: {br}{polygon_lengths}')
+        # for p in list_of_polygons:
+        #     points = list_of_polygons[p]
+        #     pol = Polygon(points)
+        #     polygons[p] = pol
+        # print(polygons)
+
+        def bla(two_polygons):
+            if polygon_lengths[two_polygons[0]] == polygon_lengths[two_polygons[1]]: # If polygons contain the same number of points, remove both
+                polygons_to_remove.extend(two_polygons)
+                
+            else:
+                if self.get_change(polygon_lengths[two_polygons[0]], polygon_lengths[two_polygons[1]]) < 40: # If difference is less than 10%, remove both
+                    print(f'Difference is less than 20% ({self.get_change(polygon_lengths[two_polygons[0]], polygon_lengths[two_polygons[1]])}). Removing {two_polygons[0]} {polygon_lengths[two_polygons[0]]} points  and {two_polygons[1]} {polygon_lengths[two_polygons[1]]} points')
+                    polygons_to_remove.extend(two_polygons)
+                else:                                
+                    if polygon_lengths[two_polygons[0]] > polygon_lengths[two_polygons[1]]: # If difference is more than 20%, remove the smallest
+                        print(f'{polygon_lengths[two_polygons[0]]} for {two_polygons[0]} more than {polygon_lengths[two_polygons[1]]} for {two_polygons[1]}. Removing smallest ({two_polygons[1]}) ({self.get_change(polygon_lengths[two_polygons[0]],polygon_lengths[two_polygons[1]])})')
+                        polygons_to_remove.append(two_polygons[1])
+                    else:
+                        print(f'{polygon_lengths[two_polygons[1]]} for {polygon_lengths[two_polygons[1]]} more than {polygon_lengths[two_polygons[0]]} for {polygon_lengths[two_polygons[0]]}. Removing smallest ({two_polygons[0]} ({self.get_change(polygon_lengths[two_polygons[1]], polygon_lengths[two_polygons[0]])})')
+                        polygons_to_remove.append(two_polygons[0])
+
+
         
[…]

### 2021-12-21 code/run.py
Run test

@@ -8,6 +8,8 @@ Created on Tue Dec 14 12:50:10 2021
 import pandas as pd
 from track import tracker
 from filtering import sieve
+
+
 import time
 
 # ===================== Run tracking on several combinations of parameters ====
@@ -64,7 +66,7 @@ import time
 # #### PATH TO TRACKS ####
 #tracks = pd.read_csv(r'../testResults/_parameterTest_NYAA-04_3\parameterTest_NYAA-04_maxDisap_0_runMean_0_maxDist_0.csv')
 # #print(tracks)
-tracks = pd.read_csv(r'U:\BITCue\Projekter\TrackingFlowers\testResults\_parameterTest_NYAA-04_3\parameterTest_NYAA-04_maxDisap_10_runMean_10_maxDist_300.csv')
+tracks = pd.read_csv(r'U:\BITCue\Projekter\TrackingFlowers\testResults\_parameterTest_NARS-13_3\parameterTest_NARS-13_maxDisap_10_runMean_10_maxDist_300.csv')
 
 # #### NORMALIZE X AND Y ####
 # #tracks['x_c'] = tracks['x_c']/6080
@@ -87,7 +89,7 @@ d,p = s.run()
 print(p)
 
 tracks_filtered = tracks[tracks['objectID'].isin(d)]
-tracks_filtered.to_csv(r'U:\BITCue\Projekter\TrackingFlowers\testResults/filtered16_NYAA-04_maxDisap_10_runMean_10_maxDist_300.csv')
+tracks_filtered.to_csv(r'U:\BITCue\Projekter\TrackingFlowers\testResults/filtered_NARS-13_maxDisap_10_runMean_10_maxDist_300.csv')
 
 
 # fig, ax1 = plt.subplots(figsize=(15,10))

### 2021-12-21 code/visualize.py
Run test

@@ -16,20 +16,41 @@ import pandas as pd
 
 #df = pd.read_csv(r"U:\BITCue\Projekter\TrackingFlowers\data\annotations\2020_04_30_NorwayAnnotations_NARS-13_IndividualAnnotations_FRCNN_Metrics.csv")
 #df = pd.read_csv(r'../testResults/_parameterTest_NYAA-04_3/parameterTest_NYAA-04_maxDisap_10_runMean_60_maxDist_300.csv')
-gt = pd.read_csv(r'../data\annotations\2021_12_13_NorwayAnnotations_NYAA-04_IndividualAnnotations_FRCNN_Metrics.csv')
+gt = pd.read_csv(r'../data\annotations\2020_04_30_NorwayAnnotations_NARS-13_IndividualAnnotations_FRCNN_Metrics.csv')
 
-df2 = pd.read_csv(r'../testResults/_parameterTest_NYAA-04_3/parameterTest_NYAA-04_maxDisap_10_runMean_10_maxDist_300.csv')
-df3 = pd.read_csv(r'../testResults/filtered16_NYAA-04_maxDisap_10_runMean_10_maxDist_300.csv')
+df2 = pd.read_csv(r'../testResults/_parameterTest_NARS-13_3/parameterTest_NARS-13_maxDisap_10_runMean_10_maxDist_300.csv')
+df3 = pd.read_csv(r'../testResults/filtered_NARS-13_maxDisap_10_runMean_10_maxDist_300.csv')
 
 #m = pd.merge(df, gt, on = ['filename', 'x_min', 'x_max', 'y_min', 'y_max'], how = 'inner')
 m2 = pd.merge(df2, gt, on = ['filename', 'x_min', 'x_max', 'y_min', 'y_max'], how = 'inner')
 m3 = pd.merge(df3, gt, on = ['filename', 'x_min', 'x_max', 'y_min', 'y_max'], how = 'inner')
 
 
-m2.to_csv(r'U:\BITCue\Projekter\TrackingFlowers\testResults/NYAA04_Merged_m2.csv')
+#m2.to_csv(r'U:\BITCue\Projekter\TrackingFlowers\testResults/NYAA04_Merged_m2_3.csv')
 
-m3.to_csv(r'U:\BITCue\Projekter\TrackingFlowers\testResults/NYAA04_Merged.csv')
+#m3.to_csv(r'U:\BITCue\Projekter\TrackingFlowers\testResults/NYAA04_Merged_m3_3.csv')
+#print(m3)
[…]

### 2021-12-21 code/DBSCANFiltering.py
Experimental code for filtering polygons with DBSCAN

@@ -1,0 +1,86 @@
+# -*- coding: utf-8 -*-
+"""
+Created on Tue Dec 21 13:28:48 2021
+
+@author: au309263
+"""
+
+import geopandas as gpd
+import pandas as pd
+from shapely.geometry import Polygon
+from sklearn.cluster import DBSCAN
+import matplotlib.pyplot as plt
+from filtering import sieve
+
+# Note, EPS_DISTANCE = 20 is a magic number and it needs to be
+# * smaller than the gap between any two islands
+# * large enough to cluster polygons in one island in same cluster
+EPS_DISTANCE = 4
+MIN_SAMPLE_POLYGONS = 1
+
+
+#tracks = pd.read_csv(r'U:\BITCue\Projekter\TrackingFlowers\testResults\_parameterTest_NARS-13_3\parameterTest_NARS-13_maxDisap_10_runMean_10_maxDist_300.csv')
+tracks = pd.read_csv(r'U:\BITCue\Projekter\TrackingFlowers\data\annotations\2020_05_15_NorwayAnnotations_THUL-01_IndividualAnnotations_FRCNN_Metrics.csv')
+### FILTER ###
+
+# s = sieve(tracks)
+
+# d,p = s.run()
+
+# print(p)
+
+# tracks_filtered = tracks[tracks['objectID'].isin(d)]
+# tracks_filtered.to_csv(r'U:\BITCue\Projekter\TrackingFlowers\testResults/filtered20_NYAA-04_maxDisap_10_runMean_10_maxDist_300.csv')
+
+polys = []
+
+
+tracks['x_c'] = (tracks['x_min'] + tracks['x_max']) / 2
+tracks['y_c'] = (tracks['y_min'] + tracks['y_max']) / 2
+ 
+grouped = tracks.groupby(['id_gt'])   
+
+
+for i,g in grouped:
+    xs = g['x_c'].tolist()
+    ys = g['x_c'].tolist()
+    k = list(zip(xs,ys))
+    polys.append(Polygon(k))
+    
+# for t in p:
+#     pointsA = p[t]
+#     polyA = Polygon(pointsA)
+#     polys.append(polyA)    
+
+print(polys)
+# p1=Polygon([(1,1), (1,3), (3,4), (4,2), (3,1)])
+# p2=Polygon([(3,3),(2,5),(5,6),(5,4)])
[…]

### 2021-12-21 code/filtering.py
Cleaning

@@ -117,9 +117,6 @@ class sieve():
                     print(f'Adding object {k} with points {polygons[k]} to lines')    
                     self.lines[k] = polygons[k] # Add the points to the line dictionary
                 
-                
-        
-            
         
 
     # def cart_product(listOfPolygons):
@@ -150,7 +147,7 @@ class sieve():
         		self.x = x
         		self.y = y
         
-        def ccw(A,B,C):
+        def ccw(A,B,C): # Look for intersection between lines
         	return (C.y-A.y)*(B.x-A.x) > (B.y-A.y)*(C.x-A.x)
         
         def intersect(A,B,C,D):
@@ -249,13 +246,8 @@ class sieve():
     def filter_polygons_on_polygons(self, list_of_polygons, polygon_lengths):
         polygons_to_remove = []
         print(f'Polygon lengths: {br}{polygon_lengths}')
-        # for p in list_of_polygons:
-        #     points = list_of_polygons[p]
-        #     pol = Polygon(points)
-        #     polygons[p] = pol
-        # print(polygons)
 
-        def bla(two_polygons):
+        def filter_two_polygons(two_polygons):
             if polygon_lengths[two_polygons[0]] == polygon_lengths[two_polygons[1]]: # If polygons contain the same number of points, remove both
                 polygons_to_remove.extend(two_polygons)
                 
@@ -311,58 +303,17 @@ class sieve():
                         print(f'Keep these: {polygonKeys[:2]}')
                         
                         polygonKeys = polygonKeys[:2]
-                        bla(polygonKeys)
+                        filter_two_polygons(polygonKeys)
                     if len(polygonKeys) == 2:
                         print("Length is two.", polygonKeys)
-                        bla(polygonKeys)
+                        filter_two_polygons(polygonKeys)
                 else:
                     overlaps = {}
+                    
         print(f'Removing {len(set(polygons_to_remove))} polygons ovelapping with polygons')
         list_of_polygons = self.remove_from_list(list_of_polygons, set(polygons_to_remove))    
         return list_of_polygons
                         
-                # for poly1, poly2 in  itertools.combinations(list_of_polygons, 2):
-                #     poly1 = Polygon(poly1)
-                #     poly2 = Polygon(poly2)
-                #     if poly1.intersects(poly2):
-                #           polygon_list_B.append(poly1)
-                #           polygon_list_B.append(poly2)
-                #  result = [geom for geom in polygon_list_A if geom not in polygon_list_B]
-        
-        
-        # for t in list_of_polygons:
[…]

### 2021-12-30 code/R/Analysis.Rmd
Added NARS-17

@@
 
+nars17 = read.csv("../../data/annotations/2021_12_30_NorwayAnnotations_NARS-17_IndividualAnnotations_FRCNN_Metrics.csv", header = T)
+
+colnames(nars17) = c("FileName", "x_min", "y_min", "x_max", "y_max", "Class")
+
+nars17 = nars17 %>%
+  mutate(FileName = str_extract(FileName, regex("[A-Z]{4,5}-\\d{0,3}_\\d{6}.JPG")))
+
 ```
@@
 nars13_meta = merge(nars13, meta_2019, by.x = "FileName")
+nars13_meta$time = nars13_meta$DOY + (nars13_meta$Hour/24)
+nars13_meta %>%
 
-nars13_meta$time = nars13_meta$DOY + (nars13_meta$Hour/24)
-
-nars13_meta %>% 
 ggplot()+
@@
 nars13_meta %>% 
+ggplot()+
+  geom_point(aes(x = flower_start, y = flower_length))
+
+nars17_meta = merge(nars17, meta_2019, by.x = "FileName")
+nars17_meta$time = nars17_meta$DOY + (nars17_meta$Hour/24)
+
+nars17_meta %>% 
+ggplot()+
+  geom_point(aes(x = time, y = Class))
+
+nars17_meta = nars17_meta %>%
+  group_by(Class) %>% 
+  mutate(flower_start = min(time),
+    flower_length = max(time) - min(time))
+
+nars17_meta %>% 
 ggplot()+
@@
 nyaa04 %>% 
+  ggplot() +
+  geom_point(aes(x = x_c, y = y_c, color = Class), size = 1) + 
+  xlim(0,6080) +
+  ylim(0,3420) + 
+  theme(legend.position = "none")
+
+nars17$x_c = (nars17$x_min + nars17$x_max) / 2
+nars17$y_c = (nars17$y_min + nars17$y_max) / 2
+
+nars17 %>% 
+  ggplot() +
+  geom_jitter(aes(x = x_c, y = y_c, color = Class), size = 5) + 
+  xlim(0,6080) +
+  ylim(0,3420) + 
+  theme(legend.position = "none")
+
+nars13$x_c = (nars13$x_min + nars13$x_max) / 2
+nars13$y_c = (nars13$y_min + nars13$y_max) / 2
+
+nars13 %>% 
   ggplot() +
@@
 
-all = bind_rows(thul01_meta[,c("flower_start","flower_length", "CameraID")], nars04_meta[,c("flower_start","flower_length", "CameraID")], nars13_meta[,c("flower_start","flower_length", "CameraID")], nyaa04_meta[,c("flower_start","flower_length", "CameraID")])
+all = bind_rows(thul01_meta[,c("flower_start","flower_length", "CameraID")], nars04_meta[,c("flower_start","flower_length", "CameraID")], nars13_meta[,c("flower_start","flower_length", "CameraID")], nyaa04_meta[,c("flower_start","flower_length", "CameraID")], nars17_meta[,c("flower_start","flower_length", "CameraID")])
 
@@
 
-```{r}
-d1 = read.csv("U:/BITCue/Projekter/TrackingFlowers/code/8temptest_eval_THUL01.csv")
-
-d1 %>% 
-  ggplot() +
-  geom_point(aes(x=runMean, y = maxDisap, size = mm, colour = mm)) + 
-  facet_wrap(vars(maxDist))+scale_color_gradient(low="blue", high="red")
-
-```
-
 # parameterTest_3
@@
 
-d1 %>% 
+NYAA04_p1 = d1 %>% 
   ggplot() +
@@
 
-d1 %>% 
+NYAA04_p2 = d1 %>% 
   arrange(mm, ascending = T) %>% 
@@
   facet_wrap(vars(maxDist))+scale_color_gradient(low="blue", high="red")
+
+NYAA04_p1
+NYAA04_p2
+
+ggsave('U:/BITCue/Projekter/TrackingFlowers/testResults/figures/NYAA-04_AllCombinations.png', NYAA04_p1)
+ggsave('U:/BITCue/Projekter/TrackingFlowers/testResults/figures/NYAA-04_TenLowest.png', NYAA04_p2)
 
@@
 
-d3 %>% 
+NARS04_p1 = d3 %>% 
   ggplot() +
@@
 
-d3 %>% 
+NARS04_p2 = d3 %>% 
   arrange(mm, ascending = T) %>% 
@@
 
+ggsave('U:/BITCue/Projekter/TrackingFlowers/testResults/figures/NARS-04_AllCombinations.png', NARS04_p1)
+ggsave('U:/BITCue/Projekter/TrackingFlowers/testResults/figures/NARS-04_TenLowest.png', NARS04_p2)
 ```
@@
 ```
+
+```{r}
+d5 = read.csv("U:/BITCue/Projekter/TrackingFlowers/testResults/parameterTest_3_NARS-17_Evaluate.csv")
+
+NARS17_p1 = d5 %>% 
+  ggplot() +
+  geom_point(aes(x=runMean, y = maxDisap, size = mm, colour = mm)) + 
+  ggtitle("NARS-13 Parameter combination") +
+  facet_wrap(vars(as.factor(maxDist))) +
+  scale_color_gradient(low="blue", high="red")
+
+NARS17_p2 = d5 %>% 
+  arrange(mm, ascending = T) %>% 
+  filter(mm == 0) %>% 
+  ggplot() +
+  geom_point(aes(x=runMean, y = maxDisap, size = mm)) + 
+  ggtitle("THUL-01 All zero mismatches") +
+  facet_wrap(vars(maxDist))+scale_color_gradient(low="blue", high="red")
+
+NARS17_p1
+NARS17_p2
+
+ggsave('U:/BITCue/Projekter/TrackingFlowers/testResults/figures/NARS-17_AllCombinations.png', NARS17_p1)
+ggsave('U:/BITCue/Projekter/TrackingFlowers/testResults/figures/NARS-17_TenLowest.png', NARS17_p2)
+
+```
+
 ```{r}

### 2021-12-30 code/analyze.py
Added NARS-17

@@ -15,7 +15,8 @@ import pandas as pd
 import numpy as np
 
 # Input ground truth annotations
-df = pd.read_csv(r"data\annotations\2020_05_23_NorwayAnnotations_NYAA-04_IndividualAnnotations_FRCNN_Metrics.csv")
+df = pd.read_csv(r"../data\annotations\2020_04_30_NorwayAnnotations_NARS-13_IndividualAnnotations_FRCNN_Metrics.csv")
+
 
 #print(thul01)
 
@@ -33,7 +34,10 @@ def dist(a,b):
     d = np.linalg.norm(a-b)
     return d
 
-#print(dist)
+def ave(lst):
+    return sum(lst) / len(lst)
+
+
 
 df_grouped = df.groupby(['id_gt'])
 
@@ -57,10 +61,45 @@ for i, g in df_grouped:
 print(f'Largest dist found: {largestDist}')
 
 # THUL-01: 449.00334074481003
-# NYAA-04: 2307.41034495384
-# NARS-13: 782.0206199327483
+# NYAA-04: 632.9178461696273
+# NARS-13: 782.0206199327483 QC
 # NARS-04: 557.953851138246
+# NARS-17: 615.5536126122565
 
+
+# Largest distance between mean of track and each point
+largestDist = 0
+
+for i, g in df_grouped:
+    #print(g)
+    x_cs = []
+    y_cs = []
+    g_points = []
+    for i,l in g.iterrows():
+        x_cs.append(l['x_c'])
+        y_cs.append(l['y_c'])
+        g_points.append(np.array([l['x_c'], l['y_c']]))
+
+    mx = ave(x_cs)
+    my = ave(y_cs)
+    
+    m = np.array([mx, my]) # Mean of x and mean of y
+    for p1 in g_points:
+        d = dist(p1, m)
+        if d > largestDist:
+            print(f'{d} is larger than {largestDist}. Storing {d}')
+            print(f'Distance calculated between {p1} and {m}')
+            largestDist = d
+            
+            
+print(f'Largest dist found: {largestDist}')
+
+
+# THUL-01: 278.2451605832456
+# NYAA-04: 
+# NARS-13: 456.82260896011803
+# NARS-04: 
+# NARS-17: 376.3443382217583
     
 
 # Find the largest distance between two points in a track. This will help us set max_distance in the tracking algorithm.

### 2021-12-30 code/evaluate.py

@@ -277,17 +277,17 @@ class evaluator():
 #dt = pd.read_csv(r"U:\BITCue\Projekter\NorwayAnnotations\Experiments\2020_06_30_DetectionPhenology\Rmd\ModelQ_Results\DT_GT\submit_boxes_NewFormat_AllTestSeries_Sampled.csv", sep=",")
 #dt = pd.read_csv(r"U:\BITCue\Projekter\NorwayAnnotations\Experiments\2020_06_30_DetectionPhenology\Rmd\ModelQ_Results\DT_GT\2020_02_01_NorwayAnnotations_AllTestSeries_IndividualAnnotations_FRCNN_Metrics_copy.csv", sep=",")
 
-gt = r"U:\BITCue\Projekter\TrackingFlowers\data\annotations\2020_04_30_NorwayAnnotations_NARS-13_IndividualAnnotations_FRCNN_Metrics.csv"
+gt = r"U:\BITCue\Projekter\TrackingFlowers\data\annotations\2021_12_30_NorwayAnnotations_NARS-17_IndividualAnnotations_FRCNN_Metrics.csv"
 #dt = r"../testResults/6testing_NYAA-04_maxDis_8_runMean_10.csv"
 #dt = r"../testResults\_parameterTest_NYAA-04_1\parameterTest_NYAA-04_maxDisap_30_runMean_30_maxDist_500.csv"
 #dt = r'../testResults\_parameterTest_NYAA-04_1\parameterTest_NYAA-04__maxDisap_30_runMean_30_maxDist_500.csv'
 
 
-path = r'../testResults\_parameterTest_NARS-13_3/'
+path = r'../testResults\_parameterTest_NARS-17_3/'
 files = [os.path.join(path,f) for f in os.listdir(path) if f.endswith('.csv')]
 print(files)
 
-results_filename = "../testResults/parameterTest_3_NARS-13_Evaluate.csv"
+results_filename = "../testResults/parameterTest_3_NARS-17_Evaluate.csv"
 
 
 write_header(results_filename)

### 2021-12-30 code/run.py

@@ -16,49 +16,49 @@ import time
 
 ### SETTINGS ###
 
-#prefix_results_filename = "parameterTest_NARS-13"
+prefix_results_filename = "parameterTest_NARS-17"
 
-#list_max_disappeared = [0,10,20,30,40,50,60,70,80,90,100,110,120,130,140,150,160] # Maximum number of frames a track can be lost before new points will be forces into a new track.
-#list_running_mean_threshold = [0,10,20,30,40,50,60,70,80,90,100,110,120,130,140,150,160] # Maximum number of frames for calculating the running mean of the position of an object If there are less than this number of frames currently in the track, a mean over what is in the track will be used..
-#list_max_distance = [0, 100, 200, 300, 400, 500, 600, 700, 800, 900, 1000] # If set to 0, this parameter will be ignored. If not zero, points that have a distance to tracked objects higher than this parameter will be forced into new tracks.
+list_max_disappeared = [0,10,20,30,40,50,60,70,80,90,100,110,120,130,140,150,160] # Maximum number of frames a track can be lost before new points will be forces into a new track.
+list_running_mean_threshold = [0,10,20,30,40,50,60,70,80,90,100,110,120,130,140,150,160] # Maximum number of frames for calculating the running mean of the position of an object If there are less than this number of frames currently in the track, a mean over what is in the track will be used..
+list_max_distance = [0, 100, 200, 300, 400, 500, 600, 700, 800, 900, 1000] # If set to 0, this parameter will be ignored. If not zero, points that have a distance to tracked objects higher than this parameter will be forced into new tracks.
 
-# list_max_disappeared = [100] # Maximum number of frames a track can be lost before new points will be forces into a new track.
-# list_running_mean_threshold = [5] # Maximum number of frames for calculating the running mean of the position of an object If there are less than this number of frames currently in the track, a mean over what is in the track will be used..
-# list_max_distance = [400] # If set to 0, this parameter will be ignored. If not zero, points that have a distance to tracked objects higher than this parameter will be forced into new tracks.
[…]

### 2021-12-30 text/ManuscriptDraft.Rmd
Small manuscript update

@@
 abstract: ABSTRACT | Often simple variables will be used to describe the flowering phenology of a population of plants, e.g.
-onset or peak of flowering and to infer respones to climate change.
+onset or peak of flowering.
 Here we show that image-based monitoring of field plots at very high temporal resolution can return information on flowering phenology at the level of indiviuals.
-Further, we present an automatic flower tracking algorithm.
+Further, we present a framework for automatically tracking, filtering, and visualizing flowers in time-lapse image series.
+We compare the results of automatic tracking with manual tracking.
 ---
@@
 
+To estimate the optimal values for max_disappeared and max_distance, we analysed the ground truth tracks of the four series.
+
+To estimate the value for max_distance, we calculated the largest distance between any two points within any track for any flower in each of the four series.
+
+The maximum number of frames a flower track was lost and subsequently reappeared were xxx.
+
 To explore the effect of the user parameters and to identify the optimal combination of parameters for our case of tracking flowers, we followed a step wise approach.

### 2022-01-04 code/DBSCANFiltering.py
DBSCAN tests

@@ -15,17 +15,16 @@ import numpy as np
 from scipy.spatial import ConvexHull
 
 
-
 # Note, EPS_DISTANCE = 20 is a magic number and it needs to be
 # * smaller than the gap between any two islands
 # * large enough to cluster polygons in one island in same cluster
-EPS_DISTANCE = 100
+EPS_DISTANCE = 281 #281
 MIN_SAMPLE_POLYGONS = 1
 
 
 #tracks = pd.read_csv(r'U:\BITCue\Projekter\TrackingFlowers\testResults\_parameterTest_THUL-01_3\parameterTest_THUL-01_maxDisap_0_runMean_0_maxDist_0.csv')
-tracks = pd.read_csv(r'U:\BITCue\Projekter\TrackingFlowers\testResults\_parameterTest_NARS-13_3\parameterTest_NARS-13_maxDisap_10_runMean_10_maxDist_400.csv')
-gt = pd.read_csv(r'U:\BITCue\Projekter\TrackingFlowers\data\annotations\2020_04_30_NorwayAnnotations_NARS-13_IndividualAnnotations_FRCNN_Metrics.csv')
+tracks = pd.read_csv(r'U:\BITCue\Projekter\TrackingFlowers\testResults\_parameterTest_NARS-04_3\parameterTest_NARS-04_maxDisap_10_runMean_10_maxDist_500.csv')
+gt = pd.read_csv(r'U:\BITCue\Projekter\TrackingFlowers\data\annotations\2020_05_17_NorwayAnnotations_NARS-04_IndividualAnnotations_FRCNN_Metrics.csv')
 
 
 # tracks = pd.read_csv(r'U:\BITCue\Projekter\TrackingFlowers\data\annotations\2020_05_15_NorwayAnnotations_THUL-01_IndividualAnnotations_FRCNN_Metrics.csv')

### 2022-01-04 code/evaluate.py
DBSCAN tests

@@ -277,17 +277,17 @@ class evaluator():
 #dt = pd.read_csv(r"U:\BITCue\Projekter\NorwayAnnotations\Experiments\2020_06_30_DetectionPhenology\Rmd\ModelQ_Results\DT_GT\submit_boxes_NewFormat_AllTestSeries_Sampled.csv", sep=",")
 #dt = pd.read_csv(r"U:\BITCue\Projekter\NorwayAnnotations\Experiments\2020_06_30_DetectionPhenology\Rmd\ModelQ_Results\DT_GT\2020_02_01_NorwayAnnotations_AllTestSeries_IndividualAnnotations_FRCNN_Metrics_copy.csv", sep=",")
 
-gt = r"U:\BITCue\Projekter\TrackingFlowers\data\annotations\2021_12_30_NorwayAnnotations_NARS-17_IndividualAnnotations_FRCNN_Metrics.csv"
+gt = r"U:\BITCue\Projekter\TrackingFlowers\data\annotations\2020_05_15_NorwayAnnotations_THUL-01_IndividualAnnotations_FRCNN_Metrics.csv"
 #dt = r"../testResults/6testing_NYAA-04_maxDis_8_runMean_10.csv"
 #dt = r"../testResults\_parameterTest_NYAA-04_1\parameterTest_NYAA-04_maxDisap_30_runMean_30_maxDist_500.csv"
 #dt = r'../testResults\_parameterTest_NYAA-04_1\parameterTest_NYAA-04__maxDisap_30_runMean_30_maxDist_500.csv'
 
 
-path = r'../testResults\_parameterTest_NARS-17_3/'
+path = r'../testResults\tempTHUL01'
 files = [os.path.join(path,f) for f in os.listdir(path) if f.endswith('.csv')]
 print(files)
 
-results_filename = "../testResults/parameterTest_3_NARS-17_Evaluate.csv"
+results_filename = "../testResults/parameterTest_3_THUL-01_temp.csv"
 
 
 write_header(results_filename)

### 2022-01-04 code/run.py
NARS-17 tracking

@@ -9,7 +9,6 @@ import pandas as pd
 from track import tracker
 from filtering import sieve
 
-
 import time
 
 # ===================== Run tracking on several combinations of parameters ====

### 2022-01-04 text/ManuscriptDraft.Rmd
Manuscript update. DBSCAN.

@@
 
-* **Flowering phenology**
-  + Importance of studying flowering phenology
-  + Responses to climate change
-  + Phenology of communities, populations, individuals
-  + Traditional methods for studying
-  + Onset of flowering says little about true distribution
-    - Even true distribition of community says little about flowering lengths of individuals and for example how it varies accross the season.
-  + Difficult to study at the individual level - requires high temporal resolution and keeping track of individuals
-* **Image based monitoring**
-  + Automatic, high temporal resolution, remote sites
-  + High temporal resolution means that we can annotate individuals and get phenology of the individual
-
-* **Tracking**
-  + Offline and online
-  + Online often coupled with CNNs that attempt to distinguish individuals from each other and recognize them through frames
-
-  + Flowers appear very similar and 
-  + Many methods for offline tracking
-  + Hungarian/Kahlmann filter
-  ++ May not be applicable for objects that move weirdly, e.g. change directions between frames.
-  + Tracking based on distance.
-    - Good but has some problems
-    - Two points always associated disregarding absolute distance
-    - Tracks lost when objects disappear
-    - Objects close to each other may swap tracks
-
-* **Our solution**
-  + Here we demonstrate a framework for automatic flower tracking and evaluation of tracking performance
-  + Ground truth tracks
-
-\newpage
-
 For the individual plant, timing of flowering is of utmost importance.
@@
 Here, we show that information on phenology at the level of individuals can be derived from image-based monitoring of flower phenology.
-Further, we present and evaluate an automatic flower tracking algorithm.
+Further, we present and evaluate an automatic flower tracking and filtering algorithm.
+
+Tracking individual flowers enables the possibility of assigning reproductive success to the individual, for example by observation of seed set.
+Thereby, it can be explored whether reproductive success is affected by timing and length of flowering.
+By simultaneously tracking flower visits, these could be assigned to the individual flower and visitation rates per flower could be calculated per flower and related to reproductive success.
+Further, any information of taxonomic grouping could refine this analyses.
+
+For complex scenes with many flowers in close vicinity to each other, we suggest a conservative filtering approach.
+The approach may remove correct tracks, but the tracks that remain will have a lower risk of tracking errors.
+The approach allows for more confidence in upscaling the method.
+E.g., when running the method on a large number of image series, it is preferable to extract individual high-confidence tracks from each series and ignore the remaining tracks.
 
@@
 
+Depending on the nature of the objects being tracked and the complexity of the scene, the user parameters can be estimated from visual examination of the tracking results.
+Often it may be preferable to manually annotate a subset of the objects in the image series and derive a set of user parameters from these results.
+
 # Evaluating tracking perfomance
@@
 
-Manually annotating a subset of the objects in the image series could help derive a good set of parameters.
+When deploying the automatic tracking algorithm on naive data without ground truth tracks, it is not possible to manually filter for correct tracks.
+Therefore, we present a conservative filtering method that extracts the most trustworthy tracks from a scene.
 
-When deploying the tracking algorithm on naive data that has not been manually tracked, it is not possible to
+We disregard single points that were not associated to a track.
+For tracks consisting of two points, we establish the straight line between the points.
+For tracks consisting of three points, we establish the triangle from the points.
+For tracks consisting of more than three points, we calculate the convex hull of all the points included in the track and derive the polygon from the vertices of the convex hull.
+
+We then apply the DBSCAN clustering algorithm on these track geometrics to remove tracks in areas with a high density of tracks as these have a high risk of tracking mismatches.
+Second, we filter overlapping tracks using the following approach.
+
+We could compare number of points in the overlapping tracks. If one is less than e.g. 10% of the other, we'll remove the small track. If more than 10% we'll remove both?
+
+We then check if any two lines intersect (for tracks with two points), if any lines intersect with any polygons, and if any polygons overlap with other polygons, and remove tracks that overlap.
 
@@
 Here we will remove overlapping tracks to reduce the risk of error.
-
-We could compare number of points in the overlapping tracks. If one is less than e.g. 10% of the other, we'll remove the small track. If more than 10% we'll remove both?
-
-We disregard single points that were not associated to a track.
-For tracks consisting of two points, we establish the straight line between the points.
-For tracks consisting of three points, we establish the triangle from the points.
-For tracks consisting of more than three points, we calculate the convex hull of all the points included in the track and derive the polygon from the vertices of the convex hull.
-
-We then check if any two lines intersect (for tracks with two points), if any lines intersect with any polygons, and if any polygons overlap with other polygons, and remove tracks that overlap.
 
@@
 
-Online tracking:
-SORT
-DeepSORT
+* **Flowering phenology**
+  + Importance of studying flowering phenology
+  + Responses to climate change
+  + Phenology of communities, populations, individuals
+  + Traditional methods for studying
+  + Onset of flowering says little about true distribution
+    - Even true distribition of community says little about flowering lengths of individuals and for example how it varies accross the season.
+  + Difficult to study at the individual level - requires high temporal resolution and keeping track of individuals
+* **Image based monitoring**
+  + Automatic, high temporal resolution, remote sites
+  + High temporal resolution means that we can annotate individuals and get phenology of the individual
+
+* **Tracking**
+  + Offline and online
+  + Online often coupled with CNNs that attempt to distinguish individuals from each other and recognize them through frames
+
+  + Flowers appear very similar and 
+  + Many methods for offline tracking
+  + Hungarian/Kahlmann filter
+  ++ May not be applicable for objects that move weirdly, e.g. change directions between frames.
+  + Tracking based on distance.
+    - Good but has some problems
+    - Two points always associated disregarding absolute distance
+    - Tracks lost when objects disappear
+    - Objects close to each other may swap tracks
+
+* **Our solution**
+  + Here we demonstrate a framework for automatic flower tracking and evaluation of tracking performance
+  + Ground truth tracks
+
+Tracking Multiple Moving Objects Using Unscented Kalman
+Filtering Techniques:
+"Kalman filtering (KF) [5] is widely used to track moving objects, with which we
+can estimate the velocity and even acceleration of an object with the measurement of its locations.
+However, the accuracy of KF is dependent on the assumption of linear motion for any object to be
+tracked. If an object takes some abrupt turns, the nonlinear movement cannot be well handled by the
+KF framework (due to the linear movement assumption of the design of KF)."
+Kalman, R. E. A New Approach to Linear Filtering and Prediction Problems. Journal of Basic
+Engineering, 1960(82), pp. 35-45.
 

### 2022-01-04 code/R/Analysis.Rmd
Outputs for presentation

@@
 ```{r}
-thul01 = read.csv("../../data/annotations/2020_05_15_NorwayAnnotations_THUL-01_IndividualAnnotations_FRCNN.csv", header = F)
+thul01 = read.csv("../../data/annotations/2020_05_15_NorwayAnnotations_THUL-01_IndividualAnnotations_FRCNN_Metrics.csv", header = T)
 
@@
 
+nars04$x_c = (nars04$x_min + nars04$x_max) / 2
+nars04$y_c = (nars04$y_min + nars04$y_max) / 2
+
 ```
@@
 
-```{r}
+```{r fig.height=4}
 d1 = read.csv("U:/BITCue/Projekter/TrackingFlowers/testResults/parameterTest_3_NYAA-04_Evaluate.csv")
@@
 ```
+
 ```{r}
@@
 
-minmm = read.csv("U:/BITCue/Projekter/TrackingFlowers/testResults/_parameterTest_NYAA-04_3/parameterTest_NYAA-04_maxDisap_10_runMean_60_maxDist_300.csv")
+nyaa04minmm = read.csv("U:/BITCue/Projekter/TrackingFlowers/testResults/_parameterTest_NYAA-04_3/parameterTest_NYAA-04_maxDisap_10_runMean_60_maxDist_300.csv")
 
@@
 
+```{r fig.height=4}
+
+nyaa04minmm %>% 
+  ggplot() +
+  geom_point(aes(x = x_c, y = y_c, color = as.factor(objectID)), size = 1) + 
+  xlim(0,6080) +
+  ylim(0,3420) + 
+  ggtitle("NYAA-04 Tracked (Minimum mm)") +
+  theme(legend.position = "none")
+
+nyaa04 %>% 
+  ggplot() +
+  geom_point(aes(x = x_c, y = y_c, color = as.factor(Class)), size = 1) + 
+  xlim(0,6080) +
+  ylim(0,3420) + 
+  ggtitle("NYAA-04 Ground Truth") +
+  theme(legend.position = "none")
+```
+
 ## THUL 01
 
-```{r}
+```{r fig.height=4}
 d2 = read.csv("U:/BITCue/Projekter/TrackingFlowers/testResults/parameterTest_3_THUL-01_Evaluate.csv")
@@
 
+```{r fig.height=4}
+
+cornoo %>% 
+  ggplot() +
+  geom_point(aes(x = x_c, y = y_c, color = as.factor(objectID)), size = 1) + 
+  xlim(0,6080) +
+  ylim(0,3420) + 
+  ggtitle("THUL-01 Tracked") +
+  theme(legend.position = "none")
+
+thul01 %>% 
+  ggplot() +
+  geom_point(aes(x = x_c, y = y_c, color = as.factor(id_gt)), size = 1, shape = 1) + 
+  xlim(0,6080) +
+  ylim(0,3420) + 
+  ggtitle("THUL-01 Ground Truth") +
+  theme(legend.position = "none")
+```
+
 ## NARS-04
 
-```{r}
+```{r fig.height=4}
+
 d3 = read.csv("U:/BITCue/Projekter/TrackingFlowers/testResults/parameterTest_3_NARS-04_Evaluate.csv")
@@
 
-ggsave('U:/BITCue/Projekter/TrackingFlowers/testResults/figures/NARS-04_AllCombinations.png', NARS04_p1)
-ggsave('U:/BITCue/Projekter/TrackingFlowers/testResults/figures/NARS-04_TenLowest.png', NARS04_p2)
+NARS04_p1
+NARS04_p2
+
+#ggsave('U:/BITCue/Projekter/TrackingFlowers/testResults/figures/NARS-04_AllCombinations.png', NARS04_p1)
+#ggsave('U:/BITCue/Projekter/TrackingFlowers/testResults/figures/NARS-04_TenLowest.png', NARS04_p2)
 ```
+
 ```{r}
-# NYAA-04 gt:
+# NARS-04 gt:
 NARS04_gt = read.csv("U:/BITCue/Projekter/TrackingFlowers/data/annotations/2020_05_17_NorwayAnnotations_NARS-04_IndividualAnnotations_FRCNN_Metrics.csv")
@@
 
-minmm = read.csv("U:/BITCue/Projekter/TrackingFlowers/testResults/_parameterTest_NARS-04_3/parameterTest_NARS-04_maxDisap_10_runMean_10_maxDist_500.csv")
+nars04minmm = read.csv("U:/BITCue/Projekter/TrackingFlowers/testResults/_parameterTest_NARS-04_3/parameterTest_NARS-04_maxDisap_10_runMean_10_maxDist_500.csv")
 
@@
 
+```
+
+```{r fig.height=4}
+
+nars04minmm %>% 
+  ggplot() +
+  geom_point(aes(x = x_c, y = y_c, color = as.factor(objectID)), size = 1) + 
+  xlim(0,6080) +
+  ylim(0,3420) + 
+  ggtitle("NARS-04 Tracked") +
+  theme(legend.position = "none")
+
+nars04 %>% 
+  ggplot() +
+  geom_point(aes(x = x_c, y = y_c, color = as.factor(Class)), size = 1) + 
+  xlim(0,6080) +
+  ylim(0,3420) + 
+  ggtitle("NARS-04 Ground Truth") +
+  theme(legend.position = "none")
 ```
@@
 ```{r}
-# NYAA-04 gt:
+# NARS-13 gt:
 NARS13_gt = read.csv("U:/BITCue/Projekter/TrackingFlowers/data/annotations/2020_04_30_NorwayAnnotations_NARS-13_IndividualAnnotations_FRCNN_Metrics.csv")
@@
 ```{r}
-# Minimum mm: 10, 500, 10, 20, 0.9883991, 32
-# All zero: 0, 0, 0, 85, 0.9506961, 31
-# Correct noo with minimum mm: 10, 500, 10, 20, 0.9883991, 32
+# Minimum mm: 10-20, 300-400, 10, 78, 0.9840393, 81+85+93+97
+# All zero: 0, 0, 0, 267, 0.9453653, 88
+# Correct noo with minimum mm: 10, 400, 10, 78, 0.9840393, 85
 
-minmm = read.csv("U:/BITCue/Projekter/TrackingFlowers/testResults/_parameterTest_NARS-13_3/parameterTest_NARS-13_maxDisap_0_runMean_0_maxDist_0.csv")
+nars13minmm = read.csv("U:/BITCue/Projekter/TrackingFlowers/testResults/_parameterTest_NARS-13_3/parameterTest_NARS-13_maxDisap_0_runMean_0_maxDist_0.csv")
 
@@
 ```
+
+```{r fig.height=4}
+
+nars13minmm %>% 
+  ggplot() +
+  geom_point(aes(x = x_c, y = y_c, color = as.factor(objectID)), size = 1) + 
+  xlim(0,6080) +
+  ylim(0,3420) + 
+  ggtitle("NARS-13 Tracked (Minimum mm)") +
+  theme(legend.position = "none")
+
+nars13 %>% 
+  ggplot() +
+  geom_point(aes(x = x_c, y = y_c, color = as.factor(Class)), size = 1) + 
+  xlim(0,6080) +
+  ylim(0,3420) + 
+  ggtitle("NARS-13 Ground Truth") +
+  theme(legend.position = "none")
+
+length(unique(nars13$Class))
+```

### 2022-01-05 code/DBSCANFiltering.py
Wrote DBSCAN class

@@ -7,7 +7,7 @@ Created on Tue Dec 21 13:28:48 2021
 
 import geopandas as gpd
 import pandas as pd
-from shapely.geometry import Polygon
+from shapely.geometry import Polygon, LineString, Point
 from sklearn.cluster import DBSCAN
 import matplotlib.pyplot as plt
 from filtering import sieve
@@ -18,72 +18,209 @@ from scipy.spatial import ConvexHull
 # Note, EPS_DISTANCE = 20 is a magic number and it needs to be
 # * smaller than the gap between any two islands
 # * large enough to cluster polygons in one island in same cluster
-EPS_DISTANCE = 281 #281
-MIN_SAMPLE_POLYGONS = 1
+#EPS_DISTANCE = 300 #281 #281
+#MIN_SAMPLE_POLYGONS = 1
+
+
 
 
 #tracks = pd.read_csv(r'U:\BITCue\Projekter\TrackingFlowers\testResults\_parameterTest_THUL-01_3\parameterTest_THUL-01_maxDisap_0_runMean_0_maxDist_0.csv')
-tracks = pd.read_csv(r'U:\BITCue\Projekter\TrackingFlowers\testResults\_parameterTest_NARS-04_3\parameterTest_NARS-04_maxDisap_10_runMean_10_maxDist_500.csv')
-gt = pd.read_csv(r'U:\BITCue\Projekter\TrackingFlowers\data\annotations\2020_05_17_NorwayAnnotations_NARS-04_IndividualAnnotations_FRCNN_Metrics.csv')
+tracks = pd.read_csv(r'U:\BITCue\Projekter\TrackingFlowers\testResults\_parameterTest_NARS-04_3\parameterTest_NARS-04_maxDisap_10_runMean_10_maxDist_100.csv')
+#gt = pd.read_csv(r'U:\BITCue\Projekter\TrackingFlowers\data\annotations\2020_05_17_NorwayAnnotations_NARS-04_IndividualAnnotations_FRCNN_Metrics.csv')
 
 
-# tracks = pd.read_csv(r'U:\BITCue\Projekter\TrackingFlowers\data\annotations\2020_05_15_NorwayAnnotations_THUL-01_IndividualAnnotations_FRCNN_Metrics.csv')
-# tracks.rename(columns={"id_gt": "objectID"}, inplace = True)
-# tracks['x_c'] = (tracks['x_min'] + tracks['x_max']) / 2
[…]

### 2022-01-05 code/DBSCANFiltering.py
Wrote DBSCAN class

@@ -168,7 +168,7 @@ class DBSCANsieve():
         dfSub.plot(ax=ax2, column = 'clusterID', marker = ".", markersize=0.2)
         
         
-        fig.savefig(f'BeforeAndAfterFiltering_eps_{self.eps_distance}.png', dpi=300)
+        fig.savefig(f'../testResults/BeforeAndAfterFiltering_eps_{self.eps_distance}.png', dpi=300)
 
     
         return 0

### 2022-01-05 text/ManuscriptDraft.Rmd
Manuscript update (DBSCAN)

@@
 
-We disregard single points that were not associated to a track.
 For tracks consisting of two points, we establish the straight line between the points.
@@
 For tracks consisting of more than three points, we calculate the convex hull of all the points included in the track and derive the polygon from the vertices of the convex hull.
+Single points are kept.
 
 We then apply the DBSCAN clustering algorithm on these track geometrics to remove tracks in areas with a high density of tracks as these have a high risk of tracking mismatches.
-Second, we filter overlapping tracks using the following approach.
+The filtering is done in two steps.
+First, DBSCAN algorithm is run with a conservatively high value for the eps parameter, meaning that tracks in close vicinity to each other will be clustered together.
+Second, all tracks that were not assigned to a unique cluster are removed.
+Tracks that are spatially isolated remains.
 
 We could compare number of points in the overlapping tracks. If one is less than e.g. 10% of the other, we'll remove the small track. If more than 10% we'll remove both?
-
-We then check if any two lines intersect (for tracks with two points), if any lines intersect with any polygons, and if any polygons overlap with other polygons, and remove tracks that overlap.
-
-From our results we see that erroneous tracks often occur as small sections consisting of only few points within the main tracks.
-
-Therefore, tracks that overlap have significant risk of errors.
-Overlapping tracks can be caused by a single flower that was erroneously assigned to several tracks, two flowers that were located sufficiently close to each other that there areas overlapped (e.g.
-when wind moves the flowers around), false positive detections close to a flower.
-Best case is two flowers that flowered in the same area but were separated by time.
-Here we will remove overlapping tracks to reduce the risk of error.
 
@@
 In cases where we are tracking perfectly with maxDisap = 0, setting it any value will not make a difference. Not quite right. Explore more...
+
+# Filtering
+
+Extracting tracks that are spatially isolated does not guarantee that the tracks are correct/without errors.
+However, as spatially isolated objects are easier to track, it increases confidence in the remaining tracks.
 

### 2022-01-07 code/R/Tables.Rmd
Added R script for tables

---
title: "Tables"
author: "Hjalte Mann"
date: "7/1/2022"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
#install.packages("gt")
library(gt)
library(tidyverse)
library(glue)
```

## R Markdown

This is an R Markdown document. Markdown is a simple formatting syntax for authoring HTML, PDF, and MS Word documents. For more details on using R Markdown see <http://rmarkdown.rstudio.com>.

When you click the **Knit** button a document will be generated that includes both content as well as the output of any embedded R code chunks within the document. You can embed an R code chunk like this:
```{r}

# Define the start and end dates for the data range
start_date <- "2010-06-07"
end_date <- "2010-06-14"

sp500

# Create a gt table based on preprocessed
# `sp500` table data
sp500 %>%
  filter(date >= start_date & date <= end_date) %>%
  select(-adj_close) %>%
  gt() %>%
  tab_header(
    title = "S&P 500",
    subtitle = glue("{start_date} to {end_date}")
  ) %>%
  fmt_date(
    columns = date,
    date_style = 3
  ) %>%
  fmt_currency(
    columns = c(open, high, low, close),
    currency = "USD"
  ) %>%
  fmt_number(
    columns = volume,
    suffixing = TRUE
  )
```
```{r}
ts = read.csv2("U:/BITCue/Projekter/TrackingFlowers/testResults/TrackingScores.csv")
print(ts)

```

### 2022-01-07 text/ManuscriptDraft.Rmd
Manuscript update

@@
 ```
+
+```{r echo=FALSE}
+# Function for creating word comments in Rmarkdown
+# Use: This text contains a `r word_comment("This is the comment", "comment.")`.
+
+word_comment <- function(comment, highlight = "") {
+  if (isTRUE(knitr:::pandoc_to() == "docx")) {
+    paste0('[', comment, ']{.comment-start id="0" author="Hjalte M."',
+           'date="2020-01"}', highlight, '[]{.comment-end id="0"}')
+  }
+}
+```
+
 \newpage
@@
 
+We demonstrate our filtering approach for the image series for which our tracking algorithm did not produce perfect results.
+
+Here we set eps DBSCAN parameter to XXX.
+Although the value could be fine-tuned for improved results for each image series individually, this value returns good results overall.
+In a naive setting, a general value could be chosen of the value could be adjusted for each series based on visual examination of performance or testing on a subset of data.
+
 # Results
 
+**Table 1:** Bla bla...
+#
+![](../figures/table1.png){ width=100% }
+
+`r word_comment("Table 1 doesn't show excactly which combinations of parameters gives the result.
+This could be shown in suppl.
+e.g.
+with a network-like graph, but I am thinking it is overkill and this table is sufficient.", "Table 1")` shows the performance of the tracking algorithm.
+
 One way to plot?: Time on x-axis, gt id on y axis, tr id as colour. This will show mismatches.
+
+Our tracking algorithm consistently returns high MOTA scores.
 
@@
 In cases where we are tracking perfectly with maxDisap = 0, setting it any value will not make a difference. Not quite right. Explore more...
+
+Our method for filtering tracks using DBSCAN on track centroids ensures that all tracks are given the same weighting in the filtering since each track is represented by a single point. In some cases 
+e.g.
+if it is given a priori that an object will always appear in a minimum of two frames, then single point tracks can be filtered out.
+However, when such a priori knowledge is not accessible, a conservative approach as the one we present is preferable.
+
+Applying the technique on detection instead of manual annotations.
+Detections introduce false positives.
+Either manual or automatic quality control to remove these before tracking or after.
+
+Our manually tracked data is ground truth, but for example when flowers periodically move out of the frame, this mimics false negatives. Similarly when one flower occludes another.
+
+Our tracking algorithm returned high MOTA scores even with all three parameters set to zero.
+
+Automatically detecting flowers would likely introduce a degree of false negatives which would decrease the MOTA score if max disappeared is set to zero.
+Introducing a value for this parameter can deal with the problem of false negatives.
+
+We ran our tracking algorithm on 3xxx combination of the three parameters.
+Using the parameters increased the performance of the tracking substantially.
+Our steps in parameter values were crude, however, and it is very likely that finer steps in these values would identify combinations that produce even better results.
+Here we do not perform this analysis, however, as our goal is to show that the parameters can be used for optimizing tracking performance in general.
 

### 2022-01-07 code/DBSCANFiltering.py
Added plotting of track filtering

@@ -13,30 +13,22 @@ import matplotlib.pyplot as plt
 from filtering import sieve
 import numpy as np
 from scipy.spatial import ConvexHull
+from datetime import datetime
 
+# Note, EPS_DISTANCE needs to be smaller than the gap between any two polygons we want to keep and large enough to cluster polygons that we want to discard together
+# EPS_DISTANCE = 281 is the highest value that keeps all tracks in THUL-01
 
-# Note, EPS_DISTANCE = 20 is a magic number and it needs to be
-# * smaller than the gap between any two islands
-# * large enough to cluster polygons in one island in same cluster
-#EPS_DISTANCE = 300 #281 #281
-#MIN_SAMPLE_POLYGONS = 1
-
-
-
-
-#tracks = pd.read_csv(r'U:\BITCue\Projekter\TrackingFlowers\testResults\_parameterTest_THUL-01_3\parameterTest_THUL-01_maxDisap_0_runMean_0_maxDist_0.csv')
-tracks = pd.read_csv(r'U:\BITCue\Projekter\TrackingFlowers\testResults\_parameterTest_NARS-04_3\parameterTest_NARS-04_maxDisap_10_runMean_10_maxDist_100.csv')
-#gt = pd.read_csv(r'U:\BITCue\Projekter\TrackingFlowers\data\annotations\2020_05_17_NorwayAnnotations_NARS-04_IndividualAnnotations_FRCNN_Metrics.csv')
-
-
+#tracks = pd.read_csv(r'U:\BITCue\Projekter\TrackingFlowers\testResults\_parameterTest_NARS-04_3\parameterTest_NARS-04_maxDisap_10_runMean_10_maxDist_500.csv')
 
 class DBSCANsieve():
     def __init__(self, tracks, eps_distance, min_sample_polygons):
-        self.tracks = tracks
+        if not isinstance(tracks, pd.DataFrame):
+            self.tracks = pd.read_csv(tracks, sep=",", header = 0)
+        else:
+            self.tracks = tracks
         
         self.eps_distance = eps_distance
         self.min_sample_polygons = min_sample_polygons
-        
         
         self.points = {}
         self.lines = {}
@@ -45,10 +37,9 @@ class DBSCANsieve():
         self.polyLengths = {}
         self.polyHulls = {}
         
-        
-    def gen(self):
+    def gen(self): # Generator that yields next group in tracks dataframe
[…]

### 2022-01-07 code/track.py
Minor update

@@ -226,9 +226,6 @@ class tracker():
 
 
 
-
-
-
 # ===================== Run a single round of tracking ========================
 # t = tracker(4, 500, 5, frames) # Instantiate the class instance and pass in the threshold for max_disappeared and the list of frames.
 # starttime = time.time()

### 2022-01-07 code/visualize.py
Minor update

@@ -18,7 +18,7 @@ import pandas as pd
 #df = pd.read_csv(r'../testResults/_parameterTest_NYAA-04_3/parameterTest_NYAA-04_maxDisap_10_runMean_60_maxDist_300.csv')
 gt = pd.read_csv(r'../data\annotations\2020_04_30_NorwayAnnotations_NARS-13_IndividualAnnotations_FRCNN_Metrics.csv')
 
-df2 = pd.read_csv(r'../testResults/_parameterTest_NARS-13_3/parameterTest_NARS-13_maxDisap_10_runMean_10_maxDist_300.csv')
+df2 = pd.read_csv(r'../testResults/_parameterTest_NARS-04_3/parameterTest_NARS-04_maxDisap_10_runMean_10_maxDist_500.csv')
 df3 = pd.read_csv(r'../testResults/filtered_NARS-13_maxDisap_10_runMean_10_maxDist_300.csv')
 
 #m = pd.merge(df, gt, on = ['filename', 'x_min', 'x_max', 'y_min', 'y_max'], how = 'inner')
@@ -57,11 +57,10 @@ m3 = pd.merge(df3, gt, on = ['filename', 'x_min', 'x_max', 'y_min', 'y_max'], ho
 #plt.scatter(m['frame'], m['id_gt'], c = m['objectID'], s = 15)
 #plt.show()
 
-plt.scatter(df2['x_c'], df2['y_c'], c = df2['objectID'], s = 5)
-plt.show()
 
-# plt.scatter(df3['x_c'], df3['y_c'], c = df3['objectID'], s = 5)
-# plt.show()
+scatter = plt.scatter(df2['x_c'], df2['y_c'], c = df2['objectID'], s = 5)
+ax = scatter.axes
+ax.invert_yaxis()
 
 
 

### 2022-01-07 code/R/Analysis.Rmd
NARS04 QC

@@
   geom_point(aes(x=runMean, y = maxDisap, size = mm, colour = mm)) + 
-  ggtitle("NARS-13 Parameter combination") +
+  ggtitle("NARS-17 Parameter combination") +
   facet_wrap(vars(as.factor(maxDist))) +
@@
   geom_point(aes(x=runMean, y = maxDisap, size = mm)) + 
-  ggtitle("THUL-01 All zero mismatches") +
+  ggtitle("NARS-17 All zero mismatches") +
   facet_wrap(vars(maxDist))+scale_color_gradient(low="blue", high="red")
@@
 ```
+
+```{r}
+tempNARS04 = read.csv("U:/BITCue/Projekter/TrackingFlowers/testResults/tempNARS04.csv")
+
+tempNARS04 %>% 
+  filter(x_c > 1100 & x_c < 1110) %>% 
+  ggplot() +
+  geom_point(aes(x = x_c, y = y_c, colour = as.factor(objectID))) + 
+  xlim(1100,1110) + 
+  ylim(2250,2320)
+
+tempNARS04 %>% 
+  ggplot() +
+  geom_point(aes(x = x_c, y = y_c, colour = as.factor(objectID)), size = 0.5)
+
+```

### 2022-01-07 code/evaluate.py
Added check for whether input is pd dataframe

@@ -14,6 +14,8 @@ br = "\n"
 
 iou_threshold = 0.5
 
+verbose = False # Set to True if you want tracking process printed to screen and False if not
+
 def write_header(results_filename):
     with open(results_filename, 'a') as resultFile: # Write the header of the output file
         header = f'maxDisap,maxDist,runMean,mm,mota,numberOfObjects{br}'
@@ -21,7 +23,7 @@ def write_header(results_filename):
 
 
 class evaluator():
-    def __init__(self, dt, gt, results_filename):
+    def __init__(self, dt, gt, results_filename, verbose):
         self.P = 0 # Positives (meaning total number of objects (e.g. flowers) across all images)
         self.CP = 0 # Correct positives (meaning total number of objects that were correclty detected)
         self.FP = 0 # False positives
@@ -36,33 +38,44 @@ class evaluator():
 
         self.dt_filename = dt
         basename = os.path.basename(self.dt_filename)
-        print(basename)
+            
         runMeanSearch = re.search('runMean_(.+?)_max', basename)
         maxDistSearch = re.search('maxDist_(.+?).csv', basename)
         maxDisapSearch = re.search('maxDisap_(.+?)_run', basename)
         
         if runMeanSearch:
             self.runMean = runMeanSearch.group(1)
-            print(f'runMean: {self.runMean}')
+
         if maxDistSearch:
             self.maxDist = maxDistSearch.group(1)
-            print(f'maxDist: {self.maxDist}')
+            
         if maxDisapSearch:
             self.maxDisap = maxDisapSearch.group(1)
-            print(f'maxDisap: {self.maxDisap}')
             
+        
+        if verbose:
+            print(f'Running on: {basename}')
+            print(f'runMean: {self.runMean}')
+            print(f'maxDist: {self.maxDist}')
+            print(f'maxDisap: {self.maxDisap}')
             
     def set_up_data(self):
         
-        dt = pd.read_csv(self.dt, sep=",", header = 0)
-        gt = pd.read_csv(self.gt, sep=",", header = 0)
+        if not isinstance(self.dt, pd.DataFrame):
+            dt = pd.read_csv(self.dt, sep=",", header = 0)
+        else:
+            dt = self.dt
+
+        if not isinstance(self.gt, pd.DataFrame):
[…]

### 2022-01-07 code/filtering.py
Comment edit

@@ -21,7 +21,7 @@ br = "\n"
 
 
 """
-TO-DO
+Pipeline
 
 - Remmove single points
 - Establish straight line between two-point tracks

### 2022-01-07 code/evaluate.py
Added to-do

@@ -16,6 +16,15 @@ iou_threshold = 0.5
 
 verbose = False # Set to True if you want tracking process printed to screen and False if not
 
+
+
+"""
+To-do
+
+If basename does not include parameters, create other filename for result and output. So evaluate can be run on other files than the parameter tests.
+
+"""
+
 def write_header(results_filename):
     with open(results_filename, 'a') as resultFile: # Write the header of the output file
         header = f'maxDisap,maxDist,runMean,mm,mota,numberOfObjects{br}'

### 2022-01-07 code/run.py
Full pipeline in run.py now. Added code for performing test of DBSCAN results

@@ -8,102 +8,276 @@ Created on Tue Dec 14 12:50:10 2021
 import pandas as pd
 from track import tracker
 from filtering import sieve
-
+from DBSCANFiltering import DBSCANsieve
+from evaluate import evaluator
+import os
 import time
+
+
+# ===================== Import ground truth data ====
+
+def impdata(camID):
+    if camID == "NARS-04":
+        gt = r"U:\BITCue\Projekter\TrackingFlowers\data\annotations\2020_05_17_NorwayAnnotations_NARS-04_IndividualAnnotations_FRCNN_Metrics.csv"
+    if camID == "NARS-13":
+        gt = r"U:\BITCue\Projekter\TrackingFlowers\data\annotations\2020_04_30_NorwayAnnotations_NARS-13_IndividualAnnotations_FRCNN_Metrics.csv"
+    if camID == "THUL-01":
+        gt = r"U:\BITCue\Projekter\TrackingFlowers\data\annotations\2020_05_15_NorwayAnnotations_THUL-01_IndividualAnnotations_FRCNN_Metrics.csv"
+    if camID == "NYAA-04":
+        gt = r"U:\BITCue\Projekter\TrackingFlowers\data\annotations\2021_12_13_NorwayAnnotations_NYAA-04_IndividualAnnotations_FRCNN_Metrics.csv"
+    if camID == "NARS-17":
+        gt = r"U:\BITCue\Projekter\TrackingFlowers\data\annotations\2021_12_30_NorwayAnnotations_NARS-17_IndividualAnnotations_FRCNN_Metrics.csv"
+    return gt
+        
 
 # ===================== Run tracking on several combinations of parameters ====
 
-### SETTINGS ###
+# ### SETTINGS ###
 
-prefix_results_filename = "parameterTest_NARS-17"
+# prefix_results_filename = "parameterTest_NARS-17"
 
-list_max_disappeared = [0,10,20,30,40,50,60,70,80,90,100,110,120,130,140,150,160] # Maximum number of frames a track can be lost before new points will be forces into a new track.
-list_running_mean_threshold = [0,10,20,30,40,50,60,70,80,90,100,110,120,130,140,150,160] # Maximum number of frames for calculating the running mean of the position of an object If there are less than this number of frames currently in the track, a mean over what is in the track will be used..
[…]

### 2022-01-11 code/evaluate.py
Produced filtering results

@@ -14,7 +14,7 @@ br = "\n"
 
 iou_threshold = 0.5
 
-verbose = False # Set to True if you want tracking process printed to screen and False if not
+#verbose = False # Set to True if you want tracking process printed to screen and False if not
 
 
 
@@ -38,7 +38,8 @@ class evaluator():
         self.FP = 0 # False positives
         self.FN = 0 # False negatives
         self.MM = 0 # Tracking mismatches
-
+        self.verbose = verbose
+        
         self.dt = dt
         self.gt = gt
         
@@ -62,13 +63,15 @@ class evaluator():
             self.maxDisap = maxDisapSearch.group(1)
             
         
-        if verbose:
+        if self.verbose:
             print(f'Running on: {basename}')
             print(f'runMean: {self.runMean}')
             print(f'maxDist: {self.maxDist}')
             print(f'maxDisap: {self.maxDisap}')
             
     def set_up_data(self):
+        dt = self.dt
+        gt = self.gt
         
         if not isinstance(self.dt, pd.DataFrame):
             dt = pd.read_csv(self.dt, sep=",", header = 0)
@@ -84,7 +87,8 @@ class evaluator():
         
         #print(f'dt type: {dt.dtypes}')
         #print(f'gt type: {gt.dtypes}')
-
+        #print(dt)
+        
         dt.rename(columns={'objectID': 'id_tr'}, inplace=True)
         gt['frame'] = gt['filename'].str.extract('(\d{6})')
         #dt['frame'] = dt['filename'].str.extract('(\d{6})')
@@ -178,7 +182,7 @@ class evaluator():
     def run(self):
         
         dt, gt, number_of_objects = self.set_up_data()
-        
+
         common = gt.merge(dt, on=['filename'], how = 'inner') # Create a dataframe of of frames that contain both groundt truths and detections. We'll also create a list of uniqe filenames in the common dataframe
         filenames_common = common['filename'].unique().tolist() # Get a list of unique filenames to iterate over
         #print("Filenames common: ",filenames_common)
@@ -259,7 +263,7 @@ class evaluator():
         PR = dt.shape[0]
         CP =  PR - self.FP
         
-        if verbose:
+        if self.verbose:
             print("\n#####", " RESULTS ", "#####\n")
             print("Number of annotated objects: ", P)
             print("Number of predictions made: ", PR )
@@ -273,7 +277,7 @@ class evaluator():
[…]

### 2022-01-11 code/run.py

@@ -83,13 +83,13 @@ def impdata(camID):
 # ===================== Evaluate tracking accuracy ====
 
 
-# Get the ground truth annotations on the format [filename, x_min, y_min, x_max, y_max, id_gt]
-# gt = impdata("NARS-04")
-# dt = r'U:\BITCue\Projekter\TrackingFlowers\testResults\_parameterTest_NARS-04_3\parameterTest_NARS-04_maxDisap_10_runMean_10_maxDist_300.csv'
-# results_filename = "../testResults/tempNARS04.csv"
+#Get the ground truth annotations on the format [filename, x_min, y_min, x_max, y_max, id_gt]
+gt = impdata("NYAA-04")
+dt = r'U:\BITCue\Projekter\TrackingFlowers\testResults\_parameterTest_NYAA-04_3\parameterTest_NYAA-04_maxDisap_10_runMean_10_maxDist_300.csv'
+results_filename = "../testResults/tempNARS04.csv"
 
-# e = evaluator(dt, gt, results_filename, verbose = False)
-# e.run()
+e = evaluator(dt, gt, results_filename, verbose = True)
+e.run()
 
 
 # # ===================== Filter with DBSCANSieve ====
@@ -131,107 +131,129 @@ def impdata(camID):
 
 # Find a good value for eps
 
-epslist = [25, 50, 75, 100, 125, 150, 175, 200, 225, 250, 275, 300, 325, 350, 375, 400, 425, 450, 475, 500, 525, 550, 600, 650, 700]
-results_filename = '../testResults/temp_epstest.csv'
+# epslist = [350]
+# results_filename = '../testResults/temp_epstest.csv'
 
-temppath = '../testResults/temp_epsExperiment_maxDisap_10_runMean_10_maxDist_300.csv'
+# temppath = '../testResults/temp_epsExperiment_maxDisap_10_runMean_10_maxDist_300.csv'
+# gt_temppath = '../testResults/gttemp.csv'
 
-for eps in epslist:
-    mmsum = 0
+# for eps in epslist:
+#     mmsum = 0
+#     flowersum = 0
+#     # Return also the number of flowers/object returned after filtering. Find good combo between this and mismatches.
+
     
-    # Return also the number of flowers/object returned after filtering. Find good combo between this and mismatches.
-
-    
[…]

### 2022-01-11 code/DBSCANFiltering.py
Produced filtering results

@@ -146,9 +146,13 @@ class DBSCANsieve():
         ax2.axes.get_xaxis().set_visible(False)  
         ax0.axes.get_xaxis().set_visible(False)
 
-        ax0.set_aspect(1)
-        ax1.set_aspect('auto')
-        ax2.set_aspect('auto')
+        # ax0.set_aspect(1)
+        # ax1.set_aspect('auto')
+        # ax2.set_aspect('auto')
+
+        # ax0.set_aspect(6080, 3420)
+        # ax1.set_aspect(6080, 3420)
+        # ax2.set_aspect(6080, 3420)
 
         scatter = ax0.scatter(self.tracks['x_c'], self.tracks['y_c'], c = self.tracks['objectID'], s = 0.2)
         ax0 = scatter.axes
@@ -159,7 +163,7 @@ class DBSCANsieve():
         
         currentTime = datetime.now() # Use this if you need to time-stamp result file
         currentTime=('%02d-%02d-%02d'%(currentTime.hour,currentTime.minute,currentTime.second))
-        fig.savefig(f'../testResults/{currentTime}_BeforeAndAfterFiltering_eps_{self.eps_distance}.png', dpi=300)
+        fig.savefig(f'../testResults/{currentTime}_BeforeAndAfterFiltering_eps_{self.eps_distance}.png', dpi=600)
         ### ####
         
         return tracksSub

### 2022-01-11 text/ManuscriptDraft.Rmd
Manuscript update

@@
 ## Study site
+Toke, perhaps you can fill in some stuff here.
 
 ## The image series
+Original time-lapse intervals, explanation and result of sampling scheme and subsequent temporal resolution.
 
@@
 
-The annotation process is described in Mann et al., in prep, but in short: We manually annotated all flowers in the sampled image series using the rectangular bounding box tool in the VIA VGG annotation software.
-Further, we assigned each individual a unique ID.
-These annotations constitue our ground truth tracks.
+We manually annotated all flowers in the sampled image series using the rectangular bounding box tool in the VIA VGG annotation software.
+Further, we assigned each individual flower a unique ID.
+These annotations constitute our ground truth tracks.
 
-# Automatic tracking
+## Automatic tracking
 
@@
 Our algorithm tracks objects based on distances between centroids of bounding boxes.
-
-# 
-![](../figures/figure_1.png){ width=100% }
-
-**Figure 1:** Simple centroid tracking may produce erroneous associations when objects move between frames.
-Blue shows detections in the current frame (bounding box and centroid point).
-Red shows centroid points for the detections in the previous frame.
-
-# 
-![](../figures/figure_22.jpg){width=100%}
-
-**Figure 2:** Simple centroid tracking may produce erroneous associations when objects move between frames.
-Basing the association on the running mean of the positions of the previous n number of tracks may alleviate this issue.
-In this case basing the association on only the previous point would produce a wrong results while basing it on the running mean would produce a correct result.
-Red points: Centroids for bounding boxes in current frame, blue points: Centroids for bounding boxes in previous frame; grey points: Centroids for bounding boxes in a number of frames before t-1; green points: Running mean of the previous n points.
-Circles delimit the two individuals.
-
-# 
-![](../figures/figure_3.png){ width=100% }
-
-**Figure 3:** Simple centroid tracking may produce erroneous associations when objects disappear periodically from the frame.
-Here the top flower moves out of frame and the bottom flower would be assigned to the track of the top flower in frame t-1.
-
-## User parameters
 
@@
 
+#
+![](../figures/figure_1.png){ width=100% }
+
+**Figure 1:** Simple centroid tracking may produce erroneous associations when objects move between frames.
+Blue shows detections in the current frame (bounding box and centroid point).
+Red shows centroid points for the detections in the previous frame.
+
+# 
+![](../figures/figure_22.jpg){width=100%}
+
+**Figure 2:** Simple centroid tracking may produce erroneous associations when objects move between frames.
+Basing the association on the running mean of the positions of the previous n number of tracks may alleviate this issue.
+In this case basing the association on only the previous point would produce a wrong results while basing it on the running mean would produce a correct result.
+Red points: Centroids for bounding boxes in current frame, blue points: Centroids for bounding boxes in previous frame; grey points: Centroids for bounding boxes in a number of frames before t-1; green points: Running mean of the previous n points.
+Circles delimit the two individuals.
+
+# 
+![](../figures/figure_3.png){ width=100% }
+
+**Figure 3:** Simple centroid tracking may produce erroneous associations when objects disappear periodically from the frame.
+Here the top flower moves out of frame and the bottom flower would be assigned to the track of the top flower in frame t-1.
+
 ### Identifying optimal user parameters
@@
 
-# Evaluating tracking perfomance
+## Evaluating tracking perfomance
 
@@
 
-We could compare number of points in the overlapping tracks. If one is less than e.g. 10% of the other, we'll remove the small track. If more than 10% we'll remove both?
-
 We evaluate the accuracy of the remaining tracks.
@@
 
-One way to plot?: Time on x-axis, gt id on y axis, tr id as colour. This will show mismatches.
+The results of the filtering algorithm on the three series in which our tracking algorithm did not return perfect results are given in table 2 and the pipeline is visualised in fig.
+4.
+Our filtering method successfully extracted 28 tracked flowers with only a single mismatch from the three series with complex scenes using a fixed value of 350 for eps and fixed values for the tracking parameters (10, 10, 10, for running mean, max disappeared and max distance, respectively).
+
+**Table 2:** Results of the filtering algorithm with an eps value of 350 applied across all three series.
+
+#
+![](../figures/FilterScores.png){ width=100% }
+
+#
+![](../figures/filtering1.png){ width=100% }
+![](../figures/filtering2.png){ width=100% }
+![](../figures/filtering3.png){ width=100% }
+
+**Figure 4:** Track filtering pipeline.
+Row 1, 2, and 3 is NARS-13, NARS-04, and NYAA-04, respectively.
+First column shows the centroid points in the given series, coloured by the track id from the centroid tracking algorithm.
+Second column shows the polygons calculated from the tracks.
+The DBSCAN clustering algorithm with eps = 350 was applied to the centroids of these polygons and the polygons are coloured by cluster id.
+Third row shows the results of the filtering where all tracks that were not assigned a unique cluster whas been removed.
+
+# Discussion
+
+## Tracking
 
@@
 
-# Discussion
+A paragraph about the ecological perspectives of being able to track the individual flowers (at scale).
 
@@
 In cases where we are tracking perfectly with maxDisap = 0, setting it any value will not make a difference. Not quite right. Explore more...
-
-Our method for filtering tracks using DBSCAN on track centroids ensures that all tracks are given the same weighting in the filtering since each track is represented by a single point. In some cases 
-e.g.
-if it is given a priori that an object will always appear in a minimum of two frames, then single point tracks can be filtered out.
-However, when such a priori knowledge is not accessible, a conservative approach as the one we present is preferable.
 
@@
 
-# Filtering
+## Filtering
 
@@
 However, as spatially isolated objects are easier to track, it increases confidence in the remaining tracks.
+
+We applied a single value for the DBSCAN eps parameter in our tracking algorithm.
+We note that this value could be finetuned for improved results for the individual series (i.e.
+more flowers extracted without increased number of mismatches).
+However, as our goal here was to show that a single conservative value can be applied across series, we do not show those results here.
+
+Our method for filtering tracks using DBSCAN on track centroids ensures that all tracks are given the same weighting in the filtering since each track is represented by a single point. In some cases 
+e.g.
+if it is given a priori that an object will always appear in a minimum of two frames, then single point tracks can be filtered out.
+However, when such a priori knowledge is not accessible, a conservative approach as the one we present is preferable.
+
+## The image series
+
+A point on the fact the these series are complex.
 

### 2022-01-13 code/run.py
Filtering results

@@ -84,12 +84,12 @@ def impdata(camID):
 
 
 #Get the ground truth annotations on the format [filename, x_min, y_min, x_max, y_max, id_gt]
-gt = impdata("NYAA-04")
-dt = r'U:\BITCue\Projekter\TrackingFlowers\testResults\_parameterTest_NYAA-04_3\parameterTest_NYAA-04_maxDisap_10_runMean_10_maxDist_300.csv'
-results_filename = "../testResults/tempNARS04.csv"
+# gt = impdata("NYAA-04")
+# dt = r'U:\BITCue\Projekter\TrackingFlowers\testResults\_parameterTest_NYAA-04_3\parameterTest_NYAA-04_maxDisap_10_runMean_10_maxDist_300.csv'
+# results_filename = "../testResults/tempNARS04.csv"
 
-e = evaluator(dt, gt, results_filename, verbose = True)
-e.run()
+# e = evaluator(dt, gt, results_filename, verbose = True)
+# e.run()
 
 
 # # ===================== Filter with DBSCANSieve ====
@@ -131,31 +131,31 @@ e.run()
 
 # Find a good value for eps
 
-# epslist = [350]
-# results_filename = '../testResults/temp_epstest.csv'
+epslist = [350]
+results_filename = '../testResults/temp_epstest.csv'
 
-# temppath = '../testResults/temp_epsExperiment_maxDisap_10_runMean_10_maxDist_300.csv'
-# gt_temppath = '../testResults/gttemp.csv'
+temppath = '../testResults/temp_epsExperiment_maxDisap_10_runMean_10_maxDist_300.csv'
+gt_temppath = '../testResults/gttemp.csv'
 
-# for eps in epslist:
-#     mmsum = 0
-#     flowersum = 0
-#     # Return also the number of flowers/object returned after filtering. Find good combo between this and mismatches.
+for eps in epslist:
+    mmsum = 0
+    flowersum = 0
+    # Return also the number of flowers/object returned after filtering. Find good combo between this and mismatches.
 
     
-#     # dt = r'U:\BITCue\Projekter\TrackingFlowers\testResults\_parameterTest_NARS-17_3\parameterTest_NARS-17_maxDisap_10_runMean_10_maxDist_300.csv'
-#     # gt = pd.read_csv(impdata("NARS-17"))
-        
-#     # d = DBSCANsieve(dt, eps_distance = eps, min_sample_polygons = 1)
[…]

### 2022-01-13 text/ManuscriptDraft.Rmd
Manuscript updates

@@
 title: |
-  **Monitoring the phenology of individual flowers using deep learning and automatic tracking**
+  **Monitoring the phenology of individual flowers using automatic tracking and filtering**
 author:
@@
 - \usepackage{amsmath}
-keywords: Tracking, flowering phenology, Arctic, 
+keywords: Tracking, flowering phenology, Arctic, DBSCAN
 bibliography: ./library.bib
@@
 Here we show that image-based monitoring of field plots at very high temporal resolution can return information on flowering phenology at the level of indiviuals.
-Further, we present a framework for automatically tracking, filtering, and visualizing flowers in time-lapse image series.
+Further, we present a framework for automatically tracking, filtering, and evaluating flowers in time-lapse image series.
 We compare the results of automatic tracking with manual tracking.
@@
 Precocious flowering means that the plant has failed to exploit the whole temporal window for accumulating resources before allocating energy to flowering.
-On the other hand, flowering too late limit the time for reproduction before the end of the growing season [@ELZINGA2007].
+Oppositely, flowering too late limits the time for reproduction before the end of the growing season [@ELZINGA2007].
 Further, flowering may need to be synchronous with pollinator activity for successful reproduction.
-Flowering phenology may plastically change as a response to abiotic cues in the environment, such as timing of spring, temperature, and photoperiod, but variation in flowering phenology is partly heritable and shaped by selective forces from the abiotic and biotic environment.
+Flowering phenology may plastically change as a response to abiotic cues in the environment, such as timing of spring, temperature, and photoperiod, but variation in flowering phenology is partly heritable and shaped by selective forces from the abiotic and biotic environment and thus sensitive to climate change.
 
-Monitoring of flower phenology at high temporal resolution is laboursome and time-consuming, particularly in logistically challenging environment such as the Arctic.
+Monitoring of flower phenology at high temporal resolution is laboursome and time-consuming, particularly in logistically challenging environments such as the Arctic.
 Consequently, simple variables are often used as proxies for the flowering phenology of a population, such as the date for onset of flowering, often derived from weekly observations of sample plots.
@@
 For example, a shortening of individual flower longevity may not be directly obvious at the population level.
-Many research questions can only be explored on the basis of individual phenology data.
-For example, investigating the association between reproductive success and timing of flowering and flower longevity requires phenology data at the level of individuals.
-Similarly, such data is necessary for investigating whether flower visitation rates and/or reproductive success depends on the timing of flowering for the individual flower.
+In fact, many research questions can only be explored on the basis of individual phenology data.
+For example, investigating the association between reproductive success and timing of flowering or flower longevity requires phenology data at the level of individuals.
+Similarly so for investigating whether flower visitation rates and/or reproductive success depends on the timing of flowering for the individual flower.
 
 Here, we show that information on phenology at the level of individuals can be derived from image-based monitoring of flower phenology.
-Further, we present and evaluate an automatic flower tracking and filtering algorithm.
+
+We present and evaluate an automatic flower tracking and filtering algorithm.
+Many tracking algorithms, such as the Simple online and realtime tracking (SORT) [@sort], use Kalman filtering to predict the future location of an object by estimating its velocity [@kalmann].
+However, Kalman filtering assumes linear motion of the object and does not handle abrupt motion well.
+As the flowers are constrained in movement by their stalk, their movement does not follow the assumption of linear motion.
+We therefore base the track assignment on the minimum distance between centroids of bounding boxes.
+However, we introduce three user-set parameters that can help optimise tracking performance.
+
+For complex scenes with many flowers in close vicinity to each other, we suggest a conservative filtering approach.
+The approach may remove correct tracks, but the tracks that remain will have a low risk of tracking errors.
+The approach allows for confident upscaling of the method.
+I.e., when running the method on a large number of image series, it is preferable to extract individual high-confidence tracks from each series and ignore the remaining tracks.
 
 Tracking individual flowers enables the possibility of assigning reproductive success to the individual, for example by observation of seed set.
-Thereby, it can be explored whether reproductive success is affected by timing and length of flowering.
+Thereby it could be explored whether reproductive success is affected by timing and length of flowering.
 By simultaneously tracking flower visits, these could be assigned to the individual flower and visitation rates per flower could be calculated per flower and related to reproductive success.
 Further, any information of taxonomic grouping could refine this analyses.
-
-For complex scenes with many flowers in close vicinity to each other, we suggest a conservative filtering approach.
-The approach may remove correct tracks, but the tracks that remain will have a lower risk of tracking errors.
-The approach allows for more confidence in upscaling the method.
-E.g., when running the method on a large number of image series, it is preferable to extract individual high-confidence tracks from each series and ignore the remaining tracks.
+We argue that the possibility of tracking phenology at the level of individuals will facilitate investigation of a novel set of research questions and is an important contribution to the shift towards efficient and automatic monitroing of ecological processes.
 
 # Material and methods
-
+#
 ## Study site
 Toke, perhaps you can fill in some stuff here.
+NARS 2018, NARS 2019, NYAA 2019, THUL 2018.
 
+#
 ## The image series
@@
 
+#
 ## Flower annotations
 
-We manually annotated all flowers in the sampled image series using the rectangular bounding box tool in the VIA VGG annotation software.
-Further, we assigned each individual flower a unique ID.
+`r word_comment("Add a figure showing example images with annotations.", "We")` manually annotated all flowers in the sampled image series using the rectangular bounding box tool in the VIA VGG annotation software and assigned each individual flower a unique ID.
 These annotations constitute our ground truth tracks.
 
+#
 ## Automatic tracking
@@
 
-Our algorithm tracks objects based on distances between centroids of bounding boxes.
-
-The tracking algorithm has a set of user adjusted parameters that can optimize tracking accuracy.
+`r word_comment("Text below is maybe more suited for intro/disc?
+Attempting to explain the reasoning of the method.", "The")` tracking algorithm tracks objects based on distances between centroids of bounding boxes.
+The algorithm has a set of user adjusted parameters that can optimize tracking accuracy.
 The parameters are particularly relevant for optimal tracking of objects that are constrained to a specific area such as flowers.
-It is important to note, however, that the tracking algorithm can be used to track any objects.
-The tracking algorithm can be applied both offline (on a set of detections/annotations that have already been produced) or online (real-time tracking frame per frame).
+It is important to note, however, that the tracking algorithm could be used to track any objects.
+The algorithm can be applied both offline (on a set of detections/annotations that have already been produced) or online (real-time tracking frame per frame).
 The speed of the tracking algorithm depends on the computational power available as well as the number of objects that are being tracked.
 The method is fast, however.
-Tracking of a series containing 85 objects ran at 0.02 seconds per frame.
+Tracking of a series containing 85 objects ran at ~0.02 seconds per frame on a standard laptop.
 
-As the wind shifts, the flower heads changes direction.
+Several things complicate the task of tracking individual flowers through the time-lapse series.
+First, as the wind shifts, the flower heads changes direction.
 This can happen instantaneously (i.e.
 between two consecutive frames).
-As they are constrained by their stalk, there is a limit to the distance they can move.
+However, as they are constrained by their stalk, there is a limit to the distance they can move.
+Establishing associations between points based on just the distance between points in the current and the previous frame can cause errors when flowers are in close vicinity of each other.
+With the tracking parameter **running mean**, we base the tracking on the distance between a point in the current frame and the running mean of the positions of the previous X points in a track.
 
-Establishing associations between points based on just the distance between points in the current and the previous frame can cause errors when flowers are in close vicinity of each other.
+Second, when the location of the flowers change they may temporarily occlude the view of other flowers and  flowers close to the edge of the frame may move in and out of view.
+This can cause a track to be lost and a new track erroneously being initiated when the flower reappears.
+However, if a flower reappears in the same area as a flower is already being tracked after disappearing in a few frames, it is a reasonable assumption that it is the same individual and not that the old flower wilted/disappeared and a new one developed.
+The parameter **max disappeared** sets the number of frames a track can be lost before a new track is initiated for points appearing in the same area.
+Concurrently, this deals with potential false negatives.
+If a given flower has not been annotated in a few frames, the track will not be lost.
+The counter for number of disappeared frames for a track is reset when a new point is associated with the track within the threshold.
 
-The flowers move around a center point because of their stalk.
-We base the tracking on the distance between a point in the current frame and the running mean of the positions of the previous X points in a track.
-
-As winds shift, flowers close to the edge of the frame may move in and out of view.
-If a flower reappears in the same area as a flower is already being tracked after disappearing in a few frames, it is a reasonable assumption that it is the same individual and not that the old flower wilted/disappeared and a new one developed.
-The parameter **max disappeared** sets the number of frames a track can be lost before a new track is initiated for points appearing in the same area.
-
-Similarly, this deals with potential false negatives. If a given flower has not been annotated in a single frame, it should not be assigned a new track.
-
-Setting **max disappeared** to 0 tracks objects based on the coordinates of the points in only the previous frame.
-The counter for number of disappeared frames is reset when a new point is associated with the track within the threshold.
+Third, a centroid tracking algorithm will associate a point with a tracking based on minimum distance only, disregarding the absolute distance.
+As flowers are constrained in the movement, we can assume a maximum distance that will occur between points and force the initiation of tracks for points that exceed this threshold.
+We do that with the parameter **max distance**.
 
@@
 
-**Figure 1:** Simple centroid tracking may produce erroneous associations when objects move between frames.
+**Figure 1:** `r word_comment("The three figures below are quick mockups to explain the challenges with tracking flowers.
+Not sure if figures are needed or text is sufficient.", "Simple")` centroid tracking may produce erroneous associations when objects move between frames.
 Blue shows detections in the current frame (bounding box and centroid point).
@@
 
+#
 ### Identifying optimal user parameters
 
-To estimate the optimal values for max_disappeared and max_distance, we analysed the ground truth tracks of the four series.
+To explore the effect of the user parameters and to identify the optimal combination of parameters for our case of tracking flowers, we ran the tracking algorithm on each of the five image series with every combination of a range of values for each parameter (3.179 combinations): Values for max disappeared and running mean were 0-160 with a step size of 10 and for max distance we used values 0-1000 with a step size of 100.
+Note that set to zero ignores the parameter altogether.
+We identified the setting(s) that returned the lowest number of track mismatches snd compare the tracking results between optimal settings and all parameters set to zero.
 
-To estimate the value for max_distance, we calculated the largest distance between any two points within any track for any flower in each of the four series.
+`r word_comment("I haven't included these results.
+Not sure if they should be.", "Additionally,")` we performed a set of analyses on the five ground tracks.
+First, we calculated the largest distance between any two points within any track for any flower in each of the five series.
+Further, we calculated the maximum number of frames a flower track was lost and subsequently reappeared for each series.
 
-The maximum number of frames a flower track was lost and subsequently reappeared were xxx.
-
-To explore the effect of the user parameters and to identify the optimal combination of parameters for our case of tracking flowers, we followed a step wise approach.
-First, we ran the tracking algorithm on each of the four image series with every combination of the following settings (3.179 combinations):
-
-max_disappeared = [0,10,20,30,40,50,60,70,80,90,100,110,120,130,140,150,160]
-
-running_mean_threshold = [0,10,20,30,40,50,60,70,80,90,100,110,120,130,140,150,160]
-
-max_distance = [0, 100, 200, 300, 400, 500, 600, 700, 800, 900, 1000]
-
-Note that max_distance set to zero.
-We identified the setting(s) that returned the lowest number of track mismatches and performed a second run with finer scaled settings for each series.
-Finally, we compare the tracking results between optimal settings and all parameters set to zero.
-
-Depending on the nature of the objects being tracked and the complexity of the scene, the user parameters can be estimated from visual examination of the tracking results.
-Often it may be preferable to manually annotate a subset of the objects in the image series and derive a set of user parameters from these results.
-
+#
 ## Evaluating tracking perfomance
@@
 To associate other information obtained in the images to the individual flower, for example flower visits, we want as much as possible of the track to be correct.
-To derive flowering length in theory we just need to track the most extreme points correctly and don't care about the intermediate points.
+To derive flowering length, in theory we just need to track the most extreme points correctly and can ignore the intermediate points.
 Lastly, we may be interested in the number of flowers that existed in a plot, in which case we want the number of tracks obtained by automatic tracking to be as close as possible to the actual number of individuals in the series.
@@
 Only the shifts in tracking identity are counted as mismatches while the number of points assigned to each track is not considered.
+We calculate the number of mismatches and the MOTA score to evaluate performance of our tracking and filtering algorithms.
+Further, we compare the number of tracks identified by the automatic tracking with the true number of flowers in a series.
+These should ideally be equal.
 
-For each series, we calculate the ratio of flowers for which the automatic tracking algorithm returns the correct flowering length compared to the ground truth tracks.
-
-Finally, we compare the number of tracks identified by the automatic tracking with the true number of flowers in a series. These should ideally be equal.
-
+#
 ## Filtering tracks
@@
 
-For tracks consisting of two points, we establish the straight line between the points.
-For tracks consisting of three points, we establish the triangle from the points.
-For tracks consisting of more than three points, we calculate the convex hull of all the points included in the track and derive the polygon from the vertices of the convex hull.
-Single points are kept.
+We base our filtering on a density based clustering of the centroid of a geometry of each track.
+The geometries are derived in the following way: For tracks consisting of only a single point, the coordinates of the point are used as the centroid.
+For tracks consisting of two points, we establish the straight line between the points and calculate the centroid of the line.
+For tracks consisting of three points, we establish the triangle from the points and calculate the centroid of the triangle.
+For tracks consisting of more than three points, we calculate the convex hull of all the points included in the track and derive the polygon from the vertices of the convex hull and calculate the centroid of this polygon.
+For tracks that contain more than two points but where the points a colinear, we establish the line through the points and calculate the centroid of the line.
 
-We then apply the DBSCAN clustering algorithm on these track geometrics to remove tracks in areas with a high density of tracks as these have a high risk of tracking mismatches.
+We then apply the filtering based on the geometry centroids to remove tracks in areas with a high density of tracks as these have a high risk of tracking mismatches.
 The filtering is done in two steps.
-First, DBSCAN algorithm is run with a conservatively high value for the eps parameter, meaning that tracks in close vicinity to each other will be clustered together.
+First, the DBSCAN clustering algorithm is run on the centroids with a conservatively high value for the eps parameter (350), meaning that tracks in close vicinity to each other will be clustered together.
 Second, all tracks that were not assigned to a unique cluster are removed.
 Tracks that are spatially isolated remains.
-
-We evaluate the accuracy of the remaining tracks.
-
+We evaluate the tracking accuracy of the remaining tracks.
 We demonstrate our filtering approach for the image series for which our tracking algorithm did not produce perfect results.
 
-Here we set eps DBSCAN parameter to XXX.
-Although the value could be fine-tuned for improved results for each image series individually, this value returns good results overall.
-In a naive setting, a general value could be chosen of the value could be adjusted for each series based on visual examination of performance or testing on a subset of data.
-
+#
 # Results
 
-**Table 1:** Bla bla...
+The results of the tracking parameter test are given in table 1.
+We present the tracking performance without the use of the three tracking parameters along with the best performance (lowest number of mismatches/highest MOTA), and the best performance where the number of tracks correspond to the number of individual flowers in the series.
+
+**Table 1:** Table description...
 #
@@
 
-The results of the filtering algorithm on the three series in which our tracking algorithm did not return perfect results are given in table 2 and the pipeline is visualised in fig.
+The results of the filtering algorithm on the three series in which our tracking algorithm did not return perfect results are given in table 2 and the filtering process is visualised in fig.
 4.
-Our filtering method successfully extracted 28 tracked flowers with only a single mismatch from the three series with complex scenes using a fixed value of 350 for eps and fixed values for the tracking parameters (10, 10, 10, for running mean, max disappeared and max distance, respectively).
+Our filtering method successfully extracted 28 tracked flowers with only a single mismatch from the three series with complex scenes using a fixed value of 350 for eps and fixed values for the tracking parameters (10, 10, 300, for running mean, max disappeared and max distance, respectively).
 
-**Table 2:** Results of the filtering algorithm with an eps value of 350 applied across all three series.
+**Table 2:** Results of the filtering algorithm with an eps value of 350 applied across all three .
+`r word_comment("Note to self.
+Table shows to what degree a tr track contains points from a single gt track, but does not show whether all points of the gt track are included in the tr track.
+We know that is the case, but should it be tested/shown?", "Table 1")` series.
 
-#
 ![](../figures/FilterScores.png){ width=100% }
@@
 
-**Figure 4:** Track filtering pipeline.
+**Figure 4:** Track filtering process.
 Row 1, 2, and 3 is NARS-13, NARS-04, and NYAA-04, respectively.
@@
 The DBSCAN clustering algorithm with eps = 350 was applied to the centroids of these polygons and the polygons are coloured by cluster id.
-Third row shows the results of the filtering where all tracks that were not assigned a unique cluster whas been removed.
+Third column shows the results of the filtering where all tracks that were not assigned a unique cluster has been removed.
 
@@
 
+#
 ## Tracking
-
-Our tracking algorithm consistently returns high MOTA scores.
-
-A paragraph about the ecological perspectives of being able to track the individual flowers (at scale).
-
-All three parameters make a difference.
-
-In cases where we are tracking perfectly with maxDisap = 0, setting it any value will not make a difference. Not quite right. Explore more...
-
-Applying the technique on detection instead of manual annotations.
-Detections introduce false positives.
-Either manual or automatic quality control to remove these before tracking or after.
-
-Our manually tracked data is ground truth, but for example when flowers periodically move out of the frame, this mimics false negatives. Similarly when one flower occludes another.
-
-Our tracking algorithm returned high MOTA scores even with all three parameters set to zero.
-
-Automatically detecting flowers would likely introduce a degree of false negatives which would decrease the MOTA score if max disappeared is set to zero.
-Introducing a value for this parameter can deal with the problem of false negatives.
 
@@
 
+Depending on the nature of the objects being tracked and the complexity of the scene, the user parameters can be estimated from visual examination of the tracking results.
+Often it may be preferable to manually annotate a subset of the objects in the image series and derive a set of user parameters from these results.
+
+Number of tracks equal to number of flowers does not necessarily mean that the tracking is correct.
+I.e., the tracking algorithm could split one ground truth track in two, but combine two other ground truth tracks into one, which would even each other out.
+
+Our tracking algorithm consistently returns high MOTA scores.
+
+For complex scenes, all three parameters make a difference.
+
+The parameters are interdependent.
+
+In cases where we are tracking perfectly with maxDisap = 0, setting it any value will not make a difference. Not quite right. Explore more...
+
+Paragraph about applying the technique on detection instead of manual annotations.
+Detections introduce false positives.
+Either manual or automatic quality control to remove these before tracking or after.
+
+Our manually tracked data is ground truth, but for example when flowers periodically move out of the frame, this mimics false negatives. Similarly when one flower occludes another.
+
+Automatically detecting flowers would likely introduce a degree of false negatives which would decrease the MOTA score if max disappeared is set to zero.
+Introducing a value for this parameter can deal with the problem of false negatives.
+
+Our tracking algorithm returned reasonably high MOTA scores even with all three parameters set to zero. The degree to which mismatches are accepted may depend on the application?
+
+#
 ## Filtering
@@
 We applied a single value for the DBSCAN eps parameter in our tracking algorithm.
-We note that this value could be finetuned for improved results for the individual series (i.e.
+We note that this value could be fine-tuned for improved results for the individual series (i.e.
 more flowers extracted without increased number of mismatches).
 However, as our goal here was to show that a single conservative value can be applied across series, we do not show those results here.
+In a naive setting, a general value could be chosen or the value could be adjusted for each series based on visual examination of performance or testing on a subset of data.
 
@@
 
+#
 ## The image series
@@
 A point on the fact the these series are complex.
+
+#
+## Conclusion
+A paragraph about the ecological perspectives of being able to track the individual flowers (at scale).
 
@@
 
-Tracking Multiple Moving Objects Using Unscented Kalman
-Filtering Techniques:
-"Kalman filtering (KF) [5] is widely used to track moving objects, with which we
-can estimate the velocity and even acceleration of an object with the measurement of its locations.
-However, the accuracy of KF is dependent on the assumption of linear motion for any object to be
-tracked. If an object takes some abrupt turns, the nonlinear movement cannot be well handled by the
-KF framework (due to the linear movement assumption of the design of KF)."
-Kalman, R. E. A New Approach to Linear Filtering and Prediction Problems. Journal of Basic
-Engineering, 1960(82), pp. 35-45.
+For each series, we calculate the ratio of flowers for which the automatic tracking algorithm returns the correct flowering length compared to the ground truth tracks.
 

### 2022-01-17 text/ManuscriptDraft.Rmd
Manuscript update rev1

@@
 title: |
-  **Monitoring the phenology of individual flowers using automatic tracking and filtering**
+  **Towards individual-based pollination ecology: Automatic tracking of life histories of individual flowers**
 author:
@@
 csl: ./journal-of-ecology.csl
-abstract: ABSTRACT | Often simple variables will be used to describe the flowering phenology of a population of plants, e.g.
+abstract: ABSTRACT |
+
+A fundamental problem in pollination ecology is that plant reproduction consequences of pollinator visits to a flower throughout its lifespan is very difficult to track.
+Often simple statistics are used to describe the flowering phenology of a population of plants, e.g.
 onset or peak of flowering.
+Similarly, flower visitation rates are based on crude observation of flower populations.
 Here we show that image-based monitoring of field plots at very high temporal resolution can return information on flowering phenology at the level of indiviuals.
@@
 We compare the results of automatic tracking with manual tracking.
+Individual flower tracking could facilitate finegrained studies on plant reproductive success.
 ---
@@
 Oppositely, flowering too late limits the time for reproduction before the end of the growing season [@ELZINGA2007].
+Both precocious and late flowering increase the risk of frost damage to the flowers.
 Further, flowering may need to be synchronous with pollinator activity for successful reproduction.
-Flowering phenology may plastically change as a response to abiotic cues in the environment, such as timing of spring, temperature, and photoperiod, but variation in flowering phenology is partly heritable and shaped by selective forces from the abiotic and biotic environment and thus sensitive to climate change.
+Flowering phenology may plastically change in response to abiotic cues in the environment, such as timing of spring, temperature, and photoperiod, but flowering phenology is partly heritable and shaped by selective forces from the abiotic and biotic environment and in any case sensitive to climate change.
 
-Monitoring of flower phenology at high temporal resolution is laboursome and time-consuming, particularly in logistically challenging environments such as the Arctic.
-Consequently, simple variables are often used as proxies for the flowering phenology of a population, such as the date for onset of flowering, often derived from weekly observations of sample plots.
-Such proxies may fail to reveal dynamics in flowering phenology for example caused by changes in climate.
+Monitoring flower phenology at high temporal resolution is time-consuming, particularly in logistically challenging environments such as the Arctic.
+Consequently, flowering phenology of a population is often characterized simply by the date for onset of flowering, typically derived from weekly observations of sample plots [@Prevey2021].
+Such inaccurate estimates may fail to reveal dynamics in flowering phenology for example caused by changes in climate.
 
-Automatic image-based monitoring of flowering phenology can return phenology data for specific species at very high temporal resolution (Mann et al., in prep), but phenological responses at the individual level may be indiscernible regardless of the temporal resolution of the data at population level.
+Automatic image-based monitoring of flowering phenology can return phenology data for specific species at very high temporal resolution (Mann et al., in review), but phenological responses at the individual level may be indiscernible regardless of the temporal resolution of the data at population level.
 For example, a shortening of individual flower longevity may not be directly obvious at the population level.
@@
 As the flowers are constrained in movement by their stalk, their movement does not follow the assumption of linear motion.
-We therefore base the track assignment on the minimum distance between centroids of bounding boxes.
-However, we introduce three user-set parameters that can help optimise tracking performance.
+We therefore base the track assignment on the minimum distance between centroids of bounding boxes and introduce three user defined parameters that can help optimise tracking performance.
 
@@
 
-Tracking individual flowers enables the possibility of assigning reproductive success to the individual, for example by observation of seed set.
-Thereby it could be explored whether reproductive success is affected by timing and length of flowering.
+Our method for tracking individual flowers facilitates the possibility of assigning indicators of reproductive success and factors driving variation in seed set to the individual.
+For instance, it could be explored whether seed set is affected by timing and length of flowering and/or abiotic conditions such as frost events.
+Hereby, the longstanding question of what drives variation in reproductive success of individual flowers could be explored.
 By simultaneously tracking flower visits, these could be assigned to the individual flower and visitation rates per flower could be calculated per flower and related to reproductive success.
-Further, any information of taxonomic grouping could refine this analyses.
-We argue that the possibility of tracking phenology at the level of individuals will facilitate investigation of a novel set of research questions and is an important contribution to the shift towards efficient and automatic monitroing of ecological processes.
+Further, any information of taxonomic grouping could facilitate comparative studies on pollination effectiveness between taxonomic groups of pollinators.
+We argue that the possibility of tracking phenology at the level of individuals will facilitate investigation of a novel set of research questions and is an important contribution to the shift towards efficient and automatic monitoring of ecological processes.
 
@@
 
-`r word_comment("Text below is maybe more suited for intro/disc?
-Attempting to explain the reasoning of the method.", "The")` tracking algorithm tracks objects based on distances between centroids of bounding boxes.
+The tracking algorithm tracks objects based on distances between centroids of bounding boxes.
 The algorithm has a set of user adjusted parameters that can optimize tracking accuracy.

### 2022-01-18 text/ManuscriptDraft.Rmd
Rev1 comments from TTH and AI implmented or comment added

@@
 Similarly, flower visitation rates are based on crude observation of flower populations.
-Here we show that image-based monitoring of field plots at very high temporal resolution can return information on flowering phenology at the level of indiviuals.
+Here we show that image-based monitoring of field plots at very high temporal resolution can return information on flowering phenology at the level of individuals.
 Further, we present a framework for automatically tracking, filtering, and evaluating flowers in time-lapse image series.
@@
 Individual flower tracking could facilitate finegrained studies on plant reproductive success.
+
+“and we show that performance achieved by following automatic tracking can be reliable enough to estimate X, Y, Z.”
 ---
@@
 
-Here, we show that information on phenology at the level of individuals can be derived from image-based monitoring of flower phenology.
-
+In this paper, we show that information on phenology at the level of individuals can be derived from image-based monitoring of flower phenology.
 We present and evaluate an automatic flower tracking and filtering algorithm.
-Many tracking algorithms, such as the Simple online and realtime tracking (SORT) [@sort], use Kalman filtering to predict the future location of an object by estimating its velocity [@kalmann].
+Many tracking algorithms, such as the Simple online and real-time tracking (SORT) [@sort], use Kalman filtering to predict the future location of an object by estimating its velocity [@kalmann].
 However, Kalman filtering assumes linear motion of the object and does not handle abrupt motion well.
@@
 
+DeepSORT and other tracking frameworks leverage convolutional neural networks (CNNs) for object classification.
+
+But in our case, flowers are not distinguishable and one flower may look less alike between two frames than two flowers in a single frame.
+
 For complex scenes with many flowers in close vicinity to each other, we suggest a conservative filtering approach.
@@
 The approach allows for confident upscaling of the method.
-I.e., when running the method on a large number of image series, it is preferable to extract individual high-confidence tracks from each series and ignore the remaining tracks.
+I.e., when applying the method on a large number of image series, it is preferable to extract individual high-confidence tracks from each series and ignore the remaining tracks.
 
@@
 
-`r word_comment("Add a figure showing example images with annotations.", "We")` manually annotated all flowers in the sampled image series using the rectangular bounding box tool in the VIA VGG annotation software and assigned each individual flower a unique ID.
+`r word_comment("Add a figure showing example images with annotations.", "We")` manually annotated all flowers in the sampled image series using the rectangular bounding box tool in the VIA VGG annotation software [@dutta2019] and assigned each individual flower a unique ID.
 These annotations constitute our ground truth tracks.
@@
 The tracking algorithm tracks objects based on distances between centroids of bounding boxes.
-The algorithm has a set of user adjusted parameters that can optimize tracking accuracy.
+The algorithm has a set of user-adjusted parameters that can optimize tracking accuracy.
 The parameters are particularly relevant for optimal tracking of objects that are constrained to a specific area such as flowers.
@@
 
-Several things complicate the task of tracking individual flowers through the time-lapse series.
+Several issues complicate the task of tracking individual flowers through the time-lapse series.
 First, as the wind shifts, the flower heads changes direction.
-This can happen instantaneously (i.e.
-between two consecutive frames).
+This can happen instantaneously (i.e., between two consecutive frames).
 However, as they are constrained by their stalk, there is a limit to the distance they can move.
@@
 
-Second, when the location of the flowers change they may temporarily occlude the view of other flowers and  flowers close to the edge of the frame may move in and out of view.
+Second, when the location of the flowers changes they may temporarily occlude the view of other flowers and  flowers close to the edge of the frame may move in and out of view.
 This can cause a track to be lost and a new track erroneously being initiated when the flower reappears.
-However, if a flower reappears in the same area as a flower is already being tracked after disappearing in a few frames, it is a reasonable assumption that it is the same individual and not that the old flower wilted/disappeared and a new one developed.
-The parameter **max disappeared** sets the number of frames a track can be lost before a new track is initiated for points appearing in the same area.
+However, if a flower reappears in the same area as a flower is already being tracked after disappearing in a few frames, it is reasonable to assume that it is the same individual and not that the old flower wilted/disappeared and a new one developed.
+The parameter **max gap size** sets the number of frames a track can be lost before a new track is initiated for points appearing in the same area.
 Concurrently, this deals with potential false negatives.
-If a given flower has not been annotated in a few frames, the track will not be lost.
+If a given flower has not been detected in a few frames, the track will not be lost.
 The counter for number of disappeared frames for a track is reset when a new point is associated with the track within the threshold.
@@
 
+# EDIT FIGURES: Zoom in, bigger dots, combine into one figure.
 #
@@
 #
+### Input data for the tracking algorithm
+
+The input data of
+
+Calculating tracking accuracy requires a ground truth ID for each bounding box.
+
+This data structure is the same that is output from manual object annotations and object detection methods with CNNs.
+
+#
 ### Identifying optimal user parameters
 
-To explore the effect of the user parameters and to identify the optimal combination of parameters for our case of tracking flowers, we ran the tracking algorithm on each of the five image series with every combination of a range of values for each parameter (3.179 combinations): Values for max disappeared and running mean were 0-160 with a step size of 10 and for max distance we used values 0-1000 with a step size of 100.
+To explore the effect of the user parameters and to identify the optimal combination of parameters for our case of tracking flowers, we applied the tracking algorithm on each of the five image-series with every combination of a range of values for each parameter (3179 combinations): Values for max gap size and running mean were 0-160 with a step size of 10 and for max distance we used values 0-1000 with a step size of 100.
 Note that set to zero ignores the parameter altogether.
-We identified the setting(s) that returned the lowest number of track mismatches snd compare the tracking results between optimal settings and all parameters set to zero.
+We identified the setting(s) that returned the lowest number of track mismatches and compare the tracking results between optimal settings and all parameters set to zero.
 
 `r word_comment("I haven't included these results.
-Not sure if they should be.", "Additionally,")` we performed a set of analyses on the five ground tracks.
+Not sure if they should be.", "Additionally,")` we performed a set of analyses on the five ground truth tracks.
 First, we calculated the largest distance between any two points within any track for any flower in each of the five series.
@@
 #
-## Evaluating tracking perfomance
+## Evaluating tracking performance
 
@@
 To associate other information obtained in the images to the individual flower, for example flower visits, we want as much as possible of the track to be correct.
-To derive flowering length, in theory we just need to track the most extreme points correctly and can ignore the intermediate points.
-Lastly, we may be interested in the number of flowers that existed in a plot, in which case we want the number of tracks obtained by automatic tracking to be as close as possible to the actual number of individuals in the series.
+To derive flowering length, in theory we just need to be able to connect the first and the last image of a flower and can ignore the intermediate points.
+Lastly, in some scenarios one may be interested in the number of flowers that existed in a plot, in which case we want the number of tracks obtained by automatic tracking to be as close as possible to the actual number of individuals in the series.
 
@@
 Only the shifts in tracking identity are counted as mismatches while the number of points assigned to each track is not considered.
-We calculate the number of mismatches and the MOTA score to evaluate performance of our tracking and filtering algorithms.
-Further, we compare the number of tracks identified by the automatic tracking with the true number of flowers in a series.
+We use three evaluation metrics: We calculate the number of mismatches and derive the MOTA score (number of mismatches in relation to the total number of points) to evaluate performance of our tracking and filtering algorithms.
+Finally, we compare the number of tracks identified by the automatic tracking with the true number of flowers in a series.
 These should ideally be equal.
@@
 
-We base our filtering on a density based clustering of the centroid of a geometry of each track.
+We base our filtering on a density-based clustering of the centroid of a geometry of each track.
 The geometries are derived in the following way: For tracks consisting of only a single point, the coordinates of the point are used as the centroid.
@@
 For tracks consisting of three points, we establish the triangle from the points and calculate the centroid of the triangle.
-For tracks consisting of more than three points, we calculate the convex hull of all the points included in the track and derive the polygon from the vertices of the convex hull and calculate the centroid of this polygon.
+For tracks consisting of three points or more, we calculate the convex hull of all the points included in the track and derive the polygon from the vertices of the convex hull and calculate the centroid of this polygon.
 For tracks that contain more than two points but where the points a colinear, we establish the line through the points and calculate the centroid of the line.
@@
 The filtering is done in two steps.
-First, the DBSCAN clustering algorithm is run on the centroids with a conservatively high value for the eps parameter (350), meaning that tracks in close vicinity to each other will be clustered together.
+First, the DBSCAN clustering algorithm [@ester1996] is run on the centroids with a conservatively high value for the eps parameter (350), meaning that tracks in close vicinity to each other will be clustered together.
+
+We identify a good value for eps by...
+
 Second, all tracks that were not assigned to a unique cluster are removed.
@@
 4.
-Our filtering method successfully extracted 28 tracked flowers with only a single mismatch from the three series with complex scenes using a fixed value of 350 for eps and fixed values for the tracking parameters (10, 10, 300, for running mean, max disappeared and max distance, respectively).
+Our filtering method successfully extracted 28 tracked flowers with only a single mismatch from the three series with complex scenes using a fixed value of 350 for eps and fixed values for the tracking parameters (10, 10, 300, for running mean, max gap size and max distance, respectively).
 
@@
 
-We ran our tracking algorithm on 3xxx combination of the three parameters.
+We applied our tracking algorithm on 3179 combinations of the three parameters.
 Using the parameters increased the performance of the tracking substantially.
@@
 Here we do not perform this analysis, however, as our goal is to show that the parameters can be used for optimizing tracking performance in general.
+
+We have demonstrated that our tracking algorithm with the three tracking parameters and the subsequent filtering algorithm can substantially increase accuracy of the tracking of flowers despite the fact that they break the assumptions of general tracking methods.
+
+In a naive setting, a priori knowledge about which settings to use is not available.
+
+As with any tracking method, accuracy can only be assessed based on visual ... or more ... based on a subsample of tracks for whih a ground truth exists.
+
+We recommend to perform a parameter test as demonstrated here on subset of ground truth tracks to derive optimal parameters for application of the method in a new setting.
 
@@
 
-Our manually tracked data is ground truth, but for example when flowers periodically move out of the frame, this mimics false negatives. Similarly when one flower occludes another.
+Our manually tracked data is ground truth, but for example when flowers periodically move out of the frame, this mimics false negatives. Similarly, when one flower occludes another.
 
-Automatically detecting flowers would likely introduce a degree of false negatives which would decrease the MOTA score if max disappeared is set to zero.
+Automatically detecting flowers would likely introduce a degree of false negatives which would decrease the MOTA score if max gap size is set to zero.
 Introducing a value for this parameter can deal with the problem of false negatives.
@@
 However, as spatially isolated objects are easier to track, it increases confidence in the remaining tracks.
+
+**Alekos: I think it would be good to add a short description or comment on what each of the mentioned parameters is used for (e.g.
+why did you need to change its value, how this affects the performance of the method on the challenges in your dataset compared to standard computer vision tracking datasets).**
 

### 2022-01-21 text/ManuscriptDraft.Rmd
Manuscript update

@@
 csl: ./journal-of-ecology.csl
-abstract: ABSTRACT |
-
-A fundamental problem in pollination ecology is that plant reproduction consequences of pollinator visits to a flower throughout its lifespan is very difficult to track.
+abstract: ABSTRACT | A fundamental problem in pollination ecology is that plant reproduction consequences of pollinator visits to a flower throughout its lifespan is very difficult to track.
 Often simple statistics are used to describe the flowering phenology of a population of plants, e.g.
@@
 We compare the results of automatic tracking with manual tracking.
-Individual flower tracking could facilitate finegrained studies on plant reproductive success.
-
-“and we show that performance achieved by following automatic tracking can be reliable enough to estimate X, Y, Z.”
+Individual flower tracking could facilitate finegrained studies on plant reproductive success....
+and we show that performance achieved by following automatic tracking can be reliable enough to estimate X, Y, Z...
 ---
@@
 
+```{r echo=FALSE}
+# Camera codes:
+# 2018 NARS-04: 2018 NARS A
+# 2018 THUL-01: 2018 THUL B
+# 2019 NARS-13: 2019 NARS C
+# 2019 NARS-17: 2019 NARS D
+# 2019 NYAA-04: 2019 NYAA E
+
+```
+
 \newpage
@@
 
+Phenology is often studied with first events as proxies for the full phenology distribution of the population .
+For flowering plants the time of first flowering is used to represent the phenology of a population of plants.
+The ideal metric for phenology depends on the topic of the question, but often, the choice of onset of flowering as the proxy for phenology is based on data availability and not because it is necessarily considered the ideal metric.
+First events, such as onset of flowering, may in some cases be particularly ill suited for estimating sensitivity to climate change [@miller2008].
+The date for first flowering represents the very extreme of the phenology distribution of a population and may be confounded with e.g.
+the size of the population considered.
+Further, for many questions the phenology distribution of a population is required, for example to quantify resource availability through the season.
+However, even detailed information on phenology at the level of a population is still an aggregation of the phenological distributions of the individual members of the population for which important variation may be masked [@inouye2019].
+Having at hand the full distribution of flowering phenology at the level of individuals allows to extract relevant characteristics of the distribution and test their sensitivity towards environmental factors, but acquiring such data requires efficient and preferably automatized methods.
+
+Pollen limitation is a common factor limiting the reproductive success of flowering plants, but the ecological mechanisms behind pollen limitation is poorly understood, at least partly because it requires information on plant reproductive success in relation to pollinator visitation rates which is difficult to acquire in a natural setting and to a satisfactory extent.
+For example, pollinator visits can vary greatly in frequency during the flowering season and the degree of pollen limitation can vary even between populations in close vicinity to each other [@larson1999; @ashman2004].
+A lot remains to be discovered about when and how the dynamics of pollination affects the reproductive success of flowering plants.
+High frequency observations of flower visits coupled with information on phenology and reproductive success at the level of individuals is required to unravel the relationship between pollinator visitation and plant reproductive success and the sensitivity of this interplay to climate.
+
 Monitoring flower phenology at high temporal resolution is time-consuming, particularly in logistically challenging environments such as the Arctic.
 Consequently, flowering phenology of a population is often characterized simply by the date for onset of flowering, typically derived from weekly observations of sample plots [@Prevey2021].
-Such inaccurate estimates may fail to reveal dynamics in flowering phenology for example caused by changes in climate.
-
+Such inaccurate estimates may fail to reveal important dynamics in flowering phenology through the season.
 Automatic image-based monitoring of flowering phenology can return phenology data for specific species at very high temporal resolution (Mann et al., in review), but phenological responses at the individual level may be indiscernible regardless of the temporal resolution of the data at population level.
@@
 As the flowers are constrained in movement by their stalk, their movement does not follow the assumption of linear motion.
-We therefore base the track assignment on the minimum distance between centroids of bounding boxes and introduce three user defined parameters that can help optimise tracking performance.
+Some tracking algorithms, such as DeepSORT [@deepsort2017], take the tracking-by-detection approach in which convolutional neural networks (CNNs) are applied to leverage appearance information in order to continuously distinguish individuals through time, thus aiding the tracking algorithm.
+In our case, however, flowers are not clearly distinguishable and one flower may look less alike between two frames than two flowers in a single frame.
 
-DeepSORT and other tracking frameworks leverage convolutional neural networks (CNNs) for object classification.
+As the data extracted from time-lapse series of flower plots through season violate the assumptions of standard tracking mechanism, we base the track assignment on the minimum distance between centroids of bounding boxes and introduce three user defined parameters that can help optimise tracking performance.
 
-But in our case, flowers are not distinguishable and one flower may look less alike between two frames than two flowers in a single frame.
-
-For complex scenes with many flowers in close vicinity to each other, we suggest a conservative filtering approach.
-The approach may remove correct tracks, but the tracks that remain will have a low risk of tracking errors.
-The approach allows for confident upscaling of the method.
-I.e., when applying the method on a large number of image series, it is preferable to extract individual high-confidence tracks from each series and ignore the remaining tracks.
+For complex scenes with many flowers in close vicinity to each other there is a sustantial risk of tracking errors regardless of the tracking method being used.
+In such cases it is preferable to extract individual high-confidence tracks from each series and ignore the remaining tracks.
+For this, we suggest a conservative filtering approach based on the density of tracks within the frame.
+The approach may remove correct tracks, but the tracks that remain will have a low risk of tracking errors, which allows for confident upscaling of the method to many series.
 
@@
 Further, any information of taxonomic grouping could facilitate comparative studies on pollination effectiveness between taxonomic groups of pollinators.
-We argue that the possibility of tracking phenology at the level of individuals will facilitate investigation of a novel set of research questions and is an important contribution to the shift towards efficient and automatic monitoring of ecological processes.
+The possibility of automatically tracking phenology at the level of individuals will facilitate investigation of a set of yet untestable research questions and is an important contribution to the shift towards efficient and detailed monitoring of ecological processes [@inouye2019].
 
@@
 #
-## Study site
-Toke, perhaps you can fill in some stuff here.
-NARS 2018, NARS 2019, NYAA 2019, THUL 2018.
+## Study site and species
+This study was based on five time-lapse image-series collected at the following three Arctic sites (sampling years given in brackets): Narsarsuaq, South Greenland (2018, 2019), Thule, North-West Greenland (2018), Ny-Ålesund, Svalbard (2019).
+The images were collected using time-lapse cameras (Moultrie Wingscapes TimelapseCam Pro, Moultrie Products, Birmingham Alabama, USA) permanently positioned approx.
+60 cm above areas with either Dryas integrifolia in Narsarsuaq and Thule or D.
+octopetala in Ny-Ålesund.
+The white flowers of these two species are insect-pollinated and held erect above cushions of foliage.
+The species are native to and widespread in Asian, North American, and European arctic and alpine regions.
+They are geographically separated apart from a possible hybrid zone in Northeast Greenland [@philipp2003].
 
 #
-## The image series
-Original time-lapse intervals, explanation and result of sampling scheme and subsequent temporal resolution.
+## The image series and flower annotations
+Images were collected throughout the full growing season.
+For this study, we limited the 1-hr interval image-series to the flowering season and randomly randomly sampled images from each series.
+The number of images for each series after sampling is given in table 1.
+All flowers in these images were manually annotated using the rectangular bounding box tool in the VIA VGG annotation software [@dutta2019] and each individual flower was assigned a unique ID.
+These annotations constitute our ground truth tracks.
+The series varied substantially in the number of flowers they contained (table 1).
+Figure 1 show annotated example images from three series (xxx) at different times during the flowering season (TO-DO).
+
+**Table 1:** Number of images and number of flowers for each of the five time-lapse image-series.
+#
+![](../figures/CameraOverview.png){ width=40% }
 
 #
-## Flower annotations
+![](../figures/SOmeFigure.png){ width=100% }
 
-`r word_comment("Add a figure showing example images with annotations.", "We")` manually annotated all flowers in the sampled image series using the rectangular bounding box tool in the VIA VGG annotation software [@dutta2019] and assigned each individual flower a unique ID.
-These annotations constitute our ground truth tracks.
+**Figure 1:** Figure text... Examples of annotated images (TO-DO)
 
 #
-## Automatic tracking
+## Automatic flower tracking
 
 We built a framework for tracking, filtering, and evaluating tracking of objects in time-lapse image series.
-
 The tracking algorithm tracks objects based on distances between centroids of bounding boxes.
@@
 Several issues complicate the task of tracking individual flowers through the time-lapse series.
-First, as the wind shifts, the flower heads changes direction.
-This can happen instantaneously (i.e., between two consecutive frames).
+First, monitoring the relatively slow process of flower development through full growing seasons (several months) requires low frame rates (several images per day).
+These frame rates capture the phenology of the flowers at very high detail, but fail to capture the full movement of flowers.
+For example, as the wind shifts, the flower heads changes direction, which in the image series happens instantaneously (i.e., between two consecutive frames).
 However, as they are constrained by their stalk, there is a limit to the distance they can move.
-Establishing associations between points based on just the distance between points in the current and the previous frame can cause errors when flowers are in close vicinity of each other.
-With the tracking parameter **running mean**, we base the tracking on the distance between a point in the current frame and the running mean of the positions of the previous X points in a track.
+With the tracking parameter **running mean**, we base the tracking on the distance between a point in the current frame and the running mean of the positions of a number of previous points in a track, taking advantage of the fact that flower move around a central point.
 
-Second, when the location of the flowers changes they may temporarily occlude the view of other flowers and  flowers close to the edge of the frame may move in and out of view.
+Second, as they move, flowers may temporarily occlude the view of other flowers and flowers close to the edge of the frame may move in and out of view.
 This can cause a track to be lost and a new track erroneously being initiated when the flower reappears.
@@
 
-The input data of
+The input data for the tracking algorithm are bounding box coordinates for each object in each frame given in the following format: frame number, x-min, y-min, x-max, y-max.
+Here, x-min, y-min is the top left corner of the bounding box and x-max, y-max is the bottom right corner.
+The bounding box centroid is calculated on the basis of these coordinates within the algorithm.
+Calculating tracking accuracy requires a ground truth ID for each bounding box which can be given in the sixth column.
 
-Calculating tracking accuracy requires a ground truth ID for each bounding box.
-
-This data structure is the same that is output from manual object annotations and object detection methods with CNNs.
+As this data structure is outputted from both manual object annotations and object detection methods with CNNs, the method can be applied on both manually and automatically derived data.
 
@@
 We identified the setting(s) that returned the lowest number of track mismatches and compare the tracking results between optimal settings and all parameters set to zero.
-
-`r word_comment("I haven't included these results.
-Not sure if they should be.", "Additionally,")` we performed a set of analyses on the five ground truth tracks.
-First, we calculated the largest distance between any two points within any track for any flower in each of the five series.
-Further, we calculated the maximum number of frames a flower track was lost and subsequently reappeared for each series.
 
@@
 First, the DBSCAN clustering algorithm [@ester1996] is run on the centroids with a conservatively high value for the eps parameter (350), meaning that tracks in close vicinity to each other will be clustered together.
+The value for eps was chosen by applying the filtering algorithm across the three series at range of eps values from x to x and a step size of x and identifying the value that returned the highest number of tracks across the three series at the lowest number of mismatches.
 
@@
 ## Tracking
+
+Fine-tuning and evaluating any tracking algorithm requires a ground dataset to compare tracking accuracy against.
 
@@
 
+The number of flowers produced by an individual is also a factor.
+
 For each series, we calculate the ratio of flowers for which the automatic tracking algorithm returns the correct flowering length compared to the ground truth tracks.
+
+`r word_comment("I haven't included these results.
+Not sure if they should be.", "Additionally,")` we performed a set of analyses on the five ground truth tracks.
+First, we calculated the largest distance between any two points within any track for any flower in each of the five series.
+Further, we calculated the maximum number of frames a flower track was lost and subsequently reappeared for each series.
 

### 2022-01-25 text/ManuscriptDraft.Rmd
Manuscript update

@@
 In our case, however, flowers are not clearly distinguishable and one flower may look less alike between two frames than two flowers in a single frame.
-
 As the data extracted from time-lapse series of flower plots through season violate the assumptions of standard tracking mechanism, we base the track assignment on the minimum distance between centroids of bounding boxes and introduce three user defined parameters that can help optimise tracking performance.
@@
 
-Several issues complicate the task of tracking individual flowers through the time-lapse series.
-First, monitoring the relatively slow process of flower development through full growing seasons (several months) requires low frame rates (several images per day).
+Several issues complicate the task of tracking individual flowers through time-lapse image-series.
+First, stand-alone monitoring of the relatively slow process of flower development through full growing seasons (several months) requires low frame rates (several images per day).
 These frame rates capture the phenology of the flowers at very high detail, but fail to capture the full movement of flowers.
@@
 The bounding box centroid is calculated on the basis of these coordinates within the algorithm.
-Calculating tracking accuracy requires a ground truth ID for each bounding box which can be given in the sixth column.
-
-As this data structure is outputted from both manual object annotations and object detection methods with CNNs, the method can be applied on both manually and automatically derived data.
+Calculating tracking accuracy requires a ground truth ID for each bounding box which can be given as the sixth column.
+As this data structure is outputted from both manual object annotations and object detection methods using CNNs, the method can be applied on both manually and automatically derived data.
+Further, as the algorithm handles frames sequentially, it can be applied both offline and online.
 
@@
 To explore the effect of the user parameters and to identify the optimal combination of parameters for our case of tracking flowers, we applied the tracking algorithm on each of the five image-series with every combination of a range of values for each parameter (3179 combinations): Values for max gap size and running mean were 0-160 with a step size of 10 and for max distance we used values 0-1000 with a step size of 100.
-Note that set to zero ignores the parameter altogether.
+Note that max distance set to zero ignores the parameter altogether.
 We identified the setting(s) that returned the lowest number of track mismatches and compare the tracking results between optimal settings and all parameters set to zero.
@@
 The optimal way of quantifying tracking performance depends on the goal of the tracking.
-To associate other information obtained in the images to the individual flower, for example flower visits, we want as much as possible of the track to be correct.
-To derive flowering length, in theory we just need to be able to connect the first and the last image of a flower and can ignore the intermediate points.
-Lastly, in some scenarios one may be interested in the number of flowers that existed in a plot, in which case we want the number of tracks obtained by automatic tracking to be as close as possible to the actual number of individuals in the series.
+To associate other information obtained in the images to the individual flower, for example flower visits, requires as much as possible of the track to be correct.
+To derive flowering length, in theory just requires correct association between the first and the last image of a flower while intermediate points can be ignored.
+Lastly, in some scenarios one may be interested in the number of flowers that existed in a plot, in which case the number of tracks obtained by automatic tracking should closely resemble the actual number of individuals in the series.
 
-The multiple object tracking accuracy (MOTA) score quantifies tracking performance based on counts of tracking mismatches [@bernardin2008].
+The commonly used multiple object tracking accuracy (MOTA) score quantifies tracking performance based on counts of tracking mismatches [@bernardin2008].
 Mismatches occur when objects swap track identity because they are in close vicinity to each other or when an object periodically disappears and is assigned a new track identity when it reappears.
@@
 These should ideally be equal.
+However, a correct number of tracks (i.e.
+number of tracks equal to number of objects) does not necessarily translate to correct tracking, since individual tracks can be erroneously split by the tracking algorithm and several tracks can be combined.
+We present the tracking performance by both the lowets number of mismatches/highest MOTA given overall and for combinations that returned a number opf tracks equal to the number of flowers in the series.
 
@@
 The value for eps was chosen by applying the filtering algorithm across the three series at range of eps values from x to x and a step size of x and identifying the value that returned the highest number of tracks across the three series at the lowest number of mismatches.
-
-We identify a good value for eps by...
 

### 2022-01-25 text/ManuscriptDraft.Rmd
Manuscript update

@@
 For example, pollinator visits can vary greatly in frequency during the flowering season and the degree of pollen limitation can vary even between populations in close vicinity to each other [@larson1999; @ashman2004].
+
+Interactions require abundance of individuals in life-history stages in which interactions occur.
+Thus, interaction strength depends on overlap in phenology between individuals in these stages.
+Ideally, investigation of the sensitivity of interaction strength to dynamics in phenology requires quantification of phenology through the season for the species in question in order to asses both their phenology and abundance [@inouye2019; @linden2018].
+
+If interacting and interdependent species, e.g.
+plants and pollinators, react to different environmental cues, or react differently to the same cues, their phenology may be desynchronized and temporal mismatches occur.
+Changes in flower phenology as a consequence of climate change may indirectly affect the abundance of flower-dependent species [@hoye2013] and these effects can be stronger than the direct effects of climate change [@ogilvie2017].
+However, the relation between flower phenology and reproductive success contain many facets and is not simply a question of temporal overlap with pollinators.
+For example, @bolmgren2015 found that for an insect-pollinated plant species early flowering was associated with increased fruit-set but late flowering was associated with higher seed mass.
+Further, periods of resource scarcity through the season may be more important than the floral sum of the season, i.e.
+dynamics in resource availability above a level of sufficiency may have a weak effect on pollinator abundance while the time above this threshold has a stronger effect [@ogilvie2017].
+
+Different phenological factors, such as onset, peak, and end of flowering may respond differently to climate change [kilder fra første manuskript] and estimation of the degree of mismatch between species may thus be sensitive to the choice of metric [@inouye2019].
+Efficient and precise methods for deriving full seasonal distributions of phenology at the level of individuals is much needed but usually prevented for logistical reasons.
+
 A lot remains to be discovered about when and how the dynamics of pollination affects the reproductive success of flowering plants.
@@
 We present and evaluate an automatic flower tracking and filtering algorithm.
-Many tracking algorithms, such as the Simple online and real-time tracking (SORT) [@sort], use Kalman filtering to predict the future location of an object by estimating its velocity [@kalmann].
-However, Kalman filtering assumes linear motion of the object and does not handle abrupt motion well.
-As the flowers are constrained in movement by their stalk, their movement does not follow the assumption of linear motion.
-Some tracking algorithms, such as DeepSORT [@deepsort2017], take the tracking-by-detection approach in which convolutional neural networks (CNNs) are applied to leverage appearance information in order to continuously distinguish individuals through time, thus aiding the tracking algorithm.
-In our case, however, flowers are not clearly distinguishable and one flower may look less alike between two frames than two flowers in a single frame.
 As the data extracted from time-lapse series of flower plots through season violate the assumptions of standard tracking mechanism, we base the track assignment on the minimum distance between centroids of bounding boxes and introduce three user defined parameters that can help optimise tracking performance.
 
-For complex scenes with many flowers in close vicinity to each other there is a sustantial risk of tracking errors regardless of the tracking method being used.
+For complex scenes with many flowers in close vicinity to each other there is a substantial risk of tracking errors regardless of the tracking method being used.
 In such cases it is preferable to extract individual high-confidence tracks from each series and ignore the remaining tracks.
@@
 Further, any information of taxonomic grouping could facilitate comparative studies on pollination effectiveness between taxonomic groups of pollinators.
-The possibility of automatically tracking phenology at the level of individuals will facilitate investigation of a set of yet untestable research questions and is an important contribution to the shift towards efficient and detailed monitoring of ecological processes [@inouye2019].
+The possibility of automatically tracking phenology at the level of individuals will facilitate investigation of a set of yet untestable research questions and is an important contribution to the shift towards efficient and detailed monitoring of ecological processes.
 
@@
 **Table 1:** Number of images and number of flowers for each of the five time-lapse image-series.
+
 #
@@
 
+Many tracking algorithms, such as the Simple online and real-time tracking (SORT) [@sort], use Kalman filtering to predict the future location of an object by estimating its velocity [@kalmann].
+However, Kalman filtering assumes linear motion of the object and does not handle abrupt motion well.
+As the flowers are constrained in movement by their stalk, their movement does not follow the assumption of linear motion.
+Some tracking algorithms, such as DeepSORT [@deepsort2017], take the tracking-by-detection approach in which convolutional neural networks (CNNs) are applied to leverage appearance information in order to continuously distinguish individuals through time, thus aiding the tracking algorithm.
+In our case, however, flowers are not clearly distinguishable and one flower may look less alike between two frames than two flowers in a single frame.
+
 Fine-tuning and evaluating any tracking algorithm requires a ground dataset to compare tracking accuracy against.

### 2022-01-26 text/ManuscriptDraft.Rmd
Manuscript update

@@
 Changes in flower phenology as a consequence of climate change may indirectly affect the abundance of flower-dependent species [@hoye2013] and these effects can be stronger than the direct effects of climate change [@ogilvie2017].
-However, the relation between flower phenology and reproductive success contain many facets and is not simply a question of temporal overlap with pollinators.
+However, the relation between flower phenology and reproductive success is multifaceted and not simply a question of temporal overlap with pollinators.
 For example, @bolmgren2015 found that for an insect-pollinated plant species early flowering was associated with increased fruit-set but late flowering was associated with higher seed mass.
@@
 
-Different phenological factors, such as onset, peak, and end of flowering may respond differently to climate change [kilder fra første manuskript] and estimation of the degree of mismatch between species may thus be sensitive to the choice of metric [@inouye2019].
-Efficient and precise methods for deriving full seasonal distributions of phenology at the level of individuals is much needed but usually prevented for logistical reasons.
-
-A lot remains to be discovered about when and how the dynamics of pollination affects the reproductive success of flowering plants.
-High frequency observations of flower visits coupled with information on phenology and reproductive success at the level of individuals is required to unravel the relationship between pollinator visitation and plant reproductive success and the sensitivity of this interplay to climate.
+Different phenological factors, such as onset, peak, and end of flowering may respond differently to climate change [@cook2012] and estimation of the degree of mismatch between species may thus be sensitive to the choice of metric [@inouye2019].
+Efficient and precise methods for deriving full seasonal distributions of phenology at the level of individuals is much needed but so far most often unfeasible for logistical reasons.
 
@@
 Similarly so for investigating whether flower visitation rates and/or reproductive success depends on the timing of flowering for the individual flower.
+
+A lot remains to be discovered about when and how the dynamics of pollination affects the reproductive success of flowering plants.
+High frequency observations of flower visits coupled with information on phenology and reproductive success at the level of individuals is required to unravel the relationship between pollinator visitation and plant reproductive success and the sensitivity of this interplay to climate.
 
@@
 Images were collected throughout the full growing season.
-For this study, we limited the 1-hr interval image-series to the flowering season and randomly randomly sampled images from each series.
+While the original time-lapse frequency was higher, the basis for this study was 1-hr interval sereis.
+We limited the 1-hr interval image-series to the flowering season and randomly randomly sampled images from each series.
 The number of images for each series after sampling is given in table 1.
-All flowers in these images were manually annotated using the rectangular bounding box tool in the VIA VGG annotation software [@dutta2019] and each individual flower was assigned a unique ID.
+All flowers in each of these images were manually annotated using the rectangular bounding box tool in the VIA VGG annotation software [@dutta2019] and each individual flower was assigned a unique ID.
 These annotations constitute our ground truth tracks.
@@
 #
+### Input data for the tracking algorithm
+
+The input data for the tracking algorithm are bounding box coordinates for each object in each frame given in the following format: frame number, x-min, y-min, x-max, y-max.
+Here, x-min, y-min is the top left corner of the bounding box and x-max, y-max is the bottom right corner.
+The bounding box centroid is calculated on the basis of these coordinates within the algorithm.
+Calculating tracking accuracy requires a ground truth ID for each bounding box which can be given as the sixth column.
+As this data structure is outputted from both manual object annotations and object detection methods using CNNs, the method can be applied on both manually and automatically derived data.
+Further, as the algorithm handles frames sequentially, it can be applied both offline and online.
+
+#
 ## Automatic flower tracking
@@
 Several issues complicate the task of tracking individual flowers through time-lapse image-series.
-First, stand-alone monitoring of the relatively slow process of flower development through full growing seasons (several months) requires low frame rates (several images per day).
+
+First, stand-alone monitoring of the relatively slow process of flower development through full growing seasons (several months) requires only low frame rates.
+Here we used several images per day.
 These frame rates capture the phenology of the flowers at very high detail, but fail to capture the full movement of flowers.
-For example, as the wind shifts, the flower heads changes direction, which in the image series happens instantaneously (i.e., between two consecutive frames).
+For example, as the wind shifts, the flower heads change direction, which in the image series happens instantaneously (i.e.
+between two consecutive frames).
 However, as they are constrained by their stalk, there is a limit to the distance they can move.
-With the tracking parameter **running mean**, we base the tracking on the distance between a point in the current frame and the running mean of the positions of a number of previous points in a track, taking advantage of the fact that flower move around a central point.
+With the tracking parameter **running mean**, we have the option to base the tracking on the distance between a point in the current frame and the running mean of the positions of a number of previous points in a track, taking advantage of the fact that flower move around a central point.
 
-Second, as they move, flowers may temporarily occlude the view of other flowers and flowers close to the edge of the frame may move in and out of view.
+Second, as flowers move, they may temporarily occlude the view of other flowers and flowers close to the edge of the frame may move in and out of view.
 This can cause a track to be lost and a new track erroneously being initiated when the flower reappears.
@@
 The parameter **max gap size** sets the number of frames a track can be lost before a new track is initiated for points appearing in the same area.
+If a given flower has not been detected in a few frames, the track will not be lost.
 Concurrently, this deals with potential false negatives.
-If a given flower has not been detected in a few frames, the track will not be lost.
 The counter for number of disappeared frames for a track is reset when a new point is associated with the track within the threshold.
@@
 As flowers are constrained in the movement, we can assume a maximum distance that will occur between points and force the initiation of tracks for points that exceed this threshold.
-We do that with the parameter **max distance**.
+We set this threshold with the parameter **max distance**.
 
@@
 #
-### Input data for the tracking algorithm
+### Identifying optimal user parameters
 
-The input data for the tracking algorithm are bounding box coordinates for each object in each frame given in the following format: frame number, x-min, y-min, x-max, y-max.
-Here, x-min, y-min is the top left corner of the bounding box and x-max, y-max is the bottom right corner.
-The bounding box centroid is calculated on the basis of these coordinates within the algorithm.
-Calculating tracking accuracy requires a ground truth ID for each bounding box which can be given as the sixth column.
-As this data structure is outputted from both manual object annotations and object detection methods using CNNs, the method can be applied on both manually and automatically derived data.
-Further, as the algorithm handles frames sequentially, it can be applied both offline and online.
+To explore the effect of the user parameters and to identify the optimal combination of parameters for our case of tracking flowers, we applied the tracking algorithm on each of the five image-series with every combination of a range of values for each parameter (3179 combinations): Values for max gap size and running mean were 0-160 with a step size of 10 and for max distance 0-1000 with a step size of 100.
+Note that max distance set to zero ignores the parameter altogether.
+We identified the setting(s) that returned the lowest number of track mismatches and compare the tracking results between optimal settings and all parameters set to zero.
 
 #
-### Identifying optimal user parameters
+## Filtering tracks
 
-To explore the effect of the user parameters and to identify the optimal combination of parameters for our case of tracking flowers, we applied the tracking algorithm on each of the five image-series with every combination of a range of values for each parameter (3179 combinations): Values for max gap size and running mean were 0-160 with a step size of 10 and for max distance we used values 0-1000 with a step size of 100.
-Note that max distance set to zero ignores the parameter altogether.
-We identified the setting(s) that returned the lowest number of track mismatches and compare the tracking results between optimal settings and all parameters set to zero.
+When deploying the automatic tracking algorithm on naive data without ground truth tracks, it is not possible to manually filter for correct tracks.
+Therefore, we present a conservative filtering method that extracts the most trustworthy tracks from a scene.
+
+We base our filtering on a density-based clustering of the centroid of a geometry of each track.
+The geometries are derived in the following way: For tracks consisting of only a single point, the coordinates of the point are used as the centroid.
+For tracks consisting of two points, we establish the straight line between the points and calculate the centroid of the line.
+For tracks consisting of three points, we establish the triangle from the points and calculate the centroid of the triangle.
+For tracks consisting of four points or more, we calculate the convex hull of all the points included in the track and derive the polygon from the vertices of the convex hull.
+Hereafter we calculate the centroid of the polygon.
+For tracks that contain more than two points but where the points a collinear, we establish the line through the points and calculate the centroid of the line.
+
+The filtering is done on the geometry centroids to remove tracks in areas with a high density of tracks as these have a high risk of tracking mismatches.
+The filtering is done in two steps.
+First, the DBSCAN clustering algorithm [@ester1996] is run on the centroids with a conservatively high value for the eps parameter, meaning that tracks in close vicinity to each other will be clustered together.
+Second, all tracks that were not assigned to a unique cluster are removed.
+Tracks that are spatially isolated remains.
+
+We demonstrate the filtering method on the three series for which our tracking algorithm did not return perfect results.
+For this, we chose a fixed combination of the tracking algorithm parameters across all three series (max gap size: 10; running mean: 10; max distance: 300).
+
+We then applied the filtering method on the three series with a range of values for eps (x-xx, step size of 10) and identified the value that returned the highest sum of tracks across the three series at the lowest number of mismatches.
+We present the tracking accuracy before filtering and on the remaining tracks after filtering.
+We underline that the results presented are thus highly conservative.
+I.e., the approach could be optimised to extract a higher number of tracks while maintaining the number of mismatches.
+For example, basing the filtering on the best performing tracking parameters for each series specifically would improve the output of the tracking algorithm.
+Similarly, the results could be improved by fine-tuning the value for eps for each series individually.
+However, as our goal here was to demonstrate that our tracking and filtering method can be applied in a naive setting with fixed values for each parameter we have not included those results.
 
@@
 number of tracks equal to number of objects) does not necessarily translate to correct tracking, since individual tracks can be erroneously split by the tracking algorithm and several tracks can be combined.
-We present the tracking performance by both the lowets number of mismatches/highest MOTA given overall and for combinations that returned a number opf tracks equal to the number of flowers in the series.
-
-#
-## Filtering tracks
-
-When deploying the automatic tracking algorithm on naive data without ground truth tracks, it is not possible to manually filter for correct tracks.
-Therefore, we present a conservative filtering method that extracts the most trustworthy tracks from a scene.
-
-We base our filtering on a density-based clustering of the centroid of a geometry of each track.
-The geometries are derived in the following way: For tracks consisting of only a single point, the coordinates of the point are used as the centroid.
-For tracks consisting of two points, we establish the straight line between the points and calculate the centroid of the line.
-For tracks consisting of three points, we establish the triangle from the points and calculate the centroid of the triangle.
-For tracks consisting of three points or more, we calculate the convex hull of all the points included in the track and derive the polygon from the vertices of the convex hull and calculate the centroid of this polygon.
-For tracks that contain more than two points but where the points a colinear, we establish the line through the points and calculate the centroid of the line.
-
-We then apply the filtering based on the geometry centroids to remove tracks in areas with a high density of tracks as these have a high risk of tracking mismatches.
-The filtering is done in two steps.
-First, the DBSCAN clustering algorithm [@ester1996] is run on the centroids with a conservatively high value for the eps parameter (350), meaning that tracks in close vicinity to each other will be clustered together.
-The value for eps was chosen by applying the filtering algorithm across the three series at range of eps values from x to x and a step size of x and identifying the value that returned the highest number of tracks across the three series at the lowest number of mismatches.
-
-Second, all tracks that were not assigned to a unique cluster are removed.
-Tracks that are spatially isolated remains.
-We evaluate the tracking accuracy of the remaining tracks.
-We demonstrate our filtering approach for the image series for which our tracking algorithm did not produce perfect results.
+We present the tracking performance by both the lowest number of mismatches/highest MOTA given overall and for combinations that returned a number of tracks equal to the number of flowers in the series.
 
@@
 We present the tracking performance without the use of the three tracking parameters along with the best performance (lowest number of mismatches/highest MOTA), and the best performance where the number of tracks correspond to the number of individual flowers in the series.
+When several values for a parameter was included in a result, we present the range of values.
+For two series, 2018 THUL B and 2019 NARS D, the tracking algorithm returned perfect results for several combinations of parameters.
+For the three remaining series, the algorithm returned consistently high MOTA scores (above 0.95).
+In all cases, except 2019 NARS D, where the tracking algorithm returned perfect results with all parameters zeroed, did the inclusion of the user set tracking parameters improve the tracking accuracy.
 
-**Table 1:** Table description...
+In one case, NARS-13, four different numbers of tracks were returned among the parameter combinations that resulted in the lowest number of mismatches.
+A number of tracks equal to the number of flowers in the series were among these four.
+
+In one case, NYAA-04, the parameter combination that produced the highest accuracy did not return a number of tracks equal to the number of flowers in the series.
+However, combinations that did produce this number of tracks also returned a higher number of mismatches.
+
+**Table 1:** Table description... To-do: Rename series to camera codes.
+
 #
@@
 4.
-Our filtering method successfully extracted 28 tracked flowers with only a single mismatch from the three series with complex scenes using a fixed value of 350 for eps and fixed values for the tracking parameters (10, 10, 300, for running mean, max gap size and max distance, respectively).
+Our filtering method extracted 28 tracked flowers from the three series with complex scenes using a fixed value of 350 for eps and fixed values for the tracking parameters (10, 10, 300, for running mean, max gap size, and max distance, respectively) with only a single mismatch included in the 28 tracks.
 
-**Table 2:** Results of the filtering algorithm with an eps value of 350 applied across all three .
-`r word_comment("Note to self.
-Table shows to what degree a tr track contains points from a single gt track, but does not show whether all points of the gt track are included in the tr track.
-We know that is the case, but should it be tested/shown?", "Table 1")` series.
+**Table 2:** Results of the filtering algorithm with an eps value of 350 applied across all three series To-do: Change image series to camera codes.
 
@@
 As the flowers are constrained in movement by their stalk, their movement does not follow the assumption of linear motion.
-Some tracking algorithms, such as DeepSORT [@deepsort2017], take the tracking-by-detection approach in which convolutional neural networks (CNNs) are applied to leverage appearance information in order to continuously distinguish individuals through time, thus aiding the tracking algorithm.
+Some tracking algorithms, such as DeepSORT [@deepsort2017], take the tracking-by-detection approach in which convolutional neural networks are applied to leverage appearance information in order to continuously distinguish individuals through time, thus aiding the tracking algorithm.
 In our case, however, flowers are not clearly distinguishable and one flower may look less alike between two frames than two flowers in a single frame.
 
-Fine-tuning and evaluating any tracking algorithm requires a ground dataset to compare tracking accuracy against.
+Finally, some tracking methods require manual initiation of a fixed number of tracks for objects in the first frame and the algorithm will track these objects in the subsequent frames [@hu2012].
+These methods does not allow new objects to enter the frame through the series.
+In our case, objects can disappear permanently from the frame, for example when flowers wilt, in which case the track should be terminated.
+Oppositely, new objects can continuously appear through the series when new flowers bloom of previously unseen flowers enter the frame.
+Additionally, objects may temporarily disappear from view and may or may not reappear at some later stage.
+These settings vastly complicate the task of tracking multiple objects.
+
+Our tracking algorithm returned consistently accuracy despite the difficult settings.
 
@@
 Our steps in parameter values were crude, however, and it is very likely that finer steps in these values would identify combinations that produce even better results.
-Here we do not perform this analysis, however, as our goal is to show that the parameters can be used for optimizing tracking performance in general.
-
-We have demonstrated that our tracking algorithm with the three tracking parameters and the subsequent filtering algorithm can substantially increase accuracy of the tracking of flowers despite the fact that they break the assumptions of general tracking methods.
+Here we do not perform this analysis, however, as our goal is to show that the parameters can be used for optimizing tracking performance in general
 
 In a naive setting, a priori knowledge about which settings to use is not available.
-
-As with any tracking method, accuracy can only be assessed based on visual ... or more ... based on a subsample of tracks for whih a ground truth exists.
-
-We recommend to perform a parameter test as demonstrated here on subset of ground truth tracks to derive optimal parameters for application of the method in a new setting.
-
-Depending on the nature of the objects being tracked and the complexity of the scene, the user parameters can be estimated from visual examination of the tracking results.
-Often it may be preferable to manually annotate a subset of the objects in the image series and derive a set of user parameters from these results.
+As with any tracking method, accuracy can only be assessed based on visual observation or more thoroughly by testing on subsample of tracks for which a ground truth exists.
+When applying the method in a new setting, we recommend performing a parameter test as demonstrated here on a subset of ground truth tracks to derive the optimal set of parameters.
 
@@
 I.e., the tracking algorithm could split one ground truth track in two, but combine two other ground truth tracks into one, which would even each other out.
+This was the case for NYAA-04 for which the parameter combination that returned a number of tracks equal to the number of flowers in the series was not the combination the produced the highest MOTA score.
 
-Our tracking algorithm consistently returns high MOTA scores.
+The two series for which the tracking algorithm produced perfect results were also the series with the lowest number of flowers.
+As the number of flowers increase so does the complexity of the task of tracking.
+A higher number of flowers within the frame increases the possibility of flowers in close vicinity to each other and of flowers overlapping and moving in and out of frame translating to a higher risk of making tracking errors.
+The very high complexity in the scenes with many flowers, especially NARS-13 with 85 flowers, renders it unlikely that any tracking method could produce perfect results.
+Although this setting can be considered an extreme case for our application, it underlines the relevance of a pipeline including both tracking and filtering.
 
-For complex scenes, all three parameters make a difference.
+#
+## Filtering
 
-The parameters are interdependent.
+We presented a filtering method based on the density of tracks within the frame which can be applied subsequent of tracking.
+Extracting tracks that are spatially isolated does not guarantee that the tracks are correct/without errors.
+However, as spatially isolated objects are easier to track, it increases confidence in the remaining tracks.
 
-In cases where we are tracking perfectly with maxDisap = 0, setting it any value will not make a difference. Not quite right. Explore more...
+The filtering algorithm successfully extracted correctly tracked flowers from each of the three series for which the tracking algorithm did not produce perfect results.
+Only a single mismatch was included in the 28 flowers that passed the filtering.
+
+We applied the filtering algorithm on tracks produced with a fixed set of tracking parameters as well as a fixed value for the DBSCAN eps parameter for the filtering algorithm.
+We note that the results could be improved (i.e.
+more flowers extracted without increased number of mismatches) by chosing the best tracking result for each series and fine-tuning the eps value for each individual series.
+However, as our goal here was to show that a single conservative value can be applied across series for which the tracking algorithm did not necessarily produce the best results, we do not show those results here.
+As in the case of the tracking parameters, we recommend setting the value for eps based on either visual examination of results or on tests on a subset of ground truth data when applying the method in a new setting.
+
+We note that our method for filtering tracks using DBSCAN on track centroids ensures that all tracks are given the same weighting in the filtering since each track is represented by a single point.
+In some cases this may be suboptimal.
+For example, if it is given a priori that an object will always appear in a minimum of two frames, then single point tracks can be filtered out.
+However, when such a priori knowledge is not accessible, a conservative approach as the one we present is preferable.
+
+## The tracking and filtering parameters
+
+Our method relies on a number of parameters set by the user.
+
+For the three series with complex scenes, applying the max gap size parameter improved tracking accuracy with values of 10-20 being optimal.
+These series all contain flowers at the edge of the frame and the higher number of flowers increases the risk of flowers temporarily occluding others.
+Correct tracking of these objects requires that an object can disappear from view without the track instantly being lost, which the max gap size allows.
+Oppositely, a higher number of flowers also increase the possibility of a new flower blooming in close vicinity to where a flower previously existed and setting the value too high may force objects such objects that are only separated in time but occur in the same space into the same track.
+
+In all cases, except NARS-17, for which perfect tracking was achieved with all parameters set to zero, did a value for running mean above 0 improve results.
+
+Interestingly, setting the value to 10 produced the best results in all these cases, except for the combination returning the lowest number of mismatches for NYAA-04.
+
+Increasing the value for running mean means that less weight is put on movements between each frame.
+
+Maybe NYAA-04 flowers move more? Maybe longer stalk?
+
+Coincidently, the flowers in NYAA-04 also flower longer on average than the other series.
+
+If the average euclidean distance that a flower moves is higher for NYAA-04 it may improve results to even out this movement by increasing the value for running mean.
+
+Look at this tomorrow....
+
+These parameters are not independent.
+For example, the distance between a point in the current frame and a point in the previous frame may not be the same as that for a point in the current frame and the mean location of the previous ten points in a track.
+This underlines the relevance of search for an optimal combination based on ground truth data.
+
+Pointe: Selvom vi kigger på mange komvinationer af parameterne, så falder kombinationerne hvor de bedste resultater produceres (laveste mismatches) indenfor forholdsvis smalle områder.
+Det tyder på at metoden med de givne parametre er generaliserbar og kan anvendes på andre kameraer med samme setting.
+
+For our application, we find that a
+
+For the three series for which the tracking algorithm failed to produce perfect results, the range
+
+Why do they make a difference:
+
+max gap size:
+
+Running mean:
+
+Max distance:
+
+eps:
+
+Offline and online tracking differ
+
+Online tracking
+
+Our tracking algorithm does not rely on information from future frames and can thus be applied both offline and online.
+
+MOT can also be categorized into online tracking and offline
+tracking.
+The difference is whether or not observations from future frames are utilized when handling the current frame.
+Online, also called causal, tracking methods only rely on the past information available up to the current frame, while
+
+It also depends on the time the flowers are spread over. Many flowers and short time is a hard problem.
 
@@
 #
-## Filtering
-
-Extracting tracks that are spatially isolated does not guarantee that the tracks are correct/without errors.
-However, as spatially isolated objects are easier to track, it increases confidence in the remaining tracks.
-
-**Alekos: I think it would be good to add a short description or comment on what each of the mentioned parameters is used for (e.g.
-why did you need to change its value, how this affects the performance of the method on the challenges in your dataset compared to standard computer vision tracking datasets).**
-
-We applied a single value for the DBSCAN eps parameter in our tracking algorithm.
-We note that this value could be fine-tuned for improved results for the individual series (i.e.
-more flowers extracted without increased number of mismatches).
-However, as our goal here was to show that a single conservative value can be applied across series, we do not show those results here.
-In a naive setting, a general value could be chosen or the value could be adjusted for each series based on visual examination of performance or testing on a subset of data.
-
-Our method for filtering tracks using DBSCAN on track centroids ensures that all tracks are given the same weighting in the filtering since each track is represented by a single point. In some cases 
-e.g.
-if it is given a priori that an object will always appear in a minimum of two frames, then single point tracks can be filtered out.
-However, when such a priori knowledge is not accessible, a conservative approach as the one we present is preferable.
-
-#
 ## The image series
@@
 A paragraph about the ecological perspectives of being able to track the individual flowers (at scale).
+
+We have demonstrated that our tracking algorithm with the three tracking parameters and the subsequent filtering algorithm can substantially increase accuracy of the tracking of flowers despite the fact that they break the assumptions of general tracking methods.
 

### 2022-01-26 code/analyze.py

@@ -43,22 +43,22 @@ df_grouped = df.groupby(['id_gt'])
 
 largestDist = 0
 
-for i, g in df_grouped:
-    #print(g)
-    g_points = []
-    for i,l in g.iterrows():
-        g_points.append(np.array([l['x_c'], l['y_c']]))
+# for i, g in df_grouped:
+#     #print(g)
+#     g_points = []
+#     for i,l in g.iterrows():
+#         g_points.append(np.array([l['x_c'], l['y_c']]))
         
-    for p1 in g_points:
-        for p2 in g_points:
-            d = dist(p1, p2)
-            if d > largestDist:
-                print(f'{d} is larger than {largestDist}. Storing {d}')
-                print(f'Distance calculated between {p1} and {p2}')
-                largestDist = d
+#     for p1 in g_points:
+#         for p2 in g_points:
+#             d = dist(p1, p2)
+#             if d > largestDist:
+#                 print(f'{d} is larger than {largestDist}. Storing {d}')
+#                 print(f'Distance calculated between {p1} and {p2}')
+#                 largestDist = d
             
             
-print(f'Largest dist found: {largestDist}')
+# print(f'Largest dist found: {largestDist}')
 
 # THUL-01: 449.00334074481003
 # NYAA-04: 632.9178461696273
@@ -70,37 +70,89 @@ print(f'Largest dist found: {largestDist}')
 # Largest distance between mean of track and each point
 largestDist = 0
 
-for i, g in df_grouped:
-    #print(g)
-    x_cs = []
-    y_cs = []
-    g_points = []
-    for i,l in g.iterrows():
-        x_cs.append(l['x_c'])
-        y_cs.append(l['y_c'])
-        g_points.append(np.array([l['x_c'], l['y_c']]))
+# for i, g in df_grouped:
+#     #print(g)
+#     x_cs = []
+#     y_cs = []
+#     g_points = []
+#     for i,l in g.iterrows():
+#         x_cs.append(l['x_c'])
+#         y_cs.append(l['y_c'])
+#         g_points.append(np.array([l['x_c'], l['y_c']]))
 
-    mx = ave(x_cs)
-    my = ave(y_cs)
+#     mx = ave(x_cs)
+#     my = ave(y_cs)
     
-    m = np.array([mx, my]) # Mean of x and mean of y
-    for p1 in g_points:
-        d = dist(p1, m)
-        if d > largestDist:
[…]

### 2022-01-27 code/R/Analysis.Rmd
Added some analyses of ground truth tracks

@@
 ggplot()+
-  geom_point(aes(x = flower_start, y = flower_length))
+  geom_point(aes(x = flower_start, y = flower_length, colour = Class))
 
@@
 
+nyaa04xx = read.csv("U:/BITCue/Projekter/TrackingFlowers/testResults/_parameterTest_NYAA-04_3/parameterTest_NYAA-04_maxDisap_10_runMean_10_maxDist_300.csv")
+
+nyaa04xx %>% 
+  ggplot() +
+  geom_point(aes(x = x_c, y = y_c, color = as.factor(objectID)), size = 1) + 
+  xlim(0,6080) +
+  ylim(0,3420) + 
+  ggtitle("NYAA-04 Tracked (xx)")  +
+  theme(legend.position = "none")
+
 nyaa04 %>% 
@@
   theme(legend.position = "none")
+
+#nyaa04_meta %>% 
+#  ggplot() +
+#  geom_point(aes(x = DOY, y = Class))
+
 ```
@@
 ```
+Largest gap in time
+
+```{r}
+x = nyaa04_meta %>%
+    group_by(Class) %>%
+    arrange(time) %>%
+    mutate(diff = time - lag(time, default = first(time)))
+
+x = nars13_meta %>%
+    group_by(Class) %>%
+    arrange(time) %>%
+    mutate(diff = time - lag(time, default = first(time)))
+
+# NYAA-04: 5.1
+# NARS-13: 1.25
+
+```
+
+Largest gap in number of frames
+
+```{r}
+
+x = nyaa04 %>% 
+  mutate(frame = as.integer(str_remove(str_extract(FileName, regex("\\d{6}")), "^0+"))) %>% 
+  group_by(Class) %>%
+  arrange(frame) %>%
+  mutate(diff = frame - lag(frame, default = first(frame)))
+
+```

### 2022-01-27 code/analyze.py

@@ -15,7 +15,7 @@ import pandas as pd
 import numpy as np
 
 # Input ground truth annotations
-df = pd.read_csv(r"../data\annotations\2020_04_30_NorwayAnnotations_NARS-13_IndividualAnnotations_FRCNN_Metrics.csv")
+df = pd.read_csv(r"../data\annotations\2021_12_13_NorwayAnnotations_NYAA-04_IndividualAnnotations_FRCNN_Metrics.csv")
 
 
 #print(thul01)
@@ -24,11 +24,6 @@ df['x_c'] = (df['x_min'] + df['x_max']) / 2
 df['y_c'] = (df['y_min'] + df['y_max']) / 2
 
 
-#print(thul01)
-
-
-a = np.array([3.4,4])
-b = np.array([6,9])
 
 def dist(a,b):
     d = np.linalg.norm(a-b)
@@ -106,20 +101,20 @@ largestDist = 0
 
 # Mean distance per flower
 
-meand = []
-for i, g in df_grouped:
-    #print(g)
-    g_points = []
-    dlist = []
-    for i,l in g.iterrows():
-        g_points.append(np.array([l['x_c'], l['y_c']]))
+# meand = []
+# for i, g in df_grouped:
+#     #print(g)
+#     g_points = []
+#     dlist = []
+#     for i,l in g.iterrows():
+#         g_points.append(np.array([l['x_c'], l['y_c']]))
         
-    for p1 in g_points:
-        for p2 in g_points:
-            d = dist(p1, p2)
-            dlist.append(d)
-    meand.append(sum(dlist))
-print(ave(meand))
+#     for p1 in g_points:
+#         for p2 in g_points:
+#             d = dist(p1, p2)
+#             dlist.append(d)
+#     meand.append(sum(dlist))
+# print(ave(meand))
 
 # THUL-01:
 # NYAA-04: 897228.1250040481
@@ -129,29 +124,136 @@ print(ave(meand))
 
     
 
-meand = []
-for i, g in df_grouped:
-    #print(g)
-    g_points = []
-    dlist = []
-    for i,l in g.iterrows():
-        g_points.append(np.array([l['x_c'], l['y_c']]))
+# meand = []
+# for i, g in df_grouped:
+#     #print(g)
+#     g_points = []
+#     dlist = []
+#     for i,l in g.iterrows():
+#         g_points.append(np.array([l['x_c'], l['y_c']]))
         
-    for p1 in g_points:
-        for p2 in g_points:
-            d = dist(p1, p2)
[…]

### 2022-01-27 text/ManuscriptDraft.Rmd
Manuscript update

@@
 Interestingly, setting the value to 10 produced the best results in all these cases, except for the combination returning the lowest number of mismatches for NYAA-04.
+Increasing the value for running mean means that less weight is put on movements between each frame.
+Changing the running mean to 10 while keeping the other two parameters equal only resulted in a slight increase in number of `r word_comment("This result is not given - should results for every combination be given in suppl.?
+It's 3179*5 rows of data, so it's quite a lot.)", "mismatches")` (from 29 to 36).
+Thus, it is a likely explanation that the improved accuracy from the large value for running mean stems from slightly better tracking of flowers that moved a lot.
 
-Increasing the value for running mean means that less weight is put on movements between each frame.
+For the three series for which the tracking algorithm did not produce perfect results, the optimal range for max distance were relatively narrow (300 - 500).
+The optimal value for this setting is one that is sufficiently high so that it does not force individual flowers into separate tracks because of movement between frames yet is low enough to separate any flowers into individual tracks that would otherwise have been grouped together.
+This explains well why the extreme values for this parameter was not included in the highest accuracy tracks.
 
-Maybe NYAA-04 flowers move more? Maybe longer stalk?
-
-Coincidently, the flowers in NYAA-04 also flower longer on average than the other series.
-
-If the average euclidean distance that a flower moves is higher for NYAA-04 it may improve results to even out this movement by increasing the value for running mean.
-
-Look at this tomorrow....
-
-These parameters are not independent.
+It is important to note that the tracking parameters are not independent.
 For example, the distance between a point in the current frame and a point in the previous frame may not be the same as that for a point in the current frame and the mean location of the previous ten points in a track.
+Thus, changing the setting for running mean could influence the optimal setting for max distance.
 This underlines the relevance of search for an optimal combination based on ground truth data.
 
-Pointe: Selvom vi kigger på mange komvinationer af parameterne, så falder kombinationerne hvor de bedste resultater produceres (laveste mismatches) indenfor forholdsvis smalle områder.
-Det tyder på at metoden med de givne parametre er generaliserbar og kan anvendes på andre kameraer med samme setting.
+In general, the optimal settings for the three tracking parameters fall into relatively narrow ranges.
+This strengthens the assumption that the method is generalisable and could be applied for example across many more cameras than what have been demonstrated here.
 
-For our application, we find that a
+The eps parameter for the filtering...
 
-For the three series for which the tracking algorithm failed to produce perfect results, the range
+Thus, increasing eps means that more distant points will be clustered together. Oppositely, lowering eps will separate points into individual clusters despite being in close vicinity.
 

### 2022-02-02 code/DistanceFiltering.py
Wrote DistanceFiltering to replace DBSCAN

@@ -1,5 +1,11 @@
 # -*- coding: utf-8 -*-
 """
+Created on Mon Jan 31 12:56:11 2022
+
+@author: au309263
+"""
+# -*- coding: utf-8 -*-
+"""
 Created on Tue Dec 21 13:28:48 2021
 
 @author: au309263
@@ -14,21 +20,22 @@ from filtering import sieve
 import numpy as np
 from scipy.spatial import ConvexHull
 from datetime import datetime
+from itertools import combinations
+
 
 # Note, EPS_DISTANCE needs to be smaller than the gap between any two polygons we want to keep and large enough to cluster polygons that we want to discard together
 # EPS_DISTANCE = 281 is the highest value that keeps all tracks in THUL-01
 
 #tracks = pd.read_csv(r'U:\BITCue\Projekter\TrackingFlowers\testResults\_parameterTest_NARS-04_3\parameterTest_NARS-04_maxDisap_10_runMean_10_maxDist_500.csv')
 
-class DBSCANsieve():
-    def __init__(self, tracks, eps_distance, min_sample_polygons):
+class distanceSieve():
+    def __init__(self, tracks, eps_distance):
         if not isinstance(tracks, pd.DataFrame):
             self.tracks = pd.read_csv(tracks, sep=",", header = 0)
         else:
             self.tracks = tracks
         
         self.eps_distance = eps_distance
-        self.min_sample_polygons = min_sample_polygons
         
         self.points = {}
         self.lines = {}
@@ -104,26 +111,58 @@ class DBSCANsieve():
                         print(f'Adding object {k} with points {polygons[k]} to lines')    
                         self.lines[k] = polygons[k] # Add the points to the line dictionary  
     
+    def dist(self,a,b):
+        d = np.linalg.norm(a-b)
+        return d
+    
+    def ave(self, lst):
+        return sum(lst) / len(lst)
+    
     def run(self):
         self.separate() # Separate the tracks into points, lines, and polygons
         self.convex_hull(self.polygons) # Calcuate the convex hull of the polygons. This will also add colinear points to lines dictionary
         
         df, coords = self.geometries() # Get the geometry dataframe containing the track polygons and the centroid coordinates of which to perform the DBSCAN clustering
-
-        dbscan = DBSCAN(eps=self.eps_distance, min_samples=self.min_sample_polygons) # Run the DBSCAN clustering
-        clusters = dbscan.fit(coords)
-
[…]

### 2022-02-02 text/SupportManuscriptDraft.Rmd
Manuscript update Rev 2

---
new_session: FALSE
title: |
  **Supporting Information: Towards individual-based pollination ecology: Automatic tracking of life histories of individual flowers**
author:
- Hjalte M. R. Mann
- Alexandros Iosifidis
- Toke T. Høye
date: "`r format(Sys.time(), '%B %d, %Y')`"
output:
  word_document: default
  pdf_document:
    fig_caption: yes
    keep_tex: yes
    latex_engine: pdflatex
  html_document:
    df_print: paged
fontfamily: mathpazo
fontsize: 12pt
geometry: margin = 1in
header-includes:
- \usepackage{setspace}\doublespacing
- \usepackage[left]{lineno}
- \linenumbers
- \usepackage{dcolumn}
- \usepackage{caption}
- \usepackage{float}
- \usepackage{afterpage}
- \usepackage{siunitx}
- \usepackage{amsmath}
keywords: Tracking, flowering phenology, Arctic
bibliography: ./library.bib
csl: ./journal-of-ecology.csl
---

```{r eval = FALSE, echo=FALSE}
library(knitr)
knit('ManuscriptDraft.Rmd')
rmarkdown::render("SupportManuscriptDraft.Rmd", output_file = "SupportManuscriptDraft.docx")
system2("open","SupportManuscriptDraft.docx")
```

```{r echo=FALSE}
# Function for creating word comments in Rmarkdown
# Use: This text contains a `r word_comment("This is the comment", "comment.")`.

word_comment <- function(comment, highlight = "") {
  if (isTRUE(knitr:::pandoc_to() == "docx")) {
    paste0('[', comment, ']{.comment-start id="0" author="Hjalte M."',
           'date="2020-01"}', highlight, '[]{.comment-end id="0"}')
  }
}
```

\newpage

**Table S1:** Results of the filtering algorithm for filtering distances 5-500 with a stepsize of 50 for all three series.

![](../figures/FullFilteringDistanceTest.png){ width=100% }

### 2022-02-02 code/R/Analysis.Rmd
Added figures and tables. Added examples of annotated images.

@@
 
+# Camera codes:
+# 2018 NARS-04: 2018 NARS A
+# 2018 THUL-01: 2018 THUL B
+# 2019 NARS-13: 2019 NARS C
+# 2019 NARS-17: 2019 NARS D
+# 2019 NYAA-04: 2019 NYAA E
+
+all = all %>% 
+  mutate("camCode" = case_when(CameraID == "THUL-01"~ "2018 THUL B", CameraID == "NARS-04"~ "2018 NARS A", CameraID == "NARS-13"~ "2019 NARS C", CameraID == "NARS-17"~ "2019 NARS D", CameraID == "NYALE-04"~ "2019 NYAA E"))
+
+axisSize = 30
+legendTitleSize = 30
+legendTextSize = 28
+
+f1 = all %>% 
+ggplot()+
+  geom_point(aes(x = flower_start, y = flower_length, fill = camCode, shape = camCode), size = 8) + 
+  xlab("Onset of flowering (DOY)") + 
+  ylab("Number of days flowering") +
+    theme_bw()  +
+  scale_fill_manual(name = "Camera ID",
+                      values = c("gray26", "brown3", "dodgerblue3", "darkorange2", "chartreuse3"), aesthetics = c("colour", "fill")) + 
+  scale_shape_manual(name = "Camera ID",
+                      values = c(21,22,23,24,25)) + 
+  theme(panel.grid.major = element_blank(), 
+        panel.grid.minor = element_blank(),
+        panel.background = element_blank(),
+        legend.title = element_text(size = legendTitleSize),
+        text = element_text(size = axisSize),
+        legend.text = element_text(size = legendTextSize),
+        axis.line = element_line(),
+        axis.text=element_text(colour="black"),
+        panel.spacing = unit(0.7, "lines"))
+
+f1
+
+ggsave("U:/BITCue/Projekter/TrackingFlowers/figures/OnsetVSLength.png", f1, width = 15, height = 10)
+
 all %>% 
@@
   ggplot() +
-  geom_point(aes(x = x_c, y = y_c, color = as.factor(id_gt)), size = 1, shape = 1) + 
+  geom_point(aes(x = x_c, y = y_c, color = as.factor(Class)), size = 1, shape = 1) + 
   xlim(0,6080) +

### 2022-02-02 text/ManuscriptDraft.Rmd
Manuscript update
Added figures and tables. Added examples of annotated images.

@@
 new_session: FALSE
-title: |
-  **Towards individual-based pollination ecology: Automatic tracking of life histories of individual flowers**
-author:
-- Hjalte M. R. Mann
-- Alexandros Iosifidis
-- Toke T. Høye
-date: "`r format(Sys.time(), '%B %d, %Y')`"
+#title: |
+#  **Towards individual-based pollination ecology: Automatic tracking of life histories of individual flowers**
+#date: "`r format(Sys.time(), '%B %d, %Y')`"
 output:
@@
 - \usepackage{amsmath}
-keywords: Tracking, flowering phenology, Arctic, DBSCAN
+#keywords: Tracking, flowering phenology, Arctic
 bibliography: ./library.bib
 csl: ./journal-of-ecology.csl
-abstract: ABSTRACT | A fundamental problem in pollination ecology is that plant reproduction consequences of pollinator visits to a flower throughout its lifespan is very difficult to track.
-Often simple statistics are used to describe the flowering phenology of a population of plants, e.g.
-onset or peak of flowering.
-Similarly, flower visitation rates are based on crude observation of flower populations.
-Here we show that image-based monitoring of field plots at very high temporal resolution can return information on flowering phenology at the level of individuals.
-Further, we present a framework for automatically tracking, filtering, and evaluating flowers in time-lapse image series.
-We compare the results of automatic tracking with manual tracking.
-Individual flower tracking could facilitate finegrained studies on plant reproductive success....
-and we show that performance achieved by following automatic tracking can be reliable enough to estimate X, Y, Z...
 ---
@@
 
+Title:
+
+**Towards individual-based pollination ecology: Automatic tracking of life histories of individual flowers**
+
+#
+Authors:
+
+Hjalte M. R. Mann^1,2^, Alexandros Iosifidis^2^, Toke T. Høye^1^
+
+#
+Author affiliations:
+
+1: Department of Ecoscience and Arctic Research Centre, Aarhus University, C.F. Møllers Allé 4-8, 8000 Aarhus C, Denmark  
+2: Department of Electrical and Computer Engineering – Signal Processing and Machine Learning, Aarhus University, Finlandsgade 22, 8200 Aarhus N, Denmark
+
+#
+Corresponding author contact information:
+
+Hjalte Mads Rosenstand Mann  
+E-mail: mann@ecos.au.dk  
+Phone: +45 31178585
+
+#
+ORCID IDs:
+
+HMRM: 0000-0002-4768-4767, AI: 0000-0003-4807-1345, TTH: 0000-0001-5387-3284
+
+#
+Keywords:
+
+Tracking, flowering phenology, Arctic, dryas octopetala, dryas integrifolia, ecological monitoring, life-history variation
+
+\newpage
+# Abstract
+
+A fundamental problem in pollination ecology is that plant reproductive consequences of pollinator visits to a flower throughout its lifespan is very difficult to track.
+Deciphering the consequences of e.g.
+shifts in flower visitation rates to plant reproductive success and to identify mismatches between interacting and interdependent species such as insect-pollinated plants and pollinators throughout the season requires full-season phenology data, preferably at the level of individuals.
+Often simple statistics are used to describe the flowering phenology of a population of plants, e.g.
+onset or peak of flowering, which may fail to reveal the true dynamics in flowering and reproduction.
+Here we show that image-based monitoring of field plots at very high temporal resolution can return information on flowering phenology at the level of individuals.
+Further, we present a framework for automatically tracking, filtering, and evaluating flowers in time-lapse image series.
+We compare the results of automatic tracking with manual tracking.
+We show that performance achieved by automatic flower tracking can be reliable in obtaining individual flower phenology.
+Our method could facilitate fine-grained studies on plant reproductive success and is an important tool for estimating the sensitivity of plant-pollinator interactions to climate change.
+Supporting this paper is a publicly available web application through which users can apply the tracking and filtering methods on their own data.
+
 \newpage
@@
 
-Phenology is often studied with first events as proxies for the full phenology distribution of the population .
-For flowering plants the time of first flowering is used to represent the phenology of a population of plants.
-The ideal metric for phenology depends on the topic of the question, but often, the choice of onset of flowering as the proxy for phenology is based on data availability and not because it is necessarily considered the ideal metric.
-First events, such as onset of flowering, may in some cases be particularly ill suited for estimating sensitivity to climate change [@miller2008].
-The date for first flowering represents the very extreme of the phenology distribution of a population and may be confounded with e.g.
-the size of the population considered.
-Further, for many questions the phenology distribution of a population is required, for example to quantify resource availability through the season.
-However, even detailed information on phenology at the level of a population is still an aggregation of the phenological distributions of the individual members of the population for which important variation may be masked [@inouye2019].
-Having at hand the full distribution of flowering phenology at the level of individuals allows to extract relevant characteristics of the distribution and test their sensitivity towards environmental factors, but acquiring such data requires efficient and preferably automatized methods.
-
 Pollen limitation is a common factor limiting the reproductive success of flowering plants, but the ecological mechanisms behind pollen limitation is poorly understood, at least partly because it requires information on plant reproductive success in relation to pollinator visitation rates which is difficult to acquire in a natural setting and to a satisfactory extent.
@@
 Ideally, investigation of the sensitivity of interaction strength to dynamics in phenology requires quantification of phenology through the season for the species in question in order to asses both their phenology and abundance [@inouye2019; @linden2018].
-
 If interacting and interdependent species, e.g.
@@
 
-Different phenological factors, such as onset, peak, and end of flowering may respond differently to climate change [@cook2012] and estimation of the degree of mismatch between species may thus be sensitive to the choice of metric [@inouye2019].
-Efficient and precise methods for deriving full seasonal distributions of phenology at the level of individuals is much needed but so far most often unfeasible for logistical reasons.
-
 Monitoring flower phenology at high temporal resolution is time-consuming, particularly in logistically challenging environments such as the Arctic.
@@
 Such inaccurate estimates may fail to reveal important dynamics in flowering phenology through the season.
-Automatic image-based monitoring of flowering phenology can return phenology data for specific species at very high temporal resolution (Mann et al., in review), but phenological responses at the individual level may be indiscernible regardless of the temporal resolution of the data at population level.
+For example, different phenological factors, such as onset, peak, and end of flowering may respond differently to climate change [@cook2012] and estimation of the degree of mismatch between species may thus be sensitive to the choice of metric [@inouye2019].
+A lot remains to be discovered about when and how the dynamics of pollination affects the reproductive success of flowering plants.
+High frequency observations of flower visits coupled with information on phenology and reproductive success at the level of individuals is required to unravel the relationship between pollinator visitation and plant reproductive success and the sensitivity of this interplay to climate.
+
+The ideal metric for phenology depends on the topic of the question, but often the choice of onset of flowering as the variable representing phenology is based on data availability and not because it is considered the ideal metric.
+First events, such as onset of flowering, may in some cases be particularly ill suited for estimating sensitivity to climate change [@miller2008].
+Onset of flowering represents the very extreme of the phenological distribution of a population and may be confounded with e.g.
+the size of the population considered.
+Further, for many questions, the phenology distribution of a population is required, for example to quantify resource availability through the season.
+However, even detailed information on phenology at the level of a population is still an aggregation of the phenological distributions of the individual members of the population for which important variation may be masked [@inouye2019].
+Having at hand the full distribution of flowering phenology at the level of individuals allows to extract relevant characteristics of the distribution and test their individual sensitivities towards environmental factors, but acquiring such data requires efficient and preferably automatized methods.
+Automatic image-based monitoring of flowering phenology can return phenology data for specific species at very high temporal resolution (Mann et al., in review (Note: I'll update this if the paper gets accepted or remove it if not)), but phenological responses at the individual level may be indiscernible regardless of the temporal resolution of the data at population level.
 For example, a shortening of individual flower longevity may not be directly obvious at the population level.
@@
 Similarly so for investigating whether flower visitation rates and/or reproductive success depends on the timing of flowering for the individual flower.
+Efficient and precise methods for deriving full seasonal distributions of phenology at the level of individuals is much needed.
 
-A lot remains to be discovered about when and how the dynamics of pollination affects the reproductive success of flowering plants.
-High frequency observations of flower visits coupled with information on phenology and reproductive success at the level of individuals is required to unravel the relationship between pollinator visitation and plant reproductive success and the sensitivity of this interplay to climate.
+Cameras have become an ubiquitous tool for non-invasive and efficient monitoring of natural processes [@howe2017; @glover2019], but the implementation of image-based collection of data has outpaced the development of reliable and easily implementable methods for processing and analysing the big data output from this transformation [@weinstein2018; @farley2018].
+Computer vision and deep learning tools and being developed and applied for automatic processing and extraction of relevant data from these very large image sets [@willi2019;  @tabak2019] and increasingly for challenging tasks in entomology [@hoye2021], such as for fine-grained species classification [@spiesman2021] and monitoring of pollination events by detecting, identifying, and counting insects [@bjerge2021].
+Collectively, the development and adoption of modern technology allows for collection and processing of ecological data at unprecedented scales.
 
-In this paper, we show that information on phenology at the level of individuals can be derived from image-based monitoring of flower phenology.
+Here, we use stand-alone time-lapse cameras to monitor the phenology of two Arctic flowering plant species Dryas integrifolia and D.
+octopetala.
+We show that information on phenology at the level of individuals can be derived from image-based monitoring of flower phenology.
 We present and evaluate an automatic flower tracking and filtering algorithm.
@@
 For instance, it could be explored whether seed set is affected by timing and length of flowering and/or abiotic conditions such as frost events.
-Hereby, the longstanding question of what drives variation in reproductive success of individual flowers could be explored.
+Hereby, the long-standing question of what drives variation in reproductive success of individual flowers could be explored.
 By simultaneously tracking flower visits, these could be assigned to the individual flower and visitation rates per flower could be calculated per flower and related to reproductive success.
 Further, any information of taxonomic grouping could facilitate comparative studies on pollination effectiveness between taxonomic groups of pollinators.
-The possibility of automatically tracking phenology at the level of individuals will facilitate investigation of a set of yet untestable research questions and is an important contribution to the shift towards efficient and detailed monitoring of ecological processes.
+The possibility of automatically tracking phenology at the level of individuals will facilitate investigation of a set of yet untestable research questions and is an important contribution to the shift towards efficient and detailed monitoring of phenology.
 
 # Material and methods
-#
 ## Study site and species
-This study was based on five time-lapse image-series collected at the following three Arctic sites (sampling years given in brackets): Narsarsuaq, South Greenland (2018, 2019), Thule, North-West Greenland (2018), Ny-Ålesund, Svalbard (2019).
+This study was based on five time-lapse image-series collected at the following three Arctic sites (sampling years given in brackets): Narsarsuaq (NARS), South Greenland (2018, 2019), Thule (THUL), North-West Greenland (2018), Ny-Ålesund (NYAA), Svalbard (2019).
 The images were collected using time-lapse cameras (Moultrie Wingscapes TimelapseCam Pro, Moultrie Products, Birmingham Alabama, USA) permanently positioned approx.
-60 cm above areas with either Dryas integrifolia in Narsarsuaq and Thule or D.
-octopetala in Ny-Ålesund.
+60 cm above areas with either *Dryas integrifolia* in Narsarsuaq and Thule or *D.
+octopetala* in Ny-Ålesund.
 The white flowers of these two species are insect-pollinated and held erect above cushions of foliage.
-The species are native to and widespread in Asian, North American, and European arctic and alpine regions.
-They are geographically separated apart from a possible hybrid zone in Northeast Greenland [@philipp2003].
+The species are native to and widespread in Asian, North American, and European Arctic and alpine regions.
+They are geographically separated apart from a possible hybrid zone in North-East Greenland [@philipp2003].
 
@@
 ## The image series and flower annotations
-Images were collected throughout the full growing season.
-While the original time-lapse frequency was higher, the basis for this study was 1-hr interval sereis.
-We limited the 1-hr interval image-series to the flowering season and randomly randomly sampled images from each series.
+Images were collected throughout the full growing seasons.
+While the original time-lapse frequency was higher, the basis for this study was 1-hr interval series.
+We limited the 1-hr image-series to the include only the flowering season and randomly randomly sampled images from each series.
 The number of images for each series after sampling is given in table 1.
-All flowers in each of these images were manually annotated using the rectangular bounding box tool in the VIA VGG annotation software [@dutta2019] and each individual flower was assigned a unique ID.
+All flowers in bloom in each of these images were manually annotated using the rectangular bounding box tool in the VIA VGG annotation software [@dutta2019] and each individual flower was assigned a unique ID.
 These annotations constitute our ground truth tracks.
 The series varied substantially in the number of flowers they contained (table 1).
-Figure 1 show annotated example images from three series (xxx) at different times during the flowering season (TO-DO).
+Examples of annotated images from the 2019 NARS C series at three different times in the flowering season are given in fig.
+1.
+Figure 2 shows the time of flowering for each individual flower and the number of days the flower was present in the images.
 
@@
 #
-![](../figures/SOmeFigure.png){ width=100% }
+![](../figures/AnnotationExamples_3.png){ width=100% }
 
-**Figure 1:** Figure text... Examples of annotated images (TO-DO)
+**Figure 1:** Examples of annotated images for 2019 NARS C.
+a: Onset of flowering at DOY 145; b: Peak flowering (52 annotated flowers) 6 days after onset (DOY 151); c: 6 days after peak flowering (DOY 157).
+Flower bounding boxes are colour coded by flower ID.
 
 #
-### Input data for the tracking algorithm
-
-The input data for the tracking algorithm are bounding box coordinates for each object in each frame given in the following format: frame number, x-min, y-min, x-max, y-max.
-Here, x-min, y-min is the top left corner of the bounding box and x-max, y-max is the bottom right corner.
-The bounding box centroid is calculated on the basis of these coordinates within the algorithm.
-Calculating tracking accuracy requires a ground truth ID for each bounding box which can be given as the sixth column.
-As this data structure is outputted from both manual object annotations and object detection methods using CNNs, the method can be applied on both manually and automatically derived data.
-Further, as the algorithm handles frames sequentially, it can be applied both offline and online.
+![](../figures/OnsetVSLength.png){ width=100% }  
+**Figure 2:** Onset of flowering (DOY) and flower lifespan (days) for each individual flower in each of the five image-series.
 
@@
 Several issues complicate the task of tracking individual flowers through time-lapse image-series.
-
 First, stand-alone monitoring of the relatively slow process of flower development through full growing seasons (several months) requires only low frame rates.
@@
 For example, as the wind shifts, the flower heads change direction, which in the image series happens instantaneously (i.e.
-between two consecutive frames).
+between two consecutive frames) (fig.
+3a).
 However, as they are constrained by their stalk, there is a limit to the distance they can move.
 With the tracking parameter **running mean**, we have the option to base the tracking on the distance between a point in the current frame and the running mean of the positions of a number of previous points in a track, taking advantage of the fact that flower move around a central point.
+Increasing the value for this parameter decreases the weighting of movement between individual frames.
 
 Second, as flowers move, they may temporarily occlude the view of other flowers and flowers close to the edge of the frame may move in and out of view.
-This can cause a track to be lost and a new track erroneously being initiated when the flower reappears.
+This can cause a track to be lost and a new track erroneously being initiated when the flower reappears (fig.
+3b).
 However, if a flower reappears in the same area as a flower is already being tracked after disappearing in a few frames, it is reasonable to assume that it is the same individual and not that the old flower wilted/disappeared and a new one developed.
@@
 
-Third, a centroid tracking algorithm will associate a point with a tracking based on minimum distance only, disregarding the absolute distance.
+Third, a standard centroid tracking algorithm will associate a point with a tracking based on minimum distance only, disregarding the absolute distance.
 As flowers are constrained in the movement, we can assume a maximum distance that will occur between points and force the initiation of tracks for points that exceed this threshold.
-We set this threshold with the parameter **max distance**.
-
-# EDIT FIGURES: Zoom in, bigger dots, combine into one figure.
-#
-![](../figures/figure_1.png){ width=100% }
-
-**Figure 1:** `r word_comment("The three figures below are quick mockups to explain the challenges with tracking flowers.
-Not sure if figures are needed or text is sufficient.", "Simple")` centroid tracking may produce erroneous associations when objects move between frames.
-Blue shows detections in the current frame (bounding box and centroid point).
-Red shows centroid points for the detections in the previous frame.
+We set this threshold with the parameter **max distance** measured as the euclidean distance in number of pixels.
 
 #
-![](../figures/figure_22.jpg){width=100%}
+![](../figures/HanddrawnFiguresIssues.png){ width=100% }
 
-**Figure 2:** Simple centroid tracking may produce erroneous associations when objects move between frames.
-Basing the association on the running mean of the positions of the previous n number of tracks may alleviate this issue.
-In this case basing the association on only the previous point would produce a wrong results while basing it on the running mean would produce a correct result.
-Red points: Centroids for bounding boxes in current frame, blue points: Centroids for bounding boxes in previous frame; grey points: Centroids for bounding boxes in a number of frames before t-1; green points: Running mean of the previous n points.
-Circles delimit the two individuals.
+**Figure 3:** a: Centroid tracking may produce erroneous associations when objects move between frames.
+Here the track initiated on the top left flower in frame t-1 would switch to the bottom right flower in frame t b: Tracking errors may occur when objects temporarily disappear from the frame.
+Here the top flower moves out of frame and the bottom flower would be assigned to the track of the top flower in frame t-1 Red shows centroid of bounding box in the current frame and blue shows centroid points from the previous frame.
 
 #
-![](../figures/figure_3.png){ width=100% }
+### Input data for the tracking algorithm
 
-**Figure 3:** Simple centroid tracking may produce erroneous associations when objects disappear periodically from the frame.
-Here the top flower moves out of frame and the bottom flower would be assigned to the track of the top flower in frame t-1.
+The input data for the tracking algorithm are bounding box coordinates for each object in each frame given in the following format: frame number, x-min, y-min, x-max, y-max.
+Here, x-min, y-min is the top left corner of the bounding box and x-max, y-max is the bottom right corner.
+The bounding box centroid is calculated on the basis of these coordinates within the algorithm.
+Calculating tracking accuracy requires a ground truth ID for each bounding box which can be given as the sixth column.
+As this data structure is outputted from both manual object annotations and object detection methods using CNNs, the method can be applied on both manually and automatically derived data.
+Further, as the algorithm handles frames sequentially, it can be applied both offline and online.
 
@@
 
-We base our filtering on a density-based clustering of the centroid of a geometry of each track.
+We base our filtering on the distance between geometry centroids the tracks.
 The geometries are derived in the following way: For tracks consisting of only a single point, the coordinates of the point are used as the centroid.
@@
 
-The filtering is done on the geometry centroids to remove tracks in areas with a high density of tracks as these have a high risk of tracking mismatches.
-The filtering is done in two steps.
-First, the DBSCAN clustering algorithm [@ester1996] is run on the centroids with a conservatively high value for the eps parameter, meaning that tracks in close vicinity to each other will be clustered together.
-Second, all tracks that were not assigned to a unique cluster are removed.
-Tracks that are spatially isolated remains.
+The filtering is done on the geometry centroids to remove tracks in close vicinity to each other as these have a high risk of tracking mismatches.
+The distances between all centroid tracks are calculated (euclidean distance in number of pixels) and only tracks that have no neighbouring centroid within the filtering distance threshold are kept.
 
@@
 For this, we chose a fixed combination of the tracking algorithm parameters across all three series (max gap size: 10; running mean: 10; max distance: 300).
-
-We then applied the filtering method on the three series with a range of values for eps (x-xx, step size of 10) and identified the value that returned the highest sum of tracks across the three series at the lowest number of mismatches.
+We then applied the filtering method on the three series with a range of values for the filtering distance (50-500, step size of 50) and identified the value that returned the highest sum of tracks across the three series at the lowest number of mismatches.
 We present the tracking accuracy before filtering and on the remaining tracks after filtering.
@@
 For example, basing the filtering on the best performing tracking parameters for each series specifically would improve the output of the tracking algorithm.
-Similarly, the results could be improved by fine-tuning the value for eps for each series individually.
+Similarly, the results could be improved by fine-tuning the filtering distance threshold for each series individually.
 However, as our goal here was to demonstrate that our tracking and filtering method can be applied in a naive setting with fixed values for each parameter we have not included those results.
@@
 #
+## Tracking and filtering web application
+
+To make the method described here easily accessible for other users, we built a web application through which the trackking and filtering method can be applied on user data.
+The web application is accessible through https://github.com/TECOLOGYxyz/FlowerTracking.
+
+#
 # Results
@@
 For two series, 2018 THUL B and 2019 NARS D, the tracking algorithm returned perfect results for several combinations of parameters.
-For the three remaining series, the algorithm returned consistently high MOTA scores (above 0.95).
+For the three remaining series, the algorithm returned consistently high MOTA scores (minimum 0.95).
 In all cases, except 2019 NARS D, where the tracking algorithm returned perfect results with all parameters zeroed, did the inclusion of the user set tracking parameters improve the tracking accuracy.
 
-In one case, NARS-13, four different numbers of tracks were returned among the parameter combinations that resulted in the lowest number of mismatches.
+In one case, 2019 NARS C, four different numbers of tracks were returned among the parameter combinations that resulted in the lowest number of mismatches.
 A number of tracks equal to the number of flowers in the series were among these four.
-
-In one case, NYAA-04, the parameter combination that produced the highest accuracy did not return a number of tracks equal to the number of flowers in the series.
+In anotehr case, 2019 NYAA E, the parameter combination that produced the highest accuracy did not return a number of tracks equal to the number of flowers in the series.
 However, combinations that did produce this number of tracks also returned a higher number of mismatches.
 
-**Table 1:** Table description... To-do: Rename series to camera codes.
-
-#
-![](../figures/table1.png){ width=100% }
+**Table 1:** Tracking performance for the automatic flower tracking algorithm for different tracking parameter combinations.
+Number of tracks in bold shows when the number of tracks equals the number of flowers in the series.
+![](../figures/TrackingScoresv22.png){ width=100% }
 
@@
 
+The value for filtering distance that returned the highest number of tracks at the lowest number of mismatches was 350 (results for all values given in supp.
+table S1).
 The results of the filtering algorithm on the three series in which our tracking algorithm did not return perfect results are given in table 2 and the filtering process is visualised in fig.
 4.
-Our filtering method extracted 28 tracked flowers from the three series with complex scenes using a fixed value of 350 for eps and fixed values for the tracking parameters (10, 10, 300, for running mean, max gap size, and max distance, respectively) with only a single mismatch included in the 28 tracks.
+Our filtering method extracted 28 tracks from the three series (2018 NARS A: 15 tracks, 1 mismatch; 2019 NARS C: 10 tracks, 0 mismatches; 2019 NYAA E: 3 tracks, 0 mismatches) with complex scenes using a fixed filtering distance value of 350 and fixed values for the tracking parameters (10, 10, 300, for running mean, max gap size, and max distance, respectively) with only a single mismatch included in the 28 tracks.
 
-**Table 2:** Results of the filtering algorithm with an eps value of 350 applied across all three series To-do: Change image series to camera codes.
+**Table 2:** Results of the filtering algorithm with a filtering distance value of 350 applied across all three series.
 
-![](../figures/FilterScores.png){ width=100% }
+![](../figures/FilterScoresv211.png){ width=100% }
 
@@
 **Figure 4:** Track filtering process.
-Row 1, 2, and 3 is NARS-13, NARS-04, and NYAA-04, respectively.
+Row 1, 2, and 3 is 2019 NARS C, 2018 NARS A, and 2019 NYAA E, respectively.
 First column shows the centroid points in the given series, coloured by the track id from the centroid tracking algorithm.
 Second column shows the polygons calculated from the tracks.
-The DBSCAN clustering algorithm with eps = 350 was applied to the centroids of these polygons and the polygons are coloured by cluster id.
-Third column shows the results of the filtering where all tracks that were not assigned a unique cluster has been removed.
+The filtering algorithm with a distance threshold of 350 was applied to the centroids of these polygons.
+Third column shows the results of the filtering where all tracks that did not pass the filtering have been removed.
 
 # Discussion  
-
-#
 ## Tracking
@@
 In our case, however, flowers are not clearly distinguishable and one flower may look less alike between two frames than two flowers in a single frame.
-
 Finally, some tracking methods require manual initiation of a fixed number of tracks for objects in the first frame and the algorithm will track these objects in the subsequent frames [@hu2012].
@@
 These settings vastly complicate the task of tracking multiple objects.
-
 Our tracking algorithm returned consistently accuracy despite the difficult settings.
@@
 In a naive setting, a priori knowledge about which settings to use is not available.
-As with any tracking method, accuracy can only be assessed based on visual observation or more thoroughly by testing on subsample of tracks for which a ground truth exists.
+As with any tracking method, accuracy can only be assessed based on visual observation or more thoroughly by testing on subset of tracks for which a ground truth exists.
 When applying the method in a new setting, we recommend performing a parameter test as demonstrated here on a subset of ground truth tracks to derive the optimal set of parameters.
@@
 A higher number of flowers within the frame increases the possibility of flowers in close vicinity to each other and of flowers overlapping and moving in and out of frame translating to a higher risk of making tracking errors.
-The very high complexity in the scenes with many flowers, especially NARS-13 with 85 flowers, renders it unlikely that any tracking method could produce perfect results.
+The very high complexity in the scenes with many flowers, especially 2019 NARS C with 85 flowers, renders it unlikely that any tracking method could produce perfect results.
 Although this setting can be considered an extreme case for our application, it underlines the relevance of a pipeline including both tracking and filtering.
@@
 
-We presented a filtering method based on the density of tracks within the frame which can be applied subsequent of tracking.
+We presented a filtering method based on the distance between track geometry centroids which can be applied subsequent of tracking.
+The filtering algorithm successfully extracted correctly tracked flowers from each of the three series for which the tracking algorithm did not produce perfect results.
+Only a single mismatch was included in the 28 flowers that passed the filtering.
 Extracting tracks that are spatially isolated does not guarantee that the tracks are correct/without errors.
@@
 
-The filtering algorithm successfully extracted correctly tracked flowers from each of the three series for which the tracking algorithm did not produce perfect results.
-Only a single mismatch was included in the 28 flowers that passed the filtering.
+We applied the filtering algorithm on tracks produced with a fixed set of tracking parameters as well as a fixed value for the filtering distance.
+We note that the results could be improved (i.e.
+more flowers extracted without increased number of mismatches) by chosing the best tracking result for each series and fine-tuning the filtering distance value for each individual series.
+However, as our goal here was to show that a single conservative value can be applied across series for which the tracking algorithm did not necessarily produce the best results, we do not show those results here.
+As in the case of the tracking parameters, we recommend setting the value for the filtering distance based on either visual examination of results or on tests on a subset of ground truth data when applying the method in a new setting.
 
-We applied the filtering algorithm on tracks produced with a fixed set of tracking parameters as well as a fixed value for the DBSCAN eps parameter for the filtering algorithm.
-We note that the results could be improved (i.e.
-more flowers extracted without increased number of mismatches) by chosing the best tracking result for each series and fine-tuning the eps value for each individual series.
-However, as our goal here was to show that a single conservative value can be applied across series for which the tracking algorithm did not necessarily produce the best results, we do not show those results here.
-As in the case of the tracking parameters, we recommend setting the value for eps based on either visual examination of results or on tests on a subset of ground truth data when applying the method in a new setting.
-
-We note that our method for filtering tracks using DBSCAN on track centroids ensures that all tracks are given the same weighting in the filtering since each track is represented by a single point.
+We note that our method for filtering tracks using track geometry centroids ensures that all tracks, despite the number of points in the track, are given the same weighting in the filtering since each track is represented by a single point.
 In some cases this may be suboptimal.
@@
 
-## The tracking and filtering parameters
+Filtering based on geometry centroids ensures that the same weight is put on all tracks despite the number of points within the track.
+Additionally, it allows for overlap between track geometries as long as the distance between the track centroids is above the threshold.
 
-Our method relies on a number of parameters set by the user.
+#
+## User-defined parameters
+
+Our method relies on a number of parameters set by the user, both for tracking and subsequent filtering.
+Our results demonstrate that the tracking accuracy is sensitive to these parameters and an optimal combination of parameters can improve results substantially.
+Below we discuss the underlying mechanisms of each of the parameters.
 
@@
 
-In all cases, except NARS-17, for which perfect tracking was achieved with all parameters set to zero, did a value for running mean above 0 improve results.
-
-Interestingly, setting the value to 10 produced the best results in all these cases, except for the combination returning the lowest number of mismatches for NYAA-04.
+In all cases, except 2019 NARS D, for which perfect tracking was achieved with all parameters set to zero, did a value for running mean above 0 improve results.
+Interestingly, a value of 10 produced the best results in all these cases, except for the combination returning the lowest number of mismatches for NYAA-04.
 Increasing the value for running mean means that less weight is put on movements between each frame.
@@
 
-For the three series for which the tracking algorithm did not produce perfect results, the optimal range for max distance were relatively narrow (300 - 500).
+For the three series for which the tracking algorithm did not produce perfect results, the optimal range for max distance were relatively narrow (300-500).
 The optimal value for this setting is one that is sufficiently high so that it does not force individual flowers into separate tracks because of movement between frames yet is low enough to separate any flowers into individual tracks that would otherwise have been grouped together.
@@
 This underlines the relevance of search for an optimal combination based on ground truth data.
-
 In general, the optimal settings for the three tracking parameters fall into relatively narrow ranges.
@@
 
-The eps parameter for the filtering...
-
-Thus, increasing eps means that more distant points will be clustered together. Oppositely, lowering eps will separate points into individual clusters despite being in close vicinity.
-
-Why do they make a difference:
-
-max gap size:
-
-Running mean:
-
-Max distance:
-
-eps:
-
-Offline and online tracking differ
-
-Online tracking
-
-Our tracking algorithm does not rely on information from future frames and can thus be applied both offline and online.
-
-MOT can also be categorized into online tracking and offline
-tracking.
-The difference is whether or not observations from future frames are utilized when handling the current frame.
-Online, also called causal, tracking methods only rely on the past information available up to the current frame, while
-
-It also depends on the time the flowers are spread over. Many flowers and short time is a hard problem.
-
-Paragraph about applying the technique on detection instead of manual annotations.
-Detections introduce false positives.
-Either manual or automatic quality control to remove these before tracking or after.
-
-Our manually tracked data is ground truth, but for example when flowers periodically move out of the frame, this mimics false negatives. Similarly, when one flower occludes another.
-
-Automatically detecting flowers would likely introduce a degree of false negatives which would decrease the MOTA score if max gap size is set to zero.
-Introducing a value for this parameter can deal with the problem of false negatives.
-
-Our tracking algorithm returned reasonably high MOTA scores even with all three parameters set to zero. The degree to which mismatches are accepted may depend on the application?
+We filter out tracks with high risk of error based on distance between track geometry centroids.
+An alternative approach could be to filter based on overlap between polygons, but this would remove tracks with even minor overlap and put weight on tracks with many points as these cover more area.
+Another alternative could be to base filtering on the centroids of the points within a track (i.e.
+mean x and mean y).
+The presented method avoids weighting track geometries based on number of included points and avoids weighting the centroid coordinate based on the point coordinates within the track.
+We note, however, that other approached may be more suitable in other settings and applications.
 
 #
-## The image series
-
-A point on the fact the these series are complex.
+## Input data format
+We demonstrated our method on manual annotations of flowers but note that the format of the input data matches the output of deep learning object detection methods.
+Thus, the method can be directly applied on such data.
+Automatically detecting flowers would likely introduce a degree of false negatives when the detection algorithm fails to detect a present object.
+The flowers in our data that periodically move out of the frame or are temporarily occluded mimics false negatives.
+Thus, it is likely that the max gap size parameter can alleviate problems with false negatives.
+Automatic detection methods may also introduce false positives.
+These detections would be included in the tracking.
+Single false positives could be removed automatically while other cases may require manual quality control.
+Depending on the application, a low degree of false positives may not problematic.
+In general, the performance of a tracking algorithm depends on the quality of input data and it is advisable to optimise detection accuracy before moving to tracking.
+The fact that our method produced perfect tracking results in simple settings could be taken into consideration when setting up future data collection schemes.
 
@@
 ## Conclusion
-A paragraph about the ecological perspectives of being able to track the individual flowers (at scale).
 
-We have demonstrated that our tracking algorithm with the three tracking parameters and the subsequent filtering algorithm can substantially increase accuracy of the tracking of flowers despite the fact that they break the assumptions of general tracking methods.
+In this paper, we have demonstrated the potential of image-based monitoring for collecting information on flower phenology at the level of individuals.
+The automatic tracking and filtering pipeline presented was able to return reliable flower tracks despite settings that violate assumptions of standard tracking algorithms and complex scenes with a high number of objects that are not permanent within the series.
+Our results show that it is possible to collect and analyse individual flower phenology at large scale at logistically challenging sites.
+The presented method is a significant contribution towards automatic and efficient phenology data collection.
 
 # Acknowledgements
+?
 
@@
 The code that supports the results in this paper will be made openly available at https://github.com/TECOLOGYxyz/FlowerTracking.
-A publicly available web application can be accessed from the Github repository through which users can run the tracking algorithm on their own data.
-Raw data as well as the trained flower detection model will be archived on https://zenodo.org/.
-
-\newpage
-**NOTES**
-
-* **Flowering phenology**
-  + Importance of studying flowering phenology
-  + Responses to climate change
-  + Phenology of communities, populations, individuals
-  + Traditional methods for studying
-  + Onset of flowering says little about true distribution
-    - Even true distribition of community says little about flowering lengths of individuals and for example how it varies accross the season.
-  + Difficult to study at the individual level - requires high temporal resolution and keeping track of individuals
-* **Image based monitoring**
-  + Automatic, high temporal resolution, remote sites
-  + High temporal resolution means that we can annotate individuals and get phenology of the individual
-
-* **Tracking**
-  + Offline and online
-  + Online often coupled with CNNs that attempt to distinguish individuals from each other and recognize them through frames
-
-  + Flowers appear very similar and 
-  + Many methods for offline tracking
-  + Hungarian/Kahlmann filter
-  ++ May not be applicable for objects that move weirdly, e.g. change directions between frames.
-  + Tracking based on distance.
-    - Good but has some problems
-    - Two points always associated disregarding absolute distance
-    - Tracks lost when objects disappear
-    - Objects close to each other may swap tracks
-
-* **Our solution**
-  + Here we demonstrate a framework for automatic flower tracking and evaluation of tracking performance
-  + Ground truth tracks
-
-The number of flowers produced by an individual is also a factor.
-
-For each series, we calculate the ratio of flowers for which the automatic tracking algorithm returns the correct flowering length compared to the ground truth tracks.
-
-`r word_comment("I haven't included these results.
-Not sure if they should be.", "Additionally,")` we performed a set of analyses on the five ground truth tracks.
-First, we calculated the largest distance between any two points within any track for any flower in each of the five series.
-Further, we calculated the maximum number of frames a flower track was lost and subsequently reappeared for each series.
+A publicly available web application can be accessed from the Github repository through which users can run the tracking and filtering algorithms on their own data.
+Raw data will be archived on https://zenodo.org/.
 

### 2022-02-02 code/run.py

@@ -9,6 +9,7 @@ import pandas as pd
 from track import tracker
 from filtering import sieve
 from DBSCANFiltering import DBSCANsieve
+from DistanceFiltering import distanceSieve
 from evaluate import evaluator
 import os
 import time
@@ -129,33 +130,33 @@ def impdata(camID):
 ###
 ####
 
-# Find a good value for eps
+#Find a good value for eps
 
-epslist = [350]
-results_filename = '../testResults/temp_epstest.csv'
+# epslist = [350] # [350]
+# results_filename = '../testResults/temp_epstest.csv'
 
-temppath = '../testResults/temp_epsExperiment_maxDisap_10_runMean_10_maxDist_300.csv'
-gt_temppath = '../testResults/gttemp.csv'
+# temppath = '../testResults/temp_epsExperiment_maxDisap_10_runMean_10_maxDist_300.csv'
+# gt_temppath = '../testResults/gttemp.csv'
 
-for eps in epslist:
-    mmsum = 0
-    flowersum = 0
-    # Return also the number of flowers/object returned after filtering. Find good combo between this and mismatches.
+# for eps in epslist:
+#     mmsum = 0
+#     flowersum = 0
+#     # Return also the number of flowers/object returned after filtering. Find good combo between this and mismatches.
 
     
-    dt = r'U:\BITCue\Projekter\TrackingFlowers\testResults\_parameterTest_NARS-17_3\parameterTest_NARS-17_maxDisap_10_runMean_10_maxDist_300.csv'
-    gt = pd.read_csv(impdata("NARS-17"))
-    print(gt)    
-    d = DBSCANsieve(dt, eps_distance = eps, min_sample_polygons = 1)
-    filteredTracks = d.run()
-    filteredTracks.to_csv(temppath)    
+#     dt = r'U:\BITCue\Projekter\TrackingFlowers\testResults\_parameterTest_NARS-17_3\parameterTest_NARS-17_maxDisap_10_runMean_10_maxDist_300.csv'
+#     gt = pd.read_csv(impdata("NARS-17"))
+#     print(gt)    
+#     d = DBSCANsieve(dt, eps_distance = eps, min_sample_polygons = 1)
+#     filteredTracks = d.run()
+    # filteredTracks.to_csv(temppath)    
     
[…]

### 2022-05-25 code/MOTMetrics.py
Transfered to using pymotmetrics for evaluating tracking accuracy. Not using evaluate.py anymore.

@@ -1,0 +1,323 @@
+# -*- coding: utf-8 -*-
+"""
+Created on Sat Apr 30 12:01:22 2022
+
+@author: au309263
+"""
+# -*- coding: utf-8 -*-
+"""
+Created on Wed Apr 27 16:29:43 2022
+
+@author: au309263
+"""
+
+
+"""
+Calculate MOT metrics using the pymotmetrics package.
+
+
+"""
+
+import pandas as pd
+from scipy.spatial import distance as dist
+import os
+import re
+import numpy as np
+
+import motmetrics as mm
+
+#sites = ['THUL-01', 'NARS-04', 'NARS-13', 'NARS-17', 'NYAA-04']
+
+# Run in several kernels
+#sites = ['THUL-01', 'NARS-04', 'NARS-17', 'NYAA-04']
+sites = ['NARS-13']
+
+#sites = ['THUL-01']
+
+
+
+def impdata(camID):
+    if camID == "NARS-04":
+        gt = r"U:\BITCue\Projekter\TrackingFlowers\data\annotations\2020_05_17_NorwayAnnotations_NARS-04_IndividualAnnotations_FRCNN_Metrics.csv"
+    if camID == "NARS-13":
+        gt = r"U:\BITCue\Projekter\TrackingFlowers\data\annotations\2020_04_30_NorwayAnnotations_NARS-13_IndividualAnnotations_FRCNN_Metrics.csv"
+    if camID == "THUL-01":
+        gt = r"U:\BITCue\Projekter\TrackingFlowers\data\annotations\2020_05_15_NorwayAnnotations_THUL-01_IndividualAnnotations_FRCNN_Metrics.csv"
+    if camID == "NYAA-04":
+        gt = r"U:\BITCue\Projekter\TrackingFlowers\data\annotations\2021_12_13_NorwayAnnotations_NYAA-04_IndividualAnnotations_FRCNN_Metrics.csv"
+    if camID == "NARS-17":
+        gt = r"U:\BITCue\Projekter\TrackingFlowers\data\annotations\2021_12_30_NorwayAnnotations_NARS-17_IndividualAnnotations_FRCNN_Metrics.csv"
+    if camID == "debug":
+        gt = r"U:\BITCue\Projekter\TrackingFlowers\data\annotations\debugGT.csv"
+    return gt
+        
+
+
+
+# ===================== Run evaluator on all files for the parameter test ====
+
+br = '\n'
+
+def write_header(results_filename):
[…]

### 2022-05-25 code/R/VisualizeTracks.Rmd
Added script for visualizing tracks to exemplify errors.

---
title: "VisualizeTracks"
author: "Hjalte Mann"
date: '2022-05-12'
output: html_document
---

```{r setup, include=FALSE}
library(ggplot2)
library(dplyr)
library(stringr)
library(GGally)
library(plotly)
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
meta_2018 = read.csv("../../data/annotations/DetectionPhenology_2018_Meta_Clean_Prepped.csv")

meta_2018 = meta_2018 %>% 
  select(FileName, DOY, DateTime)

colnames(meta_2018) = c("filename", "DOY", "time")

```

```{r}
THUL01Anno = read.csv("U:/BITCue/Projekter/TrackingFlowers/data/annotations/2020_05_17_NorwayAnnotations_NARS-04_IndividualAnnotations_FRCNN_Metrics.csv", header = T)

THUL01Tracked = read.csv("U:/BITCue/Projekter/TrackingFlowers/testResults/MMFix/_parameterTest_NARS-04_3/parameterTest_NARS-04_maxDisap_10_runMean_20_maxDist_500.csv")

m = merge(THUL01Anno, THUL01Tracked, by = c("filename","x_min", "y_min", "x_max", "y_max"))

m = merge(m, meta_2018, by = "filename")

m %>% 
  filter(id_gt == "Flower_15" | id_gt == "Flower_28"| id_gt == "Flower_25") %>% 
ggplot() + 
  geom_point(aes(x = frame, y = id_gt, colour = as.factor(objectID)), size = 4, shape = 15)

mp = m %>% 
  filter(id_gt == "Flower_15" | id_gt == "Flower_28"| id_gt == "Flower_25") %>% 
  ggplot() +
   geom_line(mapping = aes(DOY,id_gt, group=id_gt, colour = as.factor(objectID)), size = 4) +
  ylab("") +
  ggtitle("Max gap = 10 Running mean = 1 Max distance = 0") +
   theme_bw() +
   theme(legend.position = "none",
         panel.grid.minor.y = element_blank(),
         panel.grid.major.y = element_blank(),
         panel.grid.minor.x = element_blank(),
         panel.grid.major.x = element_blank()) #,
         #text = element_text(size=30))#,
         #axis.text.y=element_blank(),
         #axis.ticks.y=element_blank(),
         #axis.text=element_text(colour="black"),
         #panel.spacing.y = unit(0.7, "lines"))# +
  #theme(plot.margin = margin(0.5,1,0.5,0.5, "cm"))

```

```{r}
THUL01Anno2 = read.csv("U:/BITCue/Projekter/TrackingFlowers/data/annotations/2020_05_17_NorwayAnnotations_NARS-04_IndividualAnnotations_FRCNN_Metrics.csv", header = T)

THUL01Tracked2 = read.csv("U:/BITCue/Projekter/TrackingFlowers/testResults/MMFix/_parameterTest_NARS-04_3/parameterTest_NARS-04_maxDisap_10_runMean_1_maxDist_0.csv")

m2 = merge(THUL01Anno2, THUL01Tracked2, by = c("filename","x_min", "y_min", "x_max", "y_max"))

m2 = merge(m2, meta_2018, by = "filename")

m2 %>% 
  filter(id_gt == "Flower_15" | id_gt == "Flower_28"| id_gt == "Flower_25") %>% 
ggplot() + 
  geom_point(aes(x = frame, y = id_gt, colour = as.factor(objectID)), size = 4, shape = 15)

m2p = m2 %>% 
  filter(id_gt == "Flower_15" | id_gt == "Flower_28"| id_gt == "Flower_25") %>% 
  ggplot() +
   geom_line(mapping = aes(DOY,id_gt, group=id_gt, colour = as.factor(objectID)), size = 4) +
  ylab("") +
  ggtitle("Max gap = 10 Running mean = 1 Max distance = 0") +
   theme_bw() +
   theme(legend.position = "none",
         panel.grid.minor.y = element_blank(),
         panel.grid.major.y = element_blank(),
         panel.grid.minor.x = element_blank(),
         panel.grid.major.x = element_blank())#,
         #text = element_text(size=30))#,
         #axis.text.y=element_blank(),
         #axis.ticks.y=element_blank(),
         #axis.text=element_text(colour="black"),
         #panel.spacing.y = unit(0.7, "lines"))# +
  #theme(plot.margin = margin(0.5,1,0.5,0.5, "cm"))

```

```{r}

m$type = "Max gap = 10, Running mean = 20, Max distance = 500"
m2$type = "Max gap = 10, Running mean = 1, Max distance = 0"

m3 = bind_rows(m, m2)

m$c = "NA"

m3$c[m3$objectID == "17"] = "0"
m3$c[m3$objectID == "15"] = "0"

m3$c[m3$objectID == "13"] = "1"
m3$c[m3$objectID == "10"] = "2"

m3$c[m3$objectID == "12"] = "1"
m3$c[m3$objectID == "9"] = "2"

m3$c[m3$objectID == "27"] = "3"
m3$c[m3$objectID == "15"] = "0"

errorPlot = m3 %>% 
  filter(id_gt == "Flower_15" | id_gt == "Flower_28"| id_gt == "Flower_25") %>% 
  ggplot() +
   geom_line(mapping = aes(DOY,id_gt, group=id_gt, colour = c), size = 4) +
  ylab("Ground truth flower ID") +
  facet_wrap(vars(type), ncol = 1) +
  #ggtitle("B") +
   theme_bw() +
   theme(legend.position = "none",
         panel.grid.minor.y = element_blank(),
         panel.grid.major.y = element_blank(),
         panel.grid.minor.x = element_blank(),
         panel.grid.major.x = element_blank(),
         text = element_text(size=18))

errorPlot

ggsave("aplot.png", errorPlot, height = 4, width = 10, dpi = 800)

```

```{r}
#unique(m$objectID[m$id_gt == "Flower_32"])
#unique(m$id_gt[m$objectID == "9"])

#m %>% 
#ggplot() + 
#  geom_point(aes(x = x_c, y = y_c, colour = as.factor(id_gt)), size = 1)

#m %>% 
#ggplot() + 
#  geom_point(aes(x = x_c, y = y_c, colour = as.factor(objectID)), size = 1)

#m %>% 
  #group_by(id_gt, objectID) %>% 
  #mutate(gr = n()) %>% 
#ggplot() + 
#  geom_point(aes(x = frame, y = id_gt, colour = as.factor(objectID)), size = 4, shape = 15)

#m %>% 
# ggplot() + 
 #  geom_point(aes(x = frame, y = as.factor(objectID), colour = id_gt), size = 4, shape = 15)

#m %>% 
#  filter(id_gt == "Flower_15" | id_gt == "Flower_28"| id_gt == "Flower_25") %>% 
#ggplot() + 
#  geom_point(aes(x = frame, y = as.factor(objectID), colour = id_gt), size = 4, shape = #15)

```

```{r}
NARS13Anno = read.csv("U:/BITCue/Projekter/TrackingFlowers/data/annotations/2020_04_30_NorwayAnnotations_NARS-13_IndividualAnnotations_FRCNN_Metrics.csv", header = T)

nars13points = NARS13Anno %>% 
  mutate(x_c = (x_min + x_max)/2, y_c = (y_min + y_max)/2) %>% 
ggplot() + 
  geom_point(aes(x = x_c, y = y_c, colour = as.factor(id_gt)), size = 1, shape = 4, stroke = 1) +
  labs(x = NULL, y = NULL, title = NULL) +
  scale_x_continuous(limits = c(0,6080), expand = c(0, 0)) + 
  scale_y_reverse(limits = c(3420, 0), expand = c(0, 0)) +
   theme_bw() +
   theme(legend.position = "none",
    axis.ticks.length=unit(0, "pt"),
    #panel.background = element_rect(fill = "transparent", colour = NA),
    #plot.background = element_rect(fill = "transparent", colour = NA),
    panel.grid = element_blank())#,
    #panel.border = element_blank())

nars13points
ggsave("nars13points.png", nars13points, width = 10, height = 5.64, dpi = 800)

```

### 2022-05-25 text/ManuscriptDraft.Rmd
Manuscript revision 3

@@
 
+FIX FIGURE AND TABLE REFERENCES!
+
 Title:
@@
 #
+
 Authors:
@@
 #
+
 Author affiliations:
 
-1: Department of Ecoscience and Arctic Research Centre, Aarhus University, C.F. Møllers Allé 4-8, 8000 Aarhus C, Denmark  
-2: Department of Electrical and Computer Engineering – Signal Processing and Machine Learning, Aarhus University, Finlandsgade 22, 8200 Aarhus N, Denmark
+1: Department of Ecoscience and Arctic Research Centre, Aarhus University, C.F. Møllers Allé 8, 8000 Aarhus C, Denmark\
+2: Department of Electrical and Computer Engineering -- Signal Processing and Machine Learning, Aarhus University, Finlandsgade 22, 8200 Aarhus N, Denmark
 
 #
+
 Corresponding author contact information:
 
-Hjalte Mads Rosenstand Mann  
-E-mail: mann@ecos.au.dk  
+Hjalte Mads Rosenstand Mann\
+E-mail: [mann\@ecos.au.dk](mailto:mann@ecos.au.dk){.email}\
 Phone: +45 31178585
@@
 #
+
 ORCID IDs:
@@
 #
+
 Keywords:
@@
 \newpage
+
 # Abstract
 
-A fundamental problem in pollination ecology is that plant reproductive consequences of pollinator visits to a flower throughout its lifespan is very difficult to track.
-Deciphering the consequences of e.g.
-shifts in flower visitation rates to plant reproductive success and to identify mismatches between interacting and interdependent species such as insect-pollinated plants and pollinators throughout the season requires full-season phenology data, preferably at the level of individuals.
-Often simple statistics are used to describe the flowering phenology of a population of plants, e.g.
-onset or peak of flowering, which may fail to reveal the true dynamics in flowering and reproduction.
-Here we show that image-based monitoring of field plots at very high temporal resolution can return information on flowering phenology at the level of individuals.
-Further, we present a framework for automatically tracking, filtering, and evaluating flowers in time-lapse image series.
-We compare the results of automatic tracking with manual tracking.
-We show that performance achieved by automatic flower tracking can be reliable in obtaining individual flower phenology.
-Our method could facilitate fine-grained studies on plant reproductive success and is an important tool for estimating the sensitivity of plant-pollinator interactions to climate change.
-Supporting this paper is a publicly available web application through which users can apply the tracking and filtering methods on their own data.
+Understanding how variation in the timing of flowering, the timing and frequency of pollinator visits, and the abundance and diversity of visitors affect plant reproduction is a fundamental challenge in pollination ecology.
+Since such interactions concern individual flowers, deciphering the consequences of e.g.
+shifts in the frequency of flower visits for plant reproductive success should ideally be based on observations of individual flowers throughout their entire flowering period.
+Due to the overwhelming logistical challenge of collecting such data by direct observation, it has generally been infeasible to collect.
+Here, we show that image-based monitoring of field plots at very high temporal resolution can return information on flowering phenology at the level of individual flowers.
+Further, we present a framework for automatically tracking flowers in time-lapse image series, evaluating tracking accuracy, and filtering out unreliable tracks.
+Our results on individual flower phenology confirm that traditional weekly observations of flower plots may be insufficient to capture details in flowering dynamics.
+We compare the results of our automated method with manual tracking and show that the performance achieved can be reliable in obtaining individual flower phenology.
+Our method could facilitate fine-grained studies on plant reproductive success and could be an important tool for estimating the sensitivity of plant-pollinator interactions to climate change.
 
@@
 
-For the individual plant, timing of flowering is of utmost importance.
-Precocious flowering means that the plant has failed to exploit the whole temporal window for accumulating resources before allocating energy to flowering.
-Oppositely, flowering too late limits the time for reproduction before the end of the growing season [@ELZINGA2007].
-Both precocious and late flowering increase the risk of frost damage to the flowers.
-Further, flowering may need to be synchronous with pollinator activity for successful reproduction.
-Flowering phenology may plastically change in response to abiotic cues in the environment, such as timing of spring, temperature, and photoperiod, but flowering phenology is partly heritable and shaped by selective forces from the abiotic and biotic environment and in any case sensitive to climate change.
-
-Pollen limitation is a common factor limiting the reproductive success of flowering plants, but the ecological mechanisms behind pollen limitation is poorly understood, at least partly because it requires information on plant reproductive success in relation to pollinator visitation rates which is difficult to acquire in a natural setting and to a satisfactory extent.
-For example, pollinator visits can vary greatly in frequency during the flowering season and the degree of pollen limitation can vary even between populations in close vicinity to each other [@larson1999; @ashman2004].
-
-Interactions require abundance of individuals in life-history stages in which interactions occur.
-Thus, interaction strength depends on overlap in phenology between individuals in these stages.
-Ideally, investigation of the sensitivity of interaction strength to dynamics in phenology requires quantification of phenology through the season for the species in question in order to asses both their phenology and abundance [@inouye2019; @linden2018].
-If interacting and interdependent species, e.g.
-plants and pollinators, react to different environmental cues, or react differently to the same cues, their phenology may be desynchronized and temporal mismatches occur.
-Changes in flower phenology as a consequence of climate change may indirectly affect the abundance of flower-dependent species [@hoye2013] and these effects can be stronger than the direct effects of climate change [@ogilvie2017].
-However, the relation between flower phenology and reproductive success is multifaceted and not simply a question of temporal overlap with pollinators.
-For example, @bolmgren2015 found that for an insect-pollinated plant species early flowering was associated with increased fruit-set but late flowering was associated with higher seed mass.
-Further, periods of resource scarcity through the season may be more important than the floral sum of the season, i.e.
-dynamics in resource availability above a level of sufficiency may have a weak effect on pollinator abundance while the time above this threshold has a stronger effect [@ogilvie2017].
+The Arctic is characterized by short growing seasons and for the individual plant, timing of flowering is of critical importance.
+Flowering too early increases the risk of frost damage to the flowers, while flowering too late limits the time for seed maturation and dispersal before the end of the growing season [@pardee2019; @ELZINGA2007; @inouye2008; @iler2021].
+Further, for insect-pollinated plants, flowering must be synchronous with pollinator activity for successful reproduction.
+Both flower and insect phenology is partly driven by abiotic cues such as timing of spring and temperature and can thus be sensitive to climate change.
+Changes in flower phenology as a consequence of climate change may indirectly affect the abundance of flower-dependent insect species [@hoye2013] and these effects can be stronger than the direct effects of climate change [@ogilvie2017].
+However, the relation between flower phenology and reproductive success is not well understood.
+For example, @bolmgren2015 found that for an insect-pollinated plant species, early flowering was associated with increased fruit-set but late flowering was associated with higher seed mass.
+If we are to understand the intricate dynamics of plant-pollinator interactions, tools that aid efficient and comprehensive data collection are essential.
 
@@
 Consequently, flowering phenology of a population is often characterized simply by the date for onset of flowering, typically derived from weekly observations of sample plots [@Prevey2021].
-Such inaccurate estimates may fail to reveal important dynamics in flowering phenology through the season.
+Such crude estimates may fail to reveal important dynamics in flowering phenology through the season.
 For example, different phenological factors, such as onset, peak, and end of flowering may respond differently to climate change [@cook2012] and estimation of the degree of mismatch between species may thus be sensitive to the choice of metric [@inouye2019].
-A lot remains to be discovered about when and how the dynamics of pollination affects the reproductive success of flowering plants.
-High frequency observations of flower visits coupled with information on phenology and reproductive success at the level of individuals is required to unravel the relationship between pollinator visitation and plant reproductive success and the sensitivity of this interplay to climate.
-
-The ideal metric for phenology depends on the topic of the question, but often the choice of onset of flowering as the variable representing phenology is based on data availability and not because it is considered the ideal metric.
-First events, such as onset of flowering, may in some cases be particularly ill suited for estimating sensitivity to climate change [@miller2008].
-Onset of flowering represents the very extreme of the phenological distribution of a population and may be confounded with e.g.
+Of course, the optimal metric for phenology depends on the research question, but often the choice of recording first flowering dates to represent flowering phenology is pragmatic rather than ideal.
+In fact, first flowering dates may be particularly ill suited for estimating sensitivity to climate change [@miller2008] since they represent the very extreme of the phenological distribution of a population and may be confounded with e.g.
 the size of the population considered.
-Further, for many questions, the phenology distribution of a population is required, for example to quantify resource availability through the season.
-However, even detailed information on phenology at the level of a population is still an aggregation of the phenological distributions of the individual members of the population for which important variation may be masked [@inouye2019].
-Having at hand the full distribution of flowering phenology at the level of individuals allows to extract relevant characteristics of the distribution and test their individual sensitivities towards environmental factors, but acquiring such data requires efficient and preferably automatized methods.
-Automatic image-based monitoring of flowering phenology can return phenology data for specific species at very high temporal resolution (Mann et al., in review (Note: I'll update this if the paper gets accepted or remove it if not)), but phenological responses at the individual level may be indiscernible regardless of the temporal resolution of the data at population level.
-For example, a shortening of individual flower longevity may not be directly obvious at the population level.
-In fact, many research questions can only be explored on the basis of individual phenology data.
-For example, investigating the association between reproductive success and timing of flowering or flower longevity requires phenology data at the level of individuals.
-Similarly so for investigating whether flower visitation rates and/or reproductive success depends on the timing of flowering for the individual flower.
-Efficient and precise methods for deriving full seasonal distributions of phenology at the level of individuals is much needed.
+Rather, an accurate estimate of the phenological distribution of a population over the full season would be a preferred measure.
+Such data would allow for, for example, accurate quantification of resource availability through the season and temporal mismatches between species [@ramakers2020].
+However, even detailed information on phenology of a population is still an aggregation of the phenology of the individual members of the population for which important variation may be masked [@inouye2019].
+With the full distribution of flowering phenology at the level of individuals, it is possible to extract relevant characteristics of the distribution and test their individual sensitivities towards environmental factors, but acquiring such data requires efficient and preferably automated methods.
 
 Cameras have become an ubiquitous tool for non-invasive and efficient monitoring of natural processes [@howe2017; @glover2019], but the implementation of image-based collection of data has outpaced the development of reliable and easily implementable methods for processing and analysing the big data output from this transformation [@weinstein2018; @farley2018].
-Computer vision and deep learning tools and being developed and applied for automatic processing and extraction of relevant data from these very large image sets [@willi2019;  @tabak2019] and increasingly for challenging tasks in entomology [@hoye2021], such as for fine-grained species classification [@spiesman2021] and monitoring of pollination events by detecting, identifying, and counting insects [@bjerge2021].
-Collectively, the development and adoption of modern technology allows for collection and processing of ecological data at unprecedented scales.
+Computer vision and deep learning tools are being developed and applied for automatic processing and extraction of relevant data from these very large image sets [@willi2019; @tabak2019; @hoye2021].
+Indeed, automatic image-based monitoring of flowering phenology coupled with deep learning has facilitated collection of phenology data for specific species at very high temporal resolution [@mann2022], yet, phenological responses at the individual level may be indiscernible regardless of the temporal resolution of the data at population level.
+For example, a shortening of individual flower longevity may not be directly obvious at the population level.
+In fact, many research questions can only be directly explored on the basis of phenology data at the level of individuals.
+For example, investigating the association between reproductive success and timing of flowering and/or flower longevity requires such data.
+Individual level data would also aid the understanding of whether flower visitation rates and/or reproductive success depends on the timing of flowering for the individual flower.
+Thus, efficient and precise methods for deriving full seasonal distributions of phenology at the level of individuals are much desired.
 
-Here, we use stand-alone time-lapse cameras to monitor the phenology of two Arctic flowering plant species Dryas integrifolia and D.
-octopetala.
-We show that information on phenology at the level of individuals can be derived from image-based monitoring of flower phenology.
-We present and evaluate an automatic flower tracking and filtering algorithm.
-As the data extracted from time-lapse series of flower plots through season violate the assumptions of standard tracking mechanism, we base the track assignment on the minimum distance between centroids of bounding boxes and introduce three user defined parameters that can help optimise tracking performance.
+Here, we use time-lapse cameras to monitor the phenology of two widespread Arctic flowering plant species *Dryas integrifolia* and *D.
+octopetala*.
+We show that information on phenology at the level of individual flowers can be derived from image-based monitoring of flower plots.
+We present and evaluate a simple multiple object tracking algorithm for automatic tracking of individual flowers in the image series.
+The data extracted from the images violate common assumptions of standard tracking methods and tracking individual flowers through the season therefore represents a special case.
+For example, since flower movement is constrained by the stalk, the future location of a flower-head cannot be predicted based on previous trajectory of movement.
+Further, as individual flowers closely resemble each other, tracking cannot be based on recognition of individuals.
+Our flower tracking algorithm assigns individual flowers to tracks based on the minimum distance between centroids of bounding boxes.
+We apply three user-defined parameters to help optimise tracking performance.
 
 For complex scenes with many flowers in close vicinity to each other there is a substantial risk of tracking errors regardless of the tracking method being used.
-In such cases it is preferable to extract individual high-confidence tracks from each series and ignore the remaining tracks.
-For this, we suggest a conservative filtering approach based on the density of tracks within the frame.
-The approach may remove correct tracks, but the tracks that remain will have a low risk of tracking errors, which allows for confident upscaling of the method to many series.
-
-Our method for tracking individual flowers facilitates the possibility of assigning indicators of reproductive success and factors driving variation in seed set to the individual.
-For instance, it could be explored whether seed set is affected by timing and length of flowering and/or abiotic conditions such as frost events.
-Hereby, the long-standing question of what drives variation in reproductive success of individual flowers could be explored.
-By simultaneously tracking flower visits, these could be assigned to the individual flower and visitation rates per flower could be calculated per flower and related to reproductive success.
-Further, any information of taxonomic grouping could facilitate comparative studies on pollination effectiveness between taxonomic groups of pollinators.
-The possibility of automatically tracking phenology at the level of individuals will facilitate investigation of a set of yet untestable research questions and is an important contribution to the shift towards efficient and detailed monitoring of phenology.
+In such cases, it may still be possible to reliably track a subset of the flowers, which are spatially isolated in the frames.
+For this, we suggest a conservative filtering approach based on the distance between tracks within the frame.
 
 # Material and methods
+
 ## Study site and species
-This study was based on five time-lapse image-series collected at the following three Arctic sites (sampling years given in brackets): Narsarsuaq (NARS), South Greenland (2018, 2019), Thule (THUL), North-West Greenland (2018), Ny-Ålesund (NYAA), Svalbard (2019).
-The images were collected using time-lapse cameras (Moultrie Wingscapes TimelapseCam Pro, Moultrie Products, Birmingham Alabama, USA) permanently positioned approx.
+
+This study was based on five full-season image series collected at the following three Arctic sites (site abbreviation and sampling years in parentheses): Narsarsuaq (NARS), South Greenland (2018, 2019), Thule (THUL), North-West Greenland (2018), Ny-Ålesund (NYAA), Svalbard (2019).
+The images were captured using time-lapse cameras (Moultrie Wingscapes TimelapseCam Pro, Moultrie Products, Birmingham Alabama, USA) permanently positioned approx.
 60 cm above areas with either *Dryas integrifolia* in Narsarsuaq and Thule or *D.
@@
 
-#
 ## The image series and flower annotations
+
 Images were collected throughout the full growing seasons.
-While the original time-lapse frequency was higher, the basis for this study was 1-hr interval series.
-We limited the 1-hr image-series to the include only the flowering season and randomly randomly sampled images from each series.
-The number of images for each series after sampling is given in table 1.
-All flowers in bloom in each of these images were manually annotated using the rectangular bounding box tool in the VIA VGG annotation software [@dutta2019] and each individual flower was assigned a unique ID.
+While the original time-lapse frequency was higher, the basis for this study was 1-hour interval series.
+We limited these series to include only the flowering season and randomly sampled images from each series.
+These image sets were used for the study.
+The number of images for each series after sampling is given in Table 1.
+All flowers in bloom within each image were manually annotated using the rectangular bounding box tool in the VIA VGG annotation software [@dutta2019] and each individual flower was assigned a unique ID that was maintained throughout its lifetime.
 These annotations constitute our ground truth tracks.
-The series varied substantially in the number of flowers they contained (table 1).
-Examples of annotated images from the 2019 NARS C series at three different times in the flowering season are given in fig.
-1.
-Figure 2 shows the time of flowering for each individual flower and the number of days the flower was present in the images.
+The series varied substantially in the number of flowers they contained (Table 1).
+Examples of annotated images from the 2019 NARS C series at three different times in the flowering season are given in Figure 1.
+To demonstrate how individual flower tracking can result in detailed representations of flower phenology across the season, we use the ground truth annotations to show the duration of flowering for each individual flower across the season for each series (fig.
+2).
 
-**Table 1:** Number of images and number of flowers for each of the five time-lapse image-series.
+**Table 1:** Number of images and number of individual flowers for each of the five time-lapse image series.
 
 #
-![](../figures/CameraOverview.png){ width=40% }
+
+![](../figures/CameraOverview.png){width="40%"}
 
 #
-![](../figures/AnnotationExamples_3.png){ width=100% }
+
+![](../figures/AnnotationExamples_v2_4.png){width="100%"}
 
 **Figure 1:** Examples of annotated images for 2019 NARS C.
-a: Onset of flowering at DOY 145; b: Peak flowering (52 annotated flowers) 6 days after onset (DOY 151); c: 6 days after peak flowering (DOY 157).
-Flower bounding boxes are colour coded by flower ID.
+A: Onset of flowering at DOY 145.
+B: Peak flowering at DOY 151 (52 annotated flowers).
+C: 6 days after peak flowering (DOY 157).
+Flower bounding boxes are colour-coded by flower ID.
+D: Centroids of all flower annotations in the series, coloured by flower ID, showing how flowers may overlap in space.
 
-#
-![](../figures/OnsetVSLength.png){ width=100% }  
-**Figure 2:** Onset of flowering (DOY) and flower lifespan (days) for each individual flower in each of the five image-series.
-
-#
 ## Automatic flower tracking
@@
 We built a framework for tracking, filtering, and evaluating tracking of objects in time-lapse image series.
-The tracking algorithm tracks objects based on distances between centroids of bounding boxes.
+The tracking algorithm tracks objects based on distances between centroids of bounding boxes, i.e.
+objects with minimum displacement between frames are associated.
 The algorithm has a set of user-adjusted parameters that can optimize tracking accuracy.
 The parameters are particularly relevant for optimal tracking of objects that are constrained to a specific area such as flowers.
-It is important to note, however, that the tracking algorithm could be used to track any objects.
-The algorithm can be applied both offline (on a set of detections/annotations that have already been produced) or online (real-time tracking frame per frame).
-The speed of the tracking algorithm depends on the computational power available as well as the number of objects that are being tracked.
-The method is fast, however.
-Tracking of a series containing 85 objects ran at ~0.02 seconds per frame on a standard laptop.
+However, the tracking algorithm could be used to track any objects.
+The speed of the tracking algorithm depends on the computational power available as well as the number of objects that are being tracked, but the time for computation will be negligible in most use-cases.
+Tracking of a series containing 85 objects ran at \~0.02 seconds per frame on a standard laptop.
+This, together with the fact that the algorithm handles frames sequentially, means that it can be applied both offline (on a set of detections/annotations that have already been produced) or online (real-time tracking frame per frame).
 
-Several issues complicate the task of tracking individual flowers through time-lapse image-series.
-First, stand-alone monitoring of the relatively slow process of flower development through full growing seasons (several months) requires only low frame rates.
-Here we used several images per day.
-These frame rates capture the phenology of the flowers at very high detail, but fail to capture the full movement of flowers.
+Several issues complicate the task of tracking individual flowers through time-lapse image series.
+First, as flowers move, they may temporarily occlude the view of other flowers and flowers close to the edge of the frame may move in and out of view between images.
+This can cause a track to be lost and a new track to be erroneously initiated when the flower reappears (fig.
+1A), a tracking error known as fragmentation.
+However, if a flower reappears in the same area as a flower is already being tracked after disappearing in only a few frames, it is a reasonable assumption that it is the same individual and not that the old flower wilted/disappeared and a new one developed.
+The parameter **max gap** sets the number of frames a track can be lost before a new track is initiated for a flower appearing in the same area.
+If a given flower has not been detected in a number of frames less than or equal to the max gap size, the track will not be lost.
+Concurrently, this can deal with potential false negatives, as a track that is missing a bounding box will not be lost.
+
+Second, a standard centroid tracking algorithm will associate a point with a track based on minimum distance only, disregarding the absolute distance between centroids, which may lead to tracking errors (fig.
+1B).
+However, as flowers are constrained in their displacement, we can assume a maximum distance that will occur between points and force the initiation of tracks for points that exceed this threshold.
+We set this threshold with the parameter **max distance,** measured as the euclidean distance in number of pixels between objects in neighbouring frames.
+
+Third, with a high number of flowers in close vicinity to each other, there's a substantial risk of identity switching between flowers as they move among each other (fig.
+1C).
+Stand-alone image-based monitoring of the relatively slow process of flower development through full growing seasons (several months) is done with low frame rate, limited by power consumption and storage capacity both on and off the camera.
+It is simply not feasible to collect high frame-rate images/video across full seasons.
+Here, we used several images per day and while these images capture the phenology of the flowers at very high detail, they fail to capture the full movement of flowers.
 For example, as the wind shifts, the flower heads change direction, which in the image series happens instantaneously (i.e.
 between two consecutive frames) (fig.
-3a).
-However, as they are constrained by their stalk, there is a limit to the distance they can move.
-With the tracking parameter **running mean**, we have the option to base the tracking on the distance between a point in the current frame and the running mean of the positions of a number of previous points in a track, taking advantage of the fact that flower move around a central point.
+3C).
+The fact that the frame rate is insufficient to capture smooth movement means that bounding boxes around a single individual in two consecutive frames cannot be associated based on overlap (e.g.
+intersection over union) as there often is no overlap.
+Additionally, as the flowers are constrained by their stalk, there is a limit to the distance they can move, which complicates prediction of future location of an individual based on its previous trajectory.
+Here, we introduce the tracking parameter **running mean**, with which we can base tracking on the distance between a point in the current frame and the running mean of the positions of a number of previous points in a track, taking advantage of the fact that a flower moves around a central point.
 Increasing the value for this parameter decreases the weighting of movement between individual frames.
-
-Second, as flowers move, they may temporarily occlude the view of other flowers and flowers close to the edge of the frame may move in and out of view.
-This can cause a track to be lost and a new track erroneously being initiated when the flower reappears (fig.
-3b).
-However, if a flower reappears in the same area as a flower is already being tracked after disappearing in a few frames, it is reasonable to assume that it is the same individual and not that the old flower wilted/disappeared and a new one developed.
-The parameter **max gap size** sets the number of frames a track can be lost before a new track is initiated for points appearing in the same area.
-If a given flower has not been detected in a few frames, the track will not be lost.
-Concurrently, this deals with potential false negatives.
-The counter for number of disappeared frames for a track is reset when a new point is associated with the track within the threshold.
-
-Third, a standard centroid tracking algorithm will associate a point with a tracking based on minimum distance only, disregarding the absolute distance.
-As flowers are constrained in the movement, we can assume a maximum distance that will occur between points and force the initiation of tracks for points that exceed this threshold.
-We set this threshold with the parameter **max distance** measured as the euclidean distance in number of pixels.
+Setting the parameter to 1 means that tracking is based on the distance to centroid in the previous frame only.
 
 #
-![](../figures/HanddrawnFiguresIssues.png){ width=100% }
 
-**Figure 3:** a: Centroid tracking may produce erroneous associations when objects move between frames.
-Here the track initiated on the top left flower in frame t-1 would switch to the bottom right flower in frame t b: Tracking errors may occur when objects temporarily disappear from the frame.
-Here the top flower moves out of frame and the bottom flower would be assigned to the track of the top flower in frame t-1 Red shows centroid of bounding box in the current frame and blue shows centroid points from the previous frame.
+![](../figures/ErrorFigure2.png){width="100%"}
 
-#
+**Figure 1:** Centroid tracking may produce erroneous associations when objects move or disappear between frames.
+A: A flower temporarily disappears from view.
+If gaps in tracks are not allowed during tracking, a new track will erroneously be initiated when the flower reappears.
+B: A flower disappears from view and another reappears between two frames.
+If no threshold for maximum distance between an object and a track is set, the reappearing flower will be linked to the track of the disappearing flower based on shortest distance despite the absolute distance being large.
+C: A flower is linked to the track of another flower.
+As flowers move around a central point, basing tracking on a running mean of previous positions may be beneficial.
+Dots show position of flower in previous frame.
+Crosses show running mean of positions in x \> 1 number of previous frames.
+Dots and crosses coloured according to track identity.
+
 ### Input data for the tracking algorithm
 
-The input data for the tracking algorithm are bounding box coordinates for each object in each frame given in the following format: frame number, x-min, y-min, x-max, y-max.
+The input data for the tracking algorithm are bounding box coordinates for each object in each frame given in a commonly used format [@everingham2010]: frame number, x-min, y-min, x-max, y-max.
 Here, x-min, y-min is the top left corner of the bounding box and x-max, y-max is the bottom right corner.
@@
 Calculating tracking accuracy requires a ground truth ID for each bounding box which can be given as the sixth column.
-As this data structure is outputted from both manual object annotations and object detection methods using CNNs, the method can be applied on both manually and automatically derived data.
-Further, as the algorithm handles frames sequentially, it can be applied both offline and online.
+As this data returned from both manual object annotations and object detection methods using automatic object detection methods based on deep learning, the method can be applied on both manually and automatically derived data.
 
 #
+
 ### Identifying optimal user parameters
 
-To explore the effect of the user parameters and to identify the optimal combination of parameters for our case of tracking flowers, we applied the tracking algorithm on each of the five image-series with every combination of a range of values for each parameter (3179 combinations): Values for max gap size and running mean were 0-160 with a step size of 10 and for max distance 0-1000 with a step size of 100.
+To explore the effect of the three tracking parameters and to identify the optimal combination of parameters for our case of tracking flowers, we applied the tracking algorithm on each of the five image series with every combination of a range of values for each parameter (3179 combinations): 0-160 with a step size of 10 for max gap; 1-160 with a step size of 10 for running mean; 0-1000 with a step size of 100 for max distance.
 Note that max distance set to zero ignores the parameter altogether.
-We identified the setting(s) that returned the lowest number of track mismatches and compare the tracking results between optimal settings and all parameters set to zero.
+We identified the setting(s) that returned the most accurate tracks for each series and compare the tracking results between optimal settings and all parameters ignored (i.e.
+max gap = 0, running mean = 1, max distance = 0).
 
-#
-## Filtering tracks
-
-When deploying the automatic tracking algorithm on naive data without ground truth tracks, it is not possible to manually filter for correct tracks.
-Therefore, we present a conservative filtering method that extracts the most trustworthy tracks from a scene.
-
-We base our filtering on the distance between geometry centroids the tracks.
-The geometries are derived in the following way: For tracks consisting of only a single point, the coordinates of the point are used as the centroid.
-For tracks consisting of two points, we establish the straight line between the points and calculate the centroid of the line.
-For tracks consisting of three points, we establish the triangle from the points and calculate the centroid of the triangle.
-For tracks consisting of four points or more, we calculate the convex hull of all the points included in the track and derive the polygon from the vertices of the convex hull.
-Hereafter we calculate the centroid of the polygon.
-For tracks that contain more than two points but where the points a collinear, we establish the line through the points and calculate the centroid of the line.
-
-The filtering is done on the geometry centroids to remove tracks in close vicinity to each other as these have a high risk of tracking mismatches.
-The distances between all centroid tracks are calculated (euclidean distance in number of pixels) and only tracks that have no neighbouring centroid within the filtering distance threshold are kept.
-
-We demonstrate the filtering method on the three series for which our tracking algorithm did not return perfect results.
-For this, we chose a fixed combination of the tracking algorithm parameters across all three series (max gap size: 10; running mean: 10; max distance: 300).
-We then applied the filtering method on the three series with a range of values for the filtering distance (50-500, step size of 50) and identified the value that returned the highest sum of tracks across the three series at the lowest number of mismatches.
-We present the tracking accuracy before filtering and on the remaining tracks after filtering.
-We underline that the results presented are thus highly conservative.
-I.e., the approach could be optimised to extract a higher number of tracks while maintaining the number of mismatches.
-For example, basing the filtering on the best performing tracking parameters for each series specifically would improve the output of the tracking algorithm.
-Similarly, the results could be improved by fine-tuning the filtering distance threshold for each series individually.
-However, as our goal here was to demonstrate that our tracking and filtering method can be applied in a naive setting with fixed values for each parameter we have not included those results.
-
-#
 ## Evaluating tracking performance
@@
 The optimal way of quantifying tracking performance depends on the goal of the tracking.
-To associate other information obtained in the images to the individual flower, for example flower visits, requires as much as possible of the track to be correct.
-To derive flowering length, in theory just requires correct association between the first and the last image of a flower while intermediate points can be ignored.
-Lastly, in some scenarios one may be interested in the number of flowers that existed in a plot, in which case the number of tracks obtained by automatic tracking should closely resemble the actual number of individuals in the series.
+For example, in our case of tracking flowers, associating information obtained in the images to the individual flower, for example flower visits, would require the track to be as correct as possible (i.e.
+identity shifts should be minimized).
+On the other hand, to derive flowering length in practice just requires correct association between the first and the last image of a flower while intermediate points can be ignored.
+Here, track fragmentation can lead to underestimation of flowering duration (and overestimation of number of flowers) while track merging can lead to overestimation of flowering duration (and underestimation of number of flowers).
+In some scenarios one may be interested in the number of individual flowers that existed in a plot, in which case the number of tracks obtained by automatic tracking should closely resemble the actual number of individuals in the series.
 
-The commonly used multiple object tracking accuracy (MOTA) score quantifies tracking performance based on counts of tracking mismatches [@bernardin2008].
-Mismatches occur when objects swap track identity because they are in close vicinity to each other or when an object periodically disappears and is assigned a new track identity when it reappears.
-Only the shifts in tracking identity are counted as mismatches while the number of points assigned to each track is not considered.
-We use three evaluation metrics: We calculate the number of mismatches and derive the MOTA score (number of mismatches in relation to the total number of points) to evaluate performance of our tracking and filtering algorithms.
+Here, we apply the commonly used multiple object tracking accuracy (MOTA) score which quantifies tracking performance based on tracking mismatches derived by counting the number of identity shifts within each track [@bernardin2008].
+Mismatches occur when objects swap track identity because they are in close vicinity of each other or when an object periodically disappears and is assigned a new track identity when it reappears.
+Only the shifts in tracking identity are counted as mismatches in the MOTA score, while the number of points assigned to each track is not considered.
+For example, a ground truth track that was assigned one ID for the first half of its lifetime and another for the remaining half would only constitute one error.
+
+$$
+MOTA = 1 -  \frac{\sum{_t(fn_t + fp_t + mm_t)}}{\sum{_tg_t}},
+$$
+
+where fn, fp, and mm are the number of false negatives, false positives, and track mismatches, respectively, for frame t divided by the total number of ground truth annotations across all frames.
+The MOTA score ranges from -Inf to 1 with 1 being perfect detection and tracking.
+As tracking accuracy is commonly evaluated between manual annotations and detections, produced, for example, by a convolutional neural network (CNN), the MOTA score included false positive and false negatives as errors, but in our case, since we are tracking on ground truth data and our annotations and detections therefore are practically identical, these variables can be ignored.
+
+To evaluate tracking accuracy of our algorithm, we calculate the total number of mismatches as well as the MOTA score (number of mismatches in relation to the total number of points).
+Additionally, we derive the number of tracks that were tracked with more than 80% accuracy.
 Finally, we compare the number of tracks identified by the automatic tracking with the true number of flowers in a series.
 These should ideally be equal.
-However, a correct number of tracks (i.e.
-number of tracks equal to number of objects) does not necessarily translate to correct tracking, since individual tracks can be erroneously split by the tracking algorithm and several tracks can be combined.
-We present the tracking performance by both the lowest number of mismatches/highest MOTA given overall and for combinations that returned a number of tracks equal to the number of flowers in the series.
+However, we note that a correct number of tracks (i.e.
+number of tracks equal to number of objects) does not necessarily translate to correct tracking, since individual tracks can be erroneously split (fragmentation) by the tracking algorithm while several tracks can be merged into one.
+We present the tracking performance without the use of the three tracking parameters along with the best performance (lowest number of mismatches/highest MOTA), and the best performance where the number of tracks correspond to the number of individual flowers in the series.
+
+## Identifying the most reliably tracked flowers
+
+In order to be able to identify the most reliably tracked flowers without the need for manual annotation, we use a conservative filtering method that extracts the most trustworthy tracks from a scene based on spatial distance between tracks.
+We derive the distance between tracks based on the centroids of points included in a track produced by the tracking algorithm.
+When tracks consist of only a single point, the coordinates of the point is used as the centroid.
+For tracks consisting of two points.
+we use the midpoint of the line between them.
+For tracks consisting of three points, we establish the triangle from the points and use the centroid of the triangle.
+For tracks consisting of four points or more, we calculate the convex hull of all the points included in the track and use the centroid of this polygon.
+For tracks that contain more than two points but where the points are collinear, we use the midpoint of the line through the points.
+We hereafter refer to these centroids as track centroids.
+
+The filtering is done on the centroids to remove tracks in close vicinity to each other as these are assumed to have a high risk of tracking mismatches.
+The distances between all track centroids are calculated (euclidean distance in number of pixels) and only tracks that have no neighbouring centroid within the filtering distance threshold are kept.
+
+We demonstrate the filtering method on the three series for which our tracking algorithm did not return perfect results (i.e.
+automatic tracking identical to ground truth tracks).
+To test the filtering approach, we chose a fixed combination of the tracking algorithm parameters across all three series (max gap size: 10; running mean: 10; max distance: 300).
+We then applied the filtering method on each three series with a range of values for the filtering distance (50-500, step size of 50) and identified the value that returned the highest sum of tracks across the three series at the lowest number of mismatches.
+We present the tracking accuracy before filtering and on the remaining tracks after filtering.
+We emphasize that our results are conservative and that the approach may be optimised to extract a higher number of tracks while maintaining the number of mismatches.
+For example, basing the filtering on the best performing tracking parameters for each series specifically would improve the output of the tracking algorithm.
+Similarly, the results could be improved by fine-tuning the filtering distance threshold for each series individually.
+However, as our goal here was to demonstrate that our tracking and filtering method can be applied in a naive setting with fixed values for each parameter, we have not included those results.
+
+##
+
+# Results
+
+Manual tracking of individual flowers derived from image-based plot monitoring enabled detailed information on flower phenology at the level of individuals with substantial variation in flowering duration between flowers (fig.
+2).
+Further, our tracking algorithm was able to track individual flowers with consistently high accuracy (MOTA $\geq$ 0.981) across all series.
+In all cases, all flowers were tracked with more than 80% accuracy (table 1).
+In fact, two series were tracked without any errors (2018 THUL B and 2018 NARS D).
+Tracking with the three tracking parameters ignored (max gap = 0, max distance = 0, running mean = 1) returned high MOTA scores across series, but for the three series that were not tracked perfectly, utilizing the tracking parameters increased the MOTA score in all cases.
+
+Importantly, the tracking that returned the lowest number of mismatches did not necessarily result in a number of tracks equal to the number of flowers in a plot.
+For example, for the series with the most flowers, 2019 NARS C, one combination of parameters (max gap = 20, max distance = 600, running mean = 1), returned the lowest number of mismatches (31), but the number of tracks were lower than the actual number of flowers (71 vs.
+85).
+One combination of parameters did return a number of tracks equal to number of flowers (max gap = 10, max distance 400, running mean = 10), but also increased the number of mismatches to 91.
+For 2019 NARS D, several combinations of parameters returned a perfect MOTA score, but still some combinations returned a number of tracks lower than the number of flowers.
+This demonstrates the significance of a special case of tracking error, namely merging of tracks, which is not accounted for in the MOTA score.
+In figure 3, we exemplify the two types of errors, identity shifts and merges.
 
 #
-## Tracking and filtering web application
 
-To make the method described here easily accessible for other users, we built a web application through which the trackking and filtering method can be applied on user data.
-The web application is accessible through https://github.com/TECOLOGYxyz/FlowerTracking.
-
-#
-# Results
-
-The results of the tracking parameter test are given in table 1.
-We present the tracking performance without the use of the three tracking parameters along with the best performance (lowest number of mismatches/highest MOTA), and the best performance where the number of tracks correspond to the number of individual flowers in the series.
-When several values for a parameter was included in a result, we present the range of values.
-For two series, 2018 THUL B and 2019 NARS D, the tracking algorithm returned perfect results for several combinations of parameters.
-For the three remaining series, the algorithm returned consistently high MOTA scores (minimum 0.95).
-In all cases, except 2019 NARS D, where the tracking algorithm returned perfect results with all parameters zeroed, did the inclusion of the user set tracking parameters improve the tracking accuracy.
-
-In one case, 2019 NARS C, four different numbers of tracks were returned among the parameter combinations that resulted in the lowest number of mismatches.
-A number of tracks equal to the number of flowers in the series were among these four.
-In anotehr case, 2019 NYAA E, the parameter combination that produced the highest accuracy did not return a number of tracks equal to the number of flowers in the series.
-However, combinations that did produce this number of tracks also returned a higher number of mismatches.
+![](../figures/OnsetVSLength.png){width="100%"}\
+**Figure 2:** Flowering lifespan (days) for each individual flower in each of the five image series plotted against the onset of flowering in days after 1 January (day of the year) based on manual tracking of individual flowers.
 
@@
 Number of tracks in bold shows when the number of tracks equals the number of flowers in the series.
-![](../figures/TrackingScoresv22.png){ width=100% }
 
-`r word_comment("Table 1 doesn't show excactly which combinations of parameters gives the result.
-This could be shown in suppl.
-e.g.
-with a network-like graph, but I am thinking it is overkill and this table is sufficient.", "Table 1")` shows the performance of the tracking algorithm.
+![](../figures/tab_1.png){width="100%"}
 
-The value for filtering distance that returned the highest number of tracks at the lowest number of mismatches was 350 (results for all values given in supp.
-table S1).
-The results of the filtering algorithm on the three series in which our tracking algorithm did not return perfect results are given in table 2 and the filtering process is visualised in fig.
+![](../figures/errorExamples.png){width="100%"}
+
+**Figure 3:** Tracking of three flowers in the 2018 NARS A image series, coloured by ID assigned by the flower tracking algorithm.
+A: Track output from the flower tracking algorithm for three individual flowers using two different combinations of tracking parameters.
+In the top panel (max gap = 10, running mean = 1, max distance = 0), flower 15 and 28 have no identity shifts within each track, but the two tracks have been merged into one (see B).
+Flower 25 has been fragmented into two separate tracks causing one mismatch (see C).
+In the bottom panel, the running mean parameter has been increased to 20 and max distance to 500, which results in correct tracking of all three flowers.
+B: Frame f shows the last frame with flower 15 (blue) annotated because it has wilted.
+Two frames (5 hours) later (f+2), a new flower (flower 28; blue) blooms.
+Since max gap \> 2 and no distance threshold has been set, flower 28 is associated with the tracks of flower 15, because all other flowers are closer to their respective tracks, which causes a merging error.
+Setting a max distance threshold to 500 fixed this error.
+C: A gust of wind causes a displacement of the flowers between two frames, f and f+1, (7 hours apart), which causes an identity shift between flower 25 and flower 29.
+Setting running mean to 20 fixed this error.
+
+With our filtering method, we were able to identify tracks with no mismatch errors in all three series that were not perfectly tracked.
+The parameter value for filtering distance that returned the highest number of tracks at the lowest number of mismatches was 350 (results for all values given in supp.info).
+The results of the filtering algorithm on the three series are given in table 2 and the filtering process is visualised in fig.
 4.
-Our filtering method extracted 28 tracks from the three series (2018 NARS A: 15 tracks, 1 mismatch; 2019 NARS C: 10 tracks, 0 mismatches; 2019 NYAA E: 3 tracks, 0 mismatches) with complex scenes using a fixed filtering distance value of 350 and fixed values for the tracking parameters (10, 10, 300, for running mean, max gap size, and max distance, respectively) with only a single mismatch included in the 28 tracks.
+Our filtering method extracted 28 tracks from the three series (2018 NARS A: 15 tracks; 2019 NARS C: 10 tracks; 2019 NYAA E: 3 tracks) with complex scenes using a fixed filtering distance value of 350 and fixed values for the tracking parameters (10, 300, 10 for max gap size, max distance, and running mean, respectively) without any mismatches included in the 28 tracks.
 
-**Table 2:** Results of the filtering algorithm with a filtering distance value of 350 applied across all three series.
+**Table 2:** Results of the filtering algorithm on tracking results using the parameters 10, 300, 10 for max gap size, max distance, and running mean, respectively.
+The filtering distance value of 350 was used for all three series.
 
-![](../figures/FilterScoresv211.png){ width=100% }
+![](../figures/tableFilterscores.png){width="100%"}
 
 #
-![](../figures/filtering1.png){ width=100% }
-![](../figures/filtering2.png){ width=100% }
-![](../figures/filtering3.png){ width=100% }
 
-**Figure 4:** Track filtering process.
-Row 1, 2, and 3 is 2019 NARS C, 2018 NARS A, and 2019 NYAA E, respectively.
-First column shows the centroid points in the given series, coloured by the track id from the centroid tracking algorithm.
-Second column shows the polygons calculated from the tracks.
-The filtering algorithm with a distance threshold of 350 was applied to the centroids of these polygons.
-Third column shows the results of the filtering where all tracks that did not pass the filtering have been removed.
+![](../figures/filtering1.png){width="100%"}
+
+**Figure 4:** The track filtering process exemplified by the 2019 NARS C series.
+The first column shows the centroid points for all bounding boxes in the given series, coloured by the track id from the centroid tracking algorithm.
+The second column shows the track polygons calculated from the tracks.
+Polygons with a distance below eps = 350 between them have been grouped together.
+The polygons have been coloured according to group.
+Third column shows the results of the filtering where all tracks that were not sufficiently spatially isolated were removed.
 
 # Discussion
+
 ## Tracking
 
-Many tracking algorithms, such as the Simple online and real-time tracking (SORT) [@sort], use Kalman filtering to predict the future location of an object by estimating its velocity [@kalmann].
+In this paper, we have demonstrated the potential of image-based monitoring for collecting information on flower phenology at the level of individuals.
+Our results show substantial variation in flowering duration between individual flowers (fig.
+2).
+In fact, some flower tracks spanned less than a few days.
+As flowers may not be visible within the frame for their full lifespan, it is likely that for some flowers the full flowering has not been captured, although flowering durations of less than 24 hours has been observed from manual examination of the image series (Mann, personal observation).
+
+The automatic tracking and filtering pipeline presented was able to return reliable flower tracks despite settings that violate assumptions of standard tracking algorithms and complex scenes with a high number of objects that are not permanent within the series.
+Our tracking algorithm was capable of consistently accurate flower tracking, despite the difficult setting.
+Even with simple centroid tracking with association based solely on displacement between the current and previous frame, tracking performance was high for all series and perfect for two (table 1).
+Further, with our filtering approach based on distance between tracks, we were able to discard the least reliable tracks and thereby increase the mean tracking accuracy of the remaining flowers to a perfect score (table 2).
+
+Most multiple object tracking literature focus on tracking of humans [@luo2021; @milan2016], and the methods developed here are not necessarily directly transferable to other use-cases such as ours.
+Many tracking algorithms, such as the Simple online and real-time tracking (SORT) [@sort], use Kalman filtering to predict the future location of an object by estimating its velocity from previous frames [@kalmann].
 However, Kalman filtering assumes linear motion of the object and does not handle abrupt motion well.
-As the flowers are constrained in movement by their stalk, their movement does not follow the assumption of linear motion.
-Some tracking algorithms, such as DeepSORT [@deepsort2017], take the tracking-by-detection approach in which convolutional neural networks are applied to leverage appearance information in order to continuously distinguish individuals through time, thus aiding the tracking algorithm.
+As the flowers are constrained in movement by their stalk, their movement does not follow this assumption.
+Other tracking approaches, such as DeepSORT [@deepsort2017], use tracking-by-detection in which CNNs are applied to leverage appearance information in order to continuously distinguish individuals through time, thus aiding the tracking algorithm.
 In our case, however, flowers are not clearly distinguishable and one flower may look less alike between two frames than two flowers in a single frame.
-Finally, some tracking methods require manual initiation of a fixed number of tracks for objects in the first frame and the algorithm will track these objects in the subsequent frames [@hu2012].
-These methods does not allow new objects to enter the frame through the series.
-In our case, objects can disappear permanently from the frame, for example when flowers wilt, in which case the track should be terminated.
-Oppositely, new objects can continuously appear through the series when new flowers bloom of previously unseen flowers enter the frame.
-Additionally, objects may temporarily disappear from view and may or may not reappear at some later stage.
-These settings vastly complicate the task of tracking multiple objects.
-Our tracking algorithm returned consistently accuracy despite the difficult settings.
-
-We applied our tracking algorithm on 3179 combinations of the three parameters.
-Using the parameters increased the performance of the tracking substantially.
-Our steps in parameter values were crude, however, and it is very likely that finer steps in these values would identify combinations that produce even better results.
-Here we do not perform this analysis, however, as our goal is to show that the parameters can be used for optimizing tracking performance in general
-
-In a naive setting, a priori knowledge about which settings to use is not available.
-As with any tracking method, accuracy can only be assessed based on visual observation or more thoroughly by testing on subset of tracks for which a ground truth exists.
-When applying the method in a new setting, we recommend performing a parameter test as demonstrated here on a subset of ground truth tracks to derive the optimal set of parameters.
-
-Number of tracks equal to number of flowers does not necessarily mean that the tracking is correct.
-I.e., the tracking algorithm could split one ground truth track in two, but combine two other ground truth tracks into one, which would even each other out.
-This was the case for NYAA-04 for which the parameter combination that returned a number of tracks equal to the number of flowers in the series was not the combination the produced the highest MOTA score.
+Finally, some tracking methods require manual initiation of a fixed number of tracks for objects in the first frame and the algorithm will then track these objects in the subsequent frames [@hu2012].
+These methods do not allow new objects to enter the frame within the series.
+In our case, flowers can disappear permanently from the frame, for example when flowers wilt, in which case the track should be terminated.
+Oppositely, new objects can continuously appear through the series when new flowers bloom or previously unseen flowers enter the frame.
+Additionally, objects may disappear temporarily from view and reappear at some later stage.
+Further, as the manual tracking of flowers showed, the lifespans of flowers vary substantially (fig.
+2), which means that assumptions on flowering duration cannot be leveraged during tracking.
+These settings vastly complicate the task of tracking multiple objects [@betke2007].
+Our tracking algorithm returned consistently accuracy despite this.
 
 The two series for which the tracking algorithm produced perfect results were also the series with the lowest number of flowers.
-As the number of flowers increase so does the complexity of the task of tracking.
+While an increased number of objects does not necessarily increase tracking difficulty as this depends on the degree of isolation of tracks in time an space [@beard2018], multiple object tracking is a challenging task.
+Here, since the flowers are limited in space and time, the complexity of tracking likely increases with increasing number of flowers.
 A higher number of flowers within the frame increases the possibility of flowers in close vicinity to each other and of flowers overlapping and moving in and out of frame translating to a higher risk of making tracking errors.
-The very high complexity in the scenes with many flowers, especially 2019 NARS C with 85 flowers, renders it unlikely that any tracking method could produce perfect results.
-Although this setting can be considered an extreme case for our application, it underlines the relevance of a pipeline including both tracking and filtering.
+The very high complexity in the scenes with many flowers, especially 2019 NARS C with 85 flowers, coupled with the low frame rate renders it unlikely that any tracking method could produce perfect results across series.
+Although 2019 NARS C can be considered an extreme case for our application, it underlines the relevance of a pipeline including both tracking and filtering.
+But further, it points to the possibility of filtering image series based on an estimate of the number of flowers before tracking.
+For example, with many image series available, series with less than a given maximum of flowers detected in a single frame over a season could be selected for subsequent tracking.
+Tracks in series that exceeded this threshold could be filtered subsequent of tracking, as demonstrated here, or selected for manual validation of results.
 
-#
-## Filtering
+Our results show that optimizing for high MOTA tends to lead to an underestimate of the number of tracks (i.e.
+merging of tracks occurs) (table 1).
+This may happen, for example, because of the max gap parameter.
+Importantly, this parameter allows for temporary occlusion of a flower within a track (which would decrease the number of mismatches) but may also cause merging of two flower tracks if they bloom in the same area separated by a number of frames less than the threshold value set.
+Thus, even if the number of tracks equals the number of flowers, it does not necessarily follow that the tracking is correct.
+For instance, the tracking algorithm could split one ground truth track in two, but combine two other ground truth tracks into one, as seen in Figure 3.
+The task of evaluating multiple object tracking accuracy is a major challenge in itself [@luiten2020; @milan2013; @milan2016] and in a future study, other tracking accuracy metrics could be implemented, that account for merging errors in a better way.
+Tracking flowers may be a challenge that is especially prone to these errors since flowers can bloom in the same area at different times during the season.
 
-We presented a filtering method based on the distance between track geometry centroids which can be applied subsequent of tracking.
-The filtering algorithm successfully extracted correctly tracked flowers from each of the three series for which the tracking algorithm did not produce perfect results.
-Only a single mismatch was included in the 28 flowers that passed the filtering.
-Extracting tracks that are spatially isolated does not guarantee that the tracks are correct/without errors.
-However, as spatially isolated objects are easier to track, it increases confidence in the remaining tracks.
-
-We applied the filtering algorithm on tracks produced with a fixed set of tracking parameters as well as a fixed value for the filtering distance.
-We note that the results could be improved (i.e.
-more flowers extracted without increased number of mismatches) by chosing the best tracking result for each series and fine-tuning the filtering distance value for each individual series.
-However, as our goal here was to show that a single conservative value can be applied across series for which the tracking algorithm did not necessarily produce the best results, we do not show those results here.
-As in the case of the tracking parameters, we recommend setting the value for the filtering distance based on either visual examination of results or on tests on a subset of ground truth data when applying the method in a new setting.
-
-We note that our method for filtering tracks using track geometry centroids ensures that all tracks, despite the number of points in the track, are given the same weighting in the filtering since each track is represented by a single point.
-In some cases this may be suboptimal.
-For example, if it is given a priori that an object will always appear in a minimum of two frames, then single point tracks can be filtered out.
-However, when such a priori knowledge is not accessible, a conservative approach as the one we present is preferable.
-
-Filtering based on geometry centroids ensures that the same weight is put on all tracks despite the number of points within the track.
-Additionally, it allows for overlap between track geometries as long as the distance between the track centroids is above the threshold.
-
-#
-## User-defined parameters
-
-Our method relies on a number of parameters set by the user, both for tracking and subsequent filtering.
-Our results demonstrate that the tracking accuracy is sensitive to these parameters and an optimal combination of parameters can improve results substantially.
-Below we discuss the underlying mechanisms of each of the parameters.
-
-For the three series with complex scenes, applying the max gap size parameter improved tracking accuracy with values of 10-20 being optimal.
-These series all contain flowers at the edge of the frame and the higher number of flowers increases the risk of flowers temporarily occluding others.
+The optimal settings for the three tracking parameters fell into a narrow range for the three complex series, except for the running mean for 2019 NYAA E.
+The optimal values for the running mean for 2019 NYAA E were exceptionally high, compared to the other series.
+Interestingly, the average flower duration was also higher for this series (fig.
+2) and this could be part of the explanation, as flowers with longer lifespans can also have larger gaps in their tracks.
+Overall, the max gap parameter improved tracking accuracy with values of 10-20 being optimal.
+The three series with complex scenes all contain flowers at the edge of the frame (see fig.
+1), which means that they may occasionally be outside of view of the camera.
+Further, the higher number of flowers increase the risk of flowers temporarily occluding others.
 Correct tracking of these objects requires that an object can disappear from view without the track instantly being lost, which the max gap size allows.
-Oppositely, a higher number of flowers also increase the possibility of a new flower blooming in close vicinity to where a flower previously existed and setting the value too high may force objects such objects that are only separated in time but occur in the same space into the same track.
-
-In all cases, except 2019 NARS D, for which perfect tracking was achieved with all parameters set to zero, did a value for running mean above 0 improve results.
-Interestingly, a value of 10 produced the best results in all these cases, except for the combination returning the lowest number of mismatches for NYAA-04.
-Increasing the value for running mean means that less weight is put on movements between each frame.
-Changing the running mean to 10 while keeping the other two parameters equal only resulted in a slight increase in number of `r word_comment("This result is not given - should results for every combination be given in suppl.?
-It's 3179*5 rows of data, so it's quite a lot.)", "mismatches")` (from 29 to 36).
-Thus, it is a likely explanation that the improved accuracy from the large value for running mean stems from slightly better tracking of flowers that moved a lot.
-
-For the three series for which the tracking algorithm did not produce perfect results, the optimal range for max distance were relatively narrow (300-500).
-The optimal value for this setting is one that is sufficiently high so that it does not force individual flowers into separate tracks because of movement between frames yet is low enough to separate any flowers into individual tracks that would otherwise have been grouped together.
+Oppositely, as discussed, this setting may also lead to merging error.
+The optimal range for max distance was relatively narrow (400-600) for the parameters producing the correct number of tracks with the lowest number of mismatches.
+The optimal value for this setting is one that is sufficiently high that it does not force individual flowers into separate tracks because of movement between frames yet is low enough to separate any flowers into individual tracks that would otherwise have been grouped together.
 This explains well why the extreme values for this parameter was not included in the highest accuracy tracks.
@@
 For example, the distance between a point in the current frame and a point in the previous frame may not be the same as that for a point in the current frame and the mean location of the previous ten points in a track.
-Thus, changing the setting for running mean could influence the optimal setting for max distance.
-This underlines the relevance of search for an optimal combination based on ground truth data.
-In general, the optimal settings for the three tracking parameters fall into relatively narrow ranges.
-This strengthens the assumption that the method is generalisable and could be applied for example across many more cameras than what have been demonstrated here.
+Thus, changing the setting for running mean could influence the optimal parameter setting for max distance.
+This underlines the relevance of searching for an optimal parameter combination based on ground truth data.
+As with any tracking method, accuracy can only be assessed based on visual observation or, more rigorously, by testing on ground truth data [@dendorfer2021; @luiten2020; @milan2016].
+However, the fact that the optimal tracking parameters were overall similar across series strengthens the assumptions that a set of standard values can produce reasonable results for other similar image series with *Dryas* flowers.
+Here, a combination such as (10, 500, 10, for max gap, max distance, and running, respectively) would be a good starting point.
+When applying our method in a new setting, we recommend performing a parameter test as demonstrated here on a subset of ground truth tracks to derive the optimal set of parameters.
 
-We filter out tracks with high risk of error based on distance between track geometry centroids.
-An alternative approach could be to filter based on overlap between polygons, but this would remove tracks with even minor overlap and put weight on tracks with many points as these cover more area.
+We demonstrated our method on manual annotations of flowers but note that the format of the input data matches the output of deep learning object detection methods [@everingham2010].
+Thus, the method can be directly applied on such data.
+Automatically detecting flowers would likely introduce a proportion of false negatives when the detection algorithm fails to detect a present object.
+The flowers in our data that periodically move out of the frame or are temporarily occluded mimics false negatives.
+Thus, it is likely that the max gap parameter can alleviate problems with false negatives.
+Automatic detection methods may also introduce false positives.
+These detections would be included in the tracking on par with correct detections.
+Some cases of false positives, for example isolated single false positives, could be removed automatically while other cases may require manual quality control.
+In general, the performance of a tracking algorithm depends on the quality of input data and it is advisable to optimise detection accuracy before moving to tracking.
+The fact that our method produced perfect tracking results in the most simple settings (fewest flowers) could be taken into consideration when setting up future data collection schemes.
+
+## Filtering
+
+We presented a filtering method based on the distance between track centroids, which can be applied subsequent of tracking.
+The filtering algorithm successfully extracted 28 correct tracks from each of the three series for which the tracking algorithm did not produce perfect results without a single mismatch.
+Extracting tracks that are spatially isolated does not guarantee that the tracks are correct/without errors.
+For example, tracks of several objects may have been erroneously merged or a track could contain a small proportion of instances of an object far away, which would not have a large effect on the track centroid position (fig.
+2).
+However, as spatially isolated objects in general are easier to track, it is safe to assume that these tracks are more reliable.
+We applied the filtering algorithm on tracks produced with a fixed set of tracking parameters as well as a fixed value for the filtering distance.
+As in the case of the tracking parameters, we recommend setting the value for the filtering distance based on either visual examination of results or on tests on a subset of ground truth data when applying the method in a new setting, but for tracking *Dryas* flowers in similar image series, a value of 350 may be a good starting point.
+We note that our method for filtering tracks using track centroids ensures that all tracks, despite the number of points in the track, are given the same weighting in the filtering since each track is represented by a single point.
+In some cases, this may be suboptimal.
+For example, if it is given a priori that an object will always appear in a minimum of two frames, then single point tracks can be filtered out.
+However, when such a priori knowledge is not accessible, a conservative approach, as the one we present, is preferable.
+
+An alternative approach to filtering based on distance between track centroids could be to identify and discard track polygons that overlap.
+However, this would remove tracks with even minor overlap and put weight on tracks with many points as these typically cover a larger area.
 Another alternative could be to base filtering on the centroids of the points within a track (i.e.
 mean x and mean y).
-The presented method avoids weighting track geometries based on number of included points and avoids weighting the centroid coordinate based on the point coordinates within the track.
-We note, however, that other approached may be more suitable in other settings and applications.
+The presented method avoids weighting track geometries based on number of included points and avoids weighting the centroid coordinate based on the point coordinates within the track as the track centroid is based on the convex hull of the track points.
+Additionally, it allows for overlap between track geometries as long as the distance between the track centroids is above the threshold.
+We note, however, that other approaches may be more suitable in other settings and applications.
 
-#
-## Input data format
-We demonstrated our method on manual annotations of flowers but note that the format of the input data matches the output of deep learning object detection methods.
-Thus, the method can be directly applied on such data.
-Automatically detecting flowers would likely introduce a degree of false negatives when the detection algorithm fails to detect a present object.
-The flowers in our data that periodically move out of the frame or are temporarily occluded mimics false negatives.
-Thus, it is likely that the max gap size parameter can alleviate problems with false negatives.
-Automatic detection methods may also introduce false positives.
-These detections would be included in the tracking.
-Single false positives could be removed automatically while other cases may require manual quality control.
-Depending on the application, a low degree of false positives may not problematic.
-In general, the performance of a tracking algorithm depends on the quality of input data and it is advisable to optimise detection accuracy before moving to tracking.
-The fact that our method produced perfect tracking results in simple settings could be taken into consideration when setting up future data collection schemes.
+# Significance for ecological studies
 
-#
-## Conclusion
+Our results show that it is feasible to collect and analyse individual flower phenology at large scale at logistically challenging sites, which is a significant contribution towards automatic and efficient phenology data collection.
+The possibility of collecting individual level flower phenology is highly relevant for exploration of many research questions.
+For example, it would be possible to assign indicators of reproductive success and factors driving variation in seed set to the individual.
+With this, it could be explored how seed set is affected by timing and length of flowering and/or abiotic conditions such as frost events.
+Hereby, the long-standing question of what drives variation in reproductive success of individual flowers could be explored [@bolmgren2015].
+Further, the fact that it was not uncommon in our results that flowering lasted less than a week (fig.
+1), emphasizes that weekly observations of sample plots can be insufficient for comprehensive description of flower phenology [@Prevey2021].
 
-In this paper, we have demonstrated the potential of image-based monitoring for collecting information on flower phenology at the level of individuals.
-The automatic tracking and filtering pipeline presented was able to return reliable flower tracks despite settings that violate assumptions of standard tracking algorithms and complex scenes with a high number of objects that are not permanent within the series.
-Our results show that it is possible to collect and analyse individual flower phenology at large scale at logistically challenging sites.
-The presented method is a significant contribution towards automatic and efficient phenology data collection.
+By simultaneously tracking flower visits, these could be assigned to the individual flower and visitation rates per flower could be derived and the effects on reproductive success could be investigated [@tiusanen2016; @rafferty2012].
+Pollen limitation is a common factor limiting the reproductive success of flowering plants, but the ecological mechanisms behind pollen limitation is poorly understood, at least partly because it requires information on plant reproductive success in relation to pollinator visitation rates which is difficult to acquire in a natural setting and to a satisfactory extent.
+For example, pollinator visits can vary greatly in frequency during the flowering season and the degree of pollen limitation can vary even between populations in close vicinity to each other [@larson1999; @ashman2004].
+High frequency observations of flower visits coupled with information on phenology and reproductive success at the level of individuals could unravel the relationship between pollinator visitation and plant reproductive success and the climate sensitivity of plant-pollinator interactions.
+Further, any information on taxonomic or functional grouping could facilitate comparative studies on pollination effectiveness between such groups [@king2013; @foldesi2021].
+Importantly, the variables suggested here could be derived directly from the time-lapse images.
+For example, flower reproductive success could be estimated by observation of seed set for the individual flower and flower visitation rates could be quantified by detecting insects in flowers in high temporal resolution image series.
+
+It is clear that the images derived from monitoring of flower plots can hold substantial amounts of ecologically relevant data.
+To derive some variables, for example flower visitation rates, image series collected at high temporal frequency is required.
+As the scale of data increases as well as the number of variables to be extracted from the images increase, automated image analysis becomes relevant.
+Here, methods from deep learning and computer vision are exceedingly relevant.
+For instance, it has been demonstrated that convolutional neural networks can facilitate monitoring of pollination events by detecting, identifying, and counting insects [@bjerge2021] and perform fine-grained species classification [@spiesman2021].
+Collectively, these developments coupled with implementation of long-term high-frequency image-based monitoring will allow for collection and processing of ecological data at unprecedented scales.
+
+For flower tracking in particular, our results emphasize that this is not a trivial task.
+While we show that tracking can be done at high accuracy with our tracking algorithm, we also identify particular challenges with flower tracking, such as the violations of standard assumptions in tracking algorithms, the potentially large number of individuals overlapping in time and space, and the possible susceptibility to track merging.
+To improve tracking ability of flowers in future studies, focus should be on implementing reliable measures of different types of errors and on developing methods that can alleviate the challenges, as well as on comparison of different tracking approaches.
 
 # Acknowledgements
-?
+
+We thank Cecilie Mielec, Michael Straarup Nielsen and Maarten J.
+J.
+E.
+Loonen for set-up and maintenance of cameras in Ny-Ålesund and Peter Akers and Jeffrey M.
+Welker for set-up and maintenance of cameras in Thule.
+TTH acknowledges funding from Independent Research Fund Denmark Grant 8021-00423B.
+The authors have no conflict of interest to declare.
+`r word_comment("Hjalte: Toke, please check if this is sufficient.
+Should anyone from the Narsarsuaq fieldwork be mentioned?
+Anyone else?", "")`
+
+# Author contributions
+
+HRMR, TTH and AI conceived the idea.
+HMRM conceived and developed the method with contributions from TTH and AI.
+HMRM wrote the manuscript with contributions from TTH and AI.
+All authors gave final approval for publication.
 
@@
 
-The code that supports the results in this paper will be made openly available at https://github.com/TECOLOGYxyz/FlowerTracking.
-A publicly available web application can be accessed from the Github repository through which users can run the tracking and filtering algorithms on their own data.
-Raw data will be archived on https://zenodo.org/.
+The code that supports the results in this paper is openly available at <https://github.com/TECOLOGYxyz/FlowerTracking>. The raw data will be archived on <https://zenodo.org/>.
 
 \newpage
-# References\
+
+# References

### 2022-05-25 text/SupportManuscriptDraft.Rmd
Updated filtering results in supporting info

@@
 
-**Table S1:** Results of the filtering algorithm for filtering distances 5-500 with a stepsize of 50 for all three series.
+## Results of the filtering algorithm for filtering distances 5-500 with a stepsize of 50 for the three series.
 
-![](../figures/FullFilteringDistanceTest.png){ width=100% }
+eps = 50: 158 mismatches with 165 tracks kept 
+eps = 100: 103 mismatches with 147 tracks kept 
+eps = 150: 71 mismatches with 124 tracks kept 
+eps = 200: 47 mismatches with 100 tracks kept 
+eps = 250: 30 mismatches with 69 tracks kept 
+eps = 300: 9 mismatches with 49 tracks kept 
+eps = 350: 0 mismatches with 28 tracks kept 
+eps = 400: 0 mismatches with 18 tracks kept 
+eps = 450: 0 mismatches with 16 tracks kept 
+eps = 500: 0 mismatches with 11 tracks kept
+
+#![](../figures/FullFilteringDistanceTest.png){width="100%"}

### 2022-05-25 code/run.py
Updated to use new evaluation method (MOTMetrics) and included runs for updated results

@@ -7,15 +7,23 @@ Created on Tue Dec 14 12:50:10 2021
 
 import pandas as pd
 from track import tracker
-from filtering import sieve
+#from filtering import sieve
 from DBSCANFiltering import DBSCANsieve
-from DistanceFiltering import distanceSieve
-from evaluate import evaluator
-import os
+from distanceFiltering import distanceSieve
+#from evaluate import evaluator
+#import os
 import time
+import os
+import pandas as pd
+from scipy.spatial import distance as dist
+import os
+import re
+import numpy as np
 
+import motmetrics as mm
 
-# ===================== Import ground truth data ====
+###################
+
 
 def impdata(camID):
     if camID == "NARS-04":
@@ -28,17 +36,173 @@ def impdata(camID):
         gt = r"U:\BITCue\Projekter\TrackingFlowers\data\annotations\2021_12_13_NorwayAnnotations_NYAA-04_IndividualAnnotations_FRCNN_Metrics.csv"
     if camID == "NARS-17":
         gt = r"U:\BITCue\Projekter\TrackingFlowers\data\annotations\2021_12_30_NorwayAnnotations_NARS-17_IndividualAnnotations_FRCNN_Metrics.csv"
+    if camID == "debug":
+        gt = r"U:\BITCue\Projekter\TrackingFlowers\data\annotations\debugGT.csv"
     return gt
         
 
+
+
+# ===================== Run evaluator on all files for the parameter test ====
+
+br = '\n'
+
+def write_header(results_filename):
+    with open(results_filename, 'a') as resultFile: # Write the header of the output file
+        header = f'maxDisap,maxDist,runMean,num_frames,num_detections,num_objects,num_predictions,num_unique_objects,num_tracks,num_matches,num_switches,num_misses,num_false_positives,precision,recall,mota,motp,mostly_tracked,partially_tracked,mostly_lost{br}'
+        resultFile.write(header)
+
+def write_results_file(results_filename,maxDisap,maxDist,runMean,num_frames,num_detections,num_objects, num_predictions,num_unique_objects,num_tracks,num_matches,num_switches,num_misses,num_false_positives,precision,recall,mota,motp,mostly_tracked,partially_tracked,mostly_lost):
+    with open(results_filename, 'a') as resultFile:
[…]

### 2022-05-25 code/R/Analysis.Rmd
File cleanup

@@
 # 2018 NARS-04: 2018 NARS A
-# 2018 THUL-01: 2018 THUL B
+# 2018 THUL-01: 2018 THUL B v
 # 2019 NARS-13: 2019 NARS C
-# 2019 NARS-17: 2019 NARS D
+# 2019 NARS-17: 2019 NARS D v
 # 2019 NYAA-04: 2019 NYAA E
@@
 ```
+
+####### MMFIX PARAMETER TEST ########
+
+# NYAA-04
+```{r fig.height=4}
+NYAA04d1 = read.csv("U:/BITCue/Projekter/TrackingFlowers/testResults/MMFix/NYAA-04_parameterTest_Evaluation.csv")
+
+NYAA04_p1 = NYAA04d1 %>% 
+  ggplot() +
+  geom_point(aes(x=runMean, y = maxDisap, size = mm, colour = mm)) + 
+  ggtitle("NYAA-04 Parameter combination") +
+  facet_wrap(vars(maxDist))+scale_color_gradient(low="blue", high="red")
+
+NYAA04_p2 = NYAA04d1 %>% 
+  arrange(mm, ascending = T) %>% 
+  slice_head(n = 10) %>% 
+  ggplot() +
+  geom_point(aes(x=runMean, y = maxDisap, size = mm, colour = mm)) + 
+  ggtitle("NYAA-04 Ten Lowest Mismatches") +
+  facet_wrap(vars(maxDist))+scale_color_gradient(low="blue", high="red")
+
+NYAA04_p1
+NYAA04_p2
+
+ggsave('U:/BITCue/Projekter/TrackingFlowers/testResults/MMFix/figures/NYAA-04_AllCombinations.png', NYAA04_p1)
+ggsave('U:/BITCue/Projekter/TrackingFlowers/testResults/MMFix/figures/NYAA-04_TenLowest.png', NYAA04_p2)
+
+rm(NYAA04_p1)
+rm(NYAA04_p2)
+```
+
+# NARS-13
+```{r fig.height=4}
+NARS13d1 = read.csv("U:/BITCue/Projekter/TrackingFlowers/testResults/MMFix/NARS-13_parameterTest_Evaluation.csv")
+
+NARS13_p1 = NARS13d1 %>% 
+  ggplot() +
+  geom_point(aes(x=runMean, y = maxDisap, size = mm, colour = mm)) + 
+  ggtitle("NYAA-04 Parameter combination") +
+  facet_wrap(vars(maxDist))+scale_color_gradient(low="blue", high="red")
+
+NARS13_p2 = NARS13d1 %>% 
+  arrange(mm, ascending = T) %>% 
+  slice_head(n = 10) %>% 
+  ggplot() +
+  geom_point(aes(x=runMean, y = maxDisap, size = mm, colour = mm)) + 
+  ggtitle("NYAA-04 Ten Lowest Mismatches") +
+  facet_wrap(vars(maxDist))+scale_color_gradient(low="blue", high="red")
+
+NARS13_p1
+NARS13_p2
+
+ggsave('U:/BITCue/Projekter/TrackingFlowers/testResults/MMFix/figures/NARS-13_AllCombinations.png', NARS13_p1)
+ggsave('U:/BITCue/Projekter/TrackingFlowers/testResults/MMFix/figures/NARS-13_TenLowest.png', NARS13_p2)
+
+rm(NARS13_p1)
+rm(NARS13_p2)
+```
+
+# NARS-04
+```{r fig.height=4}
+NARS04d1 = read.csv("U:/BITCue/Projekter/TrackingFlowers/testResults/MMFix/NARS-04_parameterTest_Evaluation.csv")
+
+NARS04_p1 = NARS04d1 %>% 
+  ggplot() +
+  geom_point(aes(x=runMean, y = maxDisap, size = mm, colour = mm)) + 
+  ggtitle("NYAA-04 Parameter combination") +
+  facet_wrap(vars(maxDist))+scale_color_gradient(low="blue", high="red")
+
+NARS04_p2 = NARS04d1 %>% 
+  arrange(mm, ascending = T) %>% 
+  slice_head(n = 10) %>% 
+  ggplot() +
+  geom_point(aes(x=runMean, y = maxDisap, size = mm, colour = mm)) + 
+  ggtitle("NYAA-04 Ten Lowest Mismatches") +
+  facet_wrap(vars(maxDist))+scale_color_gradient(low="blue", high="red")
+
+NARS04_p1
+NARS04_p2
+
+ggsave('U:/BITCue/Projekter/TrackingFlowers/testResults/MMFix/figures/NARS-04_AllCombinations.png', NARS04_p1)
+ggsave('U:/BITCue/Projekter/TrackingFlowers/testResults/MMFix/figures/NARS-04_TenLowest.png', NARS04_p2)
+
+rm(NARS04_p1)
+rm(NARS04_p2)
+```
+
+# NARS-17
+```{r fig.height=4}
+NARS17d1 = read.csv("U:/BITCue/Projekter/TrackingFlowers/testResults/MMFix/NARS-17_parameterTest_Evaluation.csv")
+
+NARS17_p1 = NARS17d1 %>% 
+  ggplot() +
+  geom_point(aes(x=runMean, y = maxDisap, size = mm, colour = mm)) + 
+  ggtitle("NYAA-04 Parameter combination") +
+  facet_wrap(vars(maxDist))+scale_color_gradient(low="blue", high="red")
+
+NARS17_p2 = NARS17d1 %>% 
+  arrange(mm, ascending = T) %>% 
+  slice_head(n = 10) %>% 
+  ggplot() +
+  geom_point(aes(x=runMean, y = maxDisap, size = mm, colour = mm)) + 
+  ggtitle("NYAA-04 Ten Lowest Mismatches") +
+  facet_wrap(vars(maxDist))+scale_color_gradient(low="blue", high="red")
+
+NARS17_p1
+NARS17_p2
+
+ggsave('U:/BITCue/Projekter/TrackingFlowers/testResults/MMFix/figures/NARS-17_AllCombinations.png', NARS17_p1)
+ggsave('U:/BITCue/Projekter/TrackingFlowers/testResults/MMFix/figures/NARS-17_TenLowest.png', NARS17_p2)
+
+rm(NARS17_p1)
+rm(NARS17_p2)
+```
+
+# THUL-01
+```{r fig.height=4}
+THUL01d1 = read.csv("U:/BITCue/Projekter/TrackingFlowers/testResults/MMFix/THUL-01_parameterTest_Evaluation.csv")
+
+THUL01_p1 = THUL01d1 %>% 
+  ggplot() +
+  geom_point(aes(x=runMean, y = maxDisap, size = mm, colour = mm)) + 
+  ggtitle("NYAA-04 Parameter combination") +
+  facet_wrap(vars(maxDist))+scale_color_gradient(low="blue", high="red")
+
+THUL01_p2 = THUL01d1 %>% 
+  arrange(mm, ascending = T) %>% 
+  slice_head(n = 10) %>% 
+  ggplot() +
+  geom_point(aes(x=runMean, y = maxDisap, size = mm, colour = mm)) + 
+  ggtitle("NYAA-04 Ten Lowest Mismatches") +
+  facet_wrap(vars(maxDist))+scale_color_gradient(low="blue", high="red")
+
+THUL01_p1
+THUL01_p2
+
+ggsave('U:/BITCue/Projekter/TrackingFlowers/testResults/MMFix/figures/THUL-01_AllCombinations.png', THUL01_p1)
+ggsave('U:/BITCue/Projekter/TrackingFlowers/testResults/MMFix/figures/THUL-01_TenLowest.png', THUL01_p2)
+
+rm(THUL01_p1)
+rm(THUL01_p2)
+```
+
+### Explore combinations
+
+# NYAA-04
+```{r echo = FALSE}
+print("NYAA-04")
+nyaa04_gt = read.csv("U:/BITCue/Projekter/TrackingFlowers/data/annotations/2021_12_13_NorwayAnnotations_NYAA-04_IndividualAnnotations_FRCNN_Metrics.csv")
+
+print('True number of flowers')
+length(unique(nyaa04_gt$id_gt))
+
+# Minimum mm
+print("Lowest number of mismatches")
+min(NYAA04d1$mm)
+NYAA04Minmm = NYAA04d1[NYAA04d1$mm == min(NYAA04d1$mm),]
+
+# All zero
+NYAA04AllZero = NYAA04d1 %>% 
+  filter(maxDisap == 0 & maxDist == 0 & runMean == 0)
+
+# Correct number of flowers and minimum mm
+NYAA04CorrectflowerMinmm = NYAA04d1 %>% 
+  filter(number_of_objects == length(unique(nyaa04_gt$id_gt)))
+NYAA04CorrectflowerMinmm = NYAA04CorrectflowerMinmm[NYAA04CorrectflowerMinmm$mm == min(NYAA04CorrectflowerMinmm$mm),]
+```
+
+# NARS-13
+```{r echo = FALSE}
+print("NARS-13")
+nars13_gt = read.csv("U:/BITCue/Projekter/TrackingFlowers/data/annotations/2020_04_30_NorwayAnnotations_NARS-13_IndividualAnnotations_FRCNN_Metrics.csv")
+
+print('True number of flowers')
+length(unique(nars13_gt$id_gt))
+
+# Minimum mm
+print("Lowest number of mismatches")
+min(NARS13d1$mm)
+NARS13Minmm = NARS13d1[NARS13d1$mm == min(NARS13d1$mm),]
+
+# All zero
+NARS13AllZero = NARS13d1 %>% 
+  filter(maxDisap == 0 & maxDist == 0 & runMean == 0)
+
+# Correct number of flowers and minimum mm
+NARS13CorrectflowerMinmm = NARS13d1 %>% 
+  filter(number_of_objects == length(unique(nars13_gt$id_gt)))
+NARS13CorrectflowerMinmm = NARS13CorrectflowerMinmm[NARS13CorrectflowerMinmm$mm == min(NARS13CorrectflowerMinmm$mm),]
+```
+
+# NARS-04
+```{r echo = FALSE}
+print("NARS-04")
+nars04_gt = read.csv("U:/BITCue/Projekter/TrackingFlowers/data/annotations/2020_05_17_NorwayAnnotations_NARS-04_IndividualAnnotations_FRCNN_Metrics.csv")
+
+print('True number of flowers')
+length(unique(nars04_gt$id_gt))
+
+# Minimum mm
+print("Lowest number of mismatches")
+min(NARS04d1$mm)
+NARS04Minmm = NARS04d1[NARS04d1$mm == min(NARS04d1$mm),]
+
+# All zero
+NARS04AllZero = NARS04d1 %>% 
+  filter(maxDisap == 0 & maxDist == 0 & runMean == 0)
+
+# Correct number of flowers and minimum mm
+NARS04CorrectflowerMinmm = NARS04d1 %>% 
+  filter(number_of_objects == length(unique(nars04_gt$id_gt)))
+NARS04CorrectflowerMinmm = NARS04CorrectflowerMinmm[NARS04CorrectflowerMinmm$mm == min(NARS04CorrectflowerMinmm$mm),]
+```
+
+# NARS-17
+```{r echo = FALSE}
+print("NARS-17")
+nars17_gt = read.csv("U:/BITCue/Projekter/TrackingFlowers/data/annotations/2021_12_30_NorwayAnnotations_NARS-17_IndividualAnnotations_FRCNN_Metrics.csv")
+
+print('True number of flowers')
+length(unique(nars17_gt$id_gt))
+
+# Minimum mm
+print("Lowest number of mismatches")
+min(NARS17d1$mm)
+NARS17Minmm = NARS17d1[NARS17d1$mm == min(NARS17d1$mm),]
+
+# All zero
+NARS17AllZero = NARS17d1 %>% 
+  filter(maxDisap == 0 & maxDist == 0 & runMean == 0)
+
+# Correct number of flowers and minimum mm
+NARS17CorrectflowerMinmm = NARS17d1 %>% 
+  filter(number_of_objects == length(unique(nars17_gt$id_gt)))
+NARS17CorrectflowerMinmm = NARS17CorrectflowerMinmm[NARS17CorrectflowerMinmm$mm == min(NARS17CorrectflowerMinmm$mm),]
+```
+
+# THUL-01
+```{r echo = FALSE}
+print("THUL-01")
+thul01_gt = read.csv("U:/BITCue/Projekter/TrackingFlowers/data/annotations/2020_05_15_NorwayAnnotations_THUL-01_IndividualAnnotations_FRCNN_Metrics.csv")
+
+print('True number of flowers')
+length(unique(thul01_gt$id_gt))
+
+# Minimum mm
+print("Lowest number of mismatches")
+min(THUL01d1$mm)
+THUL01Minmm = THUL01d1[THUL01d1$mm == min(THUL01d1$mm),]
+
+# All zero
+THUL01AllZero = THUL01d1 %>% 
+  filter(maxDisap == 0 & maxDist == 0 & runMean == 0)
+
+# Correct number of flowers and minimum mm
+THUL01CorrectflowerMinmm = THUL01d1 %>% 
+  filter(number_of_objects == length(unique(thul01_gt$id_gt)))
+THUL01CorrectflowerMinmm = THUL01CorrectflowerMinmm[THUL01CorrectflowerMinmm$mm == min(THUL01CorrectflowerMinmm$mm),]
+```
+
+####### /MMFIX PARAMETER TEST ########
 
@@
 
+# After MMFix
 ```{r fig.height=4}
-d2 = read.csv("U:/BITCue/Projekter/TrackingFlowers/testResults/parameterTest_3_THUL-01_Evaluate.csv")
+d2 = read.csv("U:/BITCue/Projekter/TrackingFlowers/testResults/MMFix/THUL-01_parameterTest_Evaluation.csv")
 
@@
 
+```{r fig.height=4}
+# d2 = read.csv("U:/BITCue/Projekter/TrackingFlowers/testResults/parameterTest_3_THUL-01_Evaluate.csv")
+# 
+# 
+# THUL01_p1 = d2 %>% 
+#   ggplot() +
+#   geom_point(aes(x=runMean, y = maxDisap, size = mm, colour = mm)) + 
+#   ggtitle("THUL-01 Parameter combination") +
+#   facet_wrap(vars(as.factor(maxDist)))
+# 
+# 
+# THUL01_p2 = d2 %>% 
+#   arrange(mm, ascending = T) %>% 
+#   filter(mm == 0) %>% 
+#   ggplot() +
+#   geom_point(aes(x=runMean, y = maxDisap, size = mm)) + 
+#   ggtitle("THUL-01 All zero mismatches") +
+#   facet_wrap(vars(maxDist))+scale_color_gradient(low="blue", high="red")
+# 
+# 
+# THUL01_p1
+# THUL01_p2
+# 
+# ggsave('U:/BITCue/Projekter/TrackingFlowers/testResults/figures/THUL-01_AllCombinations.png', THUL01_p1)
+# ggsave('U:/BITCue/Projekter/TrackingFlowers/testResults/figures/THUL-01_AllZero.png', THUL01_p2)
+```
+
 ```{r}
-# NYAA-04 gt:
+# THUL-01 gt:
 thul01_gt = read.csv("U:/BITCue/Projekter/TrackingFlowers/data/annotations/2020_05_15_NorwayAnnotations_THUL-01_IndividualAnnotations_FRCNN_Metrics.csv")
@@
 
-NARS13_p1 = d4 %>% 
+NARS04_p1 = d4 %>% 
   ggplot() +
@@
 
-NARS13_p2 = d4 %>% 
+NARS04_p2 = d4 %>% 
   arrange(mm, ascending = T) %>% 
@@
 
-NARS13_p1
-NARS13_p2
+NARS04_p1
+NARS04_p2
 
-ggsave('U:/BITCue/Projekter/TrackingFlowers/testResults/figures/NARS-13_AllCombinations.png', NARS13_p1)
-ggsave('U:/BITCue/Projekter/TrackingFlowers/testResults/figures/NARS-13_TenLowest.png', NARS13_p2)
+ggsave('U:/BITCue/Projekter/TrackingFlowers/testResults/figures/NARS-13_AllCombinations.png', NARS04_p1)
+ggsave('U:/BITCue/Projekter/TrackingFlowers/testResults/figures/NARS-13_TenLowest.png', NARS04_p2)
 

### 2022-05-25 code/DBSCANFiltering.py

@@ -116,14 +116,14 @@ class DBSCANsieve():
         labels = pd.Series(clusters.labels_).rename('clusterID') # Grab the clustering labels and concatenate to df
         df = pd.concat([df, labels], axis=1)
         
-        print(df)
+        #print(df)
         grouped_df = df.groupby("clusterID") # Group by the clusterID and count the number of tracks within each cluster
         grouped_df = grouped_df.agg({"objectID": "nunique"}).reset_index()
         sub = grouped_df.loc[grouped_df['objectID'] == 1] # Make a list of clusterIDs that contain only one track
         l = sub['clusterID'].tolist()
 
         dfSub = df[df['clusterID'].isin(l)] # Subset the  df on the list so that it only includes only the isolated tracks (clusters containing a single track)
-        print("Passed DBSCAN ", l)
+        #print("Passed DBSCAN ", l)
 
         keepTracks = map(int, dfSub['objectID'].tolist()) # List of objectIDs we want to keep (map used to convert from str object to integer)
         tracksSub = self.tracks[self.tracks['objectID'].isin(keepTracks)] # Subset the tracks that passed the filtering. We'll return this dataframe from the filtering

### 2022-05-25 code/DistanceFiltering.py

@@ -56,6 +56,7 @@ class distanceSieve():
     
     def separate(self): # Take next group in tracks dataframe from the generator and split into correct dictionary according to number of points the track contains
         for oid, p in self.gen(): 
+            #print(p)
             if len(p) == 1: # Get single points
                 self.points[oid] = p
                 
@@ -121,7 +122,7 @@ class distanceSieve():
     def run(self):
         self.separate() # Separate the tracks into points, lines, and polygons
         self.convex_hull(self.polygons) # Calcuate the convex hull of the polygons. This will also add colinear points to lines dictionary
-        
+        #print("Got convex hulls")
         df, coords = self.geometries() # Get the geometry dataframe containing the track polygons and the centroid coordinates of which to perform the DBSCAN clustering
         
         #print(df)

### 2022-05-25 code/analyze.py

@@ -1,268 +1,0 @@
-# -*- coding: utf-8 -*-
-"""
-Created on Wed Dec  1 15:08:22 2021
-
-@author:
-    Hjalte M. R. Mann
-    TECOLOGY.xyz
-
-In this script we'll do some analyses of the ground truth individual annotations.
-We'll use the results for setting the parameters of our tracking algorithm.
-
-"""
-
-import pandas as pd
-import numpy as np
-
-# Input ground truth annotations
-df = pd.read_csv(r"../data\annotations\2021_12_13_NorwayAnnotations_NYAA-04_IndividualAnnotations_FRCNN_Metrics.csv")
-
-
-#print(thul01)
-
-df['x_c'] = (df['x_min'] + df['x_max']) / 2
-df['y_c'] = (df['y_min'] + df['y_max']) / 2
-
-
-
-def dist(a,b):
-    d = np.linalg.norm(a-b)
-    return d
-
-def ave(lst):
-    return sum(lst) / len(lst)
-
-
-
-df_grouped = df.groupby(['id_gt'])
-
-largestDist = 0
-
-# for i, g in df_grouped:
-#     #print(g)
-#     g_points = []
-#     for i,l in g.iterrows():
-#         g_points.append(np.array([l['x_c'], l['y_c']]))
-        
-#     for p1 in g_points:
-#         for p2 in g_points:
-#             d = dist(p1, p2)
-#             if d > largestDist:
-#                 print(f'{d} is larger than {largestDist}. Storing {d}')
-#                 print(f'Distance calculated between {p1} and {p2}')
-#                 largestDist = d
-            
-            
-# print(f'Largest dist found: {largestDist}')
-
-# THUL-01: 449.00334074481003
-# NYAA-04: 632.9178461696273
-# NARS-13: 782.0206199327483 QC
-# NARS-04: 557.953851138246
-# NARS-17: 615.5536126122565
-
-
-# Largest distance between mean of track and each point
-largestDist = 0
-
-# for i, g in df_grouped:
-#     #print(g)
-#     x_cs = []
-#     y_cs = []
-#     g_points = []
-#     for i,l in g.iterrows():
-#         x_cs.append(l['x_c'])
-#         y_cs.append(l['y_c'])
-#         g_points.append(np.array([l['x_c'], l['y_c']]))
-
-#     mx = ave(x_cs)
-#     my = ave(y_cs)
-    
-#     m = np.array([mx, my]) # Mean of x and mean of y
-#     for p1 in g_points:
[…]

### 2022-05-25 code/evaluate.py

@@ -1,312 +1,0 @@
-# -*- coding: utf-8 -*-
-"""
-Created on Mon Feb 10 15:48:25 2020
-
-Evaluating ML performance - SingleClass
-
-@author: Hjalte Mann, AU
-"""
-import pandas as pd
-import re
-import os.path
-
-br = "\n"
-
-iou_threshold = 0.5
-
-#verbose = False # Set to True if you want tracking process printed to screen and False if not
-
-
-
-"""
-To-do
-
-If basename does not include parameters, create other filename for result and output. So evaluate can be run on other files than the parameter tests.
-
-"""
-
-def write_header(results_filename):
-    with open(results_filename, 'a') as resultFile: # Write the header of the output file
-        header = f'maxDisap,maxDist,runMean,mm,mota,numberOfObjects{br}'
-        resultFile.write(header)
-
-
-class evaluator():
-    def __init__(self, dt, gt, results_filename, verbose):
-        self.P = 0 # Positives (meaning total number of objects (e.g. flowers) across all images)
-        self.CP = 0 # Correct positives (meaning total number of objects that were correclty detected)
-        self.FP = 0 # False positives
-        self.FN = 0 # False negatives
-        self.MM = 0 # Tracking mismatches
-        self.verbose = verbose
-        
-        self.dt = dt
-        self.gt = gt
-        
-        self.results_filename = results_filename
-
-
-        self.dt_filename = dt
-        basename = os.path.basename(self.dt_filename)
-            
-        runMeanSearch = re.search('runMean_(.+?)_max', basename)
-        maxDistSearch = re.search('maxDist_(.+?).csv', basename)
-        maxDisapSearch = re.search('maxDisap_(.+?)_run', basename)
-        
-        if runMeanSearch:
-            self.runMean = runMeanSearch.group(1)
-
-        if maxDistSearch:
-            self.maxDist = maxDistSearch.group(1)
-            
-        if maxDisapSearch:
-            self.maxDisap = maxDisapSearch.group(1)
-            
-        
-        if self.verbose:
-            print(f'Running on: {basename}')
-            print(f'runMean: {self.runMean}')
-            print(f'maxDist: {self.maxDist}')
-            print(f'maxDisap: {self.maxDisap}')
-            
-    def set_up_data(self):
-        dt = self.dt
-        gt = self.gt
-        
-        if not isinstance(self.dt, pd.DataFrame):
[…]

### 2022-05-25 code/filtering.py

@@ -1,350 +1,0 @@
-# -*- coding: utf-8 -*-
-"""
-Created on Sun Nov 28 20:14:12 2021
-
-@author: au309263
-"""
-from shapely.geometry import Polygon, LineString
-import pandas as pd # Replace with cudf if performance is too slow?
-from collections import OrderedDict
-import numpy as np
-from scipy.spatial import distance as dist
-from scipy.spatial import ConvexHull
-#from scipy.ndimage.filters import uniform_filter1d
-import matplotlib.pyplot as plt
-import time
-import itertools
-
-
-br = "\n"
-
-
-
-"""
-Pipeline
-
-- Remmove single points
-- Establish straight line between two-point tracks
-- Establish triangle between three-point tracks
-- Establish polygon from vertices of the convex hull of n-point tracks (n > 3)
-- Check if any lines overlap and remove both if they do
-- Check if any remaining lines overlap with any polygons, remove both line and polygon if so
-- Check if any polygons overlap and remove both if they do
-- Output filtered track file
-
-"""
-
-#### FUNCTIONS ####
-
-class sieve():
-    def __init__(self, tracks):
-        self.tracks = tracks
-        
-        self.points = {}
-        self.lines = {}
-        self.triangles = {}
-        self.polygons = {}
-        self.polyLengths = {}
-        self.polyHulls = {}
-        
-        
-    def gen(self):
-        t = self.tracks.groupby('objectID')
-        for i,g in t: # Generator that yields next group in tracks dataframe
-            oid = g['objectID'].iloc[0]
-            x_cs = g['x_c'].tolist()
-            y_cs = g['y_c'].tolist()
-            p = list(map(list, zip(x_cs, y_cs)))
-            yield oid, p
-    
-    
-    def separate(self):
-
-        for oid, p in self.gen(): # Take next group in tracks dataframe and split into correct dictionary according to number of points the track contains
-            if len(p) == 1: # Get single points
-                self.points[oid] = p
-                
-            if len(p) == 2: # Get lines
-                self.lines[oid] = p
-                
-            if len(p) == 3: # Get triangles
-                self.triangles[oid] = p
-
-            if len(p) > 3: # Get polygons
-                self.polygons[oid] = p
-                
-        for p in self.polygons:
-            self.polyLengths[p] = len(self.polygons[p])
-
-    
-        
-        #print(f'Points:{br}{points}')
[…]

### 2022-05-25 code/track.py

@@ -31,7 +31,7 @@ br = "\n"
 
 # ===================== SETTINGS ==============================================
 
-verbose = False # Set to True if you want tracking process printed to screen and False if not
+verbose = True # Set to True if you want tracking process printed to screen and False if not
 
 class tracker():
     def __init__(self, max_disappeared, max_distance, running_mean_threshold, results_filename, frames, detections, verbose):
@@ -41,8 +41,8 @@ class tracker():
         
         self.disappeared = OrderedDict() # Keeps track of how long an objectID has been lost
 
-        self.detections = detections
-        self.max_disappeared = max_disappeared # store parameters for use in the class
+        self.detections = detections  # store parameters for use in the class
+        self.max_disappeared = max_disappeared
         self.max_distance = max_distance
         self.running_mean_threshold = running_mean_threshold
         self.results_filename = results_filename
@@ -50,7 +50,7 @@ class tracker():
         
         self.tracks = [] # Create a list for storing tracking results as we go
 
-
+        
         with open(self.results_filename, 'a') as resultFile: # Write the header of the output file
             header = 'frame,filename,x_min,x_max,y_min,y_max,x_c,y_c,objectID\n'
             resultFile.write(header)
@@ -122,7 +122,7 @@ class tracker():
             
         if len(self.objects[objectID]) == self.running_mean_threshold:
             if verbose:
-                print("Length is equal to running mean threshold.")
+                print(f'Length {len(self.objects[objectID])} of {self.objects[objectID]} is equal to running mean threshold.')
                 print(f'Deleting first item in {self.objects[objectID]} ({self.objects[objectID][0]}) and appending {centroid}')
             del self.objects[objectID][0]
             self.objects[objectID].append(centroid)
@@ -131,14 +131,16 @@ class tracker():
         for key, value in self.objects.items():
             if len(value) > 1:
                 c_m = [mean([i[0] for i in value]), mean([i[1] for i in value])]
-                self.means[key] = c_m
+            if len(value) == 1:
+                c_m = value[0]
+            self.means[key] = c_m
         if verbose:
             print(f'Updated means dictionary{br}Current mean dict:{br}{self.means}')
         
     def return_tracks_webapp(self):
         return self.tracks
     
-    def track(self, frame):
+    def track(self, frame):        
         frame_detections = self.get_frame_detections(frame)#  Get the detections for the current frame
         if verbose:
             print(f'FRAME {frame}. Contains {len(frame_detections)} points.')
@@ -172,7 +174,9 @@ class tracker():
[…]

### 2022-05-25 code/visualize.py

@@ -1,344 +1,0 @@
-# # -*- coding: utf-8 -*-
-# """
-# Created on Thu Nov  7 11:07:43 2019
-
-# @author: Hjalte Mann
-
-# This script stitches images in a folder together to a time-lapse video
-# """
-
-from collections import OrderedDict
-import numpy as np
-import matplotlib.pyplot as plt
-from matplotlib.animation import FuncAnimation, PillowWriter
-import pandas as pd
-
-
-#df = pd.read_csv(r"U:\BITCue\Projekter\TrackingFlowers\data\annotations\2020_04_30_NorwayAnnotations_NARS-13_IndividualAnnotations_FRCNN_Metrics.csv")
-#df = pd.read_csv(r'../testResults/_parameterTest_NYAA-04_3/parameterTest_NYAA-04_maxDisap_10_runMean_60_maxDist_300.csv')
-gt = pd.read_csv(r'../data\annotations\2020_04_30_NorwayAnnotations_NARS-13_IndividualAnnotations_FRCNN_Metrics.csv')
-
-df2 = pd.read_csv(r'../testResults/_parameterTest_NARS-04_3/parameterTest_NARS-04_maxDisap_10_runMean_10_maxDist_500.csv')
-df3 = pd.read_csv(r'../testResults/filtered_NARS-13_maxDisap_10_runMean_10_maxDist_300.csv')
-
-#m = pd.merge(df, gt, on = ['filename', 'x_min', 'x_max', 'y_min', 'y_max'], how = 'inner')
-m2 = pd.merge(df2, gt, on = ['filename', 'x_min', 'x_max', 'y_min', 'y_max'], how = 'inner')
-m3 = pd.merge(df3, gt, on = ['filename', 'x_min', 'x_max', 'y_min', 'y_max'], how = 'inner')
-
-
-#m2.to_csv(r'U:\BITCue\Projekter\TrackingFlowers\testResults/NYAA04_Merged_m2_3.csv')
-
-#m3.to_csv(r'U:\BITCue\Projekter\TrackingFlowers\testResults/NYAA04_Merged_m3_3.csv')
-#print(m3)
-#m4 = m2.loc[m2['id_gt'] == 'FLower_01']
-
-
-#m4 = m2.loc[m2['id_gt'] == 'Flower_10']
-
-# print(m4)
-# plt.plot(m4['x_c'], m4['y_c'], c = m4['objectID'], s = 5)
-# plt.show()
-
-
[…]

### 2022-06-11 code/DBSCANFiltering.py
Adjusted plotting output

@@ -10,7 +10,7 @@ import pandas as pd
 from shapely.geometry import Polygon, LineString, Point
 from sklearn.cluster import DBSCAN
 import matplotlib.pyplot as plt
-from filtering import sieve
+#from filtering import sieve
 import numpy as np
 from scipy.spatial import ConvexHull
 from datetime import datetime
@@ -147,17 +147,27 @@ class DBSCANsieve():
         ax2.axes.get_xaxis().set_visible(False)  
         ax0.axes.get_xaxis().set_visible(False)
 
-        # ax0.set_aspect(1)
+        ax0.set_aspect(1)
         # ax1.set_aspect('auto')
         # ax2.set_aspect('auto')
 
-        # ax0.set_aspect(6080, 3420)
+        #ax0.set_aspect(6080, 3420)
         # ax1.set_aspect(6080, 3420)
         # ax2.set_aspect(6080, 3420)
-
+        
+        ax0.set_xlim((0,6080))
+        ax0.set_ylim((3420,0))
+        
+        #ax1.set_xlim((0,6080))
+        #ax1.set_ylim((3420,0))
+        
+        #ax2.set_xlim((0,6080))
+        #ax2.set_ylim((3420,0))
+        
+        print(self.tracks)
         scatter = ax0.scatter(self.tracks['x_c'], self.tracks['y_c'], c = self.tracks['objectID'], s = 0.2)
         ax0 = scatter.axes
-        ax0.invert_yaxis()
+        #ax0.invert_yaxis()
 
         df.plot(ax=ax1, column = 'clusterID', marker = ".", markersize = 0.2)
         dfSub.plot(ax=ax2, column = 'clusterID', marker = ".", markersize=0.2)

### 2022-06-11 code/R/Analysis.Rmd
Added second version of onset vs length plot

@@
 
-ggsave("U:/BITCue/Projekter/TrackingFlowers/figures/OnsetVSLength.png", f1, width = 15, height = 10)
+#ggsave("U:/BITCue/Projekter/TrackingFlowers/figures/OnsetVSLength.png", f1, width = 15, height = 10)
 
-all %>% 
+f2 = all %>% 
 ggplot()+
-  geom_point(aes(x = flower_start, y = flower_length, colour = CameraID))# +
-#geom_smooth(aes(x = flower_start, y = flower_length), method=lm)
+  geom_boxplot(aes(x = flower_start, y = flower_length, fill = camCode, shape = camCode), size = 1) + 
+  xlab("Onset of flowering (DOY)") + 
+  ylab("Number of days flowering") +
+    theme_bw()  +
+  scale_fill_manual(name = "Camera ID",
+                      values = c("gray26", "brown3", "dodgerblue3", "darkorange2", "chartreuse3"), aesthetics = c("colour", "fill")) + 
+  scale_shape_manual(name = "Camera ID",
+                      values = c(21,22,23,24,25)) + 
+  theme(panel.grid.major = element_blank(), 
+        panel.grid.minor = element_blank(),
+        panel.background = element_blank(),
+        legend.title = element_text(size = legendTitleSize),
+        text = element_text(size = axisSize),
+        legend.text = element_text(size = legendTextSize),
+        axis.line = element_line(),
+        axis.text=element_text(colour="black"),
+        panel.spacing = unit(0.7, "lines"))
 
-t.test(nyaa04_meta$flower_length, nars13_meta$flower_length)
+ggsave("U:/BITCue/Projekter/TrackingFlowers/figures/OnsetVSLengthBox.png", f2, width = 15, height = 10)
+
+# all %>% 
+# ggplot()+
+#   geom_point(aes(x = flower_start, y = flower_length, colour = CameraID))# +
+# #geom_smooth(aes(x = flower_start, y = flower_length), method=lm)
+#  
+# t.test(nyaa04_meta$flower_length, nars13_meta$flower_length)
 

### 2022-06-11 text/ManuscriptDraft.Rmd
Updates to text and suppl.

@@
 
-FIX FIGURE AND TABLE REFERENCES!
-
 Title:
@@
 
-Keywords:
-
-Tracking, flowering phenology, Arctic, dryas octopetala, dryas integrifolia, ecological monitoring, life-history variation
+Keywords: Arctic, *Dryas octopetala*, *Dryas integrifolia*, ecological monitoring, flowering phenology, life-history variation, tracking
 
@@
 
-Understanding how variation in the timing of flowering, the timing and frequency of pollinator visits, and the abundance and diversity of visitors affect plant reproduction is a fundamental challenge in pollination ecology.
-Since such interactions concern individual flowers, deciphering the consequences of e.g.
+A fundamental challenge in pollination ecology is to understand how successful pollination is affected by the timing of flowering, the timing and frequency of pollinator visits, and the abundance and diversity of visitors.
+Since plant-pollinator interactions concern individual flowers, deciphering the consequences of e.g.
 shifts in the frequency of flower visits for plant reproductive success should ideally be based on observations of individual flowers throughout their entire flowering period.
-Due to the overwhelming logistical challenge of collecting such data by direct observation, it has generally been infeasible to collect.
+Due to the overwhelming logistical challenge of collecting such data by direct observation, it has rarely been done.
 Here, we show that image-based monitoring of field plots at very high temporal resolution can return information on flowering phenology at the level of individual flowers.
-Further, we present a framework for automatically tracking flowers in time-lapse image series, evaluating tracking accuracy, and filtering out unreliable tracks.
-Our results on individual flower phenology confirm that traditional weekly observations of flower plots may be insufficient to capture details in flowering dynamics.
-We compare the results of our automated method with manual tracking and show that the performance achieved can be reliable in obtaining individual flower phenology.
-Our method could facilitate fine-grained studies on plant reproductive success and could be an important tool for estimating the sensitivity of plant-pollinator interactions to climate change.
+Further, we present a framework for automatically tracking flowers in time-lapse image series, evaluating tracking accuracy, and identifying unreliable tracks.
+Our results on individual flower phenology suggest that traditional weekly observations of flower plots are insufficient to capture details in flowering dynamics.
+We show that the results of our automated method are reliable by comparing with manual tracking of individual flower phenology.
+Our method could facilitate fine-grained studies on flower phenology and could be an important tool for estimating the sensitivity of plant-pollinator interactions to climate change.
 
@@
 Further, for insect-pollinated plants, flowering must be synchronous with pollinator activity for successful reproduction.
-Both flower and insect phenology is partly driven by abiotic cues such as timing of spring and temperature and can thus be sensitive to climate change.
+Both flower and insect phenology is partly driven by abiotic cues such as snowmelt and temperature and can thus be sensitive to climate change.
 Changes in flower phenology as a consequence of climate change may indirectly affect the abundance of flower-dependent insect species [@hoye2013] and these effects can be stronger than the direct effects of climate change [@ogilvie2017].
@@
 Monitoring flower phenology at high temporal resolution is time-consuming, particularly in logistically challenging environments such as the Arctic.
-Consequently, flowering phenology of a population is often characterized simply by the date for onset of flowering, typically derived from weekly observations of sample plots [@Prevey2021].
+Consequently, flowering phenology of a population is often characterized simply by the date for onset of flowering, typically derived from weekly observations in permanent plots [@Prevey2021].
 Such crude estimates may fail to reveal important dynamics in flowering phenology through the season.
 For example, different phenological factors, such as onset, peak, and end of flowering may respond differently to climate change [@cook2012] and estimation of the degree of mismatch between species may thus be sensitive to the choice of metric [@inouye2019].
-Of course, the optimal metric for phenology depends on the research question, but often the choice of recording first flowering dates to represent flowering phenology is pragmatic rather than ideal.
+The optimal metric for phenology depends on the research question, but often the choice of recording first flowering dates to represent flowering phenology is pragmatic rather than ideal.
 In fact, first flowering dates may be particularly ill suited for estimating sensitivity to climate change [@miller2008] since they represent the very extreme of the phenological distribution of a population and may be confounded with e.g.
 the size of the population considered.
-Rather, an accurate estimate of the phenological distribution of a population over the full season would be a preferred measure.
-Such data would allow for, for example, accurate quantification of resource availability through the season and temporal mismatches between species [@ramakers2020].
+An accurate estimate of the phenological distribution of a population over the full season would allow for, for example, accurate quantification of resource availability through the season and temporal mismatches between species [@ramakers2020].
 However, even detailed information on phenology of a population is still an aggregation of the phenology of the individual members of the population for which important variation may be masked [@inouye2019].
@@
 
-Cameras have become an ubiquitous tool for non-invasive and efficient monitoring of natural processes [@howe2017; @glover2019], but the implementation of image-based collection of data has outpaced the development of reliable and easily implementable methods for processing and analysing the big data output from this transformation [@weinstein2018; @farley2018].
-Computer vision and deep learning tools are being developed and applied for automatic processing and extraction of relevant data from these very large image sets [@willi2019; @tabak2019; @hoye2021].
-Indeed, automatic image-based monitoring of flowering phenology coupled with deep learning has facilitated collection of phenology data for specific species at very high temporal resolution [@mann2022], yet, phenological responses at the individual level may be indiscernible regardless of the temporal resolution of the data at population level.
+Cameras have become an ubiquitous tool for non-invasive and efficient monitoring of natural processes [@howe2017; @glover2019], but the implementation of image-based data collection has outpaced the development of methods for processing and analysing the big data output [@weinstein2018; @farley2018].
+Gradually, computer vision and deep learning tools are being developed and applied for automatic processing and extraction of relevant data from images [@willi2019; @tabak2019; @hoye2021].
+Indeed, automatic image-based monitoring of flowering phenology coupled with deep learning has facilitated collection of phenology data for specific species at very high temporal resolution [@mann2022].
+Yet, phenological responses at the individual level may be indiscernible regardless of the temporal resolution of the data at population level.
 For example, a shortening of individual flower longevity may not be directly obvious at the population level.
@@
 For example, investigating the association between reproductive success and timing of flowering and/or flower longevity requires such data.
-Individual level data would also aid the understanding of whether flower visitation rates and/or reproductive success depends on the timing of flowering for the individual flower.
-Thus, efficient and precise methods for deriving full seasonal distributions of phenology at the level of individuals are much desired.
+Individual level data would also aid the understanding of whether flower visitation rates and/or seed set depends on the timing of flowering for the individual flower.
+Thus, automated tracking of flowers in image-series of permanent plots to derive individual flower phenology would be ideal.
 
-Here, we use time-lapse cameras to monitor the phenology of two widespread Arctic flowering plant species *Dryas integrifolia* and *D.
+Most multiple object tracking literature focus on tracking of humans [@luo2021; @milan2016], and these methods not necessarily directly transferable to other use-cases.
+Many tracking algorithms, such as the Simple online and real-time tracking (SORT) [@sort], rely on Kalmann filtering to predict the future location of an object by estimating its velocity from previous frames [@kalmann].
+However, Kalman filtering assumes linear motion of the object and does not handle abrupt motion well.
+Flowers are constrained in movement by their stalk and violates this assumption.
+Other tracking approaches, such as DeepSORT [@deepsort2017], use tracking-by-detection in which convolutional neural networks (CNNs) are applied to continuously distinguish individuals through time, thus aiding the tracking algorithm.
+Flowers, however, are unlikely to be clearly distinguishable and one flower may look less alike between two frames than two flowers in a single frame, which prevents this approach.
+Finally, some tracking methods require manual initiation of tracks for a fixed number of objects in the first frame and the algorithm then tracks these objects in the subsequent frames [@hu2012].
+These methods do not allow objects to enter or disappear from frame during tracking.
+In full-season monitoring data, flowers will disappear from the frame when they wilt, in which case their track should be terminated.
+Oppositely, new flowers should continuously be tracked as they bloom throughout the season.
+Additionally, a flower tracking method must handle flowers disappearing and appearing at the edges of the frame.
+Clearly, individual flower tracking is a challenging task.
+
+Here, we present and evaluate a simple multiple object tracking algorithm for automatic tracking of individual flowers in image-series.
+Our flower tracking algorithm assigns individual flowers to tracks based on the minimum distance between centroids of bounding boxes and does not rely on prediction of future positions.
+Further, the method allows both initiation and termination of tracks during tracking.
+We describe three user-defined parameters to help optimise tracking of flowers in particular.
+We use time-lapse image-series from full-season monitoring of two widespread Arctic flowering plant species *Dryas integrifolia* and *D.
 octopetala*.
-We show that information on phenology at the level of individual flowers can be derived from image-based monitoring of flower plots.
-We present and evaluate a simple multiple object tracking algorithm for automatic tracking of individual flowers in the image series.
-The data extracted from the images violate common assumptions of standard tracking methods and tracking individual flowers through the season therefore represents a special case.
-For example, since flower movement is constrained by the stalk, the future location of a flower-head cannot be predicted based on previous trajectory of movement.
-Further, as individual flowers closely resemble each other, tracking cannot be based on recognition of individuals.
-Our flower tracking algorithm assigns individual flowers to tracks based on the minimum distance between centroids of bounding boxes.
-We apply three user-defined parameters to help optimise tracking performance.
-
 For complex scenes with many flowers in close vicinity to each other there is a substantial risk of tracking errors regardless of the tracking method being used.
@@
 For this, we suggest a conservative filtering approach based on the distance between tracks within the frame.
+Overall, we show that information on phenology at the level of individual flowers can be automatically derived from image-based monitoring of flower plots.
 
@@
 
-This study was based on five full-season image series collected at the following three Arctic sites (site abbreviation and sampling years in parentheses): Narsarsuaq (NARS), South Greenland (2018, 2019), Thule (THUL), North-West Greenland (2018), Ny-Ålesund (NYAA), Svalbard (2019).
+This study is based on five full-season time-lapse image series collected at the following three Arctic sites (site abbreviation and sampling years in parentheses): Narsarsuaq (NARS), South Greenland (2018, 2019), Thule (THUL), North-West Greenland (2018), Ny-Ålesund (NYAA), Svalbard (2019).
 The images were captured using time-lapse cameras (Moultrie Wingscapes TimelapseCam Pro, Moultrie Products, Birmingham Alabama, USA) permanently positioned approx.
@@
 The species are native to and widespread in Asian, North American, and European Arctic and alpine regions.
-They are geographically separated apart from a possible hybrid zone in North-East Greenland [@philipp2003].
+They are geographically separated except for a possible hybrid zone in North-East Greenland [@philipp2003].
 
@@
 While the original time-lapse frequency was higher, the basis for this study was 1-hour interval series.
-We limited these series to include only the flowering season and randomly sampled images from each series.
-These image sets were used for the study.
+We limited these series to include only the flowering season and, to limit the workload for manual annotation, randomly sampled images from each series.
 The number of images for each series after sampling is given in Table 1.
-All flowers in bloom within each image were manually annotated using the rectangular bounding box tool in the VIA VGG annotation software [@dutta2019] and each individual flower was assigned a unique ID that was maintained throughout its lifetime.
+All flowers in bloom within each of these image were manually annotated using the rectangular bounding box tool in the VIA VGG annotation software [@dutta2019] and each individual flower was assigned a unique ID that was maintained throughout its lifetime.
 These annotations constitute our ground truth tracks.
@@
 Examples of annotated images from the 2019 NARS C series at three different times in the flowering season are given in Figure 1.
-To demonstrate how individual flower tracking can result in detailed representations of flower phenology across the season, we use the ground truth annotations to show the duration of flowering for each individual flower across the season for each series (fig.
-2).
+To demonstrate how individual flower tracking can result in detailed representations of flower phenology across the season, we use the ground truth annotations to show the variation in duration of flowering for each series.
 
@@
 
-![](../figures/AnnotationExamples_v2_4.png){width="100%"}
+![](../figures/AnnotationExamples_v2_4.png){width="20%"}
 
@@
 A: Onset of flowering at DOY 145.
-B: Peak flowering at DOY 151 (52 annotated flowers).
+B: Peak flowering at day of year (DOY) 151 (52 annotated flowers).
 C: 6 days after peak flowering (DOY 157).
@@
 We built a framework for tracking, filtering, and evaluating tracking of objects in time-lapse image series.
-The tracking algorithm tracks objects based on distances between centroids of bounding boxes, i.e.
-objects with minimum displacement between frames are associated.
+The tracking algorithm (suppl.
+1) tracks objects based on distances between centroids of bounding boxes, that is, objects with minimum displacement between frames are associated.
 The algorithm has a set of user-adjusted parameters that can optimize tracking accuracy.
-The parameters are particularly relevant for optimal tracking of objects that are constrained to a specific area such as flowers.
+The parameters are particularly relevant for tracking of objects that are constrained to a specific area such as flowers.
 However, the tracking algorithm could be used to track any objects.
@@
 First, as flowers move, they may temporarily occlude the view of other flowers and flowers close to the edge of the frame may move in and out of view between images.
-This can cause a track to be lost and a new track to be erroneously initiated when the flower reappears (fig.
-1A), a tracking error known as fragmentation.
+This can cause a track to be lost and a new track to be erroneously initiated when the flower reappears (Fig.
+2A), a tracking error known as fragmentation.
 However, if a flower reappears in the same area as a flower is already being tracked after disappearing in only a few frames, it is a reasonable assumption that it is the same individual and not that the old flower wilted/disappeared and a new one developed.
-The parameter **max gap** sets the number of frames a track can be lost before a new track is initiated for a flower appearing in the same area.
-If a given flower has not been detected in a number of frames less than or equal to the max gap size, the track will not be lost.
+The parameter **max gap** sets the number of frames a flower can be absent before reappearing without the track ending and a new initiating.
+If a given flower has not been detected in a number of frames exceeding the max gap size, the track will be terminated.
 Concurrently, this can deal with potential false negatives, as a track that is missing a bounding box will not be lost.
 
-Second, a standard centroid tracking algorithm will associate a point with a track based on minimum distance only, disregarding the absolute distance between centroids, which may lead to tracking errors (fig.
-1B).
-However, as flowers are constrained in their displacement, we can assume a maximum distance that will occur between points and force the initiation of tracks for points that exceed this threshold.
+Second, a standard centroid tracking algorithm will associate a point with a track based on minimum distance only, disregarding the absolute distance between centroids, which may lead to tracking errors (Fig.
+2B).
+However, as flowers are constrained in their displacement, we can assume a maximum distance between points and force the initiation of tracks for points that exceed this threshold.
 We set this threshold with the parameter **max distance,** measured as the euclidean distance in number of pixels between objects in neighbouring frames.
 
-Third, with a high number of flowers in close vicinity to each other, there's a substantial risk of identity switching between flowers as they move among each other (fig.
-1C).
+Third, reliable tracking based on distance requires that movement of individuals are captured at sufficient frequency, such that each tracks are not switched between frames.
 Stand-alone image-based monitoring of the relatively slow process of flower development through full growing seasons (several months) is done with low frame rate, limited by power consumption and storage capacity both on and off the camera.
-It is simply not feasible to collect high frame-rate images/video across full seasons.
-Here, we used several images per day and while these images capture the phenology of the flowers at very high detail, they fail to capture the full movement of flowers.
-For example, as the wind shifts, the flower heads change direction, which in the image series happens instantaneously (i.e.
-between two consecutive frames) (fig.
-3C).
+Here, we used several images per day and while this temporal resolution captures the phenology of the flowers at very high detail, they fail to capture the full movement of flowers.
+For example, as the wind shifts, the flower heads change direction, which in the image series happens instantaneously (i.e., between two consecutive frames) (Fig.
+2C).
 The fact that the frame rate is insufficient to capture smooth movement means that bounding boxes around a single individual in two consecutive frames cannot be associated based on overlap (e.g.
 intersection over union) as there often is no overlap.
-Additionally, as the flowers are constrained by their stalk, there is a limit to the distance they can move, which complicates prediction of future location of an individual based on its previous trajectory.
+With a high number of flowers in close vicinity to each other, there's a substantial risk of identity switching between flowers as they move among each other (Fig.
+1B).
+Additionally, as the flowers are constrained by their stalk, there is a limit to the distance they can move, which prevents prediction of future location of an individual based on its previous trajectory.
 Here, we introduce the tracking parameter **running mean**, with which we can base tracking on the distance between a point in the current frame and the running mean of the positions of a number of previous points in a track, taking advantage of the fact that a flower moves around a central point.
 Increasing the value for this parameter decreases the weighting of movement between individual frames.
-Setting the parameter to 1 means that tracking is based on the distance to centroid in the previous frame only.
+Setting the parameter to 1 means that tracking is based on the distance to the centroid in the previous frame only.
 
@@
 
-![](../figures/ErrorFigure2.png){width="100%"}
-
-**Figure 1:** Centroid tracking may produce erroneous associations when objects move or disappear between frames.
+![Figure 2: Centroid tracking may produce erroneous associations when objects move or disappear between frames.
 A: A flower temporarily disappears from view.
@@
 C: A flower is linked to the track of another flower.
-As flowers move around a central point, basing tracking on a running mean of previous positions may be beneficial.
+As flowers move around a central point, it is relevant to base tracking on a running mean of previous positions.
 Dots show position of flower in previous frame.
 Crosses show running mean of positions in x \> 1 number of previous frames.
-Dots and crosses coloured according to track identity.
+Dots and crosses coloured according to track identity.](../figures/ErrorFigure2.png){width="100%"}
 
@@
 
-To explore the effect of the three tracking parameters and to identify the optimal combination of parameters for our case of tracking flowers, we applied the tracking algorithm on each of the five image series with every combination of a range of values for each parameter (3179 combinations): 0-160 with a step size of 10 for max gap; 1-160 with a step size of 10 for running mean; 0-1000 with a step size of 100 for max distance.
+To explore the effect of the three tracking parameters and to identify the optimal combination of parameters for our use-case of tracking flowers, we applied the tracking algorithm on each of the five image series with every combination of a range of values for each parameter (3179 combinations): 0-160 with a step size of 10 for max gap; 1-160 with a step size of 10 for running mean; 0-1000 with a step size of 100 for max distance.
 Note that max distance set to zero ignores the parameter altogether.
-We identified the setting(s) that returned the most accurate tracks for each series and compare the tracking results between optimal settings and all parameters ignored (i.e.
-max gap = 0, running mean = 1, max distance = 0).
+We identified the setting(s) that returned the most accurate tracks for each series and compare the tracking results between optimal settings and all parameters ignored (imax gap = 0, running mean = 1, max distance = 0).
 
@@
 The optimal way of quantifying tracking performance depends on the goal of the tracking.
-For example, in our case of tracking flowers, associating information obtained in the images to the individual flower, for example flower visits, would require the track to be as correct as possible (i.e.
-identity shifts should be minimized).
+For example, in our case of tracking flowers, associating information obtained in the images to the individual flower, for example flower visits, would require the track to be as correct as possible (i.e., identity shifts should be minimized).
 On the other hand, to derive flowering length in practice just requires correct association between the first and the last image of a flower while intermediate points can be ignored.
@@
 These should ideally be equal.
-However, we note that a correct number of tracks (i.e.
-number of tracks equal to number of objects) does not necessarily translate to correct tracking, since individual tracks can be erroneously split (fragmentation) by the tracking algorithm while several tracks can be merged into one.
+However, we note that a correct number of tracks (i.e., number of tracks equal to number of objects) does not necessarily translate to correct tracking, since individual tracks can be erroneously split (fragmentation) by the tracking algorithm while several tracks can be merged into one.
 We present the tracking performance without the use of the three tracking parameters along with the best performance (lowest number of mismatches/highest MOTA), and the best performance where the number of tracks correspond to the number of individual flowers in the series.
@@
 we use the midpoint of the line between them.
-For tracks consisting of three points, we establish the triangle from the points and use the centroid of the triangle.
-For tracks consisting of four points or more, we calculate the convex hull of all the points included in the track and use the centroid of this polygon.
-For tracks that contain more than two points but where the points are collinear, we use the midpoint of the line through the points.
+For tracks consisting of three points or more, we calculate the convex hull of all the points included in the track and use the centroid of this polygon.
+For the special case where a track contains more than two points but where the points are collinear, we use the midpoint of the line through the points.
 We hereafter refer to these centroids as track centroids.
@@
 
-We demonstrate the filtering method on the three series for which our tracking algorithm did not return perfect results (i.e.
-automatic tracking identical to ground truth tracks).
+We demonstrate the filtering method on the three series for which our tracking algorithm did not return perfect results (i.e., automatic tracking identical to ground truth tracks).
 To test the filtering approach, we chose a fixed combination of the tracking algorithm parameters across all three series (max gap size: 10; running mean: 10; max distance: 300).
@@
 
-Manual tracking of individual flowers derived from image-based plot monitoring enabled detailed information on flower phenology at the level of individuals with substantial variation in flowering duration between flowers (fig.
-2).
+Manual tracking of individual flowers derived from image-based plot monitoring enabled detailed information on flower phenology at the level of individuals with substantial variation in flowering duration between flowers (Fig.
+3).
 Further, our tracking algorithm was able to track individual flowers with consistently high accuracy (MOTA $\geq$ 0.981) across all series.
-In all cases, all flowers were tracked with more than 80% accuracy (table 1).
+In all cases, all flowers were tracked with more than 80% accuracy (Table 2).
 In fact, two series were tracked without any errors (2018 THUL B and 2018 NARS D).
@@
 This demonstrates the significance of a special case of tracking error, namely merging of tracks, which is not accounted for in the MOTA score.
-In figure 3, we exemplify the two types of errors, identity shifts and merges.
+In Figure 4, we exemplify the two types of errors, identity shifts and merges.
 
@@
 
-![](../figures/OnsetVSLength.png){width="100%"}\
-**Figure 2:** Flowering lifespan (days) for each individual flower in each of the five image series plotted against the onset of flowering in days after 1 January (day of the year) based on manual tracking of individual flowers.
+![Figure 2: Flowering lifespan (days) for each individual flower in each of the five image series plotted against the onset of flowering in days after 1 January (day of the year) based on manual tracking of individual flowers.](images/OnsetVSLengthBox.png){alt="Figure 2: Flowering lifespan (days) for each individual flower in each of the five image series plotted against the onset of flowering in days after 1 January (day of the year) based on manual tracking of individual flowers."}\
 
-**Table 1:** Tracking performance for the automatic flower tracking algorithm for different tracking parameter combinations.
+Figure 3: Variation in flowering lifespan (days) for the flowers of each of the five image series plotted against the onset of flowering in days after 1 January (day of the year, DOY) based on manual tracking of individual flowers.
+
+![Table 2: Tracking performance for the automatic flower tracking algorithm for different tracking parameter combinations.
 Number of tracks in bold shows when the number of tracks equals the number of flowers in the series.
-
-![](../figures/tab_1.png){width="100%"}
+Tracking performance is shown for parameters zeroed (PZ) (Max gap = 0; Max distance = 0; Running mean = 1), minimum number of mismatches achieved (MM), and minimum number of mismatches achieved with a number of tracks equal to the number of flowers in the seris (CF MM).
+The last type is only given if it was not equal to the results for MM.
+When several parameter combinations returned the same results, the range of parameter values included is given for each parameter.](../figures/tab_1.png){width="100%"}
 
@@
 
-**Figure 3:** Tracking of three flowers in the 2018 NARS A image series, coloured by ID assigned by the flower tracking algorithm.
+**Figure 4:** Tracking results of three flowers in the 2018 NARS A image series.
 A: Track output from the flower tracking algorithm for three individual flowers using two different combinations of tracking parameters.
+The y-axis shows the correct track IDs assigned by manual tracking.
+Each track is coloured by the IDs assigned by the flower tracking algorithm.
 In the top panel (max gap = 10, running mean = 1, max distance = 0), flower 15 and 28 have no identity shifts within each track, but the two tracks have been merged into one (see B).
@@
 In the bottom panel, the running mean parameter has been increased to 20 and max distance to 500, which results in correct tracking of all three flowers.
-B: Frame f shows the last frame with flower 15 (blue) annotated because it has wilted.
+B: Merging error between two tracks.
+Frame f shows the last frame with flower 15 (blue) annotated because it has wilted.
 Two frames (5 hours) later (f+2), a new flower (flower 28; blue) blooms.
 Since max gap \> 2 and no distance threshold has been set, flower 28 is associated with the tracks of flower 15, because all other flowers are closer to their respective tracks, which causes a merging error.
-Setting a max distance threshold to 500 fixed this error.
-C: A gust of wind causes a displacement of the flowers between two frames, f and f+1, (7 hours apart), which causes an identity shift between flower 25 and flower 29.
-Setting running mean to 20 fixed this error.
+Setting a max distance threshold to 500 fixed this error as illustrated in A.
+C: Identity shift error between two flowers.
+A gust of wind causes a displacement of the flowers between two frames, f and f+1, (7 hours apart), which causes an identity shift between flower 25 and flower 29.
+Setting running mean to 20 fixed this error as illustrated in A.
 
 With our filtering method, we were able to identify tracks with no mismatch errors in all three series that were not perfectly tracked.
-The parameter value for filtering distance that returned the highest number of tracks at the lowest number of mismatches was 350 (results for all values given in supp.info).
-The results of the filtering algorithm on the three series are given in table 2 and the filtering process is visualised in fig.
-4.
+The parameter value for filtering distance that returned the highest number of tracks at the lowest number of mismatches was 350 (results for all values given in suppl.
+2).
+The results of the filtering algorithm on the three series are given in Table 3 and the filtering process is visualised in Figure 5.
 Our filtering method extracted 28 tracks from the three series (2018 NARS A: 15 tracks; 2019 NARS C: 10 tracks; 2019 NYAA E: 3 tracks) with complex scenes using a fixed filtering distance value of 350 and fixed values for the tracking parameters (10, 300, 10 for max gap size, max distance, and running mean, respectively) without any mismatches included in the 28 tracks.
 
-**Table 2:** Results of the filtering algorithm on tracking results using the parameters 10, 300, 10 for max gap size, max distance, and running mean, respectively.
+**Table 3:** Results of the filtering algorithm on tracking results using the parameters 10, 300, 10 for max gap size, max distance, and running mean, respectively.
 The filtering distance value of 350 was used for all three series.
@@
 
-![](../figures/filtering1.png){width="100%"}
-
-**Figure 4:** The track filtering process exemplified by the 2019 NARS C series.
-The first column shows the centroid points for all bounding boxes in the given series, coloured by the track id from the centroid tracking algorithm.
-The second column shows the track polygons calculated from the tracks.
-Polygons with a distance below eps = 350 between them have been grouped together.
+![Figure 5: The track filtering process exemplified by the 2019 NARS C series.
+A: The centroid points for all bounding boxes in the series, coloured by the track id from the centroid tracking algorithm.
+B: The track convex hull polygons calculated from the tracks.
+Polygons with a euclidean distance below 350 between them have been grouped together.
 The polygons have been coloured according to group.
-Third column shows the results of the filtering where all tracks that were not sufficiently spatially isolated were removed.
+C: The results of the filtering where all tracks that were not sufficiently spatially isolated were removed, resulting in 10 remaining tracks without tracking errors.](images/filterFigLet.png)
 
@@
 In this paper, we have demonstrated the potential of image-based monitoring for collecting information on flower phenology at the level of individuals.
-Our results show substantial variation in flowering duration between individual flowers (fig.
-2).
-In fact, some flower tracks spanned less than a few days.
+The automatic tracking and filtering pipeline presented was able to return reliable flower tracks despite settings that violate assumptions of standard tracking algorithms and complex scenes with a high number of objects that are not permanent within the series.
+Our tracking algorithm was capable of accurately tracking individual flowers, despite the difficult setting.
+Even with simple centroid tracking where association is based solely on displacement between the current and previous frame, tracking performance was high for all series and perfect for two.
+Further, with our filtering approach based on distance between tracks, we were able to discard the least reliable tracks and thereby increase the mean tracking accuracy of the remaining flowers to a perfect score.
+
+Our results show substantial variation in flowering duration between individual flowers.
+In fact, flower longevity was less than a few days in some cases.
 As flowers may not be visible within the frame for their full lifespan, it is likely that for some flowers the full flowering has not been captured, although flowering durations of less than 24 hours has been observed from manual examination of the image series (Mann, personal observation).
 
-The automatic tracking and filtering pipeline presented was able to return reliable flower tracks despite settings that violate assumptions of standard tracking algorithms and complex scenes with a high number of objects that are not permanent within the series.
-Our tracking algorithm was capable of consistently accurate flower tracking, despite the difficult setting.
-Even with simple centroid tracking with association based solely on displacement between the current and previous frame, tracking performance was high for all series and perfect for two (table 1).
-Further, with our filtering approach based on distance between tracks, we were able to discard the least reliable tracks and thereby increase the mean tracking accuracy of the remaining flowers to a perfect score (table 2).
-
-Most multiple object tracking literature focus on tracking of humans [@luo2021; @milan2016], and the methods developed here are not necessarily directly transferable to other use-cases such as ours.
-Many tracking algorithms, such as the Simple online and real-time tracking (SORT) [@sort], use Kalman filtering to predict the future location of an object by estimating its velocity from previous frames [@kalmann].
-However, Kalman filtering assumes linear motion of the object and does not handle abrupt motion well.
-As the flowers are constrained in movement by their stalk, their movement does not follow this assumption.
-Other tracking approaches, such as DeepSORT [@deepsort2017], use tracking-by-detection in which CNNs are applied to leverage appearance information in order to continuously distinguish individuals through time, thus aiding the tracking algorithm.
-In our case, however, flowers are not clearly distinguishable and one flower may look less alike between two frames than two flowers in a single frame.
-Finally, some tracking methods require manual initiation of a fixed number of tracks for objects in the first frame and the algorithm will then track these objects in the subsequent frames [@hu2012].
-These methods do not allow new objects to enter the frame within the series.
-In our case, flowers can disappear permanently from the frame, for example when flowers wilt, in which case the track should be terminated.
-Oppositely, new objects can continuously appear through the series when new flowers bloom or previously unseen flowers enter the frame.
-Additionally, objects may disappear temporarily from view and reappear at some later stage.
-Further, as the manual tracking of flowers showed, the lifespans of flowers vary substantially (fig.
-2), which means that assumptions on flowering duration cannot be leveraged during tracking.
-These settings vastly complicate the task of tracking multiple objects [@betke2007].
-Our tracking algorithm returned consistently accuracy despite this.
-
 The two series for which the tracking algorithm produced perfect results were also the series with the lowest number of flowers.
-While an increased number of objects does not necessarily increase tracking difficulty as this depends on the degree of isolation of tracks in time an space [@beard2018], multiple object tracking is a challenging task.
+While an increased number of objects does not necessarily increase tracking difficulty as this depends on the degree of isolation of tracks in time and space [@beard2018], multiple object tracking is a challenging task.
 Here, since the flowers are limited in space and time, the complexity of tracking likely increases with increasing number of flowers.
-A higher number of flowers within the frame increases the possibility of flowers in close vicinity to each other and of flowers overlapping and moving in and out of frame translating to a higher risk of making tracking errors.
+A higher number of flowers within the frame increases the possibility of flowers appearing in close proximity to each other and that flowers overlap or move in and out of frame.
+All these situations translate to a higher risk of tracking errors.
 The very high complexity in the scenes with many flowers, especially 2019 NARS C with 85 flowers, coupled with the low frame rate renders it unlikely that any tracking method could produce perfect results across series.
@@
 For example, with many image series available, series with less than a given maximum of flowers detected in a single frame over a season could be selected for subsequent tracking.
-Tracks in series that exceeded this threshold could be filtered subsequent of tracking, as demonstrated here, or selected for manual validation of results.
+Tracks in series that exceeded this threshold could be filtered after tracking, as demonstrated here, or selected for manual validation of results.
 
-Our results show that optimizing for high MOTA tends to lead to an underestimate of the number of tracks (i.e.
-merging of tracks occurs) (table 1).
+Our results show that optimizing for high MOTA tends to lead to an underestimate of the number of tracks (i.e., merging of tracks occurs) (Table 2).
 This may happen, for example, because of the max gap parameter.
@@
 The optimal values for the running mean for 2019 NYAA E were exceptionally high, compared to the other series.
-Interestingly, the average flower duration was also higher for this series (fig.
-2) and this could be part of the explanation, as flowers with longer lifespans can also have larger gaps in their tracks.
+Interestingly, the average flower duration was also higher for this series and this could be part of the explanation, as longer lifespan allow for larger gaps in the tracks.
 Overall, the max gap parameter improved tracking accuracy with values of 10-20 being optimal.
-The three series with complex scenes all contain flowers at the edge of the frame (see fig.
+The three series with complex scenes all contain flowers at the edge of the frame (see Fig.
 1), which means that they may occasionally be outside of view of the camera.
@@
 Correct tracking of these objects requires that an object can disappear from view without the track instantly being lost, which the max gap size allows.
-Oppositely, as discussed, this setting may also lead to merging error.
+Oppositely, this setting may also lead to merging error.
 The optimal range for max distance was relatively narrow (400-600) for the parameters producing the correct number of tracks with the lowest number of mismatches.
 The optimal value for this setting is one that is sufficiently high that it does not force individual flowers into separate tracks because of movement between frames yet is low enough to separate any flowers into individual tracks that would otherwise have been grouped together.
-This explains well why the extreme values for this parameter was not included in the highest accuracy tracks.
+This explains why the extreme values for this parameter was not included in the highest accuracy tracks.
 
@@
 However, the fact that the optimal tracking parameters were overall similar across series strengthens the assumptions that a set of standard values can produce reasonable results for other similar image series with *Dryas* flowers.
-Here, a combination such as (10, 500, 10, for max gap, max distance, and running, respectively) would be a good starting point.
+Here, a combination such as 10, 500, 10, for max gap, max distance, and running, respectively, would be a good starting point.
 When applying our method in a new setting, we recommend performing a parameter test as demonstrated here on a subset of ground truth tracks to derive the optimal set of parameters.
@@
 Thus, the method can be directly applied on such data.
-Automatically detecting flowers would likely introduce a proportion of false negatives when the detection algorithm fails to detect a present object.
+Automatically detecting flowers would likely introduce a proportion of false negatives when the detection algorithm fails to detect an object.
 The flowers in our data that periodically move out of the frame or are temporarily occluded mimics false negatives.
@@
 Automatic detection methods may also introduce false positives.
-These detections would be included in the tracking on par with correct detections.
+These detections would be included in the tracking along with correct detections.
 Some cases of false positives, for example isolated single false positives, could be removed automatically while other cases may require manual quality control.
@@
 Extracting tracks that are spatially isolated does not guarantee that the tracks are correct/without errors.
-For example, tracks of several objects may have been erroneously merged or a track could contain a small proportion of instances of an object far away, which would not have a large effect on the track centroid position (fig.
-2).
+For example, tracks of several objects may have been erroneously merged or a track could contain a small proportion of instances of an object far away, which would not have a large effect on the track centroid position.
 However, as spatially isolated objects in general are easier to track, it is safe to assume that these tracks are more reliable.
@@
 For example, if it is given a priori that an object will always appear in a minimum of two frames, then single point tracks can be filtered out.
-However, when such a priori knowledge is not accessible, a conservative approach, as the one we present, is preferable.
+However, when such a priori knowledge is not available, a conservative approach, as the one we present, is preferable.
 
@@
 However, this would remove tracks with even minor overlap and put weight on tracks with many points as these typically cover a larger area.
-Another alternative could be to base filtering on the centroids of the points within a track (i.e.
-mean x and mean y).
+Another alternative could be to base filtering on the centroids of the points within a track (i.e., mean x and mean y).
 The presented method avoids weighting track geometries based on number of included points and avoids weighting the centroid coordinate based on the point coordinates within the track as the track centroid is based on the convex hull of the track points.
@@
 The possibility of collecting individual level flower phenology is highly relevant for exploration of many research questions.
-For example, it would be possible to assign indicators of reproductive success and factors driving variation in seed set to the individual.
-With this, it could be explored how seed set is affected by timing and length of flowering and/or abiotic conditions such as frost events.
+For example, it would be possible to assign indicators of seed set to the individual and explore how seed set is affected by timing and length of flowering and/or abiotic conditions such as frost events.
 Hereby, the long-standing question of what drives variation in reproductive success of individual flowers could be explored [@bolmgren2015].
-Further, the fact that it was not uncommon in our results that flowering lasted less than a week (fig.
-1), emphasizes that weekly observations of sample plots can be insufficient for comprehensive description of flower phenology [@Prevey2021].
+Further, the fact that it was not uncommon in our results that flowering lasted less than a week, emphasizes that weekly observations of sample plots can be insufficient for comprehensive description of flower phenology [@Prevey2021].
 
 By simultaneously tracking flower visits, these could be assigned to the individual flower and visitation rates per flower could be derived and the effects on reproductive success could be investigated [@tiusanen2016; @rafferty2012].
-Pollen limitation is a common factor limiting the reproductive success of flowering plants, but the ecological mechanisms behind pollen limitation is poorly understood, at least partly because it requires information on plant reproductive success in relation to pollinator visitation rates which is difficult to acquire in a natural setting and to a satisfactory extent.
+Pollen limitation is a common factor limiting the reproductive success of flowering plants.
+However, the ecological mechanisms behind pollen limitation is poorly understood, at least partly because it requires information on plant seed set in relation to pollinator visitation rates which is difficult to acquire in a natural setting and to a satisfactory extent.
 For example, pollinator visits can vary greatly in frequency during the flowering season and the degree of pollen limitation can vary even between populations in close vicinity to each other [@larson1999; @ashman2004].
@@
 Importantly, the variables suggested here could be derived directly from the time-lapse images.
-For example, flower reproductive success could be estimated by observation of seed set for the individual flower and flower visitation rates could be quantified by detecting insects in flowers in high temporal resolution image series.
+For example, seed set as an indicator of flower reproductive success could be estimated for individual flowers and flower visitation rates could be quantified by detecting insects in flowers in high temporal resolution image series.
 
-It is clear that the images derived from monitoring of flower plots can hold substantial amounts of ecologically relevant data.
+It is clear that the images derived from monitoring of permanent plots can hold substantial amounts of ecologically relevant data.
 To derive some variables, for example flower visitation rates, image series collected at high temporal frequency is required.
@@
 For instance, it has been demonstrated that convolutional neural networks can facilitate monitoring of pollination events by detecting, identifying, and counting insects [@bjerge2021] and perform fine-grained species classification [@spiesman2021].
-Collectively, these developments coupled with implementation of long-term high-frequency image-based monitoring will allow for collection and processing of ecological data at unprecedented scales.
+Collectively, these developments coupled with implementation of long-term high-frequency image-based monitoring will allow for collection and processing of ecological data at unprecedented scales [@hoye2021].
+For flower tracking in particular, our results emphasize that this is not a trivial task.
+The data violate common assumptions of tracking methods and tracking individual flowers through time-lapse images therefore represents a special case [@chen2022; @betke2007; @luo2021].
+For example, since flower movement is constrained by the stalk, the future location of a flower-head cannot be predicted based on previous trajectory of movement.
+As individual flowers closely resemble each other, tracking cannot be based on recognition of individuals [@chen2022; @deepsort2017].
+Further, as the manual tracking of flowers showed, the lifespans of flowers vary substantially, which means that assumptions on flowering duration cannot be leveraged during tracking.
+These settings vastly complicate the task of tracking multiple objects [@betke2007].
+Our tracking algorithm returned consistently accuracy despite this.
 
-For flower tracking in particular, our results emphasize that this is not a trivial task.
 While we show that tracking can be done at high accuracy with our tracking algorithm, we also identify particular challenges with flower tracking, such as the violations of standard assumptions in tracking algorithms, the potentially large number of individuals overlapping in time and space, and the possible susceptibility to track merging.
 To improve tracking ability of flowers in future studies, focus should be on implementing reliable measures of different types of errors and on developing methods that can alleviate the challenges, as well as on comparison of different tracking approaches.
+Alongside this, methods for extracting auxiliary information from images, such as flower seed set, should be explored.
 
@@
 E.
-Loonen for set-up and maintenance of cameras in Ny-Ålesund and Peter Akers and Jeffrey M.
-Welker for set-up and maintenance of cameras in Thule.
+Loonen for set-up and maintenance of cameras in Ny-Ålesund, Peter Akers and Jeffrey M.
+Welker for set-up and maintenance of cameras in Thule, and Rebekka Ween and Line Klausen for assistance during fieldwork in Narsarsuaq.
 TTH acknowledges funding from Independent Research Fund Denmark Grant 8021-00423B.
 The authors have no conflict of interest to declare.
-`r word_comment("Hjalte: Toke, please check if this is sufficient.
-Should anyone from the Narsarsuaq fieldwork be mentioned?
-Anyone else?", "")`
 
@@
 HMRM conceived and developed the method with contributions from TTH and AI.
+HMRM wrote all code.
 HMRM wrote the manuscript with contributions from TTH and AI.

### 2022-06-11 code/R/Analysis_flowerLifespandAndTemperature.Rmd
Updates

--- 2022-02-02 code/R/Analysis.Rmd
@@
 ---
-title: "Analysis"
+title: "flowerLifespanAndTemperature"
 author: "Hjalte Mann"
-date: "25/11/2021"
-output: html_document
----
----
-title: "IndividualPhenology"
-author: "Hjalte Mann"
-date: "2/2/2021"
+date: '2022-06-10'
 output: html_document
@@
 ```{r}
+all = bind_rows(
+thul01_meta[,c("FileName", "flower_start","flower_length", "CameraID", "Temperature", "DOY", "Class")],     nars04_meta[,c("FileName", "flower_start","flower_length", "CameraID", "Temperature", "DOY", "Class")], nars13_meta[,c("FileName", "flower_start","flower_length", "CameraID", "Temperature", "DOY", "Class")], nyaa04_meta[,c("FileName", "flower_start","flower_length", "CameraID", "Temperature", "DOY", "Class")], nars17_meta[,c("FileName", "flower_start","flower_length", "CameraID", "Temperature", "DOY", "Class")])
 
-all = bind_rows(thul01_meta[,c("flower_start","flower_length", "CameraID")], nars04_meta[,c("flower_start","flower_length", "CameraID")], nars13_meta[,c("flower_start","flower_length", "CameraID")], nyaa04_meta[,c("flower_start","flower_length", "CameraID")], nars17_meta[,c("flower_start","flower_length", "CameraID")])
+allT = all %>% 
+  distinct(FileName, .keep_all = T) %>% 
+  group_by(CameraID, DOY) %>%
+  mutate(minT = min(Temperature), maxT = max(Temperature))
 
-# Camera codes:
-# 2018 NARS-04: 2018 NARS A
-# 2018 THUL-01: 2018 THUL B
-# 2019 NARS-13: 2019 NARS C
-# 2019 NARS-17: 2019 NARS D
-# 2019 NYAA-04: 2019 NYAA E
+allT = allT %>% 
+  group_by(CameraID, DOY) %>% 
+  mutate(meanDailyT = (minT + maxT)/2)
 
+allT = allT %>% 
+  group_by(CameraID, Class) %>% 
+  mutate(meanT = mean(meanDailyT))
+
+allT %>% 
+  distinct(Class, .keep_all = T) %>% 
+  ggplot() + 
+  geom_boxplot(aes(x = meanT, y = flower_length, colour = CameraID))
+
+allT %>% 
+  distinct(Class, .keep_all = T) %>% 
+  ggplot() + 
+  geom_point(aes(x = meanT, y = flower_length, colour = CameraID))
+
+allT %>% 
+  distinct(Class, .keep_all = T) %>% 
+  ggplot() + 
+  geom_boxplot(aes(x = flower_start, y = meanT, colour = CameraID))
+
+```
+
+```{r}
 all = all %>% 
@@
 
-ggsave("U:/BITCue/Projekter/TrackingFlowers/figures/OnsetVSLength.png", f1, width = 15, height = 10)
-
-all %>% 
-ggplot()+
-  geom_point(aes(x = flower_start, y = flower_length, colour = CameraID))# +
-#geom_smooth(aes(x = flower_start, y = flower_length), method=lm)
-
-t.test(nyaa04_meta$flower_length, nars13_meta$flower_length)
+#ggsave("U:/BITCue/Projekter/TrackingFlowers/figures/OnsetVSLength.png", f1, width = 15, height = 10)
 
 ```
-
-```{r}
-all %>% 
-ggplot()+
-  geom_point(aes(x = flower_start, y = flower_length, colour = CameraID))
-```
-
-# parameterTest_3
-
-## NYAA 04
-
-```{r fig.height=4}
-d1 = read.csv("U:/BITCue/Projekter/TrackingFlowers/testResults/parameterTest_3_NYAA-04_Evaluate.csv")
-
-NYAA04_p1 = d1 %>% 
-  ggplot() +
-  geom_point(aes(x=runMean, y = maxDisap, size = mm, colour = mm)) + 
-  ggtitle("NYAA-04 Parameter combination") +
-  facet_wrap(vars(maxDist))+scale_color_gradient(low="blue", high="red")
-
-NYAA04_p2 = d1 %>% 
-  arrange(mm, ascending = T) %>% 
-  slice_head(n = 10) %>% 
-  ggplot() +
-  geom_point(aes(x=runMean, y = maxDisap, size = mm, colour = mm)) + 
-  ggtitle("NYAA-04 Ten Lowest Mismatches") +
-  facet_wrap(vars(maxDist))+scale_color_gradient(low="blue", high="red")
-
-NYAA04_p1
-NYAA04_p2
-
-ggsave('U:/BITCue/Projekter/TrackingFlowers/testResults/figures/NYAA-04_AllCombinations.png', NYAA04_p1)
-ggsave('U:/BITCue/Projekter/TrackingFlowers/testResults/figures/NYAA-04_TenLowest.png', NYAA04_p2)
-
-```
-
-```{r}
-# NYAA-04 gt:
-nyaa04_gt = read.csv("U:/BITCue/Projekter/TrackingFlowers/data/annotations/2021_12_13_NorwayAnnotations_NYAA-04_IndividualAnnotations_FRCNN_Metrics.csv")
-
-length(unique(nyaa04_gt$id_gt))
-
-```
-```{r}
-# Minimum mm: 10,300,60-70,29,0.9835414,38
-# All zero: 0,0,0,68,0.9614075,31
-# Correct noo with minimum mm:  70-160,800-1000,10,36,0.9795687,21
-
-nyaa04minmm = read.csv("U:/BITCue/Projekter/TrackingFlowers/testResults/_parameterTest_NYAA-04_3/parameterTest_NYAA-04_maxDisap_10_runMean_60_maxDist_300.csv")
-
-allzero = read.csv("U:/BITCue/Projekter/TrackingFlowers/testResults/_parameterTest_NYAA-04_3/parameterTest_NYAA-04_maxDisap_0_runMean_0_maxDist_0.csv")
-
-cornoo = read.csv("U:/BITCue/Projekter/TrackingFlowers/testResults/_parameterTest_NYAA-04_3/parameterTest_NYAA-04_maxDisap_70_runMean_10_maxDist_800.csv")
-
-```
-
-```{r fig.height=4}
-
-nyaa04minmm %>% 
-  ggplot() +
-  geom_point(aes(x = x_c, y = y_c, color = as.factor(objectID)), size = 1) + 
-  xlim(0,6080) +
-  ylim(0,3420) + 
-  ggtitle("NYAA-04 Tracked (Minimum mm)") +
-  theme(legend.position = "none")
-
-nyaa04xx = read.csv("U:/BITCue/Projekter/TrackingFlowers/testResults/_parameterTest_NYAA-04_3/parameterTest_NYAA-04_maxDisap_10_runMean_10_maxDist_300.csv")
-
-nyaa04xx %>% 
-  ggplot() +
-  geom_point(aes(x = x_c, y = y_c, color = as.factor(objectID)), size = 1) + 
-  xlim(0,6080) +
-  ylim(0,3420) + 
-  ggtitle("NYAA-04 Tracked (xx)")  +
-  theme(legend.position = "none")
-
-nyaa04 %>% 
-  ggplot() +
-  geom_point(aes(x = x_c, y = y_c, color = as.factor(Class)), size = 1) + 
-  xlim(0,6080) +
-  ylim(0,3420) + 
-  ggtitle("NYAA-04 Ground Truth") +
-  theme(legend.position = "none")
-
-#nyaa04_meta %>% 
-#  ggplot() +
-#  geom_point(aes(x = DOY, y = Class))
-
-```
-
-## THUL 01
-
-```{r fig.height=4}
-d2 = read.csv("U:/BITCue/Projekter/TrackingFlowers/testResults/parameterTest_3_THUL-01_Evaluate.csv")
-
-THUL01_p1 = d2 %>% 
-  ggplot() +
-  geom_point(aes(x=runMean, y = maxDisap, size = mm, colour = mm)) + 
-  ggtitle("THUL-01 Parameter combination") +
-  facet_wrap(vars(as.factor(maxDist)))
-
-THUL01_p2 = d2 %>% 
-  arrange(mm, ascending = T) %>% 
-  filter(mm == 0) %>% 
-  ggplot() +
-  geom_point(aes(x=runMean, y = maxDisap, size = mm)) + 
-  ggtitle("THUL-01 All zero mismatches") +
-  facet_wrap(vars(maxDist))+scale_color_gradient(low="blue", high="red")
-
-THUL01_p1
-THUL01_p2
-
-ggsave('U:/BITCue/Projekter/TrackingFlowers/testResults/figures/THUL-01_AllCombinations.png', THUL01_p1)
-ggsave('U:/BITCue/Projekter/TrackingFlowers/testResults/figures/THUL-01_AllZero.png', THUL01_p2)
-```
-
-```{r}
-# NYAA-04 gt:
-thul01_gt = read.csv("U:/BITCue/Projekter/TrackingFlowers/data/annotations/2020_05_15_NorwayAnnotations_THUL-01_IndividualAnnotations_FRCNN_Metrics.csv")
-
-length(unique(thul01_gt$id_gt))
-
-```
-```{r}
-# Minimum mm: 0-160, 0 + 200-1000, 10, 0, 1, 6
-# All zero: 0, 0, 0, 8, 0.9885551, 6
-# Correct noo with minimum mm: 0-160, 0 + 200-1000, 10, 0, 1, 6
-
-minmm = read.csv("U:/BITCue/Projekter/TrackingFlowers/testResults/_parameterTest_THUL-01_3/parameterTest_THUL-01_maxDisap_0_runMean_10_maxDist_0.csv")
-
-allzero = read.csv("U:/BITCue/Projekter/TrackingFlowers/testResults/_parameterTest_THUL-01_3/parameterTest_THUL-01_maxDisap_0_runMean_0_maxDist_0.csv")
-
-cornoo = read.csv("U:/BITCue/Projekter/TrackingFlowers/testResults/_parameterTest_THUL-01_3/parameterTest_THUL-01_maxDisap_0_runMean_10_maxDist_0.csv")
-
-```
-
-```{r fig.height=4}
-
-cornoo %>% 
-  ggplot() +
-  geom_point(aes(x = x_c, y = y_c, color = as.factor(objectID)), size = 1) + 
-  xlim(0,6080) +
-  ylim(0,3420) + 
-  ggtitle("THUL-01 Tracked") +
-  theme(legend.position = "none")
-
-thul01 %>% 
-  ggplot() +
-  geom_point(aes(x = x_c, y = y_c, color = as.factor(Class)), size = 1, shape = 1) + 
-  xlim(0,6080) +
-  ylim(0,3420) + 
-  ggtitle("THUL-01 Ground Truth") +
-  theme(legend.position = "none")
-```
-
-## NARS-04
-
-```{r fig.height=4}
-
-d3 = read.csv("U:/BITCue/Projekter/TrackingFlowers/testResults/parameterTest_3_NARS-04_Evaluate.csv")
-
-NARS04_p1 = d3 %>% 
-  ggplot() +
-  geom_point(aes(x=runMean, y = maxDisap, size = mm, colour = mm)) + 
-  ggtitle("NARS-04 Parameter combination") +
-  facet_wrap(vars(as.factor(maxDist))) +
-  scale_color_gradient(low="blue", high="red")
-
-NARS04_p2 = d3 %>% 
-  arrange(mm, ascending = T) %>% 
-  slice_head(n = 10) %>% 
-  ggplot() +
-  geom_point(aes(x=runMean, y = maxDisap, size = mm)) + 
-  ggtitle("NARS-04 Ten lowest mismatches") +
-  facet_wrap(vars(maxDist))+scale_color_gradient(low="blue", high="red")
-
-NARS04_p1
-NARS04_p2
-
-#ggsave('U:/BITCue/Projekter/TrackingFlowers/testResults/figures/NARS-04_AllCombinations.png', NARS04_p1)
-#ggsave('U:/BITCue/Projekter/TrackingFlowers/testResults/figures/NARS-04_TenLowest.png', NARS04_p2)
-```
-
-```{r}
-# NARS-04 gt:
-NARS04_gt = read.csv("U:/BITCue/Projekter/TrackingFlowers/data/annotations/2020_05_17_NorwayAnnotations_NARS-04_IndividualAnnotations_FRCNN_Metrics.csv")
-
-length(unique(NARS04_gt$id_gt))
-
-```
-
-```{r}
-# Minimum mm: 10, 500, 10, 20, 0.9883991, 32
-# All zero: 0, 0, 0, 85, 0.9506961, 31
-# Correct noo with minimum mm: 10, 500, 10, 20, 0.9883991, 32
-
-nars04minmm = read.csv("U:/BITCue/Projekter/TrackingFlowers/testResults/_parameterTest_NARS-04_3/parameterTest_NARS-04_maxDisap_10_runMean_10_maxDist_500.csv")
-
-allzero = read.csv("U:/BITCue/Projekter/TrackingFlowers/testResults/_parameterTest_NARS-04_3/parameterTest_NARS-04_maxDisap_0_runMean_0_maxDist_0.csv")
-
-cornoo = read.csv("U:/BITCue/Projekter/TrackingFlowers/testResults/_parameterTest_NARS-04_3/parameterTest_NARS-04_maxDisap_10_runMean_10_maxDist_500.csv")
-
-```
-
-```{r fig.height=4}
-
-nars04minmm %>% 
-  ggplot() +
-  geom_point(aes(x = x_c, y = y_c, color = as.factor(objectID)), size = 1) + 
-  xlim(0,6080) +
-  ylim(0,3420) + 
-  ggtitle("NARS-04 Tracked") +
-  theme(legend.position = "none")
-
-nars04 %>% 
-  ggplot() +
-  geom_point(aes(x = x_c, y = y_c, color = as.factor(Class)), size = 1) + 
-  xlim(0,6080) +
-  ylim(0,3420) + 
-  ggtitle("NARS-04 Ground Truth") +
-  theme(legend.position = "none")
-```
-
-## NARS-13
-
-```{r}
-d4 = read.csv("U:/BITCue/Projekter/TrackingFlowers/testResults/parameterTest_3_NARS-13_Evaluate.csv")
-
-NARS13_p1 = d4 %>% 
-  ggplot() +
-  geom_point(aes(x=runMean, y = maxDisap, size = mm, colour = mm)) + 
-  ggtitle("NARS-13 Parameter combination") +
-  facet_wrap(vars(as.factor(maxDist))) +
-  scale_color_gradient(low="blue", high="red")
-
-NARS13_p2 = d4 %>% 
-  arrange(mm, ascending = T) %>% 
-  slice_head(n = 10) %>% 
-  ggplot() +
-  geom_point(aes(x=runMean, y = maxDisap, size = mm)) + 
-  ggtitle("NARS-13 Ten lowest mismatches") +
-  facet_wrap(vars(maxDist))+scale_color_gradient(low="blue", high="red")
-
-NARS13_p1
-NARS13_p2
-
-ggsave('U:/BITCue/Projekter/TrackingFlowers/testResults/figures/NARS-13_AllCombinations.png', NARS13_p1)
-ggsave('U:/BITCue/Projekter/TrackingFlowers/testResults/figures/NARS-13_TenLowest.png', NARS13_p2)
-
-```
-
-```{r}
-d5 = read.csv("U:/BITCue/Projekter/TrackingFlowers/testResults/parameterTest_3_NARS-17_Evaluate.csv")
-
-NARS17_p1 = d5 %>% 
-  ggplot() +
-  geom_point(aes(x=runMean, y = maxDisap, size = mm, colour = mm)) + 
-  ggtitle("NARS-17 Parameter combination") +
-  facet_wrap(vars(as.factor(maxDist))) +
-  scale_color_gradient(low="blue", high="red")
-
-NARS17_p2 = d5 %>% 
-  arrange(mm, ascending = T) %>% 
-  filter(mm == 0) %>% 
-  ggplot() +
-  geom_point(aes(x=runMean, y = maxDisap, size = mm)) + 
-  ggtitle("NARS-17 All zero mismatches") +
-  facet_wrap(vars(maxDist))+scale_color_gradient(low="blue", high="red")
-
-NARS17_p1
-NARS17_p2
-
-ggsave('U:/BITCue/Projekter/TrackingFlowers/testResults/figures/NARS-17_AllCombinations.png', NARS17_p1)
-ggsave('U:/BITCue/Projekter/TrackingFlowers/testResults/figures/NARS-17_TenLowest.png', NARS17_p2)
-
-```
-
-```{r}
-# NARS-13 gt:
-NARS13_gt = read.csv("U:/BITCue/Projekter/TrackingFlowers/data/annotations/2020_04_30_NorwayAnnotations_NARS-13_IndividualAnnotations_FRCNN_Metrics.csv")
-
-length(unique(NARS13_gt$id_gt))
-
-```
-
-```{r}
-# Minimum mm: 10-20, 300-400, 10, 78, 0.9840393, 81+85+93+97
-# All zero: 0, 0, 0, 267, 0.9453653, 88
-# Correct noo with minimum mm: 10, 400, 10, 78, 0.9840393, 85
-
-nars13minmm = read.csv("U:/BITCue/Projekter/TrackingFlowers/testResults/_parameterTest_NARS-13_3/parameterTest_NARS-13_maxDisap_0_runMean_0_maxDist_0.csv")
-
-allzero = read.csv("U:/BITCue/Projekter/TrackingFlowers/testResults/_parameterTest_NARS-13_3/parameterTest_NARS-13_maxDisap_0_runMean_0_maxDist_0.csv")
-
-cornoo = read.csv("U:/BITCue/Projekter/TrackingFlowers/testResults/_parameterTest_NARS-13_3/parameterTest_NARS-13_maxDisap_10_runMean_10_maxDist_0.csv")
-
-```
-
-```{r fig.height=4}
-
-nars13minmm %>% 
-  ggplot() +
-  geom_point(aes(x = x_c, y = y_c, color = as.factor(objectID)), size = 1) + 
-  xlim(0,6080) +
-  ylim(0,3420) + 
-  ggtitle("NARS-13 Tracked (Minimum mm)") +
-  theme(legend.position = "none")
-
-nars13 %>% 
-  ggplot() +
-  geom_point(aes(x = x_c, y = y_c, color = as.factor(Class)), size = 1) + 
-  xlim(0,6080) +
-  ylim(0,3420) + 
-  ggtitle("NARS-13 Ground Truth") +
-  theme(legend.position = "none")
-
-length(unique(nars13$Class))
-```
-
-```{r}
-tempNARS04 = read.csv("U:/BITCue/Projekter/TrackingFlowers/testResults/tempNARS04.csv")
-
-tempNARS04 %>% 
-  filter(x_c > 1100 & x_c < 1110) %>% 
-  ggplot() +
-  geom_point(aes(x = x_c, y = y_c, colour = as.factor(objectID))) + 
-  xlim(1100,1110) + 
-  ylim(2250,2320)
-
-tempNARS04 %>% 
-  ggplot() +
-  geom_point(aes(x = x_c, y = y_c, colour = as.factor(objectID)), size = 0.5)
-
-```
-Largest gap in time
-
-```{r}
-x = nyaa04_meta %>%
-    group_by(Class) %>%
-    arrange(time) %>%
-    mutate(diff = time - lag(time, default = first(time)))
-
-x = nars13_meta %>%
-    group_by(Class) %>%
-    arrange(time) %>%
-    mutate(diff = time - lag(time, default = first(time)))
-
-# NYAA-04: 5.1
-# NARS-13: 1.25
-
-```
-
-Largest gap in number of frames
-
-```{r}
-
-x = nyaa04 %>% 
-  mutate(frame = as.integer(str_remove(str_extract(FileName, regex("\\d{6}")), "^0+"))) %>% 
-  group_by(Class) %>%
-  arrange(frame) %>%
-  mutate(diff = frame - lag(frame, default = first(frame)))
-
-```

### 2022-06-11 code/run.py

@@ -9,7 +9,7 @@ import pandas as pd
 from track import tracker
 #from filtering import sieve
 from DBSCANFiltering import DBSCANsieve
-from distanceFiltering import distanceSieve
+#from distanceFiltering import distanceSieve
 #from evaluate import evaluator
 #import os
 import time
@@ -269,8 +269,8 @@ def MOTMetrics(dt, gt,  site):
 
 #Find a good value for eps
 
-epslist = [50,100,150,200,250,300,350,400,450,500] # [350]
-#epslist = [350]
+#epslist = [50,100,150,200,250,300,350,400,450,500] # [350]
+epslist = [350]
 results_filename = '../testResults/MMFix/temp_epstest.csv'
 
 temppath = '../testResults/MMFix/temp_epsExperiment_maxDisap_10_runMean_10_maxDist_300.csv'

### 2022-06-11 code/trackForPrint.py

@@ -1,0 +1,276 @@
+# -*- coding: utf-8 -*-
+"""
+Custom multi flower tracker
+Created 2021
+
+@author:
+Hjalte M. R. Mann
+TECOLOGY.xyz
+
+Detections are in the format: x_min, y_min, x_max, y_max
+
+Max distance: Threshold for maximum distance between point and track before initiation of a new track is forced
+Max dissapeared: Max number of frames a track can be disappeared before the track is terminated and a new point in that area are considered new tracks
+Running mean: Number of previous frame on which to calculate track position (to calculate distance between point and track). If set to one, position is centroid of previous track, if above one, means of x and y are calculated.
+
+
+Pseudocode to 
+
+-Go to next frame
+    -If the frame has no detections
+        -Add one to disappeared counter for any existing tracks
+    -If the frame has detections
+        -If we are not tracking objects
+            -Initiate tracks for all points
+        -If we are tracking objects, calculate all pairwise distances (If running mean above 1, we use the mean x and y position of the track, else we just use position in the previous frame)
+            -For all distances below max distance threshold, try to associate points and tracks (shortest distance from point to object get associated and removed, then the second shortest distance etc.)
+            -For points with distances above max distance
+                -Initiate new tracks
+                -Add one to track id counter
+        -For points that did not get associated to an exisiting track
+            -Initiate new tracks
+            -Add one to track id counter
+        -Add frame tracking data to result dataframe for final output
+        -Update running mean dictionary with new data
+            -For tracks that are at the running mean threshold, remove oldest position and add current
+        -For tracks that did not get a point associated to them, add one to disappeared counter
+            -For tracks that exceed disappeared threshold, remove from active tracking storage
+            
+
+"""
+
+import pandas as pd # Replace with cudf if performance is too slow?
+from collections import OrderedDict
+import numpy as np
+from scipy.spatial import distance as dist
+import matplotlib.pyplot as plt
+import time
+from statistics import mean
+from datetime import datetime
+
+br = "\n" # Line break for use in code
+
+
+# ===================== SETTINGS ==============================================
+
+verbose = True # Set to True if you want tracking process printed to screen and False if not
+
+class tracker():
+    def __init__(self, max_disappeared, max_distance, running_mean_threshold, results_filename, frames, detections, verbose):
[…]

### 2023-01-13 src/R/Analysis.Rmd
Resubmission -first commit

--- 2022-05-25 code/R/Analysis.Rmd
@@
 ```{r setup, include=FALSE}
+
 library(ggplot2)
@@
 library(stringr)
+library(plotly)
 library(GGally)
-library(plotly)
+
 knitr::opts_chunk$set(echo = TRUE)
@@
 
-axisSize = 30
-legendTitleSize = 30
-legendTextSize = 28
+axisSize = 34
+legendTitleSize = 50
+legendTextSize = 40
 
@@
 
-ggsave("U:/BITCue/Projekter/TrackingFlowers/figures/OnsetVSLength.png", f1, width = 15, height = 10)
+#ggsave("U:/BITCue/Projekter/TrackingFlowers/figures/OnsetVSLength.png", f1, width = 15, height = 10)
 
-all %>% 
-ggplot()+
-  geom_point(aes(x = flower_start, y = flower_length, colour = CameraID))# +
-#geom_smooth(aes(x = flower_start, y = flower_length), method=lm)
+#all = unique(all)
 
+f2 = ggplot(unique(all), aes(x = camCode, y = flower_length, fill=camCode))+
+  geom_boxplot() + 
+  geom_point(position=position_jitterdodge()) +
+  xlab("Onset of flowering (DOY)") + 
+  ylab("Number of days flowering") +
+  theme_bw()  +
+  scale_fill_manual(name = "Camera ID", values = c("white", "white", "white", "white", "white"), aesthetics = c("colour", "fill")) +
+
+  #scale_shape_manual(name = "Camera ID",
+  #scale_fill_manual(name = "Camera ID", values = c("gray26", "brown3", "dodgerblue3", "darkorange2", "chartreuse3"), aesthetics = c("colour", "fill")) +
+
+  #scale_shape_manual(name = "Camera ID",
+   #                   values = c(21,22,23,24,25)) + 
+  theme(legend.position = "none",
+        panel.grid.major = element_blank(), 
+        panel.grid.minor = element_blank(),
+        panel.background = element_blank(),
+        legend.title = element_text(size = legendTitleSize),
+        text = element_text(size = axisSize),
+        legend.text = element_text(size = legendTextSize),
+        axis.line = element_line(),
+        axis.text=element_text(colour="black"),
+        panel.spacing = unit(0.7, "lines")) + 
+  guides(fill = "none", color="none")
+
+f2
+ggsave("U:/BITCue/Projekter/TrackingFlowers/figures/OnsetVSLengthBox_withPoints.pdf", f2, width = 15, height = 10)
+
+# all %>% 
+# ggplot()+
+#   geom_point(aes(x = flower_start, y = flower_length, colour = CameraID))# +
+# #geom_smooth(aes(x = flower_start, y = flower_length), method=lm)
+#  
 t.test(nyaa04_meta$flower_length, nars13_meta$flower_length)
@@
 
-ggsave('U:/BITCue/Projekter/TrackingFlowers/testResults/MMFix/figures/NYAA-04_AllCombinations.png', NYAA04_p1)
-ggsave('U:/BITCue/Projekter/TrackingFlowers/testResults/MMFix/figures/NYAA-04_TenLowest.png', NYAA04_p2)
+#ggsave('U:/BITCue/Projekter/TrackingFlowers/testResults/MMFix/figures/NYAA-04_AllCombinations.png', NYAA04_p1)
+#ggsave('U:/BITCue/Projekter/TrackingFlowers/testResults/MMFix/figures/NYAA-04_TenLowest.png', NYAA04_p2)
 
@@
 
-ggsave('U:/BITCue/Projekter/TrackingFlowers/testResults/MMFix/figures/NARS-13_AllCombinations.png', NARS13_p1)
-ggsave('U:/BITCue/Projekter/TrackingFlowers/testResults/MMFix/figures/NARS-13_TenLowest.png', NARS13_p2)
+#ggsave('U:/BITCue/Projekter/TrackingFlowers/testResults/MMFix/figures/NARS-13_AllCombinations.png', NARS13_p1)
+#ggsave('U:/BITCue/Projekter/TrackingFlowers/testResults/MMFix/figures/NARS-13_TenLowest.png', NARS13_p2)
 
@@
 
-ggsave('U:/BITCue/Projekter/TrackingFlowers/testResults/MMFix/figures/NARS-04_AllCombinations.png', NARS04_p1)
-ggsave('U:/BITCue/Projekter/TrackingFlowers/testResults/MMFix/figures/NARS-04_TenLowest.png', NARS04_p2)
+#ggsave('U:/BITCue/Projekter/TrackingFlowers/testResults/MMFix/figures/NARS-04_AllCombinations.png', NARS04_p1)
+#ggsave('U:/BITCue/Projekter/TrackingFlowers/testResults/MMFix/figures/NARS-04_TenLowest.png', NARS04_p2)
 
@@
 
-ggsave('U:/BITCue/Projekter/TrackingFlowers/testResults/MMFix/figures/NARS-17_AllCombinations.png', NARS17_p1)
-ggsave('U:/BITCue/Projekter/TrackingFlowers/testResults/MMFix/figures/NARS-17_TenLowest.png', NARS17_p2)
+#ggsave('U:/BITCue/Projekter/TrackingFlowers/testResults/MMFix/figures/NARS-17_AllCombinations.png', NARS17_p1)
+#ggsave('U:/BITCue/Projekter/TrackingFlowers/testResults/MMFix/figures/NARS-17_TenLowest.png', NARS17_p2)
 
@@
 
-ggsave('U:/BITCue/Projekter/TrackingFlowers/testResults/MMFix/figures/THUL-01_AllCombinations.png', THUL01_p1)
-ggsave('U:/BITCue/Projekter/TrackingFlowers/testResults/MMFix/figures/THUL-01_TenLowest.png', THUL01_p2)
+#ggsave('U:/BITCue/Projekter/TrackingFlowers/testResults/MMFix/figures/THUL-01_AllCombinations.png', THUL01_p1)
+#ggsave('U:/BITCue/Projekter/TrackingFlowers/testResults/MMFix/figures/THUL-01_TenLowest.png', THUL01_p2)
 
@@
 
-ggsave('U:/BITCue/Projekter/TrackingFlowers/testResults/figures/NYAA-04_AllCombinations.png', NYAA04_p1)
-ggsave('U:/BITCue/Projekter/TrackingFlowers/testResults/figures/NYAA-04_TenLowest.png', NYAA04_p2)
+#ggsave('U:/BITCue/Projekter/TrackingFlowers/testResults/figures/NYAA-04_AllCombinations.png', NYAA04_p1)
+#ggsave('U:/BITCue/Projekter/TrackingFlowers/testResults/figures/NYAA-04_TenLowest.png', NYAA04_p2)
 
@@
 
-ggsave('U:/BITCue/Projekter/TrackingFlowers/testResults/figures/THUL-01_AllCombinations.png', THUL01_p1)
-ggsave('U:/BITCue/Projekter/TrackingFlowers/testResults/figures/THUL-01_AllZero.png', THUL01_p2)
+#ggsave('U:/BITCue/Projekter/TrackingFlowers/testResults/figures/THUL-01_AllCombinations.png', THUL01_p1)
+#ggsave('U:/BITCue/Projekter/TrackingFlowers/testResults/figures/THUL-01_AllZero.png', THUL01_p2)
 ```
@@
 
-ggsave('U:/BITCue/Projekter/TrackingFlowers/testResults/figures/NARS-13_AllCombinations.png', NARS04_p1)
-ggsave('U:/BITCue/Projekter/TrackingFlowers/testResults/figures/NARS-13_TenLowest.png', NARS04_p2)
+#ggsave('U:/BITCue/Projekter/TrackingFlowers/testResults/figures/NARS-13_AllCombinations.png', NARS04_p1)
+#ggsave('U:/BITCue/Projekter/TrackingFlowers/testResults/figures/NARS-13_TenLowest.png', NARS04_p2)
 
@@
 
-ggsave('U:/BITCue/Projekter/TrackingFlowers/testResults/figures/NARS-17_AllCombinations.png', NARS17_p1)
-ggsave('U:/BITCue/Projekter/TrackingFlowers/testResults/figures/NARS-17_TenLowest.png', NARS17_p2)
+#ggsave('U:/BITCue/Projekter/TrackingFlowers/testResults/figures/NARS-17_AllCombinations.png', NARS17_p1)
+#ggsave('U:/BITCue/Projekter/TrackingFlowers/testResults/figures/NARS-17_TenLowest.png', NARS17_p2)
 

### 2023-01-13 src/R/MOTMetrics.Rmd
Resubmission -first commit

---
title: "MOTMetrics"
author: "Hjalte Mann"
date: '2022-05-02'
output: html_document
---

```{r setup, include=FALSE}
library(ggplot2)
library(dplyr)
library(stringr)
library(GGally)
library(plotly)
knitr::opts_chunk$set(echo = TRUE)
```

####### MOTMETRICS MMFix PARAMETER TEST ########

# NYAA-04
```{r fig.height=4}
NYAA04d1 = read.csv("U:/BITCue/Projekter/TrackingFlowers/testResults/MMFix/NYAA-04_parameterTest_Evaluation_20220513.csv")

NYAA04_p1 = NYAA04d1 %>% 
  ggplot() +
  geom_point(aes(x=runMean, y = maxDisap, size = num_switches, colour = num_switches)) + 
  ggtitle("NYAA-04 Parameter combination") +
  facet_wrap(vars(maxDist))+scale_color_gradient(low="blue", high="red")

NYAA04_p2 = NYAA04d1 %>% 
  arrange(num_switches, ascending = T) %>% 
  slice_head(n = 10) %>% 
  ggplot() +
  geom_point(aes(x=runMean, y = maxDisap, size = num_switches, colour = num_switches)) + 
  ggtitle("NYAA-04 Ten Lowest Mismatches") +
  facet_wrap(vars(maxDist))+scale_color_gradient(low="blue", high="red")

NYAA04_p1
NYAA04_p2

ggsave('U:/BITCue/Projekter/TrackingFlowers/testResults/MMFix/figures/NYAA-04_AllCombinations.png', NYAA04_p1)
ggsave('U:/BITCue/Projekter/TrackingFlowers/testResults/MMFix/figures/NYAA-04_TenLowest.png', NYAA04_p2)

rm(NYAA04_p1)
rm(NYAA04_p2)
```

# NARS-13

```{r fig.height=4}
NARS13d1 = read.csv("U:/BITCue/Projekter/TrackingFlowers/testResults/MMFix/NARS-13_parameterTest_Evaluation_20220513.csv")

NARS13_p1 = NARS13d1 %>%
  ggplot() +
  geom_point(aes(x=runMean, y = maxDisap, size = num_switches, colour = num_switches)) +
  ggtitle("NYAA-04 Parameter combination") +
  facet_wrap(vars(maxDist))+scale_color_gradient(low="blue", high="red")

NARS13_p2 = NARS13d1 %>%
  arrange(num_switches, ascending = T) %>%
  slice_head(n = 10) %>%
  ggplot() +
  geom_point(aes(x=runMean, y = maxDisap, size = num_switches, colour = num_switches)) +
  ggtitle("NYAA-04 Ten Lowest Mismatches") +
  facet_wrap(vars(maxDist))+scale_color_gradient(low="blue", high="red")

NARS13_p1
NARS13_p2

ggsave('U:/BITCue/Projekter/TrackingFlowers/testResults/MMFix/figures/NARS-13_AllCombinations.png', NARS13_p1)
ggsave('U:/BITCue/Projekter/TrackingFlowers/testResults/MMFix/figures/NARS-13_TenLowest.png', NARS13_p2)

rm(NARS13_p1)
rm(NARS13_p2)
```

# NARS-04
```{r fig.height=4}
NARS04d1 = read.csv("U:/BITCue/Projekter/TrackingFlowers/testResults/MMFix/NARS-04_parameterTest_Evaluation_20220513.csv")

NARS04_p1 = NARS04d1 %>% 
  ggplot() +
  geom_point(aes(x=runMean, y = maxDisap, size = num_switches, colour = num_switches)) + 
  ggtitle("NYAA-04 Parameter combination") +
  facet_wrap(vars(maxDist))+scale_color_gradient(low="blue", high="red")

NARS04_p2 = NARS04d1 %>% 
  arrange(num_switches, ascending = T) %>% 
  slice_head(n = 10) %>% 
  ggplot() +
  geom_point(aes(x=runMean, y = maxDisap, size = num_switches, colour = num_switches)) + 
  ggtitle("NYAA-04 Ten Lowest Mismatches") +
  facet_wrap(vars(maxDist))+scale_color_gradient(low="blue", high="red")

NARS04_p1
NARS04_p2

ggsave('U:/BITCue/Projekter/TrackingFlowers/testResults/MMFix/figures/NARS-04_AllCombinations.png', NARS04_p1)
ggsave('U:/BITCue/Projekter/TrackingFlowers/testResults/MMFix/figures/NARS-04_TenLowest.png', NARS04_p2)

rm(NARS04_p1)
rm(NARS04_p2)
```

# NARS-17
```{r fig.height=4}
NARS17d1 = read.csv("U:/BITCue/Projekter/TrackingFlowers/testResults/MMFix/NARS-17_parameterTest_Evaluation_20220513.csv")

NARS17_p1 = NARS17d1 %>% 
  ggplot() +
  geom_point(aes(x=runMean, y = maxDisap, size = num_switches, colour = num_switches)) + 
  ggtitle("NYAA-04 Parameter combination") +
  facet_wrap(vars(maxDist))+scale_color_gradient(low="blue", high="red")

NARS17_p2 = NARS17d1 %>% 
  arrange(num_switches, ascending = T) %>% 
  slice_head(n = 10) %>% 
  ggplot() +
  geom_point(aes(x=runMean, y = maxDisap, size = num_switches, colour = num_switches)) + 
  ggtitle("NYAA-04 Ten Lowest Mismatches") +
  facet_wrap(vars(maxDist))+scale_color_gradient(low="blue", high="red")

NARS17_p1
NARS17_p2

ggsave('U:/BITCue/Projekter/TrackingFlowers/testResults/MMFix/figures/NARS-17_AllCombinations.png', NARS17_p1)
ggsave('U:/BITCue/Projekter/TrackingFlowers/testResults/MMFix/figures/NARS-17_TenLowest.png', NARS17_p2)

rm(NARS17_p1)
rm(NARS17_p2)
```

# THUL-01
```{r fig.height=4}
THUL01d1 = read.csv("U:/BITCue/Projekter/TrackingFlowers/testResults/MMFix/THUL-01_parameterTest_Evaluation_20220513.csv")

THUL01_p1 = THUL01d1 %>% 
  ggplot() +
  geom_point(aes(x=runMean, y = maxDisap, size = num_switches, colour = num_switches)) + 
  ggtitle("NYAA-04 Parameter combination") +
  facet_wrap(vars(maxDist))+scale_color_gradient(low="blue", high="red")

THUL01_p2 = THUL01d1 %>% 
  arrange(num_switches, ascending = T) %>% 
  slice_head(n = 10) %>% 
  ggplot() +
  geom_point(aes(x=runMean, y = maxDisap, size = num_switches, colour = num_switches)) + 
  ggtitle("NYAA-04 Ten Lowest Mismatches") +
  facet_wrap(vars(maxDist))+scale_color_gradient(low="blue", high="red")

THUL01_p1
THUL01_p2

ggsave('U:/BITCue/Projekter/TrackingFlowers/testResults/MMFix/figures/THUL-01_AllCombinations.png', THUL01_p1)
ggsave('U:/BITCue/Projekter/TrackingFlowers/testResults/MMFix/figures/THUL-01_TenLowest.png', THUL01_p2)

rm(THUL01_p1)
rm(THUL01_p2)
```

### Explore combinations

# NYAA-04
```{r echo = FALSE}
print("NYAA-04")
nyaa04_gt = read.csv("U:/BITCue/Projekter/TrackingFlowers/data/annotations/2021_12_13_NorwayAnnotations_NYAA-04_IndividualAnnotations_FRCNN_Metrics.csv")

print('True number of flowers')
length(unique(nyaa04_gt$id_gt))

# Minimum num_switches
print("Lowest number of mismatches")
min(NYAA04d1$num_switches)
NYAA04Minnum_switches = NYAA04d1[NYAA04d1$num_switches == min(NYAA04d1$num_switches),]

# All zero
NYAA04AllZero = NYAA04d1 %>% 
  filter(maxDisap == 0 & maxDist == 0 & runMean == 1)

# Correct number of flowers and minimum num_switches
NYAA04CorrectflowerMinnum_switches = NYAA04d1 %>% 
  filter(num_tracks == length(unique(nyaa04_gt$id_gt)))

NYAA04CorrectflowerMinnum_switches = NYAA04CorrectflowerMinnum_switches[NYAA04CorrectflowerMinnum_switches$num_switches == min(NYAA04CorrectflowerMinnum_switches$num_switches),]

```

```{r echo = FALSE}
print("NARS-13")
nars13_gt = read.csv("U:/BITCue/Projekter/TrackingFlowers/data/annotations/2020_04_30_NorwayAnnotations_NARS-13_IndividualAnnotations_FRCNN_Metrics.csv")

print('True number of flowers')
length(unique(nars13_gt$id_gt))

# Minimum num_switches
print("Lowest number of mismatches")
min(NARS13d1$num_switches)
NARS13Minnum_switches = NARS13d1[NARS13d1$num_switches == min(NARS13d1$num_switches),]

# All zero
NARS13AllZero = NARS13d1 %>%
  filter(maxDisap == 0 & maxDist == 0 & runMean == 1)

# Correct number of flowers and minimum num_switches
NARS13CorrectflowerMinnum_switches = NARS13d1 %>%
  filter(num_tracks == length(unique(nars13_gt$id_gt)))
NARS13CorrectflowerMinnum_switches = NARS13CorrectflowerMinnum_switches[NARS13CorrectflowerMinnum_switches$num_switches == min(NARS13CorrectflowerMinnum_switches$num_switches),]
```

# NARS-04
```{r echo = FALSE}
print("NARS-04")
nars04_gt = read.csv("U:/BITCue/Projekter/TrackingFlowers/data/annotations/2020_05_17_NorwayAnnotations_NARS-04_IndividualAnnotations_FRCNN_Metrics.csv")

print('True number of flowers')
length(unique(nars04_gt$id_gt))

# Minimum num_switches
print("Lowest number of mismatches")
min(NARS04d1$num_switches)
NARS04Minnum_switches = NARS04d1[NARS04d1$num_switches == min(NARS04d1$num_switches),]

# All zero
NARS04AllZero = NARS04d1 %>% 
  filter(maxDisap == 0 & maxDist == 0 & runMean == 1)

# Correct number of flowers and minimum num_switches
NARS04CorrectflowerMinnum_switches = NARS04d1 %>% 
  filter(num_tracks == length(unique(nars04_gt$id_gt)))
NARS04CorrectflowerMinnum_switches = NARS04CorrectflowerMinnum_switches[NARS04CorrectflowerMinnum_switches$num_switches == min(NARS04CorrectflowerMinnum_switches$num_switches),]
```

# NARS-17
```{r echo = FALSE}
print("NARS-17")
nars17_gt = read.csv("U:/BITCue/Projekter/TrackingFlowers/data/annotations/2021_12_30_NorwayAnnotations_NARS-17_IndividualAnnotations_FRCNN_Metrics.csv")

print('True number of flowers')
length(unique(nars17_gt$id_gt))

# Minimum num_switches
print("Lowest number of mismatches")
min(NARS17d1$num_switches)
NARS17Minnum_switches = NARS17d1[NARS17d1$num_switches == min(NARS17d1$num_switches),]

# All zero
NARS17AllZero = NARS17d1 %>% 
  filter(maxDisap == 0 & maxDist == 0 & runMean == 1)

# Correct number of flowers and minimum num_switches
NARS17CorrectflowerMinnum_switches = NARS17d1 %>% 
  filter(num_tracks == length(unique(nars17_gt$id_gt)))
NARS17CorrectflowerMinnum_switches = NARS17CorrectflowerMinnum_switches[NARS17CorrectflowerMinnum_switches$num_switches == min(NARS17CorrectflowerMinnum_switches$num_switches),]
```

# THUL-01
```{r echo = FALSE}
print("THUL-01")
thul01_gt = read.csv("U:/BITCue/Projekter/TrackingFlowers/data/annotations/2020_05_15_NorwayAnnotations_THUL-01_IndividualAnnotations_FRCNN_Metrics.csv")

print('True number of flowers')
length(unique(thul01_gt$id_gt))

# Minimum num_switches
print("Lowest number of mismatches")
min(THUL01d1$num_switches)
THUL01Minnum_switches = THUL01d1[THUL01d1$num_switches == min(THUL01d1$num_switches),]

# All zero
THUL01AllZero = THUL01d1 %>% 
  filter(maxDisap == 0 & maxDist == 0 & runMean == 1)

# Correct number of flowers and minimum num_switches
THUL01CorrectflowerMinnum_switches = THUL01d1 %>% 
  filter(num_tracks == length(unique(thul01_gt$id_gt)))
THUL01CorrectflowerMinnum_switches = THUL01CorrectflowerMinnum_switches[THUL01CorrectflowerMinnum_switches$num_switches == min(THUL01CorrectflowerMinnum_switches$num_switches),]
```

# Table dataframe

```{r}
trackScores = setNames(data.frame(matrix(ncol = 11, nrow = 0)), c('Site', 'Type', 'Flowers','Points', 'Max gap', 'Max distance', 'Running mean', 'Mismatches', 'MOTA', '>80% Accuracy','Tracks'))

# Add THUL-01 to table
# All zero
THUL01AllZero$Site = "2018 THUL B"
THUL01AllZero$Type = "PZ"

THUL01AllZeroTable = THUL01AllZero %>% 
  select(Site,Type,num_unique_objects, num_detections, maxDisap, maxDist, runMean,num_switches, mota, mostly_tracked, num_tracks)

trackScores[nrow(trackScores) + 1,] = c(THUL01AllZeroTable[1,])

# Minimum mismatches
THUL01Minnum_switches

THUL01Minnum_switchesTable = c("2018 THUL B","MM", "6", "699", "0 - 160", "0, 200 - 1000", "1, 10", "0", "1.00", "6", "6")

trackScores[nrow(trackScores) + 1,] = c(THUL01Minnum_switchesTable)

# Correct flower, minimum mismatches
THUL01CorrectflowerMinnum_switchesTable = c("2018 THUL B","CF MM", "NA", "NA", "NA", "NA", "NA", "NA", "NA", "NA", "NA")
trackScores[nrow(trackScores) + 1,] = c(THUL01CorrectflowerMinnum_switchesTable)

# Add NARS-17 to table
# All zero
NARS17AllZero$Site = "2019 NARS D"
NARS17AllZero$Type = "PZ"

NARS17AllZeroTable = NARS17AllZero %>%
  select(Site,Type, num_unique_objects, num_detections, maxDisap, maxDist, runMean,num_switches, mota, mostly_tracked, num_tracks)

trackScores[nrow(trackScores) + 1,] = c(NARS17AllZeroTable[1,])

# Minimum mismatches
NARS17Minnum_switches

NARS17Minnum_switchesTable = c("2019 NARS D","MM", "13", "748", "0 - 160", "0, 300 - 1000", "1 - 160", "0", "1.00", "13", "7 - 13")

trackScores[nrow(trackScores) + 1,] = c(NARS17Minnum_switchesTable)

# Correct flower, minimum mismatches
NARS17CorrectflowerMinnum_switches = NARS17d1 %>% 
  filter(num_tracks == length(unique(nars17_gt$id_gt)))
NARS17CorrectflowerMinnum_switches = NARS17CorrectflowerMinnum_switches[NARS17CorrectflowerMinnum_switches$num_switches == min(NARS17CorrectflowerMinnum_switches$num_switches),]

NARS17CorrectflowerMinnum_switchesTable = c("2019 NARS D","CF MM", "13", "748", "0 - 60", "0, 300 - 1000", "1 - 160", "0", "1.00", "13", "13")

trackScores[nrow(trackScores) + 1,] = c(NARS17CorrectflowerMinnum_switchesTable)

# Add NARS-04 to table
# All zero
NARS04AllZero$Site = "2018 NARS A"
NARS04AllZero$Type = "PZ"

NARS04AllZeroTable = NARS04AllZero %>%
  select(Site,Type,num_unique_objects, num_detections, maxDisap, maxDist, runMean,num_switches, mota, mostly_tracked, num_tracks)

trackScores[nrow(trackScores) + 1,] = c(NARS04AllZeroTable[1,])

# Minimum mismatches
NARS04Minnum_switches$Site = "2018 NARS A"
NARS04Minnum_switches$Type = "MM"
NARS04Minnum_switchesTable = NARS04Minnum_switches %>%
  select(Site,Type,num_unique_objects, num_detections, maxDisap, maxDist, runMean,num_switches, mota, mostly_tracked, num_tracks)

trackScores[nrow(trackScores) + 1,] = c(NARS04Minnum_switchesTable[1,])

# Correct flower, minimum mismatches
NARS04CorrectflowerMinnum_switchesTable = c("2018 NARS A","CF MM", "32", "1724", "10", "500", "10", "19", "0.98898", "32", "32")
trackScores[nrow(trackScores) + 1,] = c(NARS04CorrectflowerMinnum_switchesTable)

# Add NYAA-04 to table
# All zero
NYAA04AllZero$Site = "2019 NYAA E"
NYAA04AllZero$Type = "PZ"

NYAA04AllZeroTable = NYAA04AllZero %>%
  select(Site,Type,num_unique_objects, num_detections, maxDisap, maxDist, runMean,num_switches, mota, mostly_tracked, num_tracks)

trackScores[nrow(trackScores) + 1,] = c(NYAA04AllZeroTable[1,])

# Minimum mismatches
NYAA04Minnum_switches

NYAA04Minnum_switchesTable = c("2019 NYAA E","MM", "21", "1762", "70 - 160", "600", "1", "17", "0.99035", "21", "21")

trackScores[nrow(trackScores) + 1,] = c(NYAA04Minnum_switchesTable)

# Correct flower, minimum mismatches
NYAA04CorrectflowerMinnum_switchesTable = c("2019 NYAA E","CF MM","NA", "NA", "NA", "NA", "NA", "NA", "NA", "NA", "NA")

trackScores[nrow(trackScores) + 1,] = c(NYAA04CorrectflowerMinnum_switchesTable)

# Add NARS-13 to table
# All zero
NARS13AllZero$Site = "2019 NARS C"
NARS13AllZero$Type = "PZ"

NARS13AllZeroTable = NARS13AllZero %>%
  select(Site,Type,num_unique_objects, num_detections, maxDisap, maxDist, runMean,num_switches, mota, mostly_tracked, num_tracks)

trackScores[nrow(trackScores) + 1,] = c(NARS13AllZeroTable[1,])

# Minimum mismatches
NARS13Minnum_switches$Site = "2019 NARS C"
NARS13Minnum_switches$Type = "MM"
NARS13Minnum_switchesTable = NARS13Minnum_switches %>%
  select(Site,Type,num_unique_objects, num_detections, maxDisap, maxDist, runMean,num_switches, mota, mostly_tracked, num_tracks)

trackScores[nrow(trackScores) + 1,] = c(NARS13Minnum_switchesTable[1,])

# Correct flower, minimum mismatches
NARS13CorrectflowerMinnum_switchesTable = c("2019 NARS C","CF MM", "85", "4887", "10", "400", "10", "91", "0.98138", "85", "85")
trackScores[nrow(trackScores) + 1,] = c(NARS13CorrectflowerMinnum_switchesTable)

```

```{r}
write.table(trackScores, "U:/BITCue/Projekter/TrackingFlowers/testResults/MMFix/trackScores.csv", row.names = F, col.names = F, quote = F, sep=";")
```

```{r}

percDiff <- function(val1, val2) {
  pdiff = abs(val1-val2) / (abs((val1+val2)/2))

  return(pdiff*100)
}

trackScoresFig = trackScores %>% 
  select(Site, Tracks, Flowers, MOTA)

trackScoresFig$flowerError = percDiff(trackScores$Tracks, trackScores$Flowers)

```

# Camera codes:
# 2018 NARS-04: 2018 NARS A
# 2018 THUL-01: 2018 THUL B
# 2019 NARS-13: 2019 NARS C
# 2019 NARS-17: 2019 NARS D
# 2019 NYAA-04: 2019 NYAA E

####### /MMFix PARAMETER TEST ########

### 2023-01-13 text/ManuscriptDraft.Rmd
Resubmission -first commit

@@
 knit('ManuscriptDraft.Rmd')
-rmarkdown::render("ManuscriptDraft.Rmd", output_file = "ManuscriptDraft.docx")
-system2("open","ManuscriptDraft.docx")
+rmarkdown::render("ManuscriptDraft.Rmd", output_file = "ManuscriptDraft_Resubmission.docx")
+system2("open","ManuscriptDraft_Resubmission.docx")
 ```
@@
 
+ 1.
 A fundamental challenge in pollination ecology is to understand how successful pollination is affected by the timing of flowering, the timing and frequency of pollinator visits, and the abundance and diversity of visitors.
@@
 Due to the overwhelming logistical challenge of collecting such data by direct observation, it has rarely been done.
+
+ 2.
 Here, we show that image-based monitoring of field plots at very high temporal resolution can return information on flowering phenology at the level of individual flowers.
 Further, we present a framework for automatically tracking flowers in time-lapse image series, evaluating tracking accuracy, and identifying unreliable tracks.
+
+ 3.
 Our results on individual flower phenology suggest that traditional weekly observations of flower plots are insufficient to capture details in flowering dynamics.
-We show that the results of our automated method are reliable by comparing with manual tracking of individual flower phenology.
-Our method could facilitate fine-grained studies on flower phenology and could be an important tool for estimating the sensitivity of plant-pollinator interactions to climate change.
+We show that the results of our automated method are reliable by comparing with manual tracking of individual floral phenology.
+
+ 4. Our method could facilitate fine-grained studies on flower phenology and could be an important tool for estimating the sensitivity of plant-pollinator interactions to climate change.
 
@@
 Both flower and insect phenology is partly driven by abiotic cues such as snowmelt and temperature and can thus be sensitive to climate change.
-Changes in flower phenology as a consequence of climate change may indirectly affect the abundance of flower-dependent insect species [@hoye2013] and these effects can be stronger than the direct effects of climate change [@ogilvie2017].
-However, the relation between flower phenology and reproductive success is not well understood.
+Changes in floral phenology as a consequence of climate change may indirectly affect the abundance of flower-dependent insect species [@hoye2013] and these effects can be stronger than the direct effects of climate change [@ogilvie2017].
+However, the relation between floral phenology and reproductive success is not well understood.
 For example, @bolmgren2015 found that for an insect-pollinated plant species, early flowering was associated with increased fruit-set but late flowering was associated with higher seed mass.
@@
 
-Monitoring flower phenology at high temporal resolution is time-consuming, particularly in logistically challenging environments such as the Arctic.
+Monitoring floral phenology at high temporal resolution is time-consuming, particularly in logistically challenging environments such as the Arctic.
 Consequently, flowering phenology of a population is often characterized simply by the date for onset of flowering, typically derived from weekly observations in permanent plots [@Prevey2021].
@@
 However, even detailed information on phenology of a population is still an aggregation of the phenology of the individual members of the population for which important variation may be masked [@inouye2019].
-With the full distribution of flowering phenology at the level of individuals, it is possible to extract relevant characteristics of the distribution and test their individual sensitivities towards environmental factors, but acquiring such data requires efficient and preferably automated methods.
+With the full distribution of flowering phenology at the level of individual flowers, it is possible to extract relevant characteristics of the distribution and test their individual sensitivities towards environmental factors, but acquiring such data requires efficient and preferably automated methods.
 
@@
 Indeed, automatic image-based monitoring of flowering phenology coupled with deep learning has facilitated collection of phenology data for specific species at very high temporal resolution [@mann2022].
-Yet, phenological responses at the individual level may be indiscernible regardless of the temporal resolution of the data at population level.
+While automatic detection and counting of flowers in image-series, as demonstrated in @mann2022, can produce detailed information about floral biology, phenological responses at the level of individual flowers may be indiscernible regardless of the temporal resolution of the data at population level.
 For example, a shortening of individual flower longevity may not be directly obvious at the population level.
-In fact, many research questions can only be directly explored on the basis of phenology data at the level of individuals.
-For example, investigating the association between reproductive success and timing of flowering and/or flower longevity requires such data.
+In fact, many research questions can only be directly explored on the basis of phenology data at the level of individual flowers and such information is critical for investigating the association between reproductive success and timing of flowering and/or flower longevity.
 Individual level data would also aid the understanding of whether flower visitation rates and/or seed set depends on the timing of flowering for the individual flower.
-Thus, automated tracking of flowers in image-series of permanent plots to derive individual flower phenology would be ideal.
+Thus, automated recording of floral phenology by tracking the presence of individual flowers in image-series of permanent plots would be ideal and the feasibility of this as a second step subsequent of flower detection was explored in this study.
+
+![Figure 1: Tracking individual flowers through time.
+A: A time-lapse sequence of five frames showing presence of five individual flowers.
+Flower a (orange) is present in all frames, while the other flowers are only present part of the sequence and/or have gaps in their presence (e.g.
+because of occlusion or because they have temporarily disappeared from frame of view).
+Correct tracking of individual flowers results in information about the distribution of each individual flower in space (B) and time (C).](../figures/flowertrackingOverview.png){width="30%"}
 
@@
 However, Kalman filtering assumes linear motion of the object and does not handle abrupt motion well.
-Flowers are constrained in movement by their stalk and violates this assumption.
+Flowers are constrained in movement by their stalk and violate this assumption.
 Other tracking approaches, such as DeepSORT [@deepsort2017], use tracking-by-detection in which convolutional neural networks (CNNs) are applied to continuously distinguish individuals through time, thus aiding the tracking algorithm.
-Flowers, however, are unlikely to be clearly distinguishable and one flower may look less alike between two frames than two flowers in a single frame, which prevents this approach.
+Individual flowers, however, are unlikely to be clearly distinguishable and one flower may look less alike between two frames than two flowers in a single frame, which prevents this approach.
 Finally, some tracking methods require manual initiation of tracks for a fixed number of objects in the first frame and the algorithm then tracks these objects in the subsequent frames [@hu2012].
@@
 Here, we present and evaluate a simple multiple object tracking algorithm for automatic tracking of individual flowers in image-series.
-Our flower tracking algorithm assigns individual flowers to tracks based on the minimum distance between centroids of bounding boxes and does not rely on prediction of future positions.
+Our tracking algorithm is based on centroid tracking and assigns individual objects to tracks based on the minimum distance between centroids of bounding boxes and does not rely on prediction of future positions.
 Further, the method allows both initiation and termination of tracks during tracking.
-We describe three user-defined parameters to help optimise tracking of flowers in particular.
-We use time-lapse image-series from full-season monitoring of two widespread Arctic flowering plant species *Dryas integrifolia* and *D.
+We describe three user-defined parameters incorporated into the algorithm, demonstrate their effect on simulated data, and discuss why they may help optimise tracking of flowers in particular.
+We apply the algorithm on annotations of real flowers in time-lapse image-series from full-season monitoring of two widespread Arctic flowering plant species *Dryas integrifolia* and *D.
 octopetala*.
-For complex scenes with many flowers in close vicinity to each other there is a substantial risk of tracking errors regardless of the tracking method being used.
-In such cases, it may still be possible to reliably track a subset of the flowers, which are spatially isolated in the frames.
-For this, we suggest a conservative filtering approach based on the distance between tracks within the frame.
-Overall, we show that information on phenology at the level of individual flowers can be automatically derived from image-based monitoring of flower plots.
+For complex scenes with many flowers in close vicinity to each other there is an increased risk of tracking errors regardless of the tracking method being used.
+For such cases, we suggest a conservative filtering approach based on the distance between flower tracks within the frame in order to identify tracks that are spatially isolated and likely to be tracked with certainty.
+Overall, we show that information on floral biology can be automatically derived from image-based monitoring of flower plots.
 
@@
 
-## Study site and species
+## Simulated data
 
-This study is based on five full-season time-lapse image series collected at the following three Arctic sites (site abbreviation and sampling years in parentheses): Narsarsuaq (NARS), South Greenland (2018, 2019), Thule (THUL), North-West Greenland (2018), Ny-Ålesund (NYAA), Svalbard (2019).
-The images were captured using time-lapse cameras (Moultrie Wingscapes TimelapseCam Pro, Moultrie Products, Birmingham Alabama, USA) permanently positioned approx.
-60 cm above areas with either *Dryas integrifolia* in Narsarsuaq and Thule or *D.
-octopetala* in Ny-Ålesund.
-The white flowers of these two species are insect-pollinated and held erect above cushions of foliage.
-The species are native to and widespread in Asian, North American, and European Arctic and alpine regions.
-They are geographically separated except for a possible hybrid zone in North-East Greenland [@philipp2003].
+To demonstrate our tracking algorithm and the effect of the three parameters, we used simulated data mimicking flower detections.To produce simulated detections of a flower, we defined a circle with radius rF and center point cF to delimit the boundaries of the flower.
+The first data point for the flower was then generated as a random point within this circle.
+Then, remaining data points (i.e.
+data points for the subsequent frames) were generated iteratively, each point constrained to be within the circle with radius rM and the center point of the previously generated point and within the flower circle.
+This approach allowed us to simulate specific scenarios on which to demonstrate our algorithm by controlling the number of flowers generated and the number of data points produced per flower, the boundaries of a flower, and the maximum distance between current and next point (i.e.
+degree of movement between frames).
 
-## The image series and flower annotations
+We generated four simulated data sets (Fig.
+2) to demonstrate each of the three tracking parameters as well as the influence of frame rate on tracking reliability.
+In simulation A (Fig.
+2A), we produced two flowers that are spatially distinct and temporally separated (i.e., one flower appears subsequent of the other).
+Each flower consists of ten data points and thus the data set spans over 20 frames.
+In simulation B, we produced four flowers that are spatially separated.
+Each flower consists of 5 data points, but the data set spans over 10 frames and each flower contains gaps of 1-3 data points at random temporal location.
+In simulation C, we produced four flowers that are in close proximity to each other and with a large degree of movement between frames.
+Each flower consist of 500 data points and the data sets spans over 500 frames.
+Finally, in simulation D, we produced four flowers consisting of a very high number of data points (10.000 per flower) but with only little movement between frames.
+Here, we used the same center point (cF) for all four flowers to increase overlap.
+Two pairs of flowers have a high degree of spatial overlap.
 
-Images were collected throughout the full growing seasons.
-While the original time-lapse frequency was higher, the basis for this study was 1-hour interval series.
-We limited these series to include only the flowering season and, to limit the workload for manual annotation, randomly sampled images from each series.
-The number of images for each series after sampling is given in Table 1.
-All flowers in bloom within each of these image were manually annotated using the rectangular bounding box tool in the VIA VGG annotation software [@dutta2019] and each individual flower was assigned a unique ID that was maintained throughout its lifetime.
-These annotations constitute our ground truth tracks.
-The series varied substantially in the number of flowers they contained (Table 1).
-Examples of annotated images from the 2019 NARS C series at three different times in the flowering season are given in Figure 1.
-To demonstrate how individual flower tracking can result in detailed representations of flower phenology across the season, we use the ground truth annotations to show the variation in duration of flowering for each series.
+For each of the simulations, we present the tracking performance of our algorithm without the use of the three parameters and subsequently show how the parameters can be used to mitigate various challenges in the data and increase tracking accuracy.
+The two objects in simulation A are adjacent in time but spatially separated.
+We therefore introduce a limit to the spatial distance that is accepted between points in adjacent frames before they are considered unique objects.
+We do that with the max distance parameter.
+The objects in simulation B overlap in time but are well separated in space.
+However, each track contains gaps which can lead to track fragmentation.
+To allow gaps in track, we set a number of frames that a track can disappear before initiating a new track for points appearing in the same location.
+In simulation c, each object is allowed to move a lot between frames.
+Each object has a large spread of points and the four objects are in close proximity to each other.
+This means that between two frames, there is a substantial risk that a point for one object is closer to a point from another object rather than the point for that object, which will lead to identity shifts.
+To mitigate this, we calculate distances based on a current point and a set number of previous points in a track using the running mean parameter.
+This reduces the weighting on the displacement of points between the current and the previous frame.
+Finally, in simulation D, we have four objects that overlap in pairs of two, but the movement allowed between frames is very small, which means that a given point from an object is likely to be closest to the point from the same object in the previous frame.
+Effectively, while the overlap is substantial, the frame rate is sufficiently high that accurate tracking is ensured.
+To demonstrate the importance of frame rate for reliable tracking, we simulate a reduction in frame rate (i.e., increased degree of movement between frames) by basing tracking on every 10th, 20th, and 30th frame, respectively, and present these results.
 
-**Table 1:** Number of images and number of individual flowers for each of the five time-lapse image series.
+![Figure 2: Simulated flower detection data for demonstration of our tracking algorithm and the effect of the three tracking parameters as well as the significance of frame rate for reliable tracking.
+A: 10 points per flower.
+rF = 200, rM = 100.
+B: 10 points per flower.
+rF = 200, rM = 100 C: 500 points per flower.
+rF = 200, rM = 200.
+D: 10,000 points per flower.
+rF = 100, rM = 1.](../figures/simulatedData.png){width="30%"}
 
-#
+## Common tracking errors and evaluating tracking performance
 
-![](../figures/CameraOverview.png){width="40%"}
-
-#
-
-![](../figures/AnnotationExamples_v2_4.png){width="20%"}
-
-**Figure 1:** Examples of annotated images for 2019 NARS C.
-A: Onset of flowering at DOY 145.
-B: Peak flowering at day of year (DOY) 151 (52 annotated flowers).
-C: 6 days after peak flowering (DOY 157).
-Flower bounding boxes are colour-coded by flower ID.
-D: Centroids of all flower annotations in the series, coloured by flower ID, showing how flowers may overlap in space.
-
-## Automatic flower tracking
-
-We built a framework for tracking, filtering, and evaluating tracking of objects in time-lapse image series.
-The tracking algorithm (suppl.
-1) tracks objects based on distances between centroids of bounding boxes, that is, objects with minimum displacement between frames are associated.
-The algorithm has a set of user-adjusted parameters that can optimize tracking accuracy.
-The parameters are particularly relevant for tracking of objects that are constrained to a specific area such as flowers.
-However, the tracking algorithm could be used to track any objects.
-The speed of the tracking algorithm depends on the computational power available as well as the number of objects that are being tracked, but the time for computation will be negligible in most use-cases.
-Tracking of a series containing 85 objects ran at \~0.02 seconds per frame on a standard laptop.
-This, together with the fact that the algorithm handles frames sequentially, means that it can be applied both offline (on a set of detections/annotations that have already been produced) or online (real-time tracking frame per frame).
-
-Several issues complicate the task of tracking individual flowers through time-lapse image series.
-First, as flowers move, they may temporarily occlude the view of other flowers and flowers close to the edge of the frame may move in and out of view between images.
-This can cause a track to be lost and a new track to be erroneously initiated when the flower reappears (Fig.
-2A), a tracking error known as fragmentation.
-However, if a flower reappears in the same area as a flower is already being tracked after disappearing in only a few frames, it is a reasonable assumption that it is the same individual and not that the old flower wilted/disappeared and a new one developed.
-The parameter **max gap** sets the number of frames a flower can be absent before reappearing without the track ending and a new initiating.
-If a given flower has not been detected in a number of frames exceeding the max gap size, the track will be terminated.
-Concurrently, this can deal with potential false negatives, as a track that is missing a bounding box will not be lost.
-
-Second, a standard centroid tracking algorithm will associate a point with a track based on minimum distance only, disregarding the absolute distance between centroids, which may lead to tracking errors (Fig.
-2B).
-However, as flowers are constrained in their displacement, we can assume a maximum distance between points and force the initiation of tracks for points that exceed this threshold.
-We set this threshold with the parameter **max distance,** measured as the euclidean distance in number of pixels between objects in neighbouring frames.
-
-Third, reliable tracking based on distance requires that movement of individuals are captured at sufficient frequency, such that each tracks are not switched between frames.
-Stand-alone image-based monitoring of the relatively slow process of flower development through full growing seasons (several months) is done with low frame rate, limited by power consumption and storage capacity both on and off the camera.
-Here, we used several images per day and while this temporal resolution captures the phenology of the flowers at very high detail, they fail to capture the full movement of flowers.
-For example, as the wind shifts, the flower heads change direction, which in the image series happens instantaneously (i.e., between two consecutive frames) (Fig.
-2C).
-The fact that the frame rate is insufficient to capture smooth movement means that bounding boxes around a single individual in two consecutive frames cannot be associated based on overlap (e.g.
-intersection over union) as there often is no overlap.
-With a high number of flowers in close vicinity to each other, there's a substantial risk of identity switching between flowers as they move among each other (Fig.
-1B).
-Additionally, as the flowers are constrained by their stalk, there is a limit to the distance they can move, which prevents prediction of future location of an individual based on its previous trajectory.
-Here, we introduce the tracking parameter **running mean**, with which we can base tracking on the distance between a point in the current frame and the running mean of the positions of a number of previous points in a track, taking advantage of the fact that a flower moves around a central point.
-Increasing the value for this parameter decreases the weighting of movement between individual frames.
-Setting the parameter to 1 means that tracking is based on the distance to the centroid in the previous frame only.
-
-#
-
-![Figure 2: Centroid tracking may produce erroneous associations when objects move or disappear between frames.
-A: A flower temporarily disappears from view.
-If gaps in tracks are not allowed during tracking, a new track will erroneously be initiated when the flower reappears.
-B: A flower disappears from view and another reappears between two frames.
-If no threshold for maximum distance between an object and a track is set, the reappearing flower will be linked to the track of the disappearing flower based on shortest distance despite the absolute distance being large.
-C: A flower is linked to the track of another flower.
-As flowers move around a central point, it is relevant to base tracking on a running mean of previous positions.
-Dots show position of flower in previous frame.
-Crosses show running mean of positions in x \> 1 number of previous frames.
-Dots and crosses coloured according to track identity.](../figures/ErrorFigure2.png){width="100%"}
-
-### Input data for the tracking algorithm
-
-The input data for the tracking algorithm are bounding box coordinates for each object in each frame given in a commonly used format [@everingham2010]: frame number, x-min, y-min, x-max, y-max.
-Here, x-min, y-min is the top left corner of the bounding box and x-max, y-max is the bottom right corner.
-The bounding box centroid is calculated on the basis of these coordinates within the algorithm.
-Calculating tracking accuracy requires a ground truth ID for each bounding box which can be given as the sixth column.
-As this data returned from both manual object annotations and object detection methods using automatic object detection methods based on deep learning, the method can be applied on both manually and automatically derived data.
-
-#
-
-### Identifying optimal user parameters
-
-To explore the effect of the three tracking parameters and to identify the optimal combination of parameters for our use-case of tracking flowers, we applied the tracking algorithm on each of the five image series with every combination of a range of values for each parameter (3179 combinations): 0-160 with a step size of 10 for max gap; 1-160 with a step size of 10 for running mean; 0-1000 with a step size of 100 for max distance.
-Note that max distance set to zero ignores the parameter altogether.
-We identified the setting(s) that returned the most accurate tracks for each series and compare the tracking results between optimal settings and all parameters ignored (imax gap = 0, running mean = 1, max distance = 0).
-
-## Evaluating tracking performance
+When tracking moving objects, two fundamental errors can occur.
+First, a single trajectory may be split into several tracks, an error known as fragmentation.
+This can happen when an object is temporarily occluded or have disappeared from view and a new track is erroneously initiated when it reappears, or when identity shifts occur between tracks, for example when they switch positions between frames.
+Fragmentation is especially prone in settings with high densities of objects.
+Second, merging occurs when several trajectories are assigned the same track identity.
+This can occur when a track is not correctly terminated at the end of a trajectory but rather continues on the trajectory of another object.
+Here, importantly, the two individual trajectories can be tracked without internal errors (i.e.
+without mismatches), but the error lies in the fact that the two trajectories have not been assigned unique tracking identities but instead are tracked as a single object.
 
@@
 The MOTA score ranges from -Inf to 1 with 1 being perfect detection and tracking.
-As tracking accuracy is commonly evaluated between manual annotations and detections, produced, for example, by a convolutional neural network (CNN), the MOTA score included false positive and false negatives as errors, but in our case, since we are tracking on ground truth data and our annotations and detections therefore are practically identical, these variables can be ignored.
+Tracking accuracy is commonly evaluated between manual annotations and detections produced, for example, by a convolutional neural network (CNN).
+Therefore, false positives and false negatives are included as errors in the MOTA score.
+In our case, we are not tracking on data produced by a CNN but rather on manually annotated tracks and do not have false positives or false negatives.
+Thus, these variables are ignored in the calculation of the MOTA score.
 
@@
 
-## Identifying the most reliably tracked flowers
+## Flower data
 
-In order to be able to identify the most reliably tracked flowers without the need for manual annotation, we use a conservative filtering method that extracts the most trustworthy tracks from a scene based on spatial distance between tracks.
+### Study site and species
+
+This study is based on five full-season time-lapse image series collected at the following three Arctic sites (site abbreviation and sampling years in parentheses): Narsarsuaq (NARS), South Greenland (2018, 2019), Thule (THUL), North-West Greenland (2018), Ny-Ålesund (NYAA), Svalbard (2019).
+The images were captured using time-lapse cameras (Moultrie Wingscapes TimelapseCam Pro, Moultrie Products, Birmingham Alabama, USA) permanently positioned approx.
+60 cm above areas with either *Dryas integrifolia* in Narsarsuaq and Thule or *D.
+octopetala* in Ny-Ålesund.
+The white flowers of these two species are insect-pollinated and held erect above cushions of foliage.
+The species are native to and widespread in Asian, North American, and European Arctic and alpine regions.
+They are geographically separated except for a possible hybrid zone in North-East Greenland [@philipp2003].
+
+### The image series and flower annotations
+
+Images were collected throughout the full growing seasons.
+While the original time-lapse frequency was higher, the basis for this study was 1-hour interval series.
+We limited these series to include only the flowering season and, to limit the workload for manual annotation, randomly sampled images from each series.
+The number of images for each series after sampling is given in Table 1.
+All flowers in bloom within each of these image were manually annotated using the rectangular bounding box tool in the VIA VGG annotation software [@dutta2019] and each individual flower was assigned a unique ID that was maintained throughout its lifetime.
+These annotations constitute our ground truth tracks.
+The series varied substantially in the number of flowers they contained (Table 1).
+Examples of annotated images from the 2019 NARS C series at three different times in the flowering season are given in figure 3.
+To demonstrate how individual flower tracking can result in detailed representations of floral phenology across the season, we use the ground truth annotations to show the variation in duration of flowering for each series.
+
+# 
+![Table 1: Number of images and number of individual flowers for each of the five time-lapse image series.](../figures/CameraOverview.png){width="40%"}
+
+#
+
+![Figure 3: Examples of annotated images for 2019 NARS C.
+A: Onset of flowering at DOY 145.
+B: Peak flowering at day of year (DOY) 151 (52 annotated flowers).
+C: 6 days after peak flowering (DOY 157).
+Flower bounding boxes are colour-coded by flower ID.
+D: Centroids of all flower annotations in the series, coloured by flower ID, showing how flowers may overlap in space.](../figures/AnnotationExamples_v2_4.png){width="20%"}
+
+### Automatic flower tracking
+
+We built a framework for tracking, filtering, and evaluating tracking of objects in time-lapse image series.
+The tracking algorithm (suppl.
+1) is based on centroid tracking [@rahman2020; @bjerge2021].
+That is, objects with minimum displacement between frames are associated.
+Centroids are derived from bounding boxes around objects stemming from manual annotation or automatic detection of objects.
+The algorithm has a set of user-adjusted parameters that can optimize tracking accuracy.
+The parameters are particularly relevant for tracking of objects that are constrained to a specific area such as flowers.
+However, the tracking algorithm could be used to track any objects.
+The speed of the tracking algorithm depends on the computational power available as well as the number of objects that are being tracked, but the time for computation will be negligible in most use-cases.
+Tracking of a series containing 85 objects ran at \~0.02 seconds per frame on a standard laptop.
+This, together with the fact that the algorithm handles frames sequentially, means that it can be applied both offline (on a set of detections/annotations that have already been produced) or online (real-time tracking frame per frame).
+
+Several issues complicate the task of tracking individual flowers through time-lapse image series.
+First, as flowers move, they may temporarily occlude the view of other flowers and flowers close to the edge of the frame may move in and out of view between images.
+This can cause a track to be lost and fragmentation to occur (Fig.
+4A).
+However, if a flower reappears in the same area as a flower is already being tracked after disappearing in only a few frames, it is a reasonable assumption that it is the same individual and not that the old flower wilted/disappeared and a new one developed.
+The parameter **max gap** sets the number of frames a flower can be absent before reappearing without the track ending and a new initiating.
+If a given flower has not been detected in a number of frames exceeding the max gap size, the track will be terminated.
+Concurrently, this can deal with potential false negatives, as a track that is missing a bounding box will not be lost.
+
+Second, a standard centroid tracking algorithm will associate a point with a track based on minimum distance only, disregarding the absolute distance between centroids, which may lead to tracking errors (Fig.
+4B).
+However, as flowers are constrained in their displacement, we can assume a maximum distance between points and force the initiation of tracks for points that exceed this threshold.
+We set this threshold with the parameter **max distance,** measured as the euclidean distance in number of pixels between objects in neighbouring frames.
+
+Third, reliable tracking based on distance requires that movement of individuals are captured at sufficient frequency, such that each tracks are not switched between frames.
+Stand-alone image-based monitoring of the relatively slow process of flower development through full growing seasons (several months) is done with low frame rate, limited by power consumption and storage capacity both on and off the camera.
+Here, we used several images per day and while this temporal resolution captures the phenology of the flowers at very high detail, they fail to capture the full movement of flowers.
+For example, as the wind shifts, the flower heads change direction, which in the image series happens instantaneously (i.e., between two consecutive frames) (Fig.
+4C).
+The fact that the frame rate is insufficient to capture smooth movement means that bounding boxes around a single individual in two consecutive frames cannot be associated based on overlap (e.g.
+intersection over union) as there often is no overlap.
+With a high number of flowers in close vicinity to each other, there's a substantial risk of identity switching between flowers as they move among each other (Fig.
+1C;4C).
+Additionally, as the flowers are constrained by their stalk, there is a limit to the distance they can move, which prevents prediction of future location of an individual based on its previous trajectory.
+Here, we introduce the tracking parameter **running mean**, with which we can base tracking on the distance between a point in the current frame and the running mean of the positions of a number of previous points in a track, taking advantage of the fact that a flower moves around a central point.
+Increasing the value for this parameter decreases the weighting of movement between individual frames.
+Setting the parameter to 1 means that tracking is based on the distance to the centroid in the previous frame only.
+
+#
+
+![Figure 4: Centroid tracking may produce erroneous associations when objects move or disappear between frames.
+A: A track may be lost when a flower temporarily disappears from view.
+If gaps in tracks are not allowed during tracking, a new track will erroneously be initiated when the flower reappears.
+B: A flower may be linked to an incorrect track despite large distance between them.
+Here, a flower disappears from view and another reappears between two frames.
+If no threshold for maximum distance between an object and a track is set, the reappearing flower will be linked to the track of the disappearing flower based on shortest distance despite the absolute distance being large.
+C: A flower may be linked to and incorrect track despite shortest distance to the running mean of the correct track.
+As flowers move around a central point, it is relevant to base tracking on a running mean of previous positions.
+Dots show position of flower in previous frame.
+Crosses show running mean of positions in x \> 1 number of previous frames.
+Dots and crosses coloured according to track identity.](../figures/errorFigure.png){width="100%"}
+
+#### Input data for the tracking algorithm
+
+The input data for the tracking algorithm are bounding box coordinates for each object in each frame given in a commonly used format [@everingham2010]: frame number, x-min, y-min, x-max, y-max.
+Here, x-min, y-min is the top left corner of the bounding box and x-max, y-max is the bottom right corner.
+The bounding box centroid is calculated on the basis of these coordinates within the algorithm.
+Calculating tracking accuracy requires a ground truth ID for each bounding box which can be given as the sixth column.
+As this data returned from both manual object annotations and object detection methods using automatic object detection methods based on deep learning, the method can be applied on both manually and automatically derived data.
+
+#### Identifying optimal user parameters
+
+To explore the effect of the three tracking parameters and to identify the optimal combination of parameters for our use-case of tracking flowers, we applied the tracking algorithm on each of the five image series with every combination of a range of values for each parameter (3179 combinations): 0-160 with a step size of 10 for max gap; 1-160 with a step size of 10 for running mean; 0-1000 with a step size of 100 for max distance.
+Note that max distance set to zero ignores the parameter altogether.
+We identified the setting(s) that returned the most accurate tracks for each series and compare the tracking results between optimal settings and all parameters ignored (max gap = 0, running mean = 1, max distance = 0).
+
+### Identifying the most reliably tracked flowers
+
+While the implemented tracking algorithm produced highly accurate flower tracks, it did not do so without errors.
+Many flowers could be tracked with certainty or near-certainty, but some settings proved challenging, especially series with many flowers or flowers in close proximity to each other.
+In order to be able to identify the most reliably tracked flowers in a series without the need for manual annotation, we use a conservative filtering method that extracts the most trustworthy tracks from a scene based on spatial distance between tracks.
 We derive the distance between tracks based on the centroids of points included in a track produced by the tracking algorithm.
 When tracks consist of only a single point, the coordinates of the point is used as the centroid.
-For tracks consisting of two points.
-we use the midpoint of the line between them.
+For tracks consisting of two points we use the midpoint of the line between them.
 For tracks consisting of three points or more, we calculate the convex hull of all the points included in the track and use the centroid of this polygon.
-For the special case where a track contains more than two points but where the points are collinear, we use the midpoint of the line through the points.
+For the special case where a track contains more than two points but where the points are collinear (i.e., lie on a single straight line), we use the midpoint of the line through the points.
 We hereafter refer to these centroids as track centroids.
@@
 
-Manual tracking of individual flowers derived from image-based plot monitoring enabled detailed information on flower phenology at the level of individuals with substantial variation in flowering duration between flowers (Fig.
-3).
+## Simulated data
+
+Tracking for simulation A without applying the three tracking parameters resulted in a merging error (i.e.
+the two objects were tracked as one).
+Setting the max distance to 200 removed this error and resulted in perfect tracking of the two objects.
+Tracking for simulation B without applying the tracking parameters returned 8 tracks and 9 mismatches (MOTA = 0.55).
+Setting max gap to 4 decreased the number of mismatches to 4 (4 tracks, MOTA = 0.8) and including a value of 200 for max distance as for simulation A removes the remaining errors and resulted in perfect tracking.
+For simulation C, tracking without parameters ignored returned the correct number of tracks (4), but 38 mismatches were made (MOTA = 0.98).
+Setting the running mean parameter to 20, 40, 60, resulted in 16, 8, and 4 mismatches, respectively, while a running mean value of 80 resulted in perfect tracking.
+A running mean value of 100 resulted in 4 mismatches.
+For simulation D, the algorithm tracked the four objects perfectly despite substantial overlap.
+Basing tracking on every 10th, 20th, and 30th frame resulted in 2, 4, and 10 mismatches, respectively.
+
+## Flowers
+Manual tracking of individual flowers derived from image-based plot monitoring enabled detailed information on floral phenology at the level of individuals with substantial variation in flowering duration between flowers (Fig.
+5).
+Further, the results show that the flowers in the series from Ny-Ålesund seem to have longer flowering durations than the other locations.
 Further, our tracking algorithm was able to track individual flowers with consistently high accuracy (MOTA $\geq$ 0.981) across all series.
@@
 This demonstrates the significance of a special case of tracking error, namely merging of tracks, which is not accounted for in the MOTA score.
-In Figure 4, we exemplify the two types of errors, identity shifts and merges.
+In figure 6, we exemplify the two types of errors, identity shifts and merges.
 
@@
 
-![Figure 2: Flowering lifespan (days) for each individual flower in each of the five image series plotted against the onset of flowering in days after 1 January (day of the year) based on manual tracking of individual flowers.](images/OnsetVSLengthBox.png){alt="Figure 2: Flowering lifespan (days) for each individual flower in each of the five image series plotted against the onset of flowering in days after 1 January (day of the year) based on manual tracking of individual flowers."}\
-
-Figure 3: Variation in flowering lifespan (days) for the flowers of each of the five image series plotted against the onset of flowering in days after 1 January (day of the year, DOY) based on manual tracking of individual flowers.
+![Figure 5: Flowering lifespan (days) for each individual flower in each of the five image series plotted against the onset of flowering in days after 1 January (day of the year) based on manual tracking of individual flowers](../figures/OnsetVSLengthBoxWithPoints.png){width="100%"}
 
@@
 
-![](../figures/errorExamples.png){width="100%"}
-
-**Figure 4:** Tracking results of three flowers in the 2018 NARS A image series.
+![Figure 6: Tracking results of three flowers in the 2018 NARS A image series.
 A: Track output from the flower tracking algorithm for three individual flowers using two different combinations of tracking parameters.
@@
 A gust of wind causes a displacement of the flowers between two frames, f and f+1, (7 hours apart), which causes an identity shift between flower 25 and flower 29.
-Setting running mean to 20 fixed this error as illustrated in A.
+Setting running mean to 20 fixed this error as illustrated in A.](../figures/errorExamples.png){width="100%"}
 
@@
 2).
-The results of the filtering algorithm on the three series are given in Table 3 and the filtering process is visualised in Figure 5.
+The results of the filtering algorithm on the three series are given in Table 3 and the filtering process is visualised in figure 7.
 Our filtering method extracted 28 tracks from the three series (2018 NARS A: 15 tracks; 2019 NARS C: 10 tracks; 2019 NYAA E: 3 tracks) with complex scenes using a fixed filtering distance value of 350 and fixed values for the tracking parameters (10, 300, 10 for max gap size, max distance, and running mean, respectively) without any mismatches included in the 28 tracks.
@@
 
-![Figure 5: The track filtering process exemplified by the 2019 NARS C series.
+![Figure 7: The track filtering process exemplified by the 2019 NARS C series.
 A: The centroid points for all bounding boxes in the series, coloured by the track id from the centroid tracking algorithm.
@@
 
-In this paper, we have demonstrated the potential of image-based monitoring for collecting information on flower phenology at the level of individuals.
-The automatic tracking and filtering pipeline presented was able to return reliable flower tracks despite settings that violate assumptions of standard tracking algorithms and complex scenes with a high number of objects that are not permanent within the series.
-Our tracking algorithm was capable of accurately tracking individual flowers, despite the difficult setting.
-Even with simple centroid tracking where association is based solely on displacement between the current and previous frame, tracking performance was high for all series and perfect for two.
+In this paper, we have demonstrated how image-based monitoring coupled with automated tracking can produce new insight into floral biology.
+We have presented a simple distance-based tracking algorithm and, simulated data, we have shown how tracking parameters can used to tune the tracking algorithm to deal with specific challenges and increase tracking performance.
+Further, we have shown that frame rate is an important consideration for reliable tracking and that sufficient frame rate can ensure accurate tracking of objects despite close proximity between them.
+We have tested our method on real flower data and demonstrated the feasibility of tracking individual flowers in images through time.
+As discussed below, challenges similar to the ones demonstrated through simulations existed in the real flower data and adjustment of the tracking parameters allowed for increased tracking accuracy of the real flowers.
+
+The automatic tracking and filtering pipeline presented was able to return reliable flower tracks despite settings that violate assumptions of standard tracking algorithms and complex scenes with a high number of objects that are not permanently visible within the series.
+Our tracking algorithm was capable of accurately tracking individual flowers, despite this difficult setting.
+Even with simple centroid tracking where association is based solely on displacement between the current and previous frame, tracking performances were high for all flower series and perfect for two.
 Further, with our filtering approach based on distance between tracks, we were able to discard the least reliable tracks and thereby increase the mean tracking accuracy of the remaining flowers to a perfect score.
@@
 As flowers may not be visible within the frame for their full lifespan, it is likely that for some flowers the full flowering has not been captured, although flowering durations of less than 24 hours has been observed from manual examination of the image series (Mann, personal observation).
+Interestingly, the flowers in Ny-Ålesund seemed to have extended flowering durations compared to other series.
+As this location is the only one included that contained *D.
+octopetala* this could be explained by species-specific differences or variation between sites and future studies could explore this further.
+Owing to the limited resolution of the cameras used in the present study, it is unlikley that detailed information about anther opening or pollen release can be recorded from these images, but with increased image quality, this may be possible.
 
@@
 While an increased number of objects does not necessarily increase tracking difficulty as this depends on the degree of isolation of tracks in time and space [@beard2018], multiple object tracking is a challenging task.
-Here, since the flowers are limited in space and time, the complexity of tracking likely increases with increasing number of flowers.
-A higher number of flowers within the frame increases the possibility of flowers appearing in close proximity to each other and that flowers overlap or move in and out of frame.
-All these situations translate to a higher risk of tracking errors.
+Here, since the flowers are limited in space and time, the complexity of tracking is likely to increase with higher numbers of flowers.
+More flowers within the frame increases the possibility of flowers appearing in close proximity to each other and that flowers overlap or move in and out of frame.
+This translates to a higher risk of tracking errors.
 The very high complexity in the scenes with many flowers, especially 2019 NARS C with 85 flowers, coupled with the low frame rate renders it unlikely that any tracking method could produce perfect results across series.
@@
 Thus, even if the number of tracks equals the number of flowers, it does not necessarily follow that the tracking is correct.
-For instance, the tracking algorithm could split one ground truth track in two, but combine two other ground truth tracks into one, as seen in Figure 3.
-The task of evaluating multiple object tracking accuracy is a major challenge in itself [@luiten2020; @milan2013; @milan2016] and in a future study, other tracking accuracy metrics could be implemented, that account for merging errors in a better way.
+For instance, the tracking algorithm could split one ground truth track in two, but merge two other ground truth tracks into one, as seen in figure 6.
+The task of evaluating multiple object tracking accuracy is a challenge in itself [@luiten2020; @milan2013; @milan2016] and in a future study, other tracking accuracy metrics could be implemented that account for merging errors in a better way.
 Tracking flowers may be a challenge that is especially prone to these errors since flowers can bloom in the same area at different times during the season.
@@
 The three series with complex scenes all contain flowers at the edge of the frame (see Fig.
-1), which means that they may occasionally be outside of view of the camera.
+3), which means that they may occasionally be outside of view of the camera.
 Further, the higher number of flowers increase the risk of flowers temporarily occluding others.
@@
 
-# Significance for ecological studies
+## Automated monitoring of ecology
 
-Our results show that it is feasible to collect and analyse individual flower phenology at large scale at logistically challenging sites, which is a significant contribution towards automatic and efficient phenology data collection.
-The possibility of collecting individual level flower phenology is highly relevant for exploration of many research questions.
+Collection of continuous observational ecological data can be a time-consuming task.
+@Estes2018 argued that as ecologists rely on conventional field techniques, the observational scales of ecological studies are generally narrow and may be insufficient to comprehend the ecological phenomena in focus.
+Development and implementation of technological tools for ecological data collection can facilitate ecological studies with larger temporal and spatial scopes [@Pimm2015; @allan2018].
+For processes that require high and continual temporal observation frequency to monitor, such as tracking individuals and their interactions, automation of data collection is critical, especially in remote places where human access is limited.
+Here, we have demonstrated the potential of image-based monitoring for collecting information on floral phenology at the level of individual flowers.
+Coupling this with automated detection of flowers as demonstrated in @mann2022 permits an automated pipeline for collection of floral biology data.
+
+The possibility of collecting individual level floral phenology is highly relevant for exploration of many research questions.
 For example, it would be possible to assign indicators of seed set to the individual and explore how seed set is affected by timing and length of flowering and/or abiotic conditions such as frost events.
 Hereby, the long-standing question of what drives variation in reproductive success of individual flowers could be explored [@bolmgren2015].
-Further, the fact that it was not uncommon in our results that flowering lasted less than a week, emphasizes that weekly observations of sample plots can be insufficient for comprehensive description of flower phenology [@Prevey2021].
+Further, the fact that it was not uncommon in our results that flowering lasted less than a week, emphasizes that weekly observations of sample plots can be insufficient for comprehensive description of floral phenology [@Prevey2021].
 
-By simultaneously tracking flower visits, these could be assigned to the individual flower and visitation rates per flower could be derived and the effects on reproductive success could be investigated [@tiusanen2016; @rafferty2012].
+It is clear that images derived from monitoring of permanent plots can hold substantial amounts of ecologically relevant data.
+To derive some variables, for example flower visitation rates, image series collected at high temporal frequency is required.
+As the scale of data increases as well as the number of variables to be extracted from the images increase, automated image analysis becomes critical.
+Here, methods from deep learning and computer vision are exceedingly relevant.
+For instance, it has been demonstrated that convolutional neural networks can facilitate monitoring of pollination events by detecting, identifying, and counting insects [@bjerge2021] and perform fine-grained species classification [@spiesman2021].
+Collectively, these developments coupled with implementation of long-term high-frequency image-based monitoring will allow for collection and processing of ecological data at unprecedented scales [@hoye2021].
+By coupling pollinator detection and classification with floral tracking, visitation rates per flower could be derived and the effects on reproductive success could be investigated [@tiusanen2016; @rafferty2012].
+
 Pollen limitation is a common factor limiting the reproductive success of flowering plants.
@@
 
-It is clear that the images derived from monitoring of permanent plots can hold substantial amounts of ecologically relevant data.
-To derive some variables, for example flower visitation rates, image series collected at high temporal frequency is required.
-As the scale of data increases as well as the number of variables to be extracted from the images increase, automated image analysis becomes relevant.
-Here, methods from deep learning and computer vision are exceedingly relevant.
-For instance, it has been demonstrated that convolutional neural networks can facilitate monitoring of pollination events by detecting, identifying, and counting insects [@bjerge2021] and perform fine-grained species classification [@spiesman2021].
-Collectively, these developments coupled with implementation of long-term high-frequency image-based monitoring will allow for collection and processing of ecological data at unprecedented scales [@hoye2021].
-For flower tracking in particular, our results emphasize that this is not a trivial task.
+Our results emphasize that this is not a trivial task.
 The data violate common assumptions of tracking methods and tracking individual flowers through time-lapse images therefore represents a special case [@chen2022; @betke2007; @luo2021].
@@
 Our tracking algorithm returned consistently accuracy despite this.
+While we show that tracking can be done at high accuracy with our tracking algorithm, we also identify particular challenges with flower tracking, such as the violations of standard assumptions in tracking algorithms, the potentially large number of individuals overlapping in time and space, and the possible susceptibility to track merging.
+To improve tracking ability of flowers in future studies, focus should be on implementing reliable measures of tracking error and on developing methods that can alleviate the challenges that exist with tracking visually similar objects, as well as on comparison of different tracking approaches.
+Alongside this, methods for extracting auxiliary information from images, such as flower seed set, should be explored.
 
-While we show that tracking can be done at high accuracy with our tracking algorithm, we also identify particular challenges with flower tracking, such as the violations of standard assumptions in tracking algorithms, the potentially large number of individuals overlapping in time and space, and the possible susceptibility to track merging.
-To improve tracking ability of flowers in future studies, focus should be on implementing reliable measures of different types of errors and on developing methods that can alleviate the challenges, as well as on comparison of different tracking approaches.
-Alongside this, methods for extracting auxiliary information from images, such as flower seed set, should be explored.
+Image-based methods facilitate automated data collection a high frequency, across diurnal cycles and throughout seasons, but the extensive amount of data that results from automated data collection also necessitates efficient and automated analysis pipelines.
+Our results show that it is feasible to collect and analyse individual floral phenology in high detail at logistically challenging sites, which is a significant contribution towards automatic and efficient phenology data collection.
+Most importantly, adoption of automated methods for data collection and analysis within ecology, and specifically within for example pollination ecology, will transform the extent and scale of studies and the questions that can be explored.
 
@@
 The authors have no conflict of interest to declare.
+
+Thank you to Adrian Rosebrock (https://pyimagesearch.com/2018/07/23/simple-object-tracking-with-opencv/, accessed on November 29th 2022) for inspiration for the design of the centroid tracking algorithm.
 

### 2023-01-13 code/DBSCANFiltering.py

@@ -1,183 +1,0 @@
-# -*- coding: utf-8 -*-
-"""
-Created on Tue Dec 21 13:28:48 2021
-
-@author: au309263
-"""
-
-import geopandas as gpd
-import pandas as pd
-from shapely.geometry import Polygon, LineString, Point
-from sklearn.cluster import DBSCAN
-import matplotlib.pyplot as plt
-#from filtering import sieve
-import numpy as np
-from scipy.spatial import ConvexHull
-from datetime import datetime
-
-# Note, EPS_DISTANCE needs to be smaller than the gap between any two polygons we want to keep and large enough to cluster polygons that we want to discard together
-# EPS_DISTANCE = 281 is the highest value that keeps all tracks in THUL-01
-
-#tracks = pd.read_csv(r'U:\BITCue\Projekter\TrackingFlowers\testResults\_parameterTest_NARS-04_3\parameterTest_NARS-04_maxDisap_10_runMean_10_maxDist_500.csv')
-
-class DBSCANsieve():
-    def __init__(self, tracks, eps_distance, min_sample_polygons):
-        if not isinstance(tracks, pd.DataFrame):
-            self.tracks = pd.read_csv(tracks, sep=",", header = 0)
-        else:
-            self.tracks = tracks
-        
-        self.eps_distance = eps_distance
-        self.min_sample_polygons = min_sample_polygons
-        
-        self.points = {}
-        self.lines = {}
-        self.triangles = {}
-        self.polygons = {}
-        self.polyLengths = {}
-        self.polyHulls = {}
-        
-    def gen(self): # Generator that yields next group in tracks dataframe
-        t = self.tracks.groupby('objectID')
-        for i,g in t: 
-            oid = g['objectID'].iloc[0]
-            x_cs = g['x_c'].tolist()
-            y_cs = g['y_c'].tolist()
-            p = list(map(list, zip(x_cs, y_cs)))
-            yield oid, p
-    
-    
-    def separate(self): # Take next group in tracks dataframe from the generator and split into correct dictionary according to number of points the track contains
-        for oid, p in self.gen(): 
-            if len(p) == 1: # Get single points
-                self.points[oid] = p
-                
-            if len(p) == 2: # Get lines
-                self.lines[oid] = p
-                
-            if len(p) == 3: # Get triangles
-                self.triangles[oid] = p
-
-            if len(p) > 3: # Get polygons
-                self.polygons[oid] = p
-                
[…]

### 2023-01-13 code/DistanceFiltering.py

@@ -1,213 +1,0 @@
-# -*- coding: utf-8 -*-
-"""
-Created on Mon Jan 31 12:56:11 2022
-
-@author: au309263
-"""
-# -*- coding: utf-8 -*-
-"""
-Created on Tue Dec 21 13:28:48 2021
-
-@author: au309263
-"""
-
-import geopandas as gpd
-import pandas as pd
-from shapely.geometry import Polygon, LineString, Point
-from sklearn.cluster import DBSCAN
-import matplotlib.pyplot as plt
-from filtering import sieve
-import numpy as np
-from scipy.spatial import ConvexHull
-from datetime import datetime
-from itertools import combinations
-
-
-# Note, EPS_DISTANCE needs to be smaller than the gap between any two polygons we want to keep and large enough to cluster polygons that we want to discard together
-# EPS_DISTANCE = 281 is the highest value that keeps all tracks in THUL-01
-
-#tracks = pd.read_csv(r'U:\BITCue\Projekter\TrackingFlowers\testResults\_parameterTest_NARS-04_3\parameterTest_NARS-04_maxDisap_10_runMean_10_maxDist_500.csv')
-
-class distanceSieve():
-    def __init__(self, tracks, eps_distance):
-        if not isinstance(tracks, pd.DataFrame):
-            self.tracks = pd.read_csv(tracks, sep=",", header = 0)
-        else:
-            self.tracks = tracks
-        
-        self.eps_distance = eps_distance
-        
-        self.points = {}
-        self.lines = {}
-        self.triangles = {}
-        self.polygons = {}
-        self.polyLengths = {}
-        self.polyHulls = {}
-        
-    def gen(self): # Generator that yields next group in tracks dataframe
-        t = self.tracks.groupby('objectID')
-        for i,g in t: 
-            oid = g['objectID'].iloc[0]
-            x_cs = g['x_c'].tolist()
-            y_cs = g['y_c'].tolist()
-            p = list(map(list, zip(x_cs, y_cs)))
-            yield oid, p
-    
-    
-    def separate(self): # Take next group in tracks dataframe from the generator and split into correct dictionary according to number of points the track contains
-        for oid, p in self.gen(): 
-            #print(p)
-            if len(p) == 1: # Get single points
-                self.points[oid] = p
-                
-            if len(p) == 2: # Get lines
-                self.lines[oid] = p
-                
-            if len(p) == 3: # Get triangles
-                self.triangles[oid] = p
-
[…]

### 2023-01-13 code/run.py

@@ -1,809 +1,0 @@
-# -*- coding: utf-8 -*-
-"""
-Created on Tue Dec 14 12:50:10 2021
-
-@author: au309263
-"""
-
-import pandas as pd
-from track import tracker
-#from filtering import sieve
-from DBSCANFiltering import DBSCANsieve
-#from distanceFiltering import distanceSieve
-#from evaluate import evaluator
-#import os
-import time
-import os
-import pandas as pd
-from scipy.spatial import distance as dist
-import os
-import re
-import numpy as np
-
-import motmetrics as mm
-
-###################
-
-
-def impdata(camID):
-    if camID == "NARS-04":
-        gt = r"U:\BITCue\Projekter\TrackingFlowers\data\annotations\2020_05_17_NorwayAnnotations_NARS-04_IndividualAnnotations_FRCNN_Metrics.csv"
-    if camID == "NARS-13":
-        gt = r"U:\BITCue\Projekter\TrackingFlowers\data\annotations\2020_04_30_NorwayAnnotations_NARS-13_IndividualAnnotations_FRCNN_Metrics.csv"
-    if camID == "THUL-01":
-        gt = r"U:\BITCue\Projekter\TrackingFlowers\data\annotations\2020_05_15_NorwayAnnotations_THUL-01_IndividualAnnotations_FRCNN_Metrics.csv"
-    if camID == "NYAA-04":
-        gt = r"U:\BITCue\Projekter\TrackingFlowers\data\annotations\2021_12_13_NorwayAnnotations_NYAA-04_IndividualAnnotations_FRCNN_Metrics.csv"
-    if camID == "NARS-17":
-        gt = r"U:\BITCue\Projekter\TrackingFlowers\data\annotations\2021_12_30_NorwayAnnotations_NARS-17_IndividualAnnotations_FRCNN_Metrics.csv"
-    if camID == "debug":
-        gt = r"U:\BITCue\Projekter\TrackingFlowers\data\annotations\debugGT.csv"
-    return gt
-        
-
-
-
-# ===================== Run evaluator on all files for the parameter test ====
-
-br = '\n'
-
-def write_header(results_filename):
-    with open(results_filename, 'a') as resultFile: # Write the header of the output file
[…]

### 2023-01-13 code/track.py

@@ -1,280 +1,0 @@
-# -*- coding: utf-8 -*-
-"""
-### Multi flower tracking algorithm ###
-Created 2021
-
-@author:
-Hjalte M. R. Mann
-TECOLOGY.xyz
-
-Detections are in the format: x_min, y_min, x_max, y_max
-
-Max distance: Threshold for maximum distance between point and track before initiation of a new track is forced
-Max dissapeared: Max number of frames a track can be disappeared before the track is terminated and a new point in that area are considered new tracks
-Running mean: Number of previous frame on which to calculate track position (to calculate distance between point and track). If set to one, position is centroid of previous track, if above one, means of x and y are calculated.
-
-The centroid tracking approach was based on: https://pyimagesearch.com/2018/07/23/simple-object-tracking-with-opencv/ (Accessed on 2022-06-10)
-
-
-### Pseudocode of tracking algorithm: ###
-
--Go to next frame
-    -If the frame has no detections
-        -Add one to disappeared counter for any existing tracks
-    -If the frame has detections
-        -If we are not tracking objects
-            -Initiate tracks for all points
-        -If we are tracking objects, calculate all pairwise distances (If running mean above 1, we use the mean x and y position of the track, else we just use position in the previous frame)
-            -For all distances below max distance threshold, try to associate points and tracks (shortest distance from point to object get associated and removed, then the second shortest distance etc.)
-            -For points with distances above max distance
-                -Initiate new tracks
-                -Add one to track id counter
-        -For points that did not get associated to an exisiting track
-            -Initiate new tracks
-            -Add one to track id counter
-        -Add frame tracking data to result dataframe for final output
-        -Update running mean dictionary with new data
-            -For tracks that are at the running mean threshold, remove oldest position and add current
-        -For tracks that did not get a point associated to them, add one to disappeared counter
-            -For tracks that exceed disappeared threshold, remove from active tracking storage
-"""
-
-import pandas as pd # Replace with cudf if performance is too slow?
-from collections import OrderedDict
-import numpy as np
-from scipy.spatial import distance as dist
-import matplotlib.pyplot as plt
-import time
-from statistics import mean
-from datetime import datetime
-
-br = "\n" # Line break for use in code
-
-
-# ===================== SETTINGS ==============================================
-
[…]

### 2023-01-13 code/trackForPrint.py

@@ -1,276 +1,0 @@
-# -*- coding: utf-8 -*-
-"""
-Custom multi flower tracker
-Created 2021
-
-@author:
-Hjalte M. R. Mann
-TECOLOGY.xyz
-
-Detections are in the format: x_min, y_min, x_max, y_max
-
-Max distance: Threshold for maximum distance between point and track before initiation of a new track is forced
-Max dissapeared: Max number of frames a track can be disappeared before the track is terminated and a new point in that area are considered new tracks
-Running mean: Number of previous frame on which to calculate track position (to calculate distance between point and track). If set to one, position is centroid of previous track, if above one, means of x and y are calculated.
-
-
-Pseudocode to 
-
--Go to next frame
-    -If the frame has no detections
-        -Add one to disappeared counter for any existing tracks
-    -If the frame has detections
-        -If we are not tracking objects
-            -Initiate tracks for all points
-        -If we are tracking objects, calculate all pairwise distances (If running mean above 1, we use the mean x and y position of the track, else we just use position in the previous frame)
-            -For all distances below max distance threshold, try to associate points and tracks (shortest distance from point to object get associated and removed, then the second shortest distance etc.)
-            -For points with distances above max distance
-                -Initiate new tracks
-                -Add one to track id counter
-        -For points that did not get associated to an exisiting track
-            -Initiate new tracks
-            -Add one to track id counter
-        -Add frame tracking data to result dataframe for final output
-        -Update running mean dictionary with new data
-            -For tracks that are at the running mean threshold, remove oldest position and add current
-        -For tracks that did not get a point associated to them, add one to disappeared counter
-            -For tracks that exceed disappeared threshold, remove from active tracking storage
-            
-
-"""
-
-import pandas as pd # Replace with cudf if performance is too slow?
-from collections import OrderedDict
-import numpy as np
-from scipy.spatial import distance as dist
-import matplotlib.pyplot as plt
-import time
-from statistics import mean
-from datetime import datetime
-
-br = "\n" # Line break for use in code
-
-
-# ===================== SETTINGS ==============================================
-
-verbose = True # Set to True if you want tracking process printed to screen and False if not
-
-class tracker():
-    def __init__(self, max_disappeared, max_distance, running_mean_threshold, results_filename, frames, detections, verbose):
[…]

### 2023-01-13 figures/flowertrackingOverview.py

@@ -1,0 +1,274 @@
+from matplotlib.markers import MarkerStyle
+import matplotlib.pyplot as plt
+from matplotlib.lines import Line2D
+from matplotlib.transforms import Affine2D
+from matplotlib.markers import JoinStyle, CapStyle
+
+import matplotlib.path as mpath
+import numpy as np
+
+
+# marker_inner = dict(markersize=35,
+#                     markerfacecolor='tab:blue',
+#                     markerfacecoloralt='lightsteelblue',
+#                     markeredgecolor='brown',
+#                     markeredgewidth=8,
+#                     )
+
+# text_style = dict(horizontalalignment='right', verticalalignment='center',
+#                   fontsize=12, fontfamily='monospace')
+# marker_style = dict(linestyle=':', color='0.8', markersize=50,
+#                     markerfacecolor="yellow", markeredgecolor="black", transform = Affine2D().rotate_deg(90) )
+
+
+# def format_axes(ax):
+#     ax.margins(0.2)
+#     ax.set_axis_off()
+#     ax.invert_yaxis()
+
+
+
+
+# star = mpath.Path.unit_regular_star(8)
+# circle = mpath.Path.unit_circle()
+# # concatenate the circle with an internal cutout of the star
+# cut_star = mpath.Path(
+#     vertices=np.concatenate([circle.vertices, star.vertices[::-1, ...]]),
+#     codes=np.concatenate([circle.codes, star.codes]))
+
+# fig, ax = plt.subplots()
+# #fig.suptitle('Path markers', fontsize=14)
+# #fig.subplots_adjust(left=0.4)
+
+# markers = {'star': star}
+
+# # for y, (name, marker) in enumerate(markers.items()):
+# #     print("y", y)
+# #     print("name ", name)
+# #     print("marker", marker)
+# #ax.text(-0.5, 0, 'star', **text_style)
+# ax.plot([0] * 1, marker=star, **marker_style)
[…]

### 2023-01-13 src/Simulation_plotting.py

@@ -1,0 +1,190 @@
+import matplotlib
+import pandas as pd
+import matplotlib.pyplot as plt
+import numpy as np
+import math
+from mpl_toolkits import mplot3d
+
+tran = {0:"a",
+        1:"b",
+        2:"c",
+        3:"d"
+        }
+
+# Read data
+
+### MAX GAP ###
+maxgap = pd.read_csv(r"U:\BITCue\Projekter\TrackingFlowers\data\simulation/maxgapExample.csv")
+
+maxgap['objectLetter'] = maxgap['object'].map(lambda x: tran[x])
+
+f, (ax1, ax2) = plt.subplots(2, 1, sharex=False, height_ratios=[4, 1])
+
+ax1.set_xlim([0, 1000])
+ax1.set_ylim([0, 1000])
+ax2.margins(y = 0.2)
+
+ax1.scatter(maxgap["x_c"], maxgap["y_c"], c=maxgap['object'])
+ax1.set_box_aspect(1)
+ax1.set_title('Max gap')
+ax2.scatter(maxgap["frame"], maxgap["objectLetter"], c = maxgap['object'])
+ax2.xaxis.set_ticks(np.arange(0, 10, 1))
+ax2.set_box_aspect(1/4)
+ax2.invert_yaxis()
+plt.savefig(r"U:\BITCue\Projekter\TrackingFlowers\data\simulation/maxgapPlot.pdf", bbox_inches='tight')
+plt.show()
+
+
+# ### MAX DIST ###
+
+maxdist = pd.read_csv(r"U:\BITCue\Projekter\TrackingFlowers\data\simulation/maxdistExample.csv")
+
+maxdist['objectLetter'] = maxdist['object'].map(lambda x: tran[x])
+print(maxdist)
+f, (ax1, ax2) = plt.subplots(2, 1, sharex=False, height_ratios=[4, 1])
+
+ax1.set_xlim([0, 1000])
+ax1.set_ylim([0, 1000])
+ax2.margins(y = 0.2)
+
+ax1.scatter(maxdist["x_c"], maxdist["y_c"], c = maxdist['object'])
+ax1.set_box_aspect(1)
+ax1.set_title('Max distance')
+ax2.scatter(maxdist["frame"], maxdist["objectLetter"], c = maxdist['object'])
+ax2.xaxis.set_ticks(np.arange(0, 20, 2))
+ax2.set_box_aspect(1/4)
+ax2.invert_yaxis()
+
[…]

### 2023-01-13 src/Simulation_produceData.py

@@ -1,0 +1,349 @@
+from sklearn.datasets import make_blobs
+import matplotlib.pyplot as plt
+import pandas as pd
+import random as rand
+import numpy as np
+import math
+from mpl_toolkits import mplot3d
+import pprint
+from operator import add, sub
+import csv
+
+
+
+# Get random point in the full image - this is the center point of the flower (in theory stalk position)
+
+# Define a radius for the flower
+
+# Generate random point within the flower circle
+
+# Define a movement radius (max movement between frames)
+
+# Generate random point within the movement radius and within the flower radius
+
+# xDim = 1
+# yDim = 1
+# flowerRadius = 0.2 #frameDiag * 0.2
+# movementRadius = 0.2 # flowerRadius * 0.001
+# numberFlowers = 4
+# pointsPerFlower = 1000
+
+rand.seed(42)
+
+
+### DEFINE SOME FUNCTIONS ###
+class simulate():
+    def __init__(self, flowerRadius, movementRadius, numberFlowers, pointsPerFlower, filename, experiment):
+        self.flowerRadius = flowerRadius
+        self.movementRadius = movementRadius
+        self.numberFlowers = numberFlowers
+        self.pointsPerFlower = pointsPerFlower
+        #self.fixedPosition = fixedPosition
+        #self.delete = delete
+        self.flowerCenters = []
+        self.flowers = []
+        self.filename = filename
+        self.experiment = experiment
+
+    def checkPointInCircle(self, flowerCenter, flowerRadius, flowerPoint):
+        # Compare radius of circle
+        # with distance of its center
+        # from given point
+
+        if ((flowerPoint[0] - flowerCenter[0]) * (flowerPoint[0] - flowerCenter[0]) +
+            (flowerPoint[1] - flowerCenter[1]) * (flowerPoint[1] - flowerCenter[1]) <= flowerRadius * flowerRadius):
+            return True
+        else:
+            return False
+
+    def getFlowerCenter(self):
+        xFlower, yFlower = rand.uniform(0,1), rand.uniform(0,1)#flowerRadius, math.floor(xDim-flowerRadius)), rand.uniform(math.floor(yDim-flowerRadius))
+        return [xFlower, yFlower]
+
+
+    def getFirstFlowerPoint(self, flowerCenter):
+        while True:
+            r = self.flowerRadius * math.sqrt(rand.uniform(0, 1))
[…]

### 2023-01-13 src/Simulation_run.py

@@ -1,0 +1,825 @@
+# -*- coding: utf-8 -*-
+"""
+Created on Tue Dec 14 12:50:10 2021
+
+@author: au309263
+"""
+import sys
+
+sys.path.append("..")
+import pandas as pd
+from track import tracker
+
+#from filtering import sieve
+#from DBSCANFiltering import DBSCANsieve
+#from distanceFiltering import distanceSieve
+#from evaluate import evaluator
+#import os
+import time
+import os
+import pandas as pd
+from scipy.spatial import distance as dist
+import os
+import re
+import numpy as np
+
+import motmetrics as mm
+br = '\n'
+###################
+
+# ===================== Run evaluator on all files for the parameter test ====
+
+#max_disappeared, max_distance, running_mean_threshold, results_filename, frames, detections, verbose
+simulated = pd.read_csv(r"U:\BITCue\Projekter\TrackingFlowers\data\simulation\maxgapExample.csv")
+simulated['x_min'] = simulated["x_c"] - 1
+simulated['x_max'] = simulated["x_c"] + 1
+simulated['y_min'] = simulated["y_c"] - 1
+simulated['y_max'] = simulated["y_c"] + 1
+simulated['frame'] = simulated['frame'].astype('int')
+simulated['filename'] = "simulated_test1"
+
+t = tracker(4, 200, 1, r"U:\BITCue\Projekter\TrackingFlowers\delete.csv", simulated, True) # Instantiate the class instance and pass in the threshold for max_disappeared and the list of frames.
+starttime = time.time()
+t.track()
+endtime = time.time()
+#print(f'Tracking done. That took {round(endtime-starttime, 3)} seconds. That is {round((endtime-starttime)/len(frames), 3)} seconds per frame.')
+#t.write_tracks_file()
+
+
+def write_header(results_filename):
+    with open(results_filename, 'a') as resultFile: # Write the header of the output file
+        header = f'maxDisap,maxDist,runMean,num_frames,num_detections,num_objects,num_predictions,num_unique_objects,num_tracks,num_matches,num_switches,num_misses,num_false_positives,precision,recall,mota,motp,mostly_tracked,partially_tracked,mostly_lost{br}'
+        resultFile.write(header)
+
+
+#results_filename = r'U:\BITCue\Projekter\TrackingFlowers\testResults/' + "framerateExample_100th" + '_0_0_1.csv'
[…]

### 2023-01-13 src/distanceFiltering.py

@@ -1,0 +1,213 @@
+# -*- coding: utf-8 -*-
+"""
+Created on Mon Jan 31 12:56:11 2022
+
+@author: au309263
+"""
+# -*- coding: utf-8 -*-
+"""
+Created on Tue Dec 21 13:28:48 2021
+
+@author: au309263
+"""
+
+import geopandas as gpd
+import pandas as pd
+from shapely.geometry import Polygon, LineString, Point
+from sklearn.cluster import DBSCAN
+import matplotlib.pyplot as plt
+#from filtering import sieve
+import numpy as np
+from scipy.spatial import ConvexHull
+from datetime import datetime
+from itertools import combinations
+
+
+# Note, EPS_DISTANCE needs to be smaller than the gap between any two polygons we want to keep and large enough to cluster polygons that we want to discard together
+# EPS_DISTANCE = 281 is the highest value that keeps all tracks in THUL-01
+
+#tracks = pd.read_csv(r'U:\BITCue\Projekter\TrackingFlowers\testResults\_parameterTest_NARS-04_3\parameterTest_NARS-04_maxDisap_10_runMean_10_maxDist_500.csv')
+
+class distanceSieve():
+    def __init__(self, tracks, eps_distance):
+        if not isinstance(tracks, pd.DataFrame):
+            self.tracks = pd.read_csv(tracks, sep=",", header = 0)
+        else:
+            self.tracks = tracks
+        
+        self.eps_distance = eps_distance
+        
+        self.points = {}
+        self.lines = {}
+        self.triangles = {}
+        self.polygons = {}
+        self.polyLengths = {}
+        self.polyHulls = {}
+        
+    def gen(self): # Generator that yields next group in tracks dataframe
+        t = self.tracks.groupby('objectID')
+        for i,g in t: 
+            oid = g['objectID'].iloc[0]
+            x_cs = g['x_c'].tolist()
+            y_cs = g['y_c'].tolist()
+            p = list(map(list, zip(x_cs, y_cs)))
+            yield oid, p
+    
+    
+    def separate(self): # Take next group in tracks dataframe from the generator and split into correct dictionary according to number of points the track contains
+        for oid, p in self.gen(): 
+            #print(p)
+            if len(p) == 1: # Get single points
+                self.points[oid] = p
+                
+            if len(p) == 2: # Get lines
+                self.lines[oid] = p
+                
+            if len(p) == 3: # Get triangles
+                self.triangles[oid] = p
+
[…]

### 2023-01-13 src/evaluate.py

@@ -1,0 +1,158 @@
+import motmetrics as mm
+import pandas as pd
+
+
+br = '\n'
+
+
+class tracker():
+    def __init__(self, dt, gt, results_filename, verbose = True):
+        self.dt = dt
+        self.gt = gt
+        self.results_filename = results_filename
+        self.verbose = verbose
+
+
+
+    def write_header(results_filename):
+        with open(results_filename, 'a') as resultFile: # Write the header of the output file
+            header = f'maxDisap,maxDist,runMean,num_frames,num_detections,num_objects,num_predictions,num_unique_objects,num_tracks,num_matches,num_switches,num_misses,num_false_positives,precision,recall,mota,motp,mostly_tracked,partially_tracked,mostly_lost{br}'
+            resultFile.write(header)
+
+    def write_results_file(results_filename,maxDisap,maxDist,runMean,num_frames,num_detections,num_objects, num_predictions,num_unique_objects,num_tracks,num_matches,num_switches,num_misses,num_false_positives,precision,recall,mota,motp,mostly_tracked,partially_tracked,mostly_lost):
+        with open(results_filename, 'a') as resultFile:
+            resultFile.write(f'{maxDisap}, {maxDist}, {runMean}, {num_frames},{num_detections},{num_objects},{num_predictions},{num_unique_objects},{num_tracks},{num_matches},{num_switches},{num_misses},{num_false_positives},{precision},{recall},{mota},{motp},{mostly_tracked},{partially_tracked},{mostly_lost}{br}')
+
+
+    ### Prepare ground truth data
+    def prepGT(gt, d):
+        gt['x_c'] = (gt['x_min'] + gt['x_max'])/2
+        gt['x_c'] = gt['x_c'].round(0).astype(int)
+        
+        gt['y_c'] = (gt['y_min'] + gt['y_max'])/2
+        gt['y_c'] = gt['y_c'].round(0).astype(int)
+
+        gt['frame'] = gt['filename'].map(d) # Create a frame column by mapping filename to dictionary
+        
+        gt['id_gt_int'] = gt.apply(lambda x: int(''.join(filter(str.isdigit, x['id_gt']))), axis=1)
+        
+        return gt
+
+    ### Prepare tracked data
+    def prepSORT(sort, d):
+        #sort.drop([6, 7, 8, 9], axis=1, inplace = True)
+        #print(sort)
[…]

### 2023-01-13 src/run.py

@@ -1,0 +1,834 @@
+# -*- coding: utf-8 -*-
+"""
+Created on Tue Dec 14 12:50:10 2021
+
+@author: au309263
+"""
+
+import pandas as pd
+from track import tracker
+#from filtering import sieve
+#from DBSCANFiltering import DBSCANsieve
+from distanceFiltering import distanceSieve
+#from evaluate import evaluator
+#import os
+import time
+import os
+from scipy.spatial import distance as dist
+import os
+import re
+import numpy as np
+
+#import motmetrics as mm
+
+###################
+
+### Function for importing data
+def impdata(camID):
+    if camID == "NARS-04":
+        gt = r"U:\BITCue\Projekter\TrackingFlowers\data\annotations\2020_05_17_NorwayAnnotations_NARS-04_IndividualAnnotations_FRCNN_Metrics.csv"
+    if camID == "NARS-13":
+        gt = r"U:\BITCue\Projekter\TrackingFlowers\data\annotations\2020_04_30_NorwayAnnotations_NARS-13_IndividualAnnotations_FRCNN_Metrics.csv"
+    if camID == "THUL-01":
+        gt = r"U:\BITCue\Projekter\TrackingFlowers\data\annotations\2020_05_15_NorwayAnnotations_THUL-01_IndividualAnnotations_FRCNN_Metrics.csv"
+    if camID == "NYAA-04":
+        gt = r"U:\BITCue\Projekter\TrackingFlowers\data\annotations\2021_12_13_NorwayAnnotations_NYAA-04_IndividualAnnotations_FRCNN_Metrics.csv"
+    if camID == "NARS-17":
+        gt = r"U:\BITCue\Projekter\TrackingFlowers\data\annotations\2021_12_30_NorwayAnnotations_NARS-17_IndividualAnnotations_FRCNN_Metrics.csv"
+    if camID == "debug":
+        gt = r"U:\BITCue\Projekter\TrackingFlowers\data\annotations\debugGT.csv"
+    return gt
+        
+
+
+# ===================== Run evaluator on all files for the parameter test ====
+
+br = '\n'
+
+# def write_header(results_filename):
+#     with open(results_filename, 'a') as resultFile: # Write the header of the output file
[…]

### 2023-01-13 src/test.py

@@ -1,0 +1,12 @@
+
+
+"""
+This runs the tracking algorithm on a set of simulated data to assert that the algorithm is running as expected.
+
+"""
+
+
+
+def test_tracking():
+    assert 1 + 1 == 2
+

### 2023-01-13 src/track.py

@@ -1,0 +1,295 @@
+# -*- coding: utf-8 -*-
+"""
+### Multi flower tracking algorithm ###
+Created 2021
+
+@author:
+Hjalte M. R. Mann
+TECOLOGY.xyz
+
+Detections are in the format: x_min, y_min, x_max, y_max
+
+Max distance: Threshold for maximum distance between point and track before initiation of a new track is forced
+Max dissapeared: Max number of frames a track can be disappeared before the track is terminated and a new point in that area are considered new tracks
+Running mean: Number of previous frame on which to calculate track position (to calculate distance between point and track). If set to one, position is centroid of previous track, if above one, means of x and y are calculated.
+
+The centroid tracking approach was based on: https://pyimagesearch.com/2018/07/23/simple-object-tracking-with-opencv/ (Accessed on 2022-06-10)
+
+
+### Pseudocode of tracking algorithm: ###
+
+-Go to next frame
+    -If the frame has no detections
+        -Add one to disappeared counter for any existing tracks
+    -If the frame has detections
+        -If we are not tracking objects
+            -Initiate tracks for all points
+        -If we are tracking objects, calculate all pairwise distances (If running mean above 1, we use the mean x and y position of the track, else we just use position in the previous frame)
+            -For all distances below max distance threshold, try to associate points and tracks (shortest distance from point to object get associated and removed, then the second shortest distance etc.)
+            -For points with distances above max distance
+                -Initiate new tracks
+                -Add one to track id counter
+        -For points that did not get associated to an exisiting track
+            -Initiate new tracks
+            -Add one to track id counter
+        -Add frame tracking data to result dataframe for final output
+        -Update running mean dictionary with new data
+            -For tracks that are at the running mean threshold, remove oldest position and add current
+        -For tracks that did not get a point associated to them, add one to disappeared counter
+            -For tracks that exceed disappeared threshold, remove from active tracking storage
+"""
+
+import pandas as pd
+from collections import OrderedDict
+import numpy as np
+from scipy.spatial import distance as dist
+import matplotlib.pyplot as plt
+import time
+from statistics import mean
+from datetime import datetime
+import sys
+import math
+
+br = "\n" # Line break for use in code
+
+
+# ===================== SETTINGS ==============================================
+
+verbose = True # Set to True if you want tracking process printed to screen and False if not
+
+
[…]

### 2023-01-13 webapp.py

@@ -1,278 +1,0 @@
-# -*- coding: utf-8 -*-
-"""
-Created on Mon Dec 13 15:30:13 2021
-
-@author: au309263
-"""
-
-import streamlit as st
-import pandas as pd
-from code.track import tracker
-from code.filtering import sieve
-import time
-import matplotlib.pyplot as plt
-from PIL import Image
-
-
-
-#image = Image.open('streamlit_logo.png')
-
-
-import base64
-
-LOGO_IMAGE = 'streamlit_logo.png'
-
-st.markdown(
-    """
-    <style>
-    .container {
-        display: flex;
-    }
-    .logo-text {
-        font-weight:600 !important;
-        font-size:50px !important;
-        color: #000000 !important;
-        padding-top: 0px !important;
-        paddinglleft: 90px;
-    }
-    .logo-img {
-        float:right;
-    }
-    </style>
-    """,
-    unsafe_allow_html=True
-)
-
-st.markdown(
-    f"""
-    <div class="container">
-        <img class="logo-img" width = 50 height = 65 src="data:image/png;base64,{base64.b64encode(open(LOGO_IMAGE, "rb").read()).decode()}" style="padding-top: 15px;">
-        <p class="logo-text">Automatic Flower Tracking</p>
-    </div>
-    """,
-    unsafe_allow_html=True
-)
-
-
-header = st.container()
-dataset = st.empty()
-plot1, plot2 = st.columns([6,6])
-plot3, plot4 = st.columns([6,6])
-results = st.container()
-
-
-
-st.sidebar.title("Upload data")
-
-uploaded_file = st.sidebar.file_uploader('Choose a file')
-
-if uploaded_file is not None:    
-    df1=pd.read_csv(uploaded_file)
-    
-    
-    df1['frame'] = df1['filename'].str.extract('(\d{6})')
-    df1['x_c'] = (df1['x_min'] + df1['x_max']) / 2
-    df1['y_c'] = (df1['y_min'] + df1['y_max']) / 2
-    df1['frame'] = df1['frame'].astype('int')
-
-    frames = list(set(df1['frame'].tolist()))
-    frames = sorted([int(i) for i in frames])
-    
-    with dataset.container():
-        st.header("Uploaded data")
-        fig0, ax0 = plt.subplots()
-        ax0.scatter(x = df1['x_c'], y = df1['y_c'], s = 15)
-
-        st.pyplot(fig0)
-    
[…]

### 2023-01-18 text/ManuscriptDraft.Rmd
Removed old figures. Updating simulated data figure

@@
   pdf_document:
+    citation_package: natbib
     fig_caption: yes
@@
 
-Hjalte M. R. Mann^1,2^, Alexandros Iosifidis^2^, Toke T. Høye^1^
+Hjalte M. R. Mann^1^, Alexandros Iosifidis^2^, Toke T. Høye^1^
 
@@
 1: Department of Ecoscience and Arctic Research Centre, Aarhus University, C.F. Møllers Allé 8, 8000 Aarhus C, Denmark\
-2: Department of Electrical and Computer Engineering -- Signal Processing and Machine Learning, Aarhus University, Finlandsgade 22, 8200 Aarhus N, Denmark
+2: Department of Electrical and Computer Engineering, Aarhus University, Finlandsgade 22, 8200 Aarhus N, Denmark
 
@@
 
+#
+
+Word count: 
+Article type: Research article
+
 \newpage
@@
 Here, we show that image-based monitoring of field plots at very high temporal resolution can return information on flowering phenology at the level of individual flowers.
-Further, we present a framework for automatically tracking flowers in time-lapse image series, evaluating tracking accuracy, and identifying unreliable tracks.
+Further, we present a framework for automatic tracking of flowers in time-lapse image series, evaluating tracking accuracy, and identifying unreliable tracks.
 
  3.
-Our results on individual flower phenology suggest that traditional weekly observations of flower plots are insufficient to capture details in flowering dynamics.
+Our results on individual floral phenology suggest that traditional weekly observations of flower plots are insufficient to capture details in flowering dynamics.
 We show that the results of our automated method are reliable by comparing with manual tracking of individual floral phenology.
 
- 4. Our method could facilitate fine-grained studies on flower phenology and could be an important tool for estimating the sensitivity of plant-pollinator interactions to climate change.
+ 4. Our method can facilitate fine-grained studies on floral phenology and could be an important tool for estimating the sensitivity of plant-pollinator interactions to climate change.
 
@@
 The Arctic is characterized by short growing seasons and for the individual plant, timing of flowering is of critical importance.
-Flowering too early increases the risk of frost damage to the flowers, while flowering too late limits the time for seed maturation and dispersal before the end of the growing season [@pardee2019; @ELZINGA2007; @inouye2008; @iler2021].
+Flowering too early increases the risk of frost damage to flowers, while flowering too late limits the time for seed maturation and dispersal before the end of the growing season [@pardee2019; @ELZINGA2007; @inouye2008; @iler2021].
 Further, for insect-pollinated plants, flowering must be synchronous with pollinator activity for successful reproduction.
@@
 ![Figure 1: Tracking individual flowers through time.
-A: A time-lapse sequence of five frames showing presence of five individual flowers.
-Flower a (orange) is present in all frames, while the other flowers are only present part of the sequence and/or have gaps in their presence (e.g.
-because of occlusion or because they have temporarily disappeared from frame of view).
+A: A time-lapse sequence of five frames showing presence of five individual flowers a (orange), b (purple), c (blue), d (black), and e (green).
+Flower a is present in all frames, while the other flowers are only present in part of the sequence and/or have gaps in their presence e.g.
+because of occlusion or because they have temporarily disappeared from frame of view.
 Correct tracking of individual flowers results in information about the distribution of each individual flower in space (B) and time (C).](../figures/flowertrackingOverview.png){width="30%"}
 
-Most multiple object tracking literature focus on tracking of humans [@luo2021; @milan2016], and these methods not necessarily directly transferable to other use-cases.
+Most multiple object tracking literature focus on tracking of humans [@luo2021; @milan2016], and these methods are not necessarily directly transferable to other use cases.
 Many tracking algorithms, such as the Simple online and real-time tracking (SORT) [@sort], rely on Kalmann filtering to predict the future location of an object by estimating its velocity from previous frames [@kalmann].
@@
 octopetala*.
-For complex scenes with many flowers in close vicinity to each other there is an increased risk of tracking errors regardless of the tracking method being used.
-For such cases, we suggest a conservative filtering approach based on the distance between flower tracks within the frame in order to identify tracks that are spatially isolated and likely to be tracked with certainty.
+For complex scenes with many flowers in close vicinity to each other, there is an increased risk of tracking errors regardless of the tracking method being used.
+For such cases, we suggest a conservative filtering approach based on the distance between flower tracks within the frame in order to identify tracks that are sufficiently spatially isolated and likely to be tracked with certainty.
 Overall, we show that information on floral biology can be automatically derived from image-based monitoring of flower plots.
@@
 Then, remaining data points (i.e.
-data points for the subsequent frames) were generated iteratively, each point constrained to be within the circle with radius rM and the center point of the previously generated point and within the flower circle.
-This approach allowed us to simulate specific scenarios on which to demonstrate our algorithm by controlling the number of flowers generated and the number of data points produced per flower, the boundaries of a flower, and the maximum distance between current and next point (i.e.
-degree of movement between frames).
+data points for the subsequent frames) were generated iteratively whith each new point being constrained to be within a circle with a center point of the position of the previously generated point and a radius rM, as well as limited to be within the boundaries of the flower.
+This approach allowed us to simulate specific scenarios on which to demonstrate our algorithm by controlling the number of flowers generated and the number of data points produced per flower, the boundaries of a flower, and the degree of movement between frames (i.e.
+the maximum distance between current and next point).
 
@@
 Each flower consist of 500 data points and the data sets spans over 500 frames.
-Finally, in simulation D, we produced four flowers consisting of a very high number of data points (10.000 per flower) but with only little movement between frames.
+Finally, in simulation D, we produced four flowers consisting of a very high number of data points (10,000 per flower) but with only little movement between frames.
 Here, we used the same center point (cF) for all four flowers to increase overlap.
@@
 rF = 200, rM = 100.
+Arrow shows the last point of flower a and the first point of flower b.
 B: 10 points per flower.
@@
 A running mean value of 100 resulted in 4 mismatches.
-For simulation D, the algorithm tracked the four objects perfectly despite substantial overlap.
-Basing tracking on every 10th, 20th, and 30th frame resulted in 2, 4, and 10 mismatches, respectively.
+For simulation D, the algorithm tracked the four objects perfectly despite substantial overlap, while basing tracking on every 10th, 20th, and 30th frame resulted in 2, 4, and 10 mismatches, respectively.
 

### 2023-01-18 figures/flowertrackingOverview.py

@@ -1,5 +1,6 @@
 from matplotlib.markers import MarkerStyle
 import matplotlib.pyplot as plt
+import matplotlib.patches as pat
 from matplotlib.lines import Line2D
 from matplotlib.transforms import Affine2D
 from matplotlib.markers import JoinStyle, CapStyle
@@ -131,7 +132,7 @@ ax1, ax2, ax3, ax4, ax5 = axes.flatten()
 # # # Frame 1
 frame1_x = [50, 235, 243]
 frame1_y = [150,125, 40]
-frame1_z = [1,5,2]
+frame1_z = [1,2,3]
 
 ax1.scatter(frame1_x, frame1_y, s = 1)
 ax1.set_xlim([0, 300])
@@ -143,7 +144,7 @@ ax1.axes.get_xaxis().set_visible(False)
 # # # Frame 2
 frame2_x = [65,30, 255]
 frame2_y = [165,34, 126]
-frame2_z = [1,3,5]
+frame2_z = [1,4,2]
 
 ax2.scatter(frame2_x, frame2_y, s = 1)
 #ax2.set_title('Frame 2', loc = "left")
@@ -157,7 +158,7 @@ ax2.axes.get_xaxis().set_visible(False)
 
 frame3_x = [30,45,229, 265]
 frame3_y = [145,54, 110, 46]
-frame3_z = [1,3,5,2]
+frame3_z = [1,4,2,3]
 
 ax3.scatter(frame3_x, frame3_y, s = 1)
 ax3.set_xlim([0, 300])
@@ -169,7 +170,7 @@ ax3.axes.get_xaxis().set_visible(False)
 # # # Frame 4
 frame4_x = [66,132, 17,275]
 frame4_y = [144,101,17, 30]
-frame4_z = [1,4,3,2]
+frame4_z = [1,5,4,3]
 
 ax4.scatter(frame4_x, frame4_y, s = 1)
 ax4.set_xlim([0, 300])
@@ -182,7 +183,7 @@ ax4.axes.get_xaxis().set_visible(False)
 # # # Frame 5
 frame5_x = [29,125, 219]
 frame5_y = [166,83, 122]
-frame5_z = [1,4, 5]
+frame5_z = [1,5,2]
 
 ax5.scatter(frame5_x, frame5_y, s = 1)
 ax5.set_xlim([0, 300])
@@ -212,8 +213,7 @@ plt.show()
 
 
 
-
-colors=["#FF4F00", "#0040A9", "#000000", "#008507", "#850080"]
[…]

### 2023-01-18 src/Simulation_plotting.py

@@ -31,7 +31,7 @@ ax2.scatter(maxgap["frame"], maxgap["objectLetter"], c = maxgap['object'])
 ax2.xaxis.set_ticks(np.arange(0, 10, 1))
 ax2.set_box_aspect(1/4)
 ax2.invert_yaxis()
-plt.savefig(r"U:\BITCue\Projekter\TrackingFlowers\data\simulation/maxgapPlot.pdf", bbox_inches='tight')
+#plt.savefig(r"U:\BITCue\Projekter\TrackingFlowers\data\simulation/maxgapPlot.pdf", bbox_inches='tight')
 plt.show()
 
 
@@ -45,6 +45,7 @@ f, (ax1, ax2) = plt.subplots(2, 1, sharex=False, height_ratios=[4, 1])
 
 ax1.set_xlim([0, 1000])
 ax1.set_ylim([0, 1000])
+ax1.set_xlabel("hello")
 ax2.margins(y = 0.2)
 
 ax1.scatter(maxdist["x_c"], maxdist["y_c"], c = maxdist['object'])
@@ -62,8 +63,11 @@ arrowYend = maxdist['y_c'][(maxdist['frame'] == min(maxdist['frame'][maxdist['ob
 
 ax1.annotate("", xy = (arrowX,arrowY), xytext = (arrowXend, arrowYend), arrowprops=dict(arrowstyle="<->"))
 
-plt.savefig(r"U:\BITCue\Projekter\TrackingFlowers\data\simulation/maxdistPlot.pdf", bbox_inches='tight')
-# plt.show()
+
+
+
+#plt.savefig(r"U:\BITCue\Projekter\TrackingFlowers\data\simulation/maxdistPlot.pdf", bbox_inches='tight')
+plt.show()
 
 
 
@@ -86,8 +90,8 @@ ax2.xaxis.set_ticks(np.arange(0, 501, 100))
 ax2.set_box_aspect(1/4)
 ax2.invert_yaxis()
 
-plt.savefig(r"U:\BITCue\Projekter\TrackingFlowers\data\simulation/runmeanPlot.pdf", bbox_inches='tight')
-#plt.show()
+#plt.savefig(r"U:\BITCue\Projekter\TrackingFlowers\data\simulation/runmeanPlot.pdf", bbox_inches='tight')
+plt.show()
 
 ### FRAME RATE ###
 
[…]

### 2023-01-19 text/ManuscriptDraft.Rmd
Reduced text manuscript version initiated.

@@
 #title: |
-#  **Towards individual-based pollination ecology: Automatic tracking of life histories of individual flowers**
+#  **Towards individual-based floral biology: Automatic tracking of life histories of individual flowers**
 #date: "`r format(Sys.time(), '%B %d, %Y')`"
@@
 library(knitr)
+
 knit('ManuscriptDraft.Rmd')
@@
 ```
-
 ```{r echo=FALSE}
@@
 
-**Towards individual-based pollination ecology: Automatic tracking of life histories of individual flowers**
+**Towards individual-based floral biology: Automatic tracking of life histories of individual flowers**
 
@@
 Consequently, flowering phenology of a population is often characterized simply by the date for onset of flowering, typically derived from weekly observations in permanent plots [@Prevey2021].
-Such crude estimates may fail to reveal important dynamics in flowering phenology through the season.
+Such crude estimates may fail to reveal important seasonal dynamics in flowering phenology..
 For example, different phenological factors, such as onset, peak, and end of flowering may respond differently to climate change [@cook2012] and estimation of the degree of mismatch between species may thus be sensitive to the choice of metric [@inouye2019].
 The optimal metric for phenology depends on the research question, but often the choice of recording first flowering dates to represent flowering phenology is pragmatic rather than ideal.
-In fact, first flowering dates may be particularly ill suited for estimating sensitivity to climate change [@miller2008] since they represent the very extreme of the phenological distribution of a population and may be confounded with e.g.
-the size of the population considered.
+In fact, first flowering dates may be particularly ill suited for estimating sensitivity to climate change since they represent the very extreme of the phenological distribution of a population and may be confounded with e.g.
+the size of the population considered [@miller2008].
 An accurate estimate of the phenological distribution of a population over the full season would allow for, for example, accurate quantification of resource availability through the season and temporal mismatches between species [@ramakers2020].
 However, even detailed information on phenology of a population is still an aggregation of the phenology of the individual members of the population for which important variation may be masked [@inouye2019].
-With the full distribution of flowering phenology at the level of individual flowers, it is possible to extract relevant characteristics of the distribution and test their individual sensitivities towards environmental factors, but acquiring such data requires efficient and preferably automated methods.
+With individual level floral phenology it is possible to extract relevant characteristics of the distribution and test their individual sensitivities towards environmental factors, but acquiring such data requires efficient and automated methods.
 
 Cameras have become an ubiquitous tool for non-invasive and efficient monitoring of natural processes [@howe2017; @glover2019], but the implementation of image-based data collection has outpaced the development of methods for processing and analysing the big data output [@weinstein2018; @farley2018].
-Gradually, computer vision and deep learning tools are being developed and applied for automatic processing and extraction of relevant data from images [@willi2019; @tabak2019; @hoye2021].
+Gradually, computer vision and deep learning tools are being developed and applied for automatic extraction of relevant data from images [@willi2019; @tabak2019; @hoye2021].
 Indeed, automatic image-based monitoring of flowering phenology coupled with deep learning has facilitated collection of phenology data for specific species at very high temporal resolution [@mann2022].
-While automatic detection and counting of flowers in image-series, as demonstrated in @mann2022, can produce detailed information about floral biology, phenological responses at the level of individual flowers may be indiscernible regardless of the temporal resolution of the data at population level.
+While automated detection and counting of flowers in image-series, as demonstrated in @mann2022, can produce detailed information about floral biology, phenological responses at the level of individual flowers may be indiscernible regardless of the temporal resolution of the data at population level.
 For example, a shortening of individual flower longevity may not be directly obvious at the population level.
-In fact, many research questions can only be directly explored on the basis of phenology data at the level of individual flowers and such information is critical for investigating the association between reproductive success and timing of flowering and/or flower longevity.
+In fact, many research questions can only be directly explored on the basis of individual level floral phenology data and such information is critical for investigating the association between reproductive success and timing of flowering and/or flower longevity.
 Individual level data would also aid the understanding of whether flower visitation rates and/or seed set depends on the timing of flowering for the individual flower.
-Thus, automated recording of floral phenology by tracking the presence of individual flowers in image-series of permanent plots would be ideal and the feasibility of this as a second step subsequent of flower detection was explored in this study.
+Thus, automated recording of floral phenology by tracking the presence of individual flowers in image-series of permanent plots would be ideal and the feasibility of this as a step subsequent of flower detection was explored in this study.
 
@@
 
-Most multiple object tracking literature focus on tracking of humans [@luo2021; @milan2016], and these methods are not necessarily directly transferable to other use cases.
-Many tracking algorithms, such as the Simple online and real-time tracking (SORT) [@sort], rely on Kalmann filtering to predict the future location of an object by estimating its velocity from previous frames [@kalmann].
-However, Kalman filtering assumes linear motion of the object and does not handle abrupt motion well.
+Most multiple object tracking (MOT) methods focus on tracking of humans [@luo2021; @milan2016], and are not necessarily directly transferable to other use cases.
+Many tracking algorithms, such as the Simple online and real-time tracking (SORT) [@sort], rely on Kalmann filtering to predict the future location of an object based on its previous velocity [@kalmann].
+However, Kalman filtering assumes linear object motion and does not handle abrupt movement well.
 Flowers are constrained in movement by their stalk and violate this assumption.
@@
 Individual flowers, however, are unlikely to be clearly distinguishable and one flower may look less alike between two frames than two flowers in a single frame, which prevents this approach.
-Finally, some tracking methods require manual initiation of tracks for a fixed number of objects in the first frame and the algorithm then tracks these objects in the subsequent frames [@hu2012].
-These methods do not allow objects to enter or disappear from frame during tracking.
-In full-season monitoring data, flowers will disappear from the frame when they wilt, in which case their track should be terminated.
-Oppositely, new flowers should continuously be tracked as they bloom throughout the season.
-Additionally, a flower tracking method must handle flowers disappearing and appearing at the edges of the frame.
+Finally, some algorithms require manual initiation of a fixed number of tracks in the first frame and do not allow objects to enter or disappear during tracking [@hu2012].
+In full-season monitoring data, flowers should continuously be tracked as they bolt and wilt throughout the season Additionally, a flower tracking method must handle flowers disappearing and appearing at the edges of the frame.
 Clearly, individual flower tracking is a challenging task.
 
-Here, we present and evaluate a simple multiple object tracking algorithm for automatic tracking of individual flowers in image-series.
-Our tracking algorithm is based on centroid tracking and assigns individual objects to tracks based on the minimum distance between centroids of bounding boxes and does not rely on prediction of future positions.
-Further, the method allows both initiation and termination of tracks during tracking.
-We describe three user-defined parameters incorporated into the algorithm, demonstrate their effect on simulated data, and discuss why they may help optimise tracking of flowers in particular.
+Here, we present and evaluate a simple MOT algorithm for automatic tracking of individual flowers in image-series.
+Our tracking algorithm is based on centroid tracking and assigns objects to tracks based on the minimum distance between centroids of bounding boxes, and allows initiation and termination of tracks during tracking.
+We describe three user-defined parameters incorporated into the algorithm, demonstrate their effect on simulated data, and discuss their applicability in flower tracking.
+For complex scenes with many flowers in close vicinity to each other, there is an increased risk of tracking errors regardless of the tracking method being used.
+For such cases, we suggest a conservative filtering approach based on the distance between flower tracks within the frame in order to identify tracks that are sufficiently spatially isolated and likely to be tracked with certainty.
 We apply the algorithm on annotations of real flowers in time-lapse image-series from full-season monitoring of two widespread Arctic flowering plant species *Dryas integrifolia* and *D.
 octopetala*.
-For complex scenes with many flowers in close vicinity to each other, there is an increased risk of tracking errors regardless of the tracking method being used.
-For such cases, we suggest a conservative filtering approach based on the distance between flower tracks within the frame in order to identify tracks that are sufficiently spatially isolated and likely to be tracked with certainty.
-Overall, we show that information on floral biology can be automatically derived from image-based monitoring of flower plots.
+Overall, we show that information on floral biology can be automatically derived from image-based flower plot monitoring.
 
 # Material and methods
+
+## Common tracking errors and evaluating tracking performance
+
+In MOT, two fundamental errors can occur.
+First, a single trajectory may be split into several tracks, an error known as fragmentation.
+This can happen when an object is temporarily occluded or have disappeared from view and a new track is erroneously initiated when it reappears, or when identity shifts occur between tracks, for example when they switch positions between frames.
+Fragmentation is especially prone in settings with high densities of objects.
+Second, merging occurs when several trajectories are assigned the same track identity.
+This can occur when a track is not correctly terminated at the end of a trajectory but rather continues on the trajectory of another object.
+Here, importantly, the two individual trajectories can be tracked without internal errors (i.e.
+without mismatches).
+
+The optimal approach for quantification of tracking performance depends on the goal of the tracking.
+For example, in our case of tracking flowers, associating information obtained in the images to the individual flower, for example flower visits, would require the track to be as correct as possible (i.e., identity shifts should be minimized).
+On the other hand, to derive flowering length in practice just requires correct association between the first and the last image of a flower while intermediate points can be ignored.
+Here, track fragmentation can lead to underestimation of flowering duration (and overestimation of number of flowers) while track merging can lead to overestimation of flowering duration (and underestimation of number of flowers).
+In some scenarios, one may be interested in the number of individual flowers that existed in a plot, in which case the number of tracks obtained by automatic tracking should closely resemble the actual number of individuals in the series.
+
+Here, we apply the commonly used MOT accuracy score (MOTA) which quantifies tracking performance based on tracking mismatches derived by counting the number of identity shifts within each track [@bernardin2008].
+Only shifts in tracking identity are counted as mismatches in the MOTA score, while the number of points assigned to each track is not considered.
+
+$$
+MOTA = 1 -  \frac{\sum{_t(fn_t + fp_t + mm_t)}}{\sum{_tg_t}},
+$$
+
+where fn, fp, and mm are the number of false negatives, false positives, and track mismatches, respectively, for frame t divided by the total number of ground truth annotations across all frames.
+The MOTA score ranges from -Inf to 1, with 1 being perfect detection and tracking.
+In our case, we can ignore the variables false positives and false negatives, as we are tracking on manually annotated tracks that are considered the ground truth.
+
+To evaluate tracking accuracy of our algorithm, we calculate the total number of mismatches as well as the MOTA score (number of mismatches in relation to the total number of points).
+Additionally, we derive the number of tracks that were tracked with more than 80% accuracy.
+Finally, we compare the number of tracks identified by the automatic tracking with the true number of flowers in a series.
+These should ideally be equal.
+However, we note that a correct number of tracks (i.e., number of tracks equal to number of objects) does not necessarily translate to correct tracking, since single tracks can be erroneously fragmented simultaneously with merging of multiple tracks.
+We present the tracking performance without the use of the three tracking parameters along with the best performance (lowest number of mismatches/highest MOTA), and the best performance where the number of tracks correspond to the number of individual flowers in the series.
 
@@
 Then, remaining data points (i.e.
-data points for the subsequent frames) were generated iteratively whith each new point being constrained to be within a circle with a center point of the position of the previously generated point and a radius rM, as well as limited to be within the boundaries of the flower.
-This approach allowed us to simulate specific scenarios on which to demonstrate our algorithm by controlling the number of flowers generated and the number of data points produced per flower, the boundaries of a flower, and the degree of movement between frames (i.e.
+data points for the subsequent frames) were generated iteratively with each new point being constrained to be within a circle with a center point of the position of the previously generated point and a radius rM, as well as limited to be within the boundaries of the flower.
+This approach allowed us to simulate specific scenarios on which to demonstrate our algorithm by controlling the number of flowers generated, their positions and boundaries, the number of data points produced per flower, and the degree of movement between frames (i.e.
 the maximum distance between current and next point).
 
-We generated four simulated data sets (Fig.
-2) to demonstrate each of the three tracking parameters as well as the influence of frame rate on tracking reliability.
+We generated four data sets (Fig.
+2) to demonstrate each tracking parameter as well as the influence of frame rate on tracking reliability.
 In simulation A (Fig.
@@
 In simulation B, we produced four flowers that are spatially separated.
-Each flower consists of 5 data points, but the data set spans over 10 frames and each flower contains gaps of 1-3 data points at random temporal location.
+Each flower consists of 5 data points, but the data set spans over 10 frames and each flower contains random gaps of 1-3 data points.
 In simulation C, we produced four flowers that are in close proximity to each other and with a large degree of movement between frames.
-Each flower consist of 500 data points and the data sets spans over 500 frames.
-Finally, in simulation D, we produced four flowers consisting of a very high number of data points (10,000 per flower) but with only little movement between frames.
+Each flower consists of 500 data points and the data set spans over 500 frames.
+Finally, in simulation D, we produced four flowers with very high number of data points (10,000 per flower) but with only little movement between frames.
 Here, we used the same center point (cF) for all four flowers to increase overlap.
@@
 This means that between two frames, there is a substantial risk that a point for one object is closer to a point from another object rather than the point for that object, which will lead to identity shifts.
-To mitigate this, we calculate distances based on a current point and a set number of previous points in a track using the running mean parameter.
-This reduces the weighting on the displacement of points between the current and the previous frame.
-Finally, in simulation D, we have four objects that overlap in pairs of two, but the movement allowed between frames is very small, which means that a given point from an object is likely to be closest to the point from the same object in the previous frame.
+To mitigate this, we calculate displacements based on the current point and a set number of previous points in a track using the running mean parameter.
+This reduces the weighting on the displacement between the current and the previous point.
+Finally, in simulation D, we have four objects that overlap in pairs of two, but with little movement between frame.
+Therefore, a point from a given object is likely to be closest to the point from the same object in the previous frame.
 Effectively, while the overlap is substantial, the frame rate is sufficiently high that accurate tracking is ensured.
-To demonstrate the importance of frame rate for reliable tracking, we simulate a reduction in frame rate (i.e., increased degree of movement between frames) by basing tracking on every 10th, 20th, and 30th frame, respectively, and present these results.
+To demonstrate the importance of frame rate for reliable tracking, we simulate a reduction in frame rate (i.e., increased degree of movement between frames) by basing tracking on every 10th, 20th, and 30th frame, respectively.
 
-![Figure 2: Simulated flower detection data for demonstration of our tracking algorithm and the effect of the three tracking parameters as well as the significance of frame rate for reliable tracking.
-A: 10 points per flower.
-rF = 200, rM = 100.
+![Figure 2: Simulated flower detection data to demonstrate the tracking algorithm and the effect of the three tracking parameters, max distance, max gap, and running mean (refer to Automatic flower tracking section for definitions), as well as the significance of frame rate for reliable tracking.
+A: 10 points per flower, rF (radius of circle delimiting flower points) = 200, rM (radius of circles delimiting distance between consecutive points) = 100.
 Arrow shows the last point of flower a and the first point of flower b.
-B: 10 points per flower.
-rF = 200, rM = 100 C: 500 points per flower.
-rF = 200, rM = 200.
-D: 10,000 points per flower.
-rF = 100, rM = 1.](../figures/simulatedData.png){width="30%"}
-
-## Common tracking errors and evaluating tracking performance
-
-When tracking moving objects, two fundamental errors can occur.
-First, a single trajectory may be split into several tracks, an error known as fragmentation.
-This can happen when an object is temporarily occluded or have disappeared from view and a new track is erroneously initiated when it reappears, or when identity shifts occur between tracks, for example when they switch positions between frames.
-Fragmentation is especially prone in settings with high densities of objects.
-Second, merging occurs when several trajectories are assigned the same track identity.
-This can occur when a track is not correctly terminated at the end of a trajectory but rather continues on the trajectory of another object.
-Here, importantly, the two individual trajectories can be tracked without internal errors (i.e.
-without mismatches), but the error lies in the fact that the two trajectories have not been assigned unique tracking identities but instead are tracked as a single object.
-
-The optimal way of quantifying tracking performance depends on the goal of the tracking.
-For example, in our case of tracking flowers, associating information obtained in the images to the individual flower, for example flower visits, would require the track to be as correct as possible (i.e., identity shifts should be minimized).
-On the other hand, to derive flowering length in practice just requires correct association between the first and the last image of a flower while intermediate points can be ignored.
-Here, track fragmentation can lead to underestimation of flowering duration (and overestimation of number of flowers) while track merging can lead to overestimation of flowering duration (and underestimation of number of flowers).
-In some scenarios one may be interested in the number of individual flowers that existed in a plot, in which case the number of tracks obtained by automatic tracking should closely resemble the actual number of individuals in the series.
-
-Here, we apply the commonly used multiple object tracking accuracy (MOTA) score which quantifies tracking performance based on tracking mismatches derived by counting the number of identity shifts within each track [@bernardin2008].
-Mismatches occur when objects swap track identity because they are in close vicinity of each other or when an object periodically disappears and is assigned a new track identity when it reappears.
-Only the shifts in tracking identity are counted as mismatches in the MOTA score, while the number of points assigned to each track is not considered.
-For example, a ground truth track that was assigned one ID for the first half of its lifetime and another for the remaining half would only constitute one error.
-
-$$
-MOTA = 1 -  \frac{\sum{_t(fn_t + fp_t + mm_t)}}{\sum{_tg_t}},
-$$
-
-where fn, fp, and mm are the number of false negatives, false positives, and track mismatches, respectively, for frame t divided by the total number of ground truth annotations across all frames.
-The MOTA score ranges from -Inf to 1 with 1 being perfect detection and tracking.
-Tracking accuracy is commonly evaluated between manual annotations and detections produced, for example, by a convolutional neural network (CNN).
-Therefore, false positives and false negatives are included as errors in the MOTA score.
-In our case, we are not tracking on data produced by a CNN but rather on manually annotated tracks and do not have false positives or false negatives.
-Thus, these variables are ignored in the calculation of the MOTA score.
-
-To evaluate tracking accuracy of our algorithm, we calculate the total number of mismatches as well as the MOTA score (number of mismatches in relation to the total number of points).
-Additionally, we derive the number of tracks that were tracked with more than 80% accuracy.
-Finally, we compare the number of tracks identified by the automatic tracking with the true number of flowers in a series.
-These should ideally be equal.
-However, we note that a correct number of tracks (i.e., number of tracks equal to number of objects) does not necessarily translate to correct tracking, since individual tracks can be erroneously split (fragmentation) by the tracking algorithm while several tracks can be merged into one.
-We present the tracking performance without the use of the three tracking parameters along with the best performance (lowest number of mismatches/highest MOTA), and the best performance where the number of tracks correspond to the number of individual flowers in the series.
+B: 10 points per flower, rF = 200, rM = 100.
+C: 500 points per flower, rF = 200, rM = 200.
+D: 10,000 points per flower, rF = 100, rM = 1.](../figures/simulatedData.png){width="30%"}
 
@@
 The images were captured using time-lapse cameras (Moultrie Wingscapes TimelapseCam Pro, Moultrie Products, Birmingham Alabama, USA) permanently positioned approx.
-60 cm above areas with either *Dryas integrifolia* in Narsarsuaq and Thule or *D.
+60 cm above areas with either *D.
+integrifolia* in Narsarsuaq and Thule or *D.
 octopetala* in Ny-Ålesund.
@@
 
-We built a framework for tracking, filtering, and evaluating tracking of objects in time-lapse image series.
+We present a framework for tracking, filtering, and evaluating tracking of objects in time-lapse image series.
 The tracking algorithm (suppl.
@@
 Centroids are derived from bounding boxes around objects stemming from manual annotation or automatic detection of objects.
+Thus, the input data for the tracking algorithm are bounding box coordinates (top left corner and bottom right corner) for each object [@everingham2010].
 The algorithm has a set of user-adjusted parameters that can optimize tracking accuracy.
 The parameters are particularly relevant for tracking of objects that are constrained to a specific area such as flowers.
-However, the tracking algorithm could be used to track any objects.
-The speed of the tracking algorithm depends on the computational power available as well as the number of objects that are being tracked, but the time for computation will be negligible in most use-cases.
-Tracking of a series containing 85 objects ran at \~0.02 seconds per frame on a standard laptop.
-This, together with the fact that the algorithm handles frames sequentially, means that it can be applied both offline (on a set of detections/annotations that have already been produced) or online (real-time tracking frame per frame).
 
@@
 4A).
-However, if a flower reappears in the same area as a flower is already being tracked after disappearing in only a few frames, it is a reasonable assumption that it is the same individual and not that the old flower wilted/disappeared and a new one developed.
-The parameter **max gap** sets the number of frames a flower can be absent before reappearing without the track ending and a new initiating.
-If a given flower has not been detected in a number of frames exceeding the max gap size, the track will be terminated.
-Concurrently, this can deal with potential false negatives, as a track that is missing a bounding box will not be lost.
+However, if a flower reappears shortly after having disappeared from view, it is a reasonable assumption that it is the same individual.
+The parameter **max gap** sets the number of frames a flower can be absent before reappearing without triggering track termination and initiation of a new track.
+Concurrently, the parameter can prevent track termination owing to false negatives.
 
@@
 However, as flowers are constrained in their displacement, we can assume a maximum distance between points and force the initiation of tracks for points that exceed this threshold.
-We set this threshold with the parameter **max distance,** measured as the euclidean distance in number of pixels between objects in neighbouring frames.
+We set this threshold with the parameter **max distance,** measured as the euclidean distance in number of pixels between objects in neighboring frames.
 
-Third, reliable tracking based on distance requires that movement of individuals are captured at sufficient frequency, such that each tracks are not switched between frames.
-Stand-alone image-based monitoring of the relatively slow process of flower development through full growing seasons (several months) is done with low frame rate, limited by power consumption and storage capacity both on and off the camera.
+Third, reliable tracking based on distance requires that movement of individuals are captured at sufficient frequency to prevent identity shifts.
+Stand-alone image-based monitoring of the relatively slow process of flower development through full growing seasons (several months) is likely to be done with low frame rate, limited by power consumption and storage capacity both on and off the camera.
 Here, we used several images per day and while this temporal resolution captures the phenology of the flowers at very high detail, they fail to capture the full movement of flowers.
@@
 intersection over union) as there often is no overlap.
-With a high number of flowers in close vicinity to each other, there's a substantial risk of identity switching between flowers as they move among each other (Fig.
+With a high number of flowers in close vicinity to each other, there is a substantial risk of identity shifts between flowers (Fig.
 1C;4C).
 Additionally, as the flowers are constrained by their stalk, there is a limit to the distance they can move, which prevents prediction of future location of an individual based on its previous trajectory.
-Here, we introduce the tracking parameter **running mean**, with which we can base tracking on the distance between a point in the current frame and the running mean of the positions of a number of previous points in a track, taking advantage of the fact that a flower moves around a central point.
+Here, we use the tracking parameter **running mean**, with which we can base tracking on the distance between a point in the current frame and the running mean of the positions of a number of previous points in a track, taking advantage of the fact that a flower moves around a central point.
 Increasing the value for this parameter decreases the weighting of movement between individual frames.
@@
 If no threshold for maximum distance between an object and a track is set, the reappearing flower will be linked to the track of the disappearing flower based on shortest distance despite the absolute distance being large.
-C: A flower may be linked to and incorrect track despite shortest distance to the running mean of the correct track.
+C: A flower may be linked to an incorrect track despite shortest distance to the running mean of the correct track.
 As flowers move around a central point, it is relevant to base tracking on a running mean of previous positions.
@@
 Crosses show running mean of positions in x \> 1 number of previous frames.
-Dots and crosses coloured according to track identity.](../figures/errorFigure.png){width="100%"}
+Flowers, dots and crosses coloured according to identity.](../figures/errorFigure.png){width="100%"}
 
-#### Input data for the tracking algorithm
-
-The input data for the tracking algorithm are bounding box coordinates for each object in each frame given in a commonly used format [@everingham2010]: frame number, x-min, y-min, x-max, y-max.
-Here, x-min, y-min is the top left corner of the bounding box and x-max, y-max is the bottom right corner.
-The bounding box centroid is calculated on the basis of these coordinates within the algorithm.
-Calculating tracking accuracy requires a ground truth ID for each bounding box which can be given as the sixth column.
-As this data returned from both manual object annotations and object detection methods using automatic object detection methods based on deep learning, the method can be applied on both manually and automatically derived data.
-
-#### Identifying optimal user parameters
+#### Testing the use of the tracking parameters
 
@@
 Note that max distance set to zero ignores the parameter altogether.
-We identified the setting(s) that returned the most accurate tracks for each series and compare the tracking results between optimal settings and all parameters ignored (max gap = 0, running mean = 1, max distance = 0).
+We present the tracking accuracies for three different scenarios: Parameters zeroed (PZ) (Max gap = 0; Max distance = 0; Running mean = 1) (hereafter referred to as naive tracking), minimum number of mismatches achieved, and minimum number of mismatches achieved with a number of tracks equal to the number of flowers in the series.
 
@@
 
-While the implemented tracking algorithm produced highly accurate flower tracks, it did not do so without errors.
-Many flowers could be tracked with certainty or near-certainty, but some settings proved challenging, especially series with many flowers or flowers in close proximity to each other.
-In order to be able to identify the most reliably tracked flowers in a series without the need for manual annotation, we use a conservative filtering method that extracts the most trustworthy tracks from a scene based on spatial distance between tracks.
-We derive the distance between tracks based on the centroids of points included in a track produced by the tracking algorithm.
+With the substantial variation in number of flowers per series as well as distances between flowers, it is likely that some can tracked with certainty or near-certainty while others, especially flowers in close proximity to each other, will prove challenging.
+To facilitate identification of the most reliably tracked flowers in a series without the need for manual annotation, we use a conservative filtering method that extracts the most trustworthy tracks from a scene based on spatial distance between tracks.
+We derive the distance between tracks based on the centroids of points included in a track.
 When tracks consist of only a single point, the coordinates of the point is used as the centroid.
@@
 The filtering is done on the centroids to remove tracks in close vicinity to each other as these are assumed to have a high risk of tracking mismatches.
-The distances between all track centroids are calculated (euclidean distance in number of pixels) and only tracks that have no neighbouring centroid within the filtering distance threshold are kept.
-
+The distances between all track centroids are calculated (euclidean distance in number of pixels) and only tracks that have no neighboring track centroid within the filtering distance threshold are kept.
 We demonstrate the filtering method on the three series for which our tracking algorithm did not return perfect results (i.e., automatic tracking identical to ground truth tracks).
 To test the filtering approach, we chose a fixed combination of the tracking algorithm parameters across all three series (max gap size: 10; running mean: 10; max distance: 300).
-We then applied the filtering method on each three series with a range of values for the filtering distance (50-500, step size of 50) and identified the value that returned the highest sum of tracks across the three series at the lowest number of mismatches.
+We then applied the filtering method on each three series with a range of values for the filtering distance (50-500, step size of 50) and identified the value that returned the highest sum of tracks across the three series at zero mismatches.
 We present the tracking accuracy before filtering and on the remaining tracks after filtering.
-We emphasize that our results are conservative and that the approach may be optimised to extract a higher number of tracks while maintaining the number of mismatches.
-For example, basing the filtering on the best performing tracking parameters for each series specifically would improve the output of the tracking algorithm.
-Similarly, the results could be improved by fine-tuning the filtering distance threshold for each series individually.
-However, as our goal here was to demonstrate that our tracking and filtering method can be applied in a naive setting with fixed values for each parameter, we have not included those results.
-
-##
+We emphasize that our results are conservative and that the approach may be optimised to extract a higher number of tracks while maintaining the number of mismatches, for example by basing the filtering on the optimal tracking results for each series or by fine-tuning the filtering distance threshold for each series.
+As our goal here was to demonstrate that our tracking and filtering method can be applied in a naive setting with fixed values for each parameter, we have not included those results.
 
@@
 
-Tracking for simulation A without applying the three tracking parameters resulted in a merging error (i.e.
+Naive tracking for simulation A resulted in a merging error (i.e.
 the two objects were tracked as one).
-Setting the max distance to 200 removed this error and resulted in perfect tracking of the two objects.
-Tracking for simulation B without applying the tracking parameters returned 8 tracks and 9 mismatches (MOTA = 0.55).
-Setting max gap to 4 decreased the number of mismatches to 4 (4 tracks, MOTA = 0.8) and including a value of 200 for max distance as for simulation A removes the remaining errors and resulted in perfect tracking.
-For simulation C, tracking without parameters ignored returned the correct number of tracks (4), but 38 mismatches were made (MOTA = 0.98).
+A max distance value of 200 removed this error and resulted in perfect tracking of the two objects.
+Naive tracking for simulation B returned eight tracks and nine mismatches (MOTA = 0.55).
+Setting max gap to 4 decreased the number of mismatches to four (4 tracks, MOTA = 0.8) and including a value of 200 for max distance as for simulation A removed the remaining errors and resulted in perfect tracking.
+For simulation C, naive tracking returned the correct number of tracks (4), but 38 mismatches were made (MOTA = 0.98).
 Setting the running mean parameter to 20, 40, 60, resulted in 16, 8, and 4 mismatches, respectively, while a running mean value of 80 resulted in perfect tracking.
 A running mean value of 100 resulted in 4 mismatches.
-For simulation D, the algorithm tracked the four objects perfectly despite substantial overlap, while basing tracking on every 10th, 20th, and 30th frame resulted in 2, 4, and 10 mismatches, respectively.
+For simulation D, the algorithm tracked the four objects perfectly despite substantial overlap, while tracking on every 10th, 20th, and 30th frame resulted in 2, 4, and 10 mismatches, respectively.
 
@@
 5).
-Further, the results show that the flowers in the series from Ny-Ålesund seem to have longer flowering durations than the other locations.
-Further, our tracking algorithm was able to track individual flowers with consistently high accuracy (MOTA $\geq$ 0.981) across all series.
-In all cases, all flowers were tracked with more than 80% accuracy (Table 2).
-In fact, two series were tracked without any errors (2018 THUL B and 2018 NARS D).
-Tracking with the three tracking parameters ignored (max gap = 0, max distance = 0, running mean = 1) returned high MOTA scores across series, but for the three series that were not tracked perfectly, utilizing the tracking parameters increased the MOTA score in all cases.
+Our tracking algorithm was able to track individual flowers with consistently high accuracy (MOTA $\geq$ 0.981) across all series.
+In all cases, all flowers were tracked with more than 80% accuracy (Table 2) and two series were tracked without errors (2018 THUL B and 2018 NARS D).
+Naive tracking returned high MOTA scores across series, but for all three series that were not tracked perfectly, utilizing the tracking parameters increased the MOTA score.
 
@@
 85).
-One combination of parameters did return a number of tracks equal to number of flowers (max gap = 10, max distance 400, running mean = 10), but also increased the number of mismatches to 91.
+One paramereter combination did return a number of tracks equal to number of flowers (max gap = 10, max distance 400, running mean = 10), but also increased the number of mismatches to 91.
 For 2019 NARS D, several combinations of parameters returned a perfect MOTA score, but still some combinations returned a number of tracks lower than the number of flowers.
-This demonstrates the significance of a special case of tracking error, namely merging of tracks, which is not accounted for in the MOTA score.
-In figure 6, we exemplify the two types of errors, identity shifts and merges.
-
-#
+This demonstrates the significance of merging errors, which are not accounted for in the MOTA score.
+In figure 6, we exemplify the two types of errors, identity shifts and merges, in the flower data.
 
@@
 ![Table 2: Tracking performance for the automatic flower tracking algorithm for different tracking parameter combinations.
-Number of tracks in bold shows when the number of tracks equals the number of flowers in the series.
-Tracking performance is shown for parameters zeroed (PZ) (Max gap = 0; Max distance = 0; Running mean = 1), minimum number of mismatches achieved (MM), and minimum number of mismatches achieved with a number of tracks equal to the number of flowers in the seris (CF MM).
-The last type is only given if it was not equal to the results for MM.
-When several parameter combinations returned the same results, the range of parameter values included is given for each parameter.](../figures/tab_1.png){width="100%"}
+Number of tracks in bold signifies a number of tracks equal to the number of flowers in the series.
+Tracking performance is shown for three different tracking settings: Parameters zeroed (PZ) (Max gap = 0; Max distance = 0; Running mean = 1), minimum number of mismatches achieved (MM), and minimum number of mismatches achieved with a number of tracks equal to the number of flowers in the seris (CF MM).
+The last type is not given if the results were equal too those for MM.
+When several parameter combinations returned the same results, the range of parameter values included is given.](../figures/tab_1.png){width="100%"}
 
@@
 Each track is coloured by the IDs assigned by the flower tracking algorithm.
-In the top panel (max gap = 10, running mean = 1, max distance = 0), flower 15 and 28 have no identity shifts within each track, but the two tracks have been merged into one (see B).
-Flower 25 has been fragmented into two separate tracks causing one mismatch (see C).
-In the bottom panel, the running mean parameter has been increased to 20 and max distance to 500, which results in correct tracking of all three flowers.
+The top panel (max gap = 10, running mean = 1, max distance = 0) shows a merging error of flower 15 and 28 (see B), while flower 25 has been fragmented into two separate tracks causing one mismatch (see C).
+The bottom panel show correct tracking output returned when using the parameters max gap = 10, running mean = 20, max distance = 50.
 B: Merging error between two tracks.
-Frame f shows the last frame with flower 15 (blue) annotated because it has wilted.
-Two frames (5 hours) later (f+2), a new flower (flower 28; blue) blooms.
-Since max gap \> 2 and no distance threshold has been set, flower 28 is associated with the tracks of flower 15, because all other flowers are closer to their respective tracks, which causes a merging error.
+Frame f shows the last frame containing flower 15 (blue).
+Two frames (5 hours) later (f+2), flower 28 (blue) has  bolted.
+Since max gap \> 2 and no distance threshold has been set, flower 28 is associated with the track of flower 15.
 Setting a max distance threshold to 500 fixed this error as illustrated in A.
-C: Identity shift error between two flowers.
-A gust of wind causes a displacement of the flowers between two frames, f and f+1, (7 hours apart), which causes an identity shift between flower 25 and flower 29.
+C: Displacement of flowers 25 and 29 between two frames, f and f+1, (7 hours apart), causes an identity shift.
 Setting running mean to 20 fixed this error as illustrated in A.](../figures/errorExamples.png){width="100%"}
 
-With our filtering method, we were able to identify tracks with no mismatch errors in all three series that were not perfectly tracked.
-The parameter value for filtering distance that returned the highest number of tracks at the lowest number of mismatches was 350 (results for all values given in suppl.
+With our filtering method, we were able to identify 28 tracks with no mismatches from the three series (2018 NARS A: 15 tracks; 2019 NARS C: 10 tracks; 2019 NYAA E: 3 tracks) with complex scenes using a fixed filtering distance value of 350 and fixed values for the tracking parameters (suppl.
 2).
 The results of the filtering algorithm on the three series are given in Table 3 and the filtering process is visualised in figure 7.
-Our filtering method extracted 28 tracks from the three series (2018 NARS A: 15 tracks; 2019 NARS C: 10 tracks; 2019 NYAA E: 3 tracks) with complex scenes using a fixed filtering distance value of 350 and fixed values for the tracking parameters (10, 300, 10 for max gap size, max distance, and running mean, respectively) without any mismatches included in the 28 tracks.
 
-**Table 3:** Results of the filtering algorithm on tracking results using the parameters 10, 300, 10 for max gap size, max distance, and running mean, respectively.
+**Table 3:** Results of the filtering algorithm on tracking results using the parameters max gap = 10, max distance = 300, running mean =  10.
 The filtering distance value of 350 was used for all three series.
@@
 B: The track convex hull polygons calculated from the tracks.
-Polygons with a euclidean distance below 350 between them have been grouped together.
-The polygons have been coloured according to group.
-C: The results of the filtering where all tracks that were not sufficiently spatially isolated were removed, resulting in 10 remaining tracks without tracking errors.](images/filterFigLet.png)
+Polygons are coloured according to group and polygons with a euclidean distance below 350 between them have been grouped together.
+C: The results of the filtering where all tracks that were not sufficiently spatially isolated have been removed, resulting in 10 remaining tracks without tracking errors.](images/filterFigLet.png)
 
@@
 
-In this paper, we have demonstrated how image-based monitoring coupled with automated tracking can produce new insight into floral biology.
-We have presented a simple distance-based tracking algorithm and, simulated data, we have shown how tracking parameters can used to tune the tracking algorithm to deal with specific challenges and increase tracking performance.
-Further, we have shown that frame rate is an important consideration for reliable tracking and that sufficient frame rate can ensure accurate tracking of objects despite close proximity between them.
-We have tested our method on real flower data and demonstrated the feasibility of tracking individual flowers in images through time.
-As discussed below, challenges similar to the ones demonstrated through simulations existed in the real flower data and adjustment of the tracking parameters allowed for increased tracking accuracy of the real flowers.
+In this paper, we have demonstrated the potential of automated tracking to monitor floral phenology of individual flowers from bolting to withing and produce new insight into floral biology.
+We have presented a simple distance-based tracking algorithm and, with simulated data, we have shown how tracking parameters can be used to tune the tracking algorithm to deal with specific challenges and increase tracking performance.
+We have tested our method on real flower and demonstrated the feasibility of tracking individual flowers in images through time.
+Challenges similar to the ones demonstrated through simulations existed in the real flower data and adjustment of the tracking parameters allowed for increased tracking accuracy of the real flowers.
 
-The automatic tracking and filtering pipeline presented was able to return reliable flower tracks despite settings that violate assumptions of standard tracking algorithms and complex scenes with a high number of objects that are not permanently visible within the series.
-Our tracking algorithm was capable of accurately tracking individual flowers, despite this difficult setting.
+The automatic tracking and filtering pipeline presented was able to return reliable flower tracks despite complex scenes with a high number of objects with sporadic visibility and violation of standard tracking algorithm assumptions.
 Even with simple centroid tracking where association is based solely on displacement between the current and previous frame, tracking performances were high for all flower series and perfect for two.
-Further, with our filtering approach based on distance between tracks, we were able to discard the least reliable tracks and thereby increase the mean tracking accuracy of the remaining flowers to a perfect score.
-
-Our results show substantial variation in flowering duration between individual flowers.
-In fact, flower longevity was less than a few days in some cases.
-As flowers may not be visible within the frame for their full lifespan, it is likely that for some flowers the full flowering has not been captured, although flowering durations of less than 24 hours has been observed from manual examination of the image series (Mann, personal observation).
-Interestingly, the flowers in Ny-Ålesund seemed to have extended flowering durations compared to other series.
-As this location is the only one included that contained *D.
-octopetala* this could be explained by species-specific differences or variation between sites and future studies could explore this further.
-Owing to the limited resolution of the cameras used in the present study, it is unlikley that detailed information about anther opening or pollen release can be recorded from these images, but with increased image quality, this may be possible.
+Further, coupling this the user-defined tracking parameters and our filtering approach, we were able to optimize tracking performance and extract reliable tracks.
 
 The two series for which the tracking algorithm produced perfect results were also the series with the lowest number of flowers.
-While an increased number of objects does not necessarily increase tracking difficulty as this depends on the degree of isolation of tracks in time and space [@beard2018], multiple object tracking is a challenging task.
-Here, since the flowers are limited in space and time, the complexity of tracking is likely to increase with higher numbers of flowers.
-More flowers within the frame increases the possibility of flowers appearing in close proximity to each other and that flowers overlap or move in and out of frame.
-This translates to a higher risk of tracking errors.
-The very high complexity in the scenes with many flowers, especially 2019 NARS C with 85 flowers, coupled with the low frame rate renders it unlikely that any tracking method could produce perfect results across series.
-Although 2019 NARS C can be considered an extreme case for our application, it underlines the relevance of a pipeline including both tracking and filtering.
-But further, it points to the possibility of filtering image series based on an estimate of the number of flowers before tracking.
-For example, with many image series available, series with less than a given maximum of flowers detected in a single frame over a season could be selected for subsequent tracking.
-Tracks in series that exceeded this threshold could be filtered after tracking, as demonstrated here, or selected for manual validation of results.
+While an increased number of objects does not necessarily increase tracking difficulty as this depends on the degree of isolation of tracks in time and space [@beard2018], here, since the flowers are limited in space and time, the complexity of tracking is likely to increase with higher numbers of flowers as this increases the possibility of flowers overlapping or moving in and out of frame.
+The high complexity in the scenes with many flowers, especially 2019 NARS C with 85 flowers, coupled with the low frame rate renders it unlikely that any tracking method could produce perfect results across series.
+Although 2019 NARS C may be considered an extreme case for our application, it underlines the relevance of a pipeline including both tracking and filtering, but also points to the possibility of filtering image series based on an estimate of the number of flowers before automated tracking, which could be relevant in some cases.
 
-Our results show that optimizing for high MOTA tends to lead to an underestimate of the number of tracks (i.e., merging of tracks occurs) (Table 2).
-This may happen, for example, because of the max gap parameter.
-Importantly, this parameter allows for temporary occlusion of a flower within a track (which would decrease the number of mismatches) but may also cause merging of two flower tracks if they bloom in the same area separated by a number of frames less than the threshold value set.
-Thus, even if the number of tracks equals the number of flowers, it does not necessarily follow that the tracking is correct.
-For instance, the tracking algorithm could split one ground truth track in two, but merge two other ground truth tracks into one, as seen in figure 6.
-The task of evaluating multiple object tracking accuracy is a challenge in itself [@luiten2020; @milan2013; @milan2016] and in a future study, other tracking accuracy metrics could be implemented that account for merging errors in a better way.
-Tracking flowers may be a challenge that is especially prone to these errors since flowers can bloom in the same area at different times during the season.
+Importantly, optimizing for high MOTA can lead to an underestimate of the number of tracks (i.e., merging of tracks occurs) (Table 2).
+While accepting gaps in tracks allows for temporary occlusion of a flower within a track it may also lead to merging of separate tracks.
+Flower tracking may be especially prone to these errors since flowers can bolt in the same area at different times during the season.
+The task of evaluating multiple object tracking accuracy is a challenge in itself [@luiten2020; @milan2013; @milan2016] and other tracking accuracy metrics that account for merging errors in a better way should be further studied.
 
-The optimal settings for the three tracking parameters fell into a narrow range for the three complex series, except for the running mean for 2019 NYAA E.
-The optimal values for the running mean for 2019 NYAA E were exceptionally high, compared to the other series.
+The optimal settings for the three tracking parameters fell into a narrow range for the three complex series, except for the running mean for 2019 NYAA E, which were exceptionally high compared to the other series.
 Interestingly, the average flower duration was also higher for this series and this could be part of the explanation, as longer lifespan allow for larger gaps in the tracks.
+The three series with complex scenes all contain flowers at the edge of the frame that may occasionally be outside the view of the camera (see Fig.
+3) and the high number of flowers increase the risk of flowers temporarily occluding others.
+Correct tracking of these objects requires a max gap setting that optimizes the trade-off between fragmentation and merging errors.
 Overall, the max gap parameter improved tracking accuracy with values of 10-20 being optimal.
-The three series with complex scenes all contain flowers at the edge of the frame (see Fig.
-3), which means that they may occasionally be outside of view of the camera.
-Further, the higher number of flowers increase the risk of flowers temporarily occluding others.
-Correct tracking of these objects requires that an object can disappear from view without the track instantly being lost, which the max gap size allows.
-Oppositely, this setting may also lead to merging error.
+
 The optimal range for max distance was relatively narrow (400-600) for the parameters producing the correct number of tracks with the lowest number of mismatches.
-The optimal value for this setting is one that is sufficiently high that it does not force individual flowers into separate tracks because of movement between frames yet is low enough to separate any flowers into individual tracks that would otherwise have been grouped together.
+The optimal value for this setting is one that is sufficiently high that it does not force individual flowers into separate tracks because of movement between frames, yet is low enough to separate flowers into individual tracks that would otherwise have been grouped together.
 This explains why the extreme values for this parameter was not included in the highest accuracy tracks.
@@
 It is important to note that the tracking parameters are not independent.
-For example, the distance between a point in the current frame and a point in the previous frame may not be the same as that for a point in the current frame and the mean location of the previous ten points in a track.
-Thus, changing the setting for running mean could influence the optimal parameter setting for max distance.
+For example, changing the setting for running mean could influence the optimal value for max distance.
 This underlines the relevance of searching for an optimal parameter combination based on ground truth data.
-As with any tracking method, accuracy can only be assessed based on visual observation or, more rigorously, by testing on ground truth data [@dendorfer2021; @luiten2020; @milan2016].
+As with any tracking method, accuracy can only be assessed based on ground truth data [@dendorfer2021; @luiten2020; @milan2016].
 However, the fact that the optimal tracking parameters were overall similar across series strengthens the assumptions that a set of standard values can produce reasonable results for other similar image series with *Dryas* flowers.
-Here, a combination such as 10, 500, 10, for max gap, max distance, and running, respectively, would be a good starting point.
+Here, a combination such as 10, 500, 10, for max gap, max distance, and running mean, respectively, would be a good starting point.
 When applying our method in a new setting, we recommend performing a parameter test as demonstrated here on a subset of ground truth tracks to derive the optimal set of parameters.
@@
 We demonstrated our method on manual annotations of flowers but note that the format of the input data matches the output of deep learning object detection methods [@everingham2010].
-Thus, the method can be directly applied on such data.
-Automatically detecting flowers would likely introduce a proportion of false negatives when the detection algorithm fails to detect an object.
-The flowers in our data that periodically move out of the frame or are temporarily occluded mimics false negatives.
-Thus, it is likely that the max gap parameter can alleviate problems with false negatives.
+Thus, the method can be directly applied on such data, for instance as produced by the method described in @mann2022.
+Automatic detecting of flowers would likely introduce false negatives when the detection algorithm fails to detect an object, but here the max gap parameter may alleviate such problems.
 Automatic detection methods may also introduce false positives.
@@
 We presented a filtering method based on the distance between track centroids, which can be applied subsequent of tracking.
-The filtering algorithm successfully extracted 28 correct tracks from each of the three series for which the tracking algorithm did not produce perfect results without a single mismatch.
-Extracting tracks that are spatially isolated does not guarantee that the tracks are correct/without errors.
-For example, tracks of several objects may have been erroneously merged or a track could contain a small proportion of instances of an object far away, which would not have a large effect on the track centroid position.
-However, as spatially isolated objects in general are easier to track, it is safe to assume that these tracks are more reliable.
+The filtering algorithm successfully extracted 28 correct tracks from the three series for which the tracking algorithm did not produce perfect results.
+Extracting tracks that are spatially isolated does not guarantee that the tracks are without errors.
+For example, tracks of several objects may have been erroneously merged or a track could contain few instances of another object.
+However, it is reasonable to assume that spatially isolated objects are in general easier to track.
 We applied the filtering algorithm on tracks produced with a fixed set of tracking parameters as well as a fixed value for the filtering distance.
 As in the case of the tracking parameters, we recommend setting the value for the filtering distance based on either visual examination of results or on tests on a subset of ground truth data when applying the method in a new setting, but for tracking *Dryas* flowers in similar image series, a value of 350 may be a good starting point.
-We note that our method for filtering tracks using track centroids ensures that all tracks, despite the number of points in the track, are given the same weighting in the filtering since each track is represented by a single point.
-In some cases, this may be suboptimal.
-For example, if it is given a priori that an object will always appear in a minimum of two frames, then single point tracks can be filtered out.
-However, when such a priori knowledge is not available, a conservative approach, as the one we present, is preferable.
-
-An alternative approach to filtering based on distance between track centroids could be to identify and discard track polygons that overlap.
-However, this would remove tracks with even minor overlap and put weight on tracks with many points as these typically cover a larger area.
-Another alternative could be to base filtering on the centroids of the points within a track (i.e., mean x and mean y).
-The presented method avoids weighting track geometries based on number of included points and avoids weighting the centroid coordinate based on the point coordinates within the track as the track centroid is based on the convex hull of the track points.
+The presented method avoids weighting track geometries based on number of included points and also weighting the centroid coordinate based on the point coordinates within the track as the track centroid is based on the convex hull of the track points.
 Additionally, it allows for overlap between track geometries as long as the distance between the track centroids is above the threshold.
@@
 For processes that require high and continual temporal observation frequency to monitor, such as tracking individuals and their interactions, automation of data collection is critical, especially in remote places where human access is limited.
-Here, we have demonstrated the potential of image-based monitoring for collecting information on floral phenology at the level of individual flowers.
+Here, we have demonstrated the potential of image-based monitoring for collecting information on individual-level floral phenology.
 Coupling this with automated detection of flowers as demonstrated in @mann2022 permits an automated pipeline for collection of floral biology data.
-
-The possibility of collecting individual level floral phenology is highly relevant for exploration of many research questions.
-For example, it would be possible to assign indicators of seed set to the individual and explore how seed set is affected by timing and length of flowering and/or abiotic conditions such as frost events.
-Hereby, the long-standing question of what drives variation in reproductive success of individual flowers could be explored [@bolmgren2015].
-Further, the fact that it was not uncommon in our results that flowering lasted less than a week, emphasizes that weekly observations of sample plots can be insufficient for comprehensive description of floral phenology [@Prevey2021].
 
@@
 Here, methods from deep learning and computer vision are exceedingly relevant.
-For instance, it has been demonstrated that convolutional neural networks can facilitate monitoring of pollination events by detecting, identifying, and counting insects [@bjerge2021] and perform fine-grained species classification [@spiesman2021].
-Collectively, these developments coupled with implementation of long-term high-frequency image-based monitoring will allow for collection and processing of ecological data at unprecedented scales [@hoye2021].
+For instance, it has been demonstrated that CNNs can facilitate monitoring of pollination events by detecting, identifying, and counting insects [@bjerge2021] and perform fine-grained species classification [@spiesman2021].
+Collectively, these developments coupled with implementation of long-term high-frequency image-based monitoring will allow for collection and processing of ecological data at unprecedented scales [@hoye2021; @besson2022].
 By coupling pollinator detection and classification with floral tracking, visitation rates per flower could be derived and the effects on reproductive success could be investigated [@tiusanen2016; @rafferty2012].
+
+The possibility of collecting individual level floral phenology is relevant for many research questions.
+For example, coupling such information with indicators of seed set could facilitate investigation of the effect of timing and length of flowering and/or abiotic conditions such as frost events on seed set.
+Hereby, the long-standing question of what drives variation in reproductive success of individual flowers could be explored [@bolmgren2015].
+Further, the fact that it was not uncommon in our results that flowering lasted less than a week emphasizes that weekly observations of sample plots can be insufficient for comprehensive description of floral phenology [@Prevey2021].
 
@@
 
-Our results emphasize that this is not a trivial task.
+Our results emphasize that floral tracking is not a trivial task.
 The data violate common assumptions of tracking methods and tracking individual flowers through time-lapse images therefore represents a special case [@chen2022; @betke2007; @luo2021].
-For example, since flower movement is constrained by the stalk, the future location of a flower-head cannot be predicted based on previous trajectory of movement.
+Since flower movement is constrained, the future location of a flower-head cannot be predicted based on the previous movement trajectory.
 As individual flowers closely resemble each other, tracking cannot be based on recognition of individuals [@chen2022; @deepsort2017].
-Further, as the manual tracking of flowers showed, the lifespans of flowers vary substantially, which means that assumptions on flowering duration cannot be leveraged during tracking.
+Further, flowers have varied lifespans and may have sporadic visibility within the frame.
 These settings vastly complicate the task of tracking multiple objects [@betke2007].
-Our tracking algorithm returned consistently accuracy despite this.
-While we show that tracking can be done at high accuracy with our tracking algorithm, we also identify particular challenges with flower tracking, such as the violations of standard assumptions in tracking algorithms, the potentially large number of individuals overlapping in time and space, and the possible susceptibility to track merging.
-To improve tracking ability of flowers in future studies, focus should be on implementing reliable measures of tracking error and on developing methods that can alleviate the challenges that exist with tracking visually similar objects, as well as on comparison of different tracking approaches.
+Our tracking algorithm returned consistently high accuracy despite this, but were unable to reliably track flowers in some cases.
+To further improve tracking ability of flowers in future studies, focus should be on implementing reliable measures of tracking error and on developing methods that can alleviate the challenges that exist with tracking visually similar objects, as well as on comparison of different tracking approaches.
 Alongside this, methods for extracting auxiliary information from images, such as flower seed set, should be explored.
@@
 Most importantly, adoption of automated methods for data collection and analysis within ecology, and specifically within for example pollination ecology, will transform the extent and scale of studies and the questions that can be explored.
+
+Owing to the limited resolution of the cameras used in the present study, it is unlikley that detailed information about anther opening or pollen release can be recorded from these images, but with increased image quality, this may be possible.
 
@@
 The authors have no conflict of interest to declare.
-
 Thank you to Adrian Rosebrock (https://pyimagesearch.com/2018/07/23/simple-object-tracking-with-opencv/, accessed on November 29th 2022) for inspiration for the design of the centroid tracking algorithm.
@@
 
-The code that supports the results in this paper is openly available at <https://github.com/TECOLOGYxyz/FlowerTracking>. The raw data will be archived on <https://zenodo.org/>.
+The code and data that supports the results in this paper is openly available at <https://github.com/TECOLOGYxyz/FlowerTracking>.
 

### 2023-01-19 src/Simulation_plotting.py

@@ -19,19 +19,28 @@ maxgap = pd.read_csv(r"U:\BITCue\Projekter\TrackingFlowers\data\simulation/maxga
 maxgap['objectLetter'] = maxgap['object'].map(lambda x: tran[x])
 
 f, (ax1, ax2) = plt.subplots(2, 1, sharex=False, height_ratios=[4, 1])
+plt.subplots_adjust(hspace=0.25)
+
 
 ax1.set_xlim([0, 1000])
 ax1.set_ylim([0, 1000])
+ax1.xaxis.set_ticks([0, 1000])
+ax1.yaxis.set_ticks([0, 1000])
+ax1.set_xlabel("X", labelpad=-12)
+ax1.set_ylabel("Y", labelpad=-20)
 ax2.margins(y = 0.2)
 
 ax1.scatter(maxgap["x_c"], maxgap["y_c"], c=maxgap['object'])
 ax1.set_box_aspect(1)
-ax1.set_title('Max gap')
+ax1.set_title('Max gap', fontweight="bold")
+
 ax2.scatter(maxgap["frame"], maxgap["objectLetter"], c = maxgap['object'])
 ax2.xaxis.set_ticks(np.arange(0, 10, 1))
 ax2.set_box_aspect(1/4)
 ax2.invert_yaxis()
-#plt.savefig(r"U:\BITCue\Projekter\TrackingFlowers\data\simulation/maxgapPlot.pdf", bbox_inches='tight')
+ax2.set_xlabel("Frame number", labelpad=0)
+ax2.set_ylabel("Object ID", labelpad=0)
+plt.savefig(r"U:\BITCue\Projekter\TrackingFlowers\data\simulation/maxgapPlot.pdf", bbox_inches='tight')
 plt.show()
 
 
@@ -42,19 +51,25 @@ maxdist = pd.read_csv(r"U:\BITCue\Projekter\TrackingFlowers\data\simulation/maxd
 maxdist['objectLetter'] = maxdist['object'].map(lambda x: tran[x])
 print(maxdist)
 f, (ax1, ax2) = plt.subplots(2, 1, sharex=False, height_ratios=[4, 1])
+plt.subplots_adjust(hspace=0.25)
 
 ax1.set_xlim([0, 1000])
 ax1.set_ylim([0, 1000])
-ax1.set_xlabel("hello")
+ax1.xaxis.set_ticks([0, 1000])
+ax1.yaxis.set_ticks([0, 1000])
+ax1.set_xlabel("X", labelpad=-12)
[…]

### 2023-01-19 text/ManuscriptDraft.Rmd
Reduced text manuscript version initiated.

@@
 library(knitr)
-
 knit('ManuscriptDraft.Rmd')

### 2023-01-20 src/R/tableGenerator.Rmd
Submitted revised edition of manuscript

---
title: "TableGenerator"
author: "Hjalte Mann"
date: '2022-05-02'
#output: pdf
pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = F)
library(ggplot2)
library(dplyr)
library(stringr)
library(GGally)
library(plotly)
library(kableExtra)
library(gt)
install.packages("webshot2")
library(webshot2)

```

```{r echo = F}

trackScores = read.csv2("U:/BITCue/Projekter/TrackingFlowers/testResults/MMFix/trackScores.csv", header = F)

colnames(trackScores) = c('Site','Type', 'Flowers','Points', 'Max gap', 'Max distance', 'Running mean', 'Mismatches', 'MOTA', '>80% Accuracy','Tracks')

trackScores$MOTA = signif(as.numeric(trackScores$MOTA), digits = 3)

trackScores %>% 
  gt() %>%
  tab_spanner(
    label = "Tracking parameters",
    columns = c('Max gap', 'Max distance', 'Running mean'
    )
  )%>%
  tab_spanner(
    label = "Result",
    columns = c('Mismatches', 'MOTA', '>80% Accuracy', 'Tracks'
    )
  ) %>% 
   cols_align(
    align = "left"
  ) %>%
  tab_row_group(
    label = "2018 THUL B",
    rows = Site == "2018 THUL B"
  )  %>%
  tab_row_group(
    label = "2019 NARS D",
    rows = Site == "2019 NARS D"
  ) %>%
  tab_row_group(
    label = "2018 NARS A",
    rows = Site == "2018 NARS A"
  ) %>%
  tab_row_group(
    label = "2019 NYAA E",
    rows = Site == "2019 NYAA E"
  ) %>%
  tab_row_group(
    label = "2019 NARS C",
    rows = Site == "2019 NARS C"
  ) %>% tab_style(
    style = list(
      cell_fill(color = "#FFFFFF"),
      cell_text(color = "#FFFFFF")
      ),
    locations = cells_body(
      columns = Site
    )
  ) %>% 
  row_group_order(groups = c("2018 THUL B","2019 NARS D","2018 NARS A","2019 NYAA E", "2019 NARS C" )) %>%
  fmt_missing(
    columns = 1:11,
    missing_text = "-"
  ) %>%
  tab_style(
    style = list(
      cell_text(weight = "bold")
      ),
    locations = cells_body(
      columns = Tracks,
      rows = c(1,2,4,6,9,11,12,15)
    )
  )%>% 
  text_transform(locations = cells_body(columns = Tracks), 
                 fn = function(x) {
                   x <- gsub("7 - 13", "7 - <span style=\"font-weight:bold;\">13</span>", x)
                 }) %>% 
  gtsave("tab_1_2.png", expand = 10)

```

```{r}

filterScores = read.csv2("U:/BITCue/Projekter/TrackingFlowers/testResults/MMFix/FilterScoresForTable.csv", header = T)

colnames(filterScores) = c('Image series','Flowers', 'Tracks','Mismatches', 'MOTA', 'Tracks ', 'Mismatches ', 'MOTA ')

filterScores$MOTA = signif(as.numeric(filterScores$MOTA), digits = 3)

filterScores %>% 
  gt() %>%
  tab_spanner(
    label = "Before filtering",
    columns = c('Tracks','Mismatches', 'MOTA'
    )
  )%>%
  tab_spanner(
    label = "After filtering",
    columns = c('Tracks ','Mismatches ', 'MOTA '
    )
  ) %>% 
   cols_align(
    align = "left"
  ) %>% 
  fmt_missing(
  columns = everything(),
  rows = everything(),
  missing_text = ""
) %>% tab_style(
    style = cell_borders(
      sides = c("top", "bottom"),
      color = "lightgrey",
      weight = px(2),
      style = "solid"
    ),
    locations = cells_body(
      columns = everything(),
      rows = 4
    )
  ) %>%
  tab_style(
    style = list(
      cell_text(weight = "bold")
      ),
    locations = cells_body(
      columns = everything(),
      rows = 4
    )
  ) %>% 
  gtsave("tableFilterscores.png", expand = 10)
```

```{r}

seriesOverview = read.csv2("U:/BITCue/Projekter/TrackingFlowers/data/annotations/seriesOverview.csv", header = T)
colnames(seriesOverview) = c('Image series','Number of images', 'Number of flowers')
seriesOverview

seriesOverview  %>% 
  gt() %>%
  cols_align(
    align = "left") %>% 
    gtsave("U:/BITCue/Projekter/TrackingFlowers/src/R/seriesOverview.png", expand = 10)

```

### 2023-01-20 text/ManuscriptDraft.Rmd
Submitted revised edition of manuscript

--- 2023-01-19 text/ManuscriptDraft.Rmd
@@
 ```
+
 ```{r echo=FALSE}
@@
 
-Keywords: Arctic, *Dryas octopetala*, *Dryas integrifolia*, ecological monitoring, flowering phenology, life-history variation, tracking
+Keywords: Arctic, *Dryas octopetala*, *Dryas integrifolia*, ecological monitoring, floral phenology, life-history variation, tracking
 
@@
 
-Word count: 
 Article type: Research article
+
+Word count: 8674
 
@@
 1.
-A fundamental challenge in pollination ecology is to understand how successful pollination is affected by the timing of flowering, the timing and frequency of pollinator visits, and the abundance and diversity of visitors.
+A fundamental challenge in floral biology is to understand how reproductive success is affected by the timing of flowering, the timing and frequency of pollinator visits, and the abundance and diversity of visitors.
 Since plant-pollinator interactions concern individual flowers, deciphering the consequences of e.g.
@@
 Here, we show that image-based monitoring of field plots at very high temporal resolution can return information on flowering phenology at the level of individual flowers.
-Further, we present a framework for automatic tracking of flowers in time-lapse image series, evaluating tracking accuracy, and identifying unreliable tracks.
+Further, we present a framework for automatic tracking of flowers in time-lapse image series, evaluating tracking accuracy, and identifying reliable tracks.
 
@@
 The Arctic is characterized by short growing seasons and for the individual plant, timing of flowering is of critical importance.
-Flowering too early increases the risk of frost damage to flowers, while flowering too late limits the time for seed maturation and dispersal before the end of the growing season [@pardee2019; @ELZINGA2007; @inouye2008; @iler2021].
+Flowering too early increases the risk of frost damage to flowers, while flowering too late limits the time for seed maturation and dispersal before the end of the growing season [@pardee2019; @inouye2008; @iler2021].
 Further, for insect-pollinated plants, flowering must be synchronous with pollinator activity for successful reproduction.
@@
 Cameras have become an ubiquitous tool for non-invasive and efficient monitoring of natural processes [@howe2017; @glover2019], but the implementation of image-based data collection has outpaced the development of methods for processing and analysing the big data output [@weinstein2018; @farley2018].
-Gradually, computer vision and deep learning tools are being developed and applied for automatic extraction of relevant data from images [@willi2019; @tabak2019; @hoye2021].
-Indeed, automatic image-based monitoring of flowering phenology coupled with deep learning has facilitated collection of phenology data for specific species at very high temporal resolution [@mann2022].
-While automated detection and counting of flowers in image-series, as demonstrated in @mann2022, can produce detailed information about floral biology, phenological responses at the level of individual flowers may be indiscernible regardless of the temporal resolution of the data at population level.
+Gradually, computer vision and deep learning tools are being developed and applied for automatic extraction of relevant data from images [@willi2019; @hoye2021].
+Indeed, @mann2022 demonstrated that automatic image-based monitoring of flowering phenology coupled with deep learning can facilitate collection of phenology data for specific species at very high temporal resolution.
+While such automated detection of flowers can produce detailed information about floral biology, phenological responses at the level of individual flowers may be indiscernible regardless of the temporal resolution of the data at population level.
 For example, a shortening of individual flower longevity may not be directly obvious at the population level.
@@
 Individual level data would also aid the understanding of whether flower visitation rates and/or seed set depends on the timing of flowering for the individual flower.
-Thus, automated recording of floral phenology by tracking the presence of individual flowers in image-series of permanent plots would be ideal and the feasibility of this as a step subsequent of flower detection was explored in this study.
+Thus, automated recording of individual floral phenology by tracking the presence of individual flowers through time would be ideal and the feasibility of this as a step subsequent of flower detection was explored in this study.
 
@@
 
-Most multiple object tracking (MOT) methods focus on tracking of humans [@luo2021; @milan2016], and are not necessarily directly transferable to other use cases.
+Many multiple object tracking (MOT) methods focus on tracking of humans [@luo2021], and are not necessarily directly transferable to other use cases.
 Many tracking algorithms, such as the Simple online and real-time tracking (SORT) [@sort], rely on Kalmann filtering to predict the future location of an object based on its previous velocity [@kalmann].
@@
 Finally, some algorithms require manual initiation of a fixed number of tracks in the first frame and do not allow objects to enter or disappear during tracking [@hu2012].
-In full-season monitoring data, flowers should continuously be tracked as they bolt and wilt throughout the season Additionally, a flower tracking method must handle flowers disappearing and appearing at the edges of the frame.
+In full-season monitoring data, flowers should continuously be tracked as they bolt and wilt throughout the season (fig.
+1).
+Additionally, a flower tracking method must handle sporadic visibility of flowers because of occlusion or temporary movement outside of frame view.
 Clearly, individual flower tracking is a challenging task.
@@
 
-In MOT, two fundamental errors can occur.
+In MOT, tracking accuracy is quantified by comparing tracking output with a ground truth [@dendorfer2021; @luiten2020; @milan2013] and two fundamental errors can occur.
 First, a single trajectory may be split into several tracks, an error known as fragmentation.
@@
 C: 500 points per flower, rF = 200, rM = 200.
-D: 10,000 points per flower, rF = 100, rM = 1.](../figures/simulatedData.png){width="30%"}
+D: 10,000 points per flower, rF = 100, rM = 1.](../figures/simulatedData_v2.png)
 
@@
 #
-![Table 1: Number of images and number of individual flowers for each of the five time-lapse image series.](../figures/CameraOverview.png){width="40%"}
+
+![Table 1: Number of images and number of individual flowers in each of the five time-lapse image series.](../figures/seriesOverview.png)
 
@@
 Flower bounding boxes are colour-coded by flower ID.
-D: Centroids of all flower annotations in the series, coloured by flower ID, showing how flowers may overlap in space.](../figures/AnnotationExamples_v2_4.png){width="20%"}
+D: Centroids of all flower annotations in the series, coloured by flower ID, showing how flowers may overlap in space.](../figures/AnnotationExamples_v2_5.png)
 
@@
 The algorithm has a set of user-adjusted parameters that can optimize tracking accuracy.
-The parameters are particularly relevant for tracking of objects that are constrained to a specific area such as flowers.
+The parameters are particularly relevant for tracking of objects that are constrained to a specific area such as plants.
 
@@
 First, as flowers move, they may temporarily occlude the view of other flowers and flowers close to the edge of the frame may move in and out of view between images.
-This can cause a track to be lost and fragmentation to occur (Fig.
-4A).
+This can cause a track to be lost and fragmentation to occur.
 However, if a flower reappears shortly after having disappeared from view, it is a reasonable assumption that it is the same individual.
@@
 
-Second, a standard centroid tracking algorithm will associate a point with a track based on minimum distance only, disregarding the absolute distance between centroids, which may lead to tracking errors (Fig.
-4B).
+Second, a standard centroid tracking algorithm will associate a point with a track based on minimum distance only, disregarding the absolute distance between centroids, which may lead to tracking errors.
 However, as flowers are constrained in their displacement, we can assume a maximum distance between points and force the initiation of tracks for points that exceed this threshold.
@@
 Here, we used several images per day and while this temporal resolution captures the phenology of the flowers at very high detail, they fail to capture the full movement of flowers.
-For example, as the wind shifts, the flower heads change direction, which in the image series happens instantaneously (i.e., between two consecutive frames) (Fig.
-4C).
-The fact that the frame rate is insufficient to capture smooth movement means that bounding boxes around a single individual in two consecutive frames cannot be associated based on overlap (e.g.
-intersection over union) as there often is no overlap.
-With a high number of flowers in close vicinity to each other, there is a substantial risk of identity shifts between flowers (Fig.
-1C;4C).
+For example, as the wind shifts, the flower heads change direction, which in the image series happens instantaneously (i.e., between two consecutive frames).
+With a high number of flowers in close vicinity to each other, this leads to a substantial risk of identity shifts between flowers (Fig.
+2C).
 Additionally, as the flowers are constrained by their stalk, there is a limit to the distance they can move, which prevents prediction of future location of an individual based on its previous trajectory.
-Here, we use the tracking parameter **running mean**, with which we can base tracking on the distance between a point in the current frame and the running mean of the positions of a number of previous points in a track, taking advantage of the fact that a flower moves around a central point.
-Increasing the value for this parameter decreases the weighting of movement between individual frames.
+Here, we use the tracking parameter **running mean**, with which we can base tracking on the distance between a point in the current frame and the running mean of the positions of a number of previous points in a track, facilitating a reduction of the weighting on movement between individual frames.
 Setting the parameter to 1 means that tracking is based on the distance to the centroid in the previous frame only.
-
-#
-
-![Figure 4: Centroid tracking may produce erroneous associations when objects move or disappear between frames.
-A: A track may be lost when a flower temporarily disappears from view.
-If gaps in tracks are not allowed during tracking, a new track will erroneously be initiated when the flower reappears.
-B: A flower may be linked to an incorrect track despite large distance between them.
-Here, a flower disappears from view and another reappears between two frames.
-If no threshold for maximum distance between an object and a track is set, the reappearing flower will be linked to the track of the disappearing flower based on shortest distance despite the absolute distance being large.
-C: A flower may be linked to an incorrect track despite shortest distance to the running mean of the correct track.
-As flowers move around a central point, it is relevant to base tracking on a running mean of previous positions.
-Dots show position of flower in previous frame.
-Crosses show running mean of positions in x \> 1 number of previous frames.
-Flowers, dots and crosses coloured according to identity.](../figures/errorFigure.png){width="100%"}
 
@@
 ## Flowers
+
 Manual tracking of individual flowers derived from image-based plot monitoring enabled detailed information on floral phenology at the level of individuals with substantial variation in flowering duration between flowers (Fig.
-5).
+4).
 Our tracking algorithm was able to track individual flowers with consistently high accuracy (MOTA $\geq$ 0.981) across all series.
@@
 This demonstrates the significance of merging errors, which are not accounted for in the MOTA score.
-In figure 6, we exemplify the two types of errors, identity shifts and merges, in the flower data.
+In figure 5, we exemplify the two types of errors, identity shifts and merges, in the flower data.
 
-![Figure 5: Flowering lifespan (days) for each individual flower in each of the five image series plotted against the onset of flowering in days after 1 January (day of the year) based on manual tracking of individual flowers](../figures/OnsetVSLengthBoxWithPoints.png){width="100%"}
+![Figure 4: Flowering lifespan (days) for each individual flower in each of the five image series plotted against the onset of flowering in days after 1 January (day of the year) based on manual tracking of individual flowers](../figures/OnsetVSLengthBoxWithPoints.png){width="100%"}
 
@@
 
-![Figure 6: Tracking results of three flowers in the 2018 NARS A image series.
+![Figure 5: Tracking results of three flowers in the 2018 NARS A image series.
 A: Track output from the flower tracking algorithm for three individual flowers using two different combinations of tracking parameters.
-The y-axis shows the correct track IDs assigned by manual tracking.
+The y-axis shows the correct track IDs , that is the ground truth IDs assigned by manual tracking.
 Each track is coloured by the IDs assigned by the flower tracking algorithm.
+For correct algorithmic tracking, a whole groun truth track should have one unique color.
 The top panel (max gap = 10, running mean = 1, max distance = 0) shows a merging error of flower 15 and 28 (see B), while flower 25 has been fragmented into two separate tracks causing one mismatch (see C).
@@
 2).
-The results of the filtering algorithm on the three series are given in Table 3 and the filtering process is visualised in figure 7.
+The results of the filtering algorithm on the three series are given in Table 3 and the filtering process is visualised in figure 6.
 
@@
 
-![Figure 7: The track filtering process exemplified by the 2019 NARS C series.
+![Figure 6: The track filtering process exemplified by the 2019 NARS C series.
 A: The centroid points for all bounding boxes in the series, coloured by the track id from the centroid tracking algorithm.
@@
 
-In this paper, we have demonstrated the potential of automated tracking to monitor floral phenology of individual flowers from bolting to withing and produce new insight into floral biology.
-We have presented a simple distance-based tracking algorithm and, with simulated data, we have shown how tracking parameters can be used to tune the tracking algorithm to deal with specific challenges and increase tracking performance.
-We have tested our method on real flower and demonstrated the feasibility of tracking individual flowers in images through time.
-Challenges similar to the ones demonstrated through simulations existed in the real flower data and adjustment of the tracking parameters allowed for increased tracking accuracy of the real flowers.
-
-The automatic tracking and filtering pipeline presented was able to return reliable flower tracks despite complex scenes with a high number of objects with sporadic visibility and violation of standard tracking algorithm assumptions.
-Even with simple centroid tracking where association is based solely on displacement between the current and previous frame, tracking performances were high for all flower series and perfect for two.
-Further, coupling this the user-defined tracking parameters and our filtering approach, we were able to optimize tracking performance and extract reliable tracks.
+In this paper, we have demonstrated the potential of automated tracking to monitor floral phenology of individual flowers from bolting to withing to produce new insight into floral biology.
+We have presented a simple distance-based tracking algorithm and, with simulated data, we have shown how a set of tracking parameters can be used to tune the tracking algorithm to deal with specific challenges and increase tracking performance.
+We have tested our method on real flower data and demonstrated the feasibility of tracking individual flowers in images through time.
+Challenges similar to the ones demonstrated through simulations existed in the real flower data and while even simple centroid tracking resulted in high accuracy across all flower series and perfect accuracy for two, adjustment of the tracking parameters allowed for increased tracking accuracy of the flowers.
+Further, by coupling tracking with our suggested filtering approach, we were able to extract reliable tracks in complex scenes.
+Overall, the automatic tracking and filtering pipeline presented was able to return reliable flower tracks despite complex scenes with a high number of objects with sporadic visibility and violation of standard tracking algorithm assumptions.
 
@@
 Although 2019 NARS C may be considered an extreme case for our application, it underlines the relevance of a pipeline including both tracking and filtering, but also points to the possibility of filtering image series based on an estimate of the number of flowers before automated tracking, which could be relevant in some cases.
+The fact that our method produced perfect tracking results in the most simple settings (fewest flowers) could be taken into consideration when setting up future data collection schemes.
 
@@
 Flower tracking may be especially prone to these errors since flowers can bolt in the same area at different times during the season.
-The task of evaluating multiple object tracking accuracy is a challenge in itself [@luiten2020; @milan2013; @milan2016] and other tracking accuracy metrics that account for merging errors in a better way should be further studied.
+The task of evaluating MOT accuracy is a challenge in itself [@luiten2020; @milan2013] and other tracking accuracy metrics that account for merging errors in a better way should be further studied.
 
@@
 Interestingly, the average flower duration was also higher for this series and this could be part of the explanation, as longer lifespan allow for larger gaps in the tracks.
-The three series with complex scenes all contain flowers at the edge of the frame that may occasionally be outside the view of the camera (see Fig.
+The three series with complex scenes all contain flowers at the edge of the frame that may occasionally be outside the view of the camera (see fig.
 3) and the high number of flowers increase the risk of flowers temporarily occluding others.
@@
 The optimal range for max distance was relatively narrow (400-600) for the parameters producing the correct number of tracks with the lowest number of mismatches.
-The optimal value for this setting is one that is sufficiently high that it does not force individual flowers into separate tracks because of movement between frames, yet is low enough to separate flowers into individual tracks that would otherwise have been grouped together.
-This explains why the extreme values for this parameter was not included in the highest accuracy tracks.
+The optimal value for this setting is one that is sufficiently high that it does not force individual flowers into separate tracks because of movement between frames, yet is low enough to separate flowers into individual tracks that would otherwise have been grouped together, which explains why the extreme values for this parameter was not included in the highest accuracy tracks.
+
+The best values for running mean were either 1 or 10, the two lowest possible values.
+The parameter may have had an effect in regards to identifying the correct number of tracks, but this is not immediately clear from the results, since it always differed in combination with others.
+In future studies it may be relevant to include values between 0 and 10.
 
@@
 This underlines the relevance of searching for an optimal parameter combination based on ground truth data.
-As with any tracking method, accuracy can only be assessed based on ground truth data [@dendorfer2021; @luiten2020; @milan2016].
 However, the fact that the optimal tracking parameters were overall similar across series strengthens the assumptions that a set of standard values can produce reasonable results for other similar image series with *Dryas* flowers.
 Here, a combination such as 10, 500, 10, for max gap, max distance, and running mean, respectively, would be a good starting point.
-When applying our method in a new setting, we recommend performing a parameter test as demonstrated here on a subset of ground truth tracks to derive the optimal set of parameters.
+When applying our method in a new setting, we recommend performing a parameter test as demonstrated on a subset of ground truth tracks to derive the optimal set of parameters.
 
@@
 These detections would be included in the tracking along with correct detections.
-Some cases of false positives, for example isolated single false positives, could be removed automatically while other cases may require manual quality control.
+Some cases of false positives, for example isolated single false positives, could be filtered automatically while other cases may require manual quality control.
 In general, the performance of a tracking algorithm depends on the quality of input data and it is advisable to optimise detection accuracy before moving to tracking.
-The fact that our method produced perfect tracking results in the most simple settings (fewest flowers) could be taken into consideration when setting up future data collection schemes.
 
@@
 For example, tracks of several objects may have been erroneously merged or a track could contain few instances of another object.
-However, it is reasonable to assume that spatially isolated objects are in general easier to track.
+However, it is a reasonable assumption that spatially isolated objects are generally easier to track.
 We applied the filtering algorithm on tracks produced with a fixed set of tracking parameters as well as a fixed value for the filtering distance.
 As in the case of the tracking parameters, we recommend setting the value for the filtering distance based on either visual examination of results or on tests on a subset of ground truth data when applying the method in a new setting, but for tracking *Dryas* flowers in similar image series, a value of 350 may be a good starting point.
-The presented method avoids weighting track geometries based on number of included points and also weighting the centroid coordinate based on the point coordinates within the track as the track centroid is based on the convex hull of the track points.
-Additionally, it allows for overlap between track geometries as long as the distance between the track centroids is above the threshold.
-We note, however, that other approaches may be more suitable in other settings and applications.
 
@@
 @Estes2018 argued that as ecologists rely on conventional field techniques, the observational scales of ecological studies are generally narrow and may be insufficient to comprehend the ecological phenomena in focus.
-Development and implementation of technological tools for ecological data collection can facilitate ecological studies with larger temporal and spatial scopes [@Pimm2015; @allan2018].
-For processes that require high and continual temporal observation frequency to monitor, such as tracking individuals and their interactions, automation of data collection is critical, especially in remote places where human access is limited.
+Development and implementation of technological tools for ecological data collection and analysis can facilitate ecological studies with larger temporal and spatial scopes [@Pimm2015; @allan2018; @stears2022].
+For processes that require monitoring at high and continual temporal observation frequency, such as tracking individuals and their interactions, automation of data collection is critical, especially in remote places where human access is limited.
 Here, we have demonstrated the potential of image-based monitoring for collecting information on individual-level floral phenology.
@@
 It is clear that images derived from monitoring of permanent plots can hold substantial amounts of ecologically relevant data.
-To derive some variables, for example flower visitation rates, image series collected at high temporal frequency is required.
-As the scale of data increases as well as the number of variables to be extracted from the images increase, automated image analysis becomes critical.
+To derive some variables, for example flower visitation rates of insects, image series high frequency image collection is required.
+As the scale of data as well as the number of variables to be extracted from the images increase, automated image analysis becomes critical.
 Here, methods from deep learning and computer vision are exceedingly relevant.
 For instance, it has been demonstrated that CNNs can facilitate monitoring of pollination events by detecting, identifying, and counting insects [@bjerge2021] and perform fine-grained species classification [@spiesman2021].
-Collectively, these developments coupled with implementation of long-term high-frequency image-based monitoring will allow for collection and processing of ecological data at unprecedented scales [@hoye2021; @besson2022].
-By coupling pollinator detection and classification with floral tracking, visitation rates per flower could be derived and the effects on reproductive success could be investigated [@tiusanen2016; @rafferty2012].
-
-The possibility of collecting individual level floral phenology is relevant for many research questions.
-For example, coupling such information with indicators of seed set could facilitate investigation of the effect of timing and length of flowering and/or abiotic conditions such as frost events on seed set.
+Coupling pollinator detection and classification with floral tracking can facilitate investigation of pollinator visitation rates on flower reproductive success [@rafferty2012; @nagai2022], while information on indicators of seed set could facilitate investigation of the effect of timing and length of flowering and/or abiotic conditions such as frost events on seed set.
 Hereby, the long-standing question of what drives variation in reproductive success of individual flowers could be explored [@bolmgren2015].
-Further, the fact that it was not uncommon in our results that flowering lasted less than a week emphasizes that weekly observations of sample plots can be insufficient for comprehensive description of floral phenology [@Prevey2021].
-
-Pollen limitation is a common factor limiting the reproductive success of flowering plants.
-However, the ecological mechanisms behind pollen limitation is poorly understood, at least partly because it requires information on plant seed set in relation to pollinator visitation rates which is difficult to acquire in a natural setting and to a satisfactory extent.
-For example, pollinator visits can vary greatly in frequency during the flowering season and the degree of pollen limitation can vary even between populations in close vicinity to each other [@larson1999; @ashman2004].
-High frequency observations of flower visits coupled with information on phenology and reproductive success at the level of individuals could unravel the relationship between pollinator visitation and plant reproductive success and the climate sensitivity of plant-pollinator interactions.
-Further, any information on taxonomic or functional grouping could facilitate comparative studies on pollination effectiveness between such groups [@king2013; @foldesi2021].
-Importantly, the variables suggested here could be derived directly from the time-lapse images.
-For example, seed set as an indicator of flower reproductive success could be estimated for individual flowers and flower visitation rates could be quantified by detecting insects in flowers in high temporal resolution image series.
+Owing to the limited resolution of the cameras used in the present study, it is unlikely that detailed information about anther opening or pollen release can be recorded from our images, but with increased image quality, this may be possible.
+Collectively, these developments coupled with implementation of long-term high-frequency image-based monitoring will allow for collection and processing of ecological data at unprecedented scales [@besson2022].
 
@@
 Our results show that it is feasible to collect and analyse individual floral phenology in high detail at logistically challenging sites, which is a significant contribution towards automatic and efficient phenology data collection.
+Further, the fact that it was not uncommon in our results that flowering lasted less than a week emphasizes that weekly observations of sample plots can be insufficient for comprehensive description of floral phenology [@Prevey2021].
 Most importantly, adoption of automated methods for data collection and analysis within ecology, and specifically within for example pollination ecology, will transform the extent and scale of studies and the questions that can be explored.
-
-Owing to the limited resolution of the cameras used in the present study, it is unlikley that detailed information about anther opening or pollen release can be recorded from these images, but with increased image quality, this may be possible.
 
@@
 The authors have no conflict of interest to declare.
-Thank you to Adrian Rosebrock (https://pyimagesearch.com/2018/07/23/simple-object-tracking-with-opencv/, accessed on November 29th 2022) for inspiration for the design of the centroid tracking algorithm.
+Thank you to Adrian Rosebrock (<https://pyimagesearch.com/2018/07/23/simple-object-tracking-with-opencv/>, accessed on November 29th 2022) for inspiration for the design of the centroid tracking algorithm.
 

### 2023-01-20 text/SupportManuscriptDraft.Rmd
Updates to text and suppl.
Submitted revised edition of manuscript

--- 2022-02-02 text/SupportManuscriptDraft.Rmd
@@
 title: |
-  **Supporting Information: Towards individual-based pollination ecology: Automatic tracking of life histories of individual flowers**
+  **Towards individual-based floral biology: Automatic tracking of life histories of individual flowers**
 author:
@@
 
-**Table S1:** Results of the filtering algorithm for filtering distances 5-500 with a stepsize of 50 for all three series.
+# 2
 
-![](../figures/FullFilteringDistanceTest.png){ width=100% }
+## Results of the filtering algorithm for filtering distances (eps) 50-500 with a stepsize of 50 for the three series.
+
+eps = 50: 158 mismatches with 171 tracks kept 
+eps = 100: 103 mismatches with 153 tracks kept 
+eps = 150: 71 mismatches with 130 tracks kept 
+eps = 200: 47 mismatches with 106 tracks kept 
+eps = 250: 30 mismatches with 75 tracks kept 
+eps = 300: 9 mismatches with 51 tracks kept 
+eps = 350: 0 mismatches with 28 tracks kept 
+eps = 400: 0 mismatches with 18 tracks kept 
+eps = 450: 0 mismatches with 16 tracks kept 
+eps = 500: 0 mismatches with 11 tracks kept
+
+#![](../figures/FullFilteringDistanceTest.png){width="100%"}

### 2023-01-20 src/Simulation_plotting.py

@@ -124,10 +124,10 @@ framerate['objectLetter'] = framerate['object'].map(lambda x: tran[x])
 f, (ax1, ax2) = plt.subplots(2, 1, sharex=False, height_ratios=[4, 1])
 plt.subplots_adjust(hspace=0.25)
 
-ax1.set_xlim([0, 210])
-ax1.set_ylim([0, 210])
-ax1.xaxis.set_ticks([0, 200])
-ax1.yaxis.set_ticks([0, 200])
+ax1.set_xlim([0, 250])
+ax1.set_ylim([0, 250])
+ax1.xaxis.set_ticks([0, 250])
+ax1.yaxis.set_ticks([0, 250])
 ax1.set_xlabel("X", labelpad=-12)
 ax1.set_ylabel("Y", labelpad=-20)
 ax2.margins(y = 0.2)

### 2023-01-20 src/runExperiments.py

@@ -7,176 +7,36 @@ from evaluate import evaluator
 
 # Import global packages
 import pandas as pd
+import time
+import os
+import re
 
-
-#import os
-#import time
-#import os
-#from scipy.spatial import distance as dist
-#import os
-#import re
-#import numpy as np
 br = '\n'
 
 
 ###################
 
-### Function for importing data
-# def impdata(camID):
-#     if camID == "NARS-04":
-#         gt = r".\data\annotations\2020_05_17_NorwayAnnotations_NARS-04_IndividualAnnotations_FRCNN_Metrics.csv"
-#     if camID == "NARS-13":
-#         gt = r".\data\annotations\2020_04_30_NorwayAnnotations_NARS-13_IndividualAnnotations_FRCNN_Metrics.csv"
-#     if camID == "THUL-01":
-#         gt = r".\data\annotations\2020_05_15_NorwayAnnotations_THUL-01_IndividualAnnotations_FRCNN_Metrics.csv"
-#     if camID == "NYAA-04":
-#         gt = r".\data\annotations\2021_12_13_NorwayAnnotations_NYAA-04_IndividualAnnotations_FRCNN_Metrics.csv"
-#     if camID == "NARS-17":
-#         gt = r".\data\annotations\2021_12_30_NorwayAnnotations_NARS-17_IndividualAnnotations_FRCNN_Metrics.csv"
-#     if camID == "debug":
-#         gt = r".\data\annotations\debugGT.csv"
-#     return gt
+## Function for importing data
+def impdata(camID):
+    if camID == "NARS-04":
+        gt = r".\data\annotations\2020_05_17_NorwayAnnotations_NARS-04_IndividualAnnotations_FRCNN_Metrics.csv"
+    if camID == "NARS-13":
+        gt = r".\data\annotations\2020_04_30_NorwayAnnotations_NARS-13_IndividualAnnotations_FRCNN_Metrics.csv"
+    if camID == "THUL-01":
+        gt = r".\data\annotations\2020_05_15_NorwayAnnotations_THUL-01_IndividualAnnotations_FRCNN_Metrics.csv"
+    if camID == "NYAA-04":
+        gt = r".\data\annotations\2021_12_13_NorwayAnnotations_NYAA-04_IndividualAnnotations_FRCNN_Metrics.csv"
+    if camID == "NARS-17":
+        gt = r".\data\annotations\2021_12_30_NorwayAnnotations_NARS-17_IndividualAnnotations_FRCNN_Metrics.csv"
+    if camID == "debug":
[…]
